### 2017-09-06 docs/build/latex/cis_interface.tex
rebuilt docs

%% Generated by Sphinx.
\def\sphinxdocclass{report}
\documentclass[letterpaper,10pt,english]{sphinxmanual}
\ifdefined\pdfpxdimen
   \let\sphinxpxdimen\pdfpxdimen\else\newdimen\sphinxpxdimen
\fi \sphinxpxdimen=.75bp\relax

\usepackage[utf8]{inputenc}
\ifdefined\DeclareUnicodeCharacter
 \ifdefined\DeclareUnicodeCharacterAsOptional\else
  \DeclareUnicodeCharacter{00A0}{\nobreakspace}
\fi\fi
\usepackage{cmap}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amstext}
\usepackage{babel}
\usepackage{times}
\usepackage[Bjarne]{fncychap}
\usepackage{longtable}
\usepackage{sphinx}

\usepackage{geometry}
\usepackage{multirow}
\usepackage{eqparbox}

% Include hyperref last.
\usepackage{hyperref}
% Fix anchor placement for figures with captions.
\usepackage{hypcap}% it must be loaded after hyperref.
% Set up styles of URL: it should be placed after hyperref.
\urlstyle{same}
\addto\captionsenglish{\renewcommand{\contentsname}{Contents:}}

\addto\captionsenglish{\renewcommand{\figurename}{Fig.}}
\addto\captionsenglish{\renewcommand{\tablename}{Table}}
\addto\captionsenglish{\renewcommand{\literalblockname}{Listing}}

\addto\extrasenglish{\def\pageautorefname{page}}

\setcounter{tocdepth}{1}

\title{cis\_interface Documentation}
\date{Sep 06, 2017}
\release{0.0.0}
\author{Meagan Lang, David Raila}
\newcommand{\sphinxlogo}{}
\renewcommand{\releasename}{Release}
\makeindex

\begin{document}

\maketitle
\sphinxtableofcontents
\phantomsection\label{\detokenize{index::doc}}

\chapter{Model Integration and Communications Architecture}
\label{\detokenize{includeme:model-integration-and-communications-architecture}}\label{\detokenize{includeme::doc}}\label{\detokenize{includeme:welcome-to-cis-interface-s-documentation}}

\section{Overview}
\label{\detokenize{includeme:overview}}
The CiS model system provides integration and communications support for
model providers. Modelers add simple communications interfaces within
their model code, and provide a simple declarative specification that
informs the system of the names and types of inputs and outputs of the
model and the commands required to run the model.

The system uses the specification to configure the communications
channels and bind them into the model. The complexity of the particular
communications system is managed by a model driver that performns
communication setup, binds the communications to a simple interface
within the model, and manages the model's exection. The complexities of
model registration and discovery, as well as the complexities of setup
and management of the communications system are handled by the model
driver under direction of the model specification, freeing the model
programmer from implementing communications protocols of any particular
messaging infrastructure. The model programmer is provided with a very
simple and easy to use abstract send/receive message interface using
basic IPC mechanisms available in that binds in virtually every language
and platform.

The model communications interface consists of 2 pairs of very simple
functions - one pair to create channels and one pair to accomplish data
tranfer. Each pair represents a read-side and a write-side.

\begin{sphinxVerbatim}[commandchars=\\\{\}]
\PYG{n}{psi\PYGZus{}input}\PYG{p}{(}\PYG{n}{name}\PYG{p}{)}  \PYG{o}{\PYGZhy{}} \PYG{n}{create} \PYG{n}{a} \PYG{n}{named} \PYG{n+nb}{input} \PYG{n}{channel}
\PYG{n}{psi\PYGZus{}output}\PYG{p}{(}\PYG{n}{name}\PYG{p}{)} \PYG{o}{\PYGZhy{}} \PYG{n}{create} \PYG{n}{a} \PYG{n}{named} \PYG{n}{output} \PYG{n}{channel}

\PYG{n}{psi\PYGZus{}send}\PYG{p}{(}\PYG{n}{name}\PYG{p}{,} \PYG{n}{data}\PYG{p}{)}   \PYG{o}{\PYGZhy{}} \PYG{n}{send} \PYG{n}{a} \PYG{n}{message} \PYG{n}{through} \PYG{n}{the} \PYG{n}{managed} \PYG{n}{channel}
\PYG{n}{data} \PYG{o}{=} \PYG{n}{psi\PYGZus{}recv}\PYG{p}{(}\PYG{n}{name}\PYG{p}{)}  \PYG{o}{\PYGZhy{}} \PYG{n}{receive} \PYG{n}{message} \PYG{k+kn}{from} \PYG{n+nn}{channel} \PYG{o}{\PYGZhy{}} \PYG{n}{block} \PYG{k}{if} \PYG{n}{empty} \PYG{n}{channel}
\end{sphinxVerbatim}

C-style is shown, but as these interfaces are thin wrappers around
System V IPC message queues, equivalent thin-wrappers are implemented
easily in almost any language. Models may create and use as many
channels as necessary, depending on the number of interfaces that they
require.

The only other model requirement is to provide a YAML specification file
that declares the channels that the model uses, and the system command
required to start the model executing.

\begin{sphinxVerbatim}[commandchars=\\\{\}]
\PYG{n}{model}\PYG{p}{:}
  \PYG{n}{name}\PYG{p}{:} \PYG{n}{hellopython}
  \PYG{n}{driver}\PYG{p}{:} \PYG{n}{ModelDriver}
  \PYG{n}{args}\PYG{p}{:} \PYG{p}{[} \PYG{l+s+s2}{\PYGZdq{}}\PYG{l+s+s2}{python}\PYG{l+s+s2}{\PYGZdq{}}\PYG{p}{,} \PYG{l+s+s2}{\PYGZdq{}}\PYG{l+s+s2}{hellopython.py}\PYG{l+s+s2}{\PYGZdq{}} \PYG{p}{]}

  \PYG{n}{inputs}\PYG{p}{:}
    \PYG{o}{\PYGZhy{}} \PYG{n}{name}\PYG{p}{:} \PYG{n}{inputFile}
      \PYG{n}{driver}\PYG{p}{:} \PYG{n}{FileInputDriver}
      \PYG{n}{args}\PYG{p}{:} \PYG{n}{inputdata}\PYG{o}{/}\PYG{n+nb}{input}\PYG{o}{.}\PYG{n}{txt}
      \PYG{n}{onexit}\PYG{p}{:} \PYG{n}{delete}    \PYG{c+c1}{\PYGZsh{} delete the queue when the model exits}
    \PYG{o}{\PYGZhy{}} \PYG{n}{name}\PYG{p}{:} \PYG{n}{fooInputChannel}
      \PYG{n}{driver}\PYG{p}{:} \PYG{n}{RMQInputDriver}
      \PYG{n}{args}\PYG{p}{:} \PYG{n}{fooqueue}
      \PYG{n}{onexit}\PYG{p}{:} \PYG{n}{delete}    \PYG{c+c1}{\PYGZsh{} delete the queue when the model exits}

  \PYG{n}{outputs}\PYG{p}{:}
    \PYG{o}{\PYGZhy{}} \PYG{n}{name}\PYG{p}{:} \PYG{n}{fooOutput}
      \PYG{n}{driver}\PYG{p}{:} \PYG{n}{RMQOutputDriver}
\end{sphinxVerbatim}

This example tells the system to run your model as if you typed \sphinxquotedblleft{}python
hellopython.py\sphinxquotedblright{} and to connect the file inputdata/input.txt to the
psi\_input named `inputFile' in your model, so when you recv() on it,
you'll get the file contents. Likewise when you send(data\_to\_send) to
the channel you made in your code with psi\_output(`fooOutput') the data
will go to the message queue names FooOutput.

You may have multiple inputs and multiple outputs as you need. If you
send to a RMQOutput named `fred' your another model can use fred as a
RMQInput to read from and get your messages. By keeping multiple yml
files around with different inputs and outputs but the same model stanza
you can test with different data inputs and outputs conveniently without
changing the model code.

The finel part is the model wrapper, PsiRun.py - to which you pass your
yml file. The model wrapper reads the specification, arranges for your
inputs and outputs to come and go as you've specified, and then calls
your model with the arguments as specified in the model stanza. The
model runs, using the send and recv calls as needed, and when the model
exits, the communications get cleaned up bu the wrapper. The exmaple
hellopython.py under examples is run like this (in examples/python):

\begin{sphinxVerbatim}[commandchars=\\\{\}]
\PYG{n}{python} \PYG{o}{.}\PYG{o}{.}\PYG{o}{/}\PYG{o}{.}\PYG{o}{.}\PYG{o}{/}\PYG{n}{interface}\PYG{o}{/}\PYG{n}{PsiRun}\PYG{o}{.}\PYG{n}{py} \PYG{n}{hellopython}\PYG{o}{.}\PYG{n}{yml}
\end{sphinxVerbatim}

The code in hellopython.py uses the channels:

\begin{sphinxVerbatim}[commandchars=\\\{\}]
\PYG{n}{inf} \PYG{o}{=} \PYG{n}{PsiInput}\PYG{p}{(}\PYG{l+s+s1}{\PYGZsq{}}\PYG{l+s+s1}{infile}\PYG{l+s+s1}{\PYGZsq{}}\PYG{p}{)}
\PYG{n}{outf} \PYG{o}{=} \PYG{n}{PsiOutput}\PYG{p}{(}\PYG{l+s+s1}{\PYGZsq{}}\PYG{l+s+s1}{outfile}\PYG{l+s+s1}{\PYGZsq{}}\PYG{p}{)}
\PYG{n}{inq} \PYG{o}{=} \PYG{n}{PsiInput}\PYG{p}{(}\PYG{l+s+s1}{\PYGZsq{}}\PYG{l+s+s1}{helloQueue}\PYG{l+s+s1}{\PYGZsq{}}\PYG{p}{)}
\PYG{n}{outq} \PYG{o}{=} \PYG{n}{PsiOutput}\PYG{p}{(}\PYG{l+s+s1}{\PYGZsq{}}\PYG{l+s+s1}{helloQueue}\PYG{l+s+s1}{\PYGZsq{}}\PYG{p}{)}

\PYG{n}{buf} \PYG{o}{=} \PYG{n}{inf}\PYG{o}{.}\PYG{n}{recv}\PYG{p}{(}\PYG{p}{)}
\PYG{n}{outq}\PYG{o}{.}\PYG{n}{send}\PYG{p}{(}\PYG{n}{buf}\PYG{p}{)}
\PYG{n}{buf} \PYG{o}{=} \PYG{n}{inq}\PYG{o}{.}\PYG{n}{recv}\PYG{p}{(}\PYG{p}{)}
\PYG{n}{outf}\PYG{o}{.}\PYG{n}{send}\PYG{p}{(}\PYG{n}{buf}\PYG{p}{)}
\end{sphinxVerbatim}

The yaml file says to run the model, and connects the inputs and
outputs:

\begin{sphinxVerbatim}[commandchars=\\\{\}]
\PYG{n}{model}\PYG{p}{:}
  \PYG{n}{name}\PYG{p}{:} \PYG{n}{hellopython}
  \PYG{n}{driver}\PYG{p}{:} \PYG{n}{ModelDriver}
  \PYG{n}{args}\PYG{p}{:} \PYG{p}{[} \PYG{l+s+s2}{\PYGZdq{}}\PYG{l+s+s2}{python}\PYG{l+s+s2}{\PYGZdq{}}\PYG{p}{,} \PYG{l+s+s2}{\PYGZdq{}}\PYG{l+s+s2}{hellopython.py}\PYG{l+s+s2}{\PYGZdq{}} \PYG{p}{]}

  \PYG{n}{inputs}\PYG{p}{:}
    \PYG{o}{\PYGZhy{}} \PYG{n}{name}\PYG{p}{:} \PYG{n}{infile}
      \PYG{n}{driver}\PYG{p}{:} \PYG{n}{FileInputDriver}
      \PYG{n}{filename}\PYG{p}{:} \PYG{n+nb}{input}\PYG{o}{.}\PYG{n}{txt}
    \PYG{o}{\PYGZhy{}} \PYG{n}{name}\PYG{p}{:} \PYG{n}{helloQueue}
      \PYG{n}{driver}\PYG{p}{:} \PYG{n}{RMQInputDriver}

  \PYG{n}{outputs}\PYG{p}{:}
    \PYG{o}{\PYGZhy{}} \PYG{n}{name}\PYG{p}{:} \PYG{n}{helloQueue}
      \PYG{n}{driver}\PYG{p}{:} \PYG{n}{RMQOutputDriver}
    \PYG{o}{\PYGZhy{}} \PYG{n}{name}\PYG{p}{:} \PYG{n}{outFile}
      \PYG{n}{driver}\PYG{p}{:} \PYG{n}{FileOutputDriver}
      \PYG{n}{filename}\PYG{p}{:} \PYG{n}{output}\PYG{o}{.}\PYG{n}{txt}
\end{sphinxVerbatim}

\section{Installation notes}
\label{\detokenize{includeme:installation-notes}}\begin{itemize}
\item {} 
Matlab

\item {} 
Matlab uses the python interface directly

\item {} 
Remove the matlab crypto and ssl libraries and install the python
engine sudo -s rm -f
./R2015b/toolbox/compiler\_sdk/mps\_clients/c/glnxa64/lib/libcrypto.so.1
rm -f
./R2015b/toolbox/compiler\_sdk/mps\_clients/c/glnxa64/lib/libcrypto.so
rm -f
./R2015b/toolbox/compiler\_sdk/mps\_clients/c/glnxa64/lib/libcrypto.so.1.0.0
rm -f ./R2015b/bin/glnxa64/libcrypto.so.1 rm -f
./R2015b/bin/glnxa64/libcrypto.so.1.0.0 rm -f
./R2015b/toolbox/compiler\_sdk/mps\_clients/c/glnxa64/lib/libssl.so.1
rm -f
./R2015b/toolbox/compiler\_sdk/mps\_clients/c/glnxa64/lib/libssl.so.1.0.0
rm -f
./R2015b/toolbox/compiler\_sdk/mps\_clients/c/glnxa64/lib/libssl.so
rm -f
./R2015b/sys/jxbrowser/glnxa64/xulrunner/xulrunner-linux-64/libssl3.so
rm -f ./R2015b/bin/glnxa64/libssl.so.1 rm -f
./R2015b/bin/glnxa64/libssl.so.1.0.0 cd
/usr/local/MATLAB/R2015b/extern/engines/python python setup.py

\end{itemize}

\section{Contents}
\label{\detokenize{includeme:contents}}\begin{itemize}
\item {} 
\sphinxstylestrong{build} - model build help and container builders

\item {} 
docker - build containers (in progress)

\item {} 
\sphinxstylestrong{edge-example} - full example in python, matlab, C(++)

\item {} 
models - models and specifications

\item {} 
app

\item {} 
\sphinxstylestrong{lib} - Common services and drivers

\item {} 
drivers - inputs, outputs, messaging, model-wrapper

\item {} 
yml - yaml specification support

\item {} 
controller - cli PSi system setup/run

\end{itemize}

\section{Functional Architecture}
\label{\detokenize{includeme:functional-architecture}}\begin{itemize}
\item {} 
Models

\item {} 
Provide callable functions (ml, c, python)
\begin{itemize}
\item {} 
Function(s) take an input, returns an output

\item {} 
A function to set optional parameters

\end{itemize}

\item {} 
Provide a specification - yaml file
\begin{itemize}
\item {} 
Describes the model, inputs, outputs, function names, files to
use, filenames to store output

\end{itemize}

\item {} 
Driver

\item {} 
A driver is a `wrapper' or adapter that encapsulates other objects or
functions, perhaps in other languages or platforms, into a callable
python object - for example the MatlabCaller encapsulates the Matlab
runtime and callable function in a clean python object.

\item {} 
Controller - A python cli program that runs and feeds the models

\item {} 
Reads the specifications

\item {} 
Setup and connect the queues

\item {} 
Setup the models inside model drivers

\item {} 
Setup the input and output drivers

\item {} 
Sets all drivers to running
\begin{itemize}
\item {} 
Input drivers send data into queues

\item {} 
Model drivers receive inputs, call model processing functions,
send outputs to downstream queue drivers and/or output drivers

\end{itemize}

\end{itemize}

\section{Advantages}
\label{\detokenize{includeme:advantages}}\begin{itemize}
\item {} 
The demands on model providers is greatly simplified

\item {} 
Model contains only domain-specific model code - focused

\item {} 
No coding based on unfamilar techniques, languages, etc.
\begin{itemize}
\item {} 
System code, protocols, services, and languages outside of a
modelers area of expertise

\end{itemize}

\item {} 
The specification is much simpler to create and maintain than code

\item {} 
Declare what the model wants, take inputs, produce outputs and let
the system do the rest

\item {} 
Single system modules and drivers are easier to maintain and evolve

\item {} 
Messaging, data-handling, and system run-time facilities become
shared and isolated, bounded in their impact

\item {} 
Easier to understand, easier to maintain
\begin{itemize}
\item {} 
Fixing a shared driver is preferrable to system code embedded in
multiple models

\end{itemize}

\end{itemize}

\section{Development Notes: Installation and test code fragments}
\label{\detokenize{includeme:development-notes-installation-and-test-code-fragments}}
Python to matlabs setup: cd
/usr/local/MATLAB/R2015b/extern/engines/python sudo python setup.py
install.

\subsection{Python protocols Matlab}
\label{\detokenize{includeme:python-protocols-matlab}}
from psimsg import * import matlab.engine eng =
matlab.engine.connect\_matlab()

\subsection{Matlab to Python}
\label{\detokenize{includeme:matlab-to-python}}
Matlab only seems limited to storing basic data (ints/floats/arrays),
unable to receive objects from python for storage/calling P =
py.sys.path; append(P, pwd); import psimsg.*

\chapter{Installation}
\label{\detokenize{install:installation}}\label{\detokenize{install::doc}}

\section{Basic Installation}
\label{\detokenize{install:basic-installation}}
..todo:: Links to PyPI release

{\color{red}\bfseries{}\textbar{}cis\_interface\textbar{}} can be installed from either PyPI using \sphinxcode{pip}:

\begin{sphinxVerbatim}[commandchars=\\\{\}]
\PYGZdl{} pip install cis\PYGZus{}interface
\end{sphinxVerbatim}

or by cloning the \sphinxhref{https://git-scm.com/}{Git} repository on \sphinxhref{https://github.com/cropsinsilico/cis\_interface}{Github}:

\begin{sphinxVerbatim}[commandchars=\\\{\}]
\PYGZdl{} git clone https://github.com/cropsinsilico/cis\PYGZus{}interface.git
\end{sphinxVerbatim}

and then building the distribution.:

\begin{sphinxVerbatim}[commandchars=\\\{\}]
\PYGZdl{} cd cis\PYGZus{}interface
\PYGZdl{} python setup.py install
\end{sphinxVerbatim}

If you do not have admin privileges on the target machine, \sphinxcode{-{-}user} can be added to either the \sphinxcode{pip} or \sphinxcode{setup.py} installation commands.

\section{Additional Steps for Matlab Models}
\label{\detokenize{install:additional-steps-for-matlab-models}}
To run Matlab models, you will also need to install the Matlab engine for
Python. This requires that you have an existing Matlab installation and license.

Instructions for installing the Matlab engine as a python package can be found
\sphinxhref{https://www.mathworks.com/help/matlab/matlab\_external/install-the-matlab-engine-for-python.html}{\textless{}here}.

\section{Additional Steps for RabbitMQ Message Passing}
\label{\detokenize{install:additional-steps-for-rabbitmq-message-passing}}

\chapter{Code}
\label{\detokenize{code:code}}\label{\detokenize{code::doc}}\phantomsection\label{\detokenize{code:module-cis_interface}}\index{cis\_interface (module)}
This package provides a framework for integrating models across languages
such that they can be run simultaneously, passing input back and forth.
\index{interface (in module cis\_interface)}

\begin{fulllineitems}
\phantomsection\label{\detokenize{code:cis_interface.interface}}\pysigline{\sphinxcode{cis\_interface.}\sphinxbfcode{interface}}
alias of {\hyperref[\detokenize{code:cis_interface.interface}]{\sphinxcrossref{\sphinxcode{cis\_interface.interface}}}}

\end{fulllineitems}

\index{drivers (in module cis\_interface)}

\begin{fulllineitems}
\phantomsection\label{\detokenize{code:cis_interface.drivers}}\pysigline{\sphinxcode{cis\_interface.}\sphinxbfcode{drivers}}
alias of {\hyperref[\detokenize{code:cis_interface.drivers}]{\sphinxcrossref{\sphinxcode{cis\_interface.drivers}}}}

\end{fulllineitems}

\index{dataio (in module cis\_interface)}

\begin{fulllineitems}
\phantomsection\label{\detokenize{code:cis_interface.dataio}}\pysigline{\sphinxcode{cis\_interface.}\sphinxbfcode{dataio}}
alias of {\hyperref[\detokenize{code:cis_interface.dataio}]{\sphinxcrossref{\sphinxcode{cis\_interface.dataio}}}}

\end{fulllineitems}

\chapter{Indices and tables}
\label{\detokenize{index:indices-and-tables}}\begin{itemize}
\item {} 
\DUrole{xref,std,std-ref}{genindex}

\item {} 
\DUrole{xref,std,std-ref}{modindex}

\item {} 
\DUrole{xref,std,std-ref}{search}

\end{itemize}

\renewcommand{\indexname}{Python Module Index}
\begin{sphinxtheindex}
\def\bigletter#1{{\Large\sffamily#1}\nopagebreak\vspace{1mm}}
\bigletter{c}
\item {\sphinxstyleindexentry{cis\_interface}}\sphinxstyleindexpageref{code:\detokenize{module-cis_interface}}
\end{sphinxtheindex}

\renewcommand{\indexname}{Index}
\printindex
\end{document}

### 2017-09-06 cis_interface/__init__.py
rebuilt docs

@@ -1,30 +1,0 @@
-r"""This package provides a framework for integrating models across languages
-such that they can be run simultaneously, passing input back and forth."""
-import os
-import backwards
-import config
-import tools
-import interface
-import drivers
-import dataio
-import tests
-import examples
-
-
-# Set paths so that c headers are located
-# TODO: Only the CIS_INCLUDE environment variable should be used
-cis_base = os.path.dirname(__file__)
-cis_include = os.path.join(cis_base, 'interface')
-os.environ['CIS_BASE'] = cis_base
-os.environ['CIS_INCLUDE'] = cis_include
-path = os.environ.get('PATH', cis_include)
-cpath = os.environ.get('CPATH', cis_include)
-if cis_include not in path:
-    os.environ['PATH'] = cis_include + ':' + path
-if cis_include not in cpath:
-    os.environ['CPATH'] = cis_include + ':' + cpath
-
-
-__all__ = ['backwards', 'config', 'tools',
-           'interface', 'drivers', 'dataio',
-           'tests', 'examples']

### 2017-09-06 cis_interface/backwards.py

@@ -1,14 +1,0 @@
-r"""This module allows for backward compatibility."""
-import sys
-PY2 = (sys.version_info[0] == 2)
-if PY2:  # pragma: Python 2
-    import cPickle as pickle
-    import ConfigParser as configparser
-    import StringIO as sio
-else:  # pragma: Python 3
-    import pickle
-    import configparser
-    import sio
-
-    
-__all__ = ['pickle', 'configparser', 'sio']

### 2017-09-06 cis_interface/dataio/AsciiFile.h

@@ -1,84 +1,0 @@
-#include <string.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <stdarg.h>
-
-#define LINE_SIZE_MAX 1024*64
-
-typedef struct AsciiFile {
-  const char *filepath;
-  char io_mode[64];
-  char comment[64];
-  char newline[64];
-  FILE *fd;
-} AsciiFile;
-
-static inline
-int af_is_open(const AsciiFile t) {
-  if (t.fd == NULL)
-    return 0;
-  else
-    return 1;
-};
-
-static inline
-int af_open(AsciiFile *t) {
-  int ret = -1;
-  if (af_is_open(*t) == 0) {
-    (*t).fd = fopen((*t).filepath, (*t).io_mode);
-    if ((*t).fd != NULL)
-      ret = 0;
-  }
-  return ret;
-};
-
-static inline
-void af_close(AsciiFile *t) {
-  if (af_is_open(*t) == 1) {
-    fclose((*t).fd);
-    (*t).fd = NULL;
-  }
-};
-
-static inline
-int af_is_comment(AsciiFile t, char *line) {
-  if (strncmp(line, t.comment, strlen(t.comment)) == 0)
-    return 1;
-  else
-    return 0;
-};
-
-static inline
-int af_readline_full(const AsciiFile t, char **line, size_t *n) {
-  if (af_is_open(t) == 1) {
-    return getline(line, n, t.fd);
-  }
-  return -1;
-};
-
-static inline
-int af_writeline_full(const AsciiFile t, const char *line) {
-  if (af_is_open(t) == 1)
-    return fwrite(line, 1, strlen(line), t.fd);
-  return -1;
-};
-
-static inline
-AsciiFile ascii_file(const char *filepath, const char *io_mode,
-		     char *comment, char *newline) {
-  AsciiFile t;
-  t.fd = NULL;
-  t.filepath = filepath;
-  strcpy(t.io_mode, io_mode);
-  // Set defaults for optional parameters
-  if (comment == NULL)
-    strcpy(t.comment, "# ");
-  else
-    strcpy(t.comment, comment);
-  if (newline == NULL)
-    strcpy(t.newline, "\n");
-  else
-    strcpy(t.newline, newline);
-  return t;
-};
-

### 2017-09-06 cis_interface/dataio/AsciiFile.py

@@ -1,135 +1,0 @@
-import os
-
-
-_default_args = {'comment': '#',
-                 'newline': '\n'}
-
-
-class AsciiFile(object):
-    def __init__(self, filepath, io_mode, **kwargs):
-        r"""Class for reading/writing an ASCII file.
-
-        Args:
-            filepath (str): Full path to the file that should be read from
-                or written to.
-            io_mode (str): Mode that should be used to open the file. Valid
-                values include 'r', 'w', and None. None can be used to
-                indicate an in memory table that will not be read from or
-                written to a file.
-            comment (str, optional): String that should be used to identify
-                comments. Defaults to '#'.
-            newline (str, optional): String that should be used to identify
-                the end of a line. Defaults to '\n'.
-
-        Raises:
-            TypeError: If filepath is not a string.
-            ValueError: If io_mode is not one of the allowed values.
-            ValueError: If filepath is not a valid path and io_mode is 'r'.
-
-        """
-        if not isinstance(filepath, str):
-            raise TypeError('File path must be provided as a string.')
-        self.filepath = os.path.abspath(filepath)
-        if io_mode not in ['r', 'w', None]:
-            raise ValueError("IO specifier must be 'r' or 'w'.")
-        self.io_mode = io_mode
-        if self.io_mode == 'r' and not os.path.isfile(filepath):
-            raise ValueError("File does not exist.")
-        for k, v in kwargs.items():
-            setattr(self, k, v)
-        for k, v in _default_args.items():
-            if not hasattr(self, k):
-                setattr(self, k, v)
-        self.fd = None
-
-    @property
-    def is_open(self):
-        r"""Returns True if the file descriptor is open."""
-        return (self.fd is not None)
-
-    def open(self):
-        r"""Open the associated file descriptor if it is not already open."""
-        if self.io_mode is None:
-            raise Exception("Cannot open in memory table.")
-        if not self.is_open:
-            self.fd = open(self.filepath, self.io_mode)
-
-    def close(self):
-        r"""Close the associated file descriptor if it is open."""
-        if self.is_open:
-            self.fd.close()
-            self.fd = None
-
-    def readline(self):
-        r"""Continue reading lines until a valid line (uncommented) is
-        encountered
-        
-        Returns:
-            tuple (bool, str): End of file flag and the line that was read
-                (None if the end of file was encountered).
-
-        """
-        eof, line = False, None
-        while (not eof) and (line is None):
-            eof, line = self.readline_full()
-        if eof:
-            line = None
-        return eof, line
-
[…]

### 2017-09-06 cis_interface/dataio/AsciiTable.py

@@ -1,689 +1,0 @@
-import numpy as np
-from cis_interface.interface.scanf import scanf
-from AsciiFile import AsciiFile
-try:
-    from astropy.io import ascii as apy_ascii
-    from astropy.table import Table as apy_Table
-    _use_astropy = True
-except:  # pragma: no cover
-    apy_ascii, apy_Table = None, None
-    print("astropy is not installed, reading/writing as an array will be " +
-          "disabled. astropy can be installed using 'pip install astropy'.")
-    _use_astropy = False
-
-
-_default_args = {'column': '\t'}
-
-
-def nptype2cformat(nptype):
-    r"""Convert a numpy data type to a c format string.
-
-    Args:
-        nptype (str or numpy.dtype): Numpy data type that should be converted.
-
-    Returns:
-        str: Corresponding c format specification string.
-
-    Raises:
-        TypeError: If nptype is not a string or numpy.dtype.
-        ValueError: If a matching format string cannot be determined.
-
-    """
-    if isinstance(nptype, np.dtype):
-        t = nptype
-    elif isinstance(nptype, str):
-        t = np.dtype(nptype)
-    else:
-        raise TypeError("Input must be a string or a numpy.dtype")
-    if t in [np.dtype(x) for x in ["float_", "float16", "float32", "float64"]]:
-        cfmt = "%g"  # Ensures readability
-    elif t == np.dtype("int8"):
-        cfmt = "%hhd"
-    elif t == np.dtype("short"):
-        cfmt = "%hd"
-    elif t == np.dtype("intc"):
-        cfmt = "%d"
-    elif t == np.dtype("int_"):
-        cfmt = "%ld"
-    elif t == np.dtype("longlong"):  # pragma: no cover
-        # If it is different than C long
-        cfmt = "%lld"
-    elif t == np.dtype("uint8"):
-        cfmt = "%hhu"
-    elif t == np.dtype("ushort"):
-        cfmt = "%hu"
-    elif t == np.dtype("uintc"):
-        cfmt = "%u"
-    elif t == np.dtype("uint64"):  # Platform dependent
-        cfmt = "%lu"
-    elif t == np.dtype("ulonglong"):  # pragma: no cover
-        cfmt = "%llu"
-    elif np.issubdtype(t, np.dtype("S")):
-        # cfmt = '%s'
-        if t.itemsize is 0:
-            cfmt = '%s'
-        else:
-            cfmt = "%" + str(t.itemsize) + "s"
-    else:
[…]

### 2017-09-06 cis_interface/dataio/__init__.py

@@ -1,7 +1,0 @@
-r"""Routines for standardized I/O."""
-
-import AsciiFile
-import AsciiTable
-import tests
-
-__all__ = ['AsciiFile', 'AsciiTable', 'tests']

### 2017-09-06 cis_interface/dataio/tests/test_AsciiFile.py

@@ -1,133 +1,0 @@
-import os
-import tempfile
-from nose.tools import assert_raises, assert_equal
-from cis_interface.dataio import AsciiFile
-from cis_interface.tests import data
-
-
-input_file = data['txt']
-output_dir = tempfile.gettempdir()
-output_file = os.path.join(output_dir, os.path.basename(input_file))
-
-input_ncomments = 1
-input_nlines = 3
-output_ncomments = 0
-output_nlines = input_nlines
-
-mode_list = ['r', 'w', None]
-mode2file = {'r': input_file,
-             'w': output_file,
-             None: 'null'}
-
-
-def test_AsciiFile():
-    for mode in mode_list:
-        AsciiFile.AsciiFile(mode2file[mode], mode)
-    assert_raises(TypeError, AsciiFile.AsciiFile, 0, 'r')
-    assert_raises(ValueError, AsciiFile.AsciiFile, input_file, 0)
-    assert_raises(ValueError, AsciiFile.AsciiFile, 'null', 'r')
-
-
-def test_AsciiFile_open_close():
-    for mode in mode_list:
-        AF = AsciiFile.AsciiFile(mode2file[mode], mode)
-        assert(not AF.is_open)
-        if mode is None:
-            assert_raises(Exception, AF.open)
-        else:
-            AF.open()
-            assert(AF.is_open)
-            AF.close()
-        assert(not AF.is_open)
-
-        
-def test_AsciiFile_line_full():
-    AF_in = AsciiFile.AsciiFile(input_file, 'r')
-    AF_out = AsciiFile.AsciiFile(output_file, 'w')
-    # Read/write before open returns None
-    eof, line = AF_in.readline_full()
-    AF_in.writeline_full(line)
-    assert(eof)
-    assert_equal(line, None)
-    # Read/write all lines
-    AF_in.open()
-    AF_out.open()
-    count_lines = 0
-    count_comments = 0
-    assert_raises(TypeError, AF_in.writeline_full, 0)
-    eof, line = False, None
-    while not eof:
-        eof, line = AF_in.readline_full()
-        if not eof:
-            if line is None:
-                count_comments += 1
-            else:
-                AF_out.writeline_full(line)
-                count_lines += 1
-    AF_in.close()
-    AF_out.close()
-    assert_equal(count_lines, input_nlines)
-    assert_equal(count_comments, input_ncomments)
-    # Read output file to make sure it has lines
[…]

### 2017-09-06 cis_interface/dataio/tests/test_AsciiTable.py

@@ -1,394 +1,0 @@
-import os
-import numpy as np
-import tempfile
-from nose.tools import assert_raises, assert_equal
-from cis_interface.dataio import AsciiTable
-from cis_interface.tests import data
-
-
-input_file = data['table']
-output_dir = tempfile.gettempdir()
-output_file = os.path.join(output_dir, os.path.basename(input_file))
-
-ncols = 3
-nrows = 3
-input_ncomments = 2  # Names & formats
-input_nlines = nrows
-output_ncomments = 2  # Just formats
-output_nlines = nrows
-format_str = "%5s\t%d\t%f\n"
-column_names = ["name", "number", "value"]
-
-
-mode_list = ['r', 'w', None]
-mode2file = {'r': input_file,
-             'w': output_file,
-             None: 'null'}
-mode2kws = {'r': {},
-            'w': {'format_str': format_str, 'column_names': column_names},
-            None: {'format_str': format_str, 'column_names': column_names}}
-
-unsupported_nptype = ['complex_', 'complex64', 'complex128', 'bool_']
-map_nptype2cformat = [
-    (['float_', 'float16', 'float32', 'float64'], '%g'),
-    # (['int8', 'short', 'intc', 'int_', 'longlong'], '%d'),
-    # (['uint8', 'ushort', 'uintc', 'uint64', 'ulonglong'], '%u'),
-    ('int8', '%hhd'), ('short', '%hd'), ('intc', '%d'), ('int_', '%ld'),
-    ('uint8', '%hhu'), ('ushort', '%hu'), ('uintc', '%u'), ('uint64', '%lu'),
-    ('S', '%s')]
-# This is for when default int is 32bit
-if np.dtype('int_') != np.dtype('longlong'):  # pragma: no cover
-    map_nptype2cformat.append(('longlong', '%lld'))
-    map_nptype2cformat.append(('ulonglong', '%llu'))
-map_cformat2pyscanf = [(['%hhd', '%hd', '%d', '%ld', '%lld'], '%d'),
-                       (['%hhu', '%hu', '%u', '%lu', '%llu'], '%u'),
-                       (['%5s', '%s'], '%s'),
-                       ('%s', '%s')]
-
-unsupported_cfmt = ['a', 'A', 'p', 'n', '']
-map_cformat2nptype = [(['f', 'F', 'e', 'E', 'g', 'G'], 'float64'),
-                      (['hhd', 'hhi'], 'int8'),
[…]

### 2017-09-06 cis_interface/drivers/AsciiFileOutputDriver.py

@@ -1,100 +1,0 @@
-import os
-from cis_interface.interface.PsiInterface import PSI_MSG_EOF
-from cis_interface.drivers.FileOutputDriver import FileOutputDriver
-from cis_interface.dataio.AsciiFile import AsciiFile
-
-
-class AsciiFileOutputDriver(FileOutputDriver):
-    r"""Class to handle output line by line to an ASCII file.
-
-    Args:
-        name (str): Name of the output queue to receive messages from.
-        args (str or dict): Path to the file that messages should be written to
-            or dictionary containing the filepath and other keyword arguments
-            to be passed to the created AsciiFile object.
-        skip_AsciiFile (bool, optional): If True, the AsciiFile instance is not
-            created. Defaults to False.
-        \*\*kwargs: Additional keyword arguments are passed to parent class's
-            __init__ method.
-
-    Attributes (in addition to parent class's):
-        file_kwargs (dict): Arguments used to create AsciiFile instance.
-        file (:class:`AsciiFile.AsciiFile`): Associated special class for ASCII
-            file.
-
-    """
-    
-    def __init__(self, name, args, skip_AsciiFile=False, **kwargs):
-        filepath = None
-        self.args_ignored = []
-        if isinstance(args, str):
-            filepath = args
-            args = {}
-        elif isinstance(args, list):
-            if isinstance(args[0], str):
-                filepath = args.pop(0)
-            args_new = {}
-            for a in args:
-                if isinstance(a, dict):
-                    args_new.update(**a)
-                else:
-                    self.args_ignored.append(a)
-            args = args_new
-        elif not isinstance(args, dict):  # pragma: debug
-            raise TypeError("args is incorrect type, check the yaml.")
-        if filepath is None:
-            filepath = args.pop('filename', None)
-            filepath = args.pop('filepath', filepath)
-        super(AsciiFileOutputDriver, self).__init__(name, filepath, **kwargs)
-        self.debug('(%s)', filepath)
-        for a in self.args_ignored:
-            self.info(": Ignoring argument '%s'", str(a))
-        self.file_kwargs = args
-        self.file_kwargs['format_str'] = ''
-        if skip_AsciiFile:
-            self.file = None
-        else:
-            self.file = AsciiFile(filepath, 'w', **self.file_kwargs)
-        self.debug('(%s): done with init', args)
-
-    @property
-    def eof_msg(self):
-        r"""str: Message indicating end of file."""
[…]

### 2017-09-06 cis_interface/drivers/AsciiTableInputDriver.py

@@ -1,64 +1,0 @@
-import os
-from cis_interface.drivers.AsciiFileInputDriver import AsciiFileInputDriver
-from cis_interface.dataio.AsciiTable import AsciiTable
-
-
-class AsciiTableInputDriver(AsciiFileInputDriver):
-    r"""Class to handle input from an ASCII table.
-
-    Args:
-        name (str): Name of the input queue to send messages to.
-        args (str or dict): Path to the file that messages should be read from
-            or dictionary containing the filepath and other keyword arguments
-            to be passed to the created AsciiTable object.
-        \*\*kwargs: Additional keyword arguments are passed to parent class's
-            __init__ method.
-
-    Attributes (in additon to parent class's):
-        file (:class:`AsciiTable.AsciiTable`): Associated special class for
-            ASCII table.
-        as_array (bool): If True, the table contents are sent all at once as an
-            array. Defaults to False if not set in args dict.
-
-    """
-    def __init__(self, name, args, **kwargs):
-        super(AsciiTableInputDriver, self).__init__(
-            name, args, skip_AsciiFile=True, **kwargs)
-        self.debug('(%s)', args)
-        self.as_array = self.file_kwargs.pop('as_array', False)
-        self.file = AsciiTable(self.args, 'r', **self.file_kwargs)
-        self.debug('(%s): done with init', args)
-
-    def run(self):
-        r"""Run the driver. The file is opened and then data is read from the
-        file and sent to the message queue until eof is encountered or the file
-        is closed.
-        """
-        self.debug(':run in %s', os.getcwd())
-        self.ipc_send(self.file.format_str)
-        if self.as_array:
-            with self.lock:
-                data = self.file.read_bytes(order='F')
-            self.debug(':run: read: %d bytes', len(data))
-            self.ipc_send_nolimit(data)
-        else:
-            with self.lock:
-                self.file.open()
-            nread = 0
-            while self.file.is_open:
-                with self.lock:
-                    if self.file.is_open:
-                        eof, data = self.file.readline_full()
-                    else:  # pragma: debug
-                        break
-                if eof:
-                    self.debug(':run, End of file encountered')
-                    self.ipc_send_nolimit(self.eof_msg)
-                    break
-                elif (data is not None):
-                    self.debug(':run: read: %d bytes', len(data))
-                    self.ipc_send_nolimit(data)
-                    nread += 1
-            if nread == 0:  # pragma: debug
-                self.debug(':run, no input')
-        self.debug(':run returned')

### 2017-09-06 cis_interface/drivers/FileInputDriver.py

@@ -1,63 +1,0 @@
-import os
-from IODriver import IODriver
-
-
-class FileInputDriver(IODriver):
-    r"""Class that sends messages read from a file.
-
-    Args:
-        name (str): Name of the queue that messages should be sent to.
-        args (str): Path to the file that messages should be read from.
-        \*\*kwargs: Additional keyword arguments are passed to parent class's
-            __init__ method.
-
-    Attributes (in additon to parent class's):
-        args (str): Path to the file that messages should be read from.
-        fd (file-like): File descriptor for the input file while it is open.
-        lock (:class:`threading.Lock`): Lock to be used when accessing file.
-
-    """
-    def __init__(self, name, args, **kwargs):
-        super(FileInputDriver, self).__init__(name, "_IN", **kwargs)
-        self.debug('(%s)', args)
-        self.args = os.path.abspath(args)
-        self.fd = None
-        self.debug('(%s): done with init', args)
-
-    def close_file(self):
-        r"""Close the file."""
-        self.debug(':close_file()')
-        with self.lock:
-            if self.fd:
-                self.fd.close()
-            self.fd = None
-
-    def terminate(self):
-        r"""Terminate the driver. The file is closed as necessary."""
-        self.debug(':terminate()')
-        super(FileInputDriver, self).terminate()
-        self.close_file()
-
-    def run(self):
-        r"""Run the driver. The file is opened and then data is read from the
-        file and sent to the message queue until eof is encountered or the file
-        is closed.
-        """
-        self.debug(':run in %s', os.getcwd())
-        try:
-            with self.lock:
-                self.fd = open(self.args, 'rb')
-        except:  # pragma: debug
-            self.exception('Could not open file.')
-            return
-        with self.lock:
-            if self.fd is None:  # pragma: debug
-                data = ''
-            else:
-                data = self.fd.read()
-        self.debug(':run: read: %d bytes', len(data))
-        if len(data) == 0:  # pragma: debug
-            self.debug(':run, no input')
-        else:
-            self.ipc_send(data)
-        self.debug(':run returned')

### 2017-09-06 cis_interface/drivers/FileOutputDriver.py

@@ -1,67 +1,0 @@
-import os
-from IODriver import IODriver
-
-
-class FileOutputDriver(IODriver):
-    r"""Class to handle output of received messages to a file.
-
-    Args:
-        name (str): Name of the output queue to receive messages from.
-        args (str): Path to the file that messages should be written to.
-        \*\*kwargs: Additional keyword arguments are passed to parent class's
-            __init__ method.
-
-    Attributes (in addition to parent class's):
-        args (str): Path to the file that messages should be written to.
-        fd (file-like): File descriptor for the target file if open.
-        lock (:class:`threading.Lock`): Lock to be used when accessing file.
-
-    """
-    def __init__(self, name, args, **kwargs):
-        super(FileOutputDriver, self).__init__(name, "_OUT", **kwargs)
-        self.debug('(%s)', args)
-        self.args = os.path.abspath(args)
-        self.fd = None
-        self.debug('(%s): done with init', args)
-
-    def close_file(self):
-        r"""Close the file."""
-        self.debug(':close_file()')
-        with self.lock:
-            if self.fd:
-                self.fd.close()
-            self.fd = None
-
-    def terminate(self):
-        r"""Terminate the driver, closeing the file as necessary."""
-        self.debug(':terminate()')
-        super(FileOutputDriver, self).terminate()
-        self.close_file()
-
-    def run(self):
-        r"""Run the driver. The driver will open the file and write receieved
-        messages to the file as they are received until the file is closed.
-        """
-        self.debug(':run in %s', os.getcwd())
-        try:
-            with self.lock:
-                self.fd = open(self.args, 'w+')
-        except:  # pragma: debug
-            self.exception('Could not open file.')
-            return
-        while self.fd is not None:
-            data = self.ipc_recv()
-            if data is None:
-                self.debug(':recv: closed')
-                break
-            self.debug(':recvd %s bytes', len(data))
-            if len(data) > 0:
-                with self.lock:
-                    if self.fd is None:
-                        break  # pragma: debug
-                    else:
-                        self.fd.write(data)
-            else:
-                self.debug(':recv: no data')
-                self.sleep()
-        self.debug(':run returns')

### 2017-09-06 cis_interface/drivers/GCCModelDriver.py

@@ -1,74 +1,0 @@
-#
-# This should not be used directly by modelers
-#
-from ModelDriver import ModelDriver
-import subprocess
-import os
-_top_dir = os.path.normpath(os.path.join(os.path.dirname(__file__), '../'))
-_incl_interface = os.path.join(_top_dir, 'interface')
-_incl_io = os.path.join(_top_dir, 'io')
-
-
-class GCCModelDriver(ModelDriver):
-    r"""Class from running gcc compiled driveres.
-
-    Args:
-        name (str): Driver name.
-        args (str or list): Argument(s) for running the model on the command
-            line. If the first element ends with '.c', the driver attempts to
-            compile the code with the necessary interface include directories.
-            Additional arguments that start with '-I' are included in the
-            compile command. Others are assumed to be runtime arguments.
-        \*\*kwargs: Additional keyword arguments are passed to parent class's
-            __init__ method.
-
-    Attributes (in additon to parent class's):
-        compiled (bool): True if the compilation was succesful. False otherwise.
-
-    """
-
-    def __init__(self, name, args, **kwargs):
-        super(GCCModelDriver, self).__init__(name, args, **kwargs)
-        self.debug()
-
-        # Prepare arguments to compile the file
-        # TODO: Allow user to provide a makefile
-        cfile = self.args.pop(0)
-        assert(cfile.endswith('.c'))
-        execname = os.path.splitext(cfile)[0] + '.out'
-        compile_args = ["gcc", "-g", "-Wall"]
-        for x in [_incl_interface, _incl_io]:
-            compile_args += ["-I" + x]
-        run_args = [os.path.join(".", execname)]
-        for arg in self.args:
-            if arg.startswith("-I"):
-                compile_args.append(arg)
-            else:
-                run_args.append(arg)
-        compile_args += ["-o", execname, cfile]
-        # Compile in a new process
-        self.args = run_args
-        self.compiled = True
-        try:
-            self.debug("::compiling")
-            comp_process = subprocess.Popen(['stdbuf', '-o0'] + compile_args,
-                                            bufsize=0, stdin=subprocess.PIPE,
-                                            stderr=subprocess.STDOUT,
-                                            stdout=subprocess.PIPE)
-            output, err = comp_process.communicate()
-            exit_code = comp_process.returncode
-            if exit_code != 0:  # pragma: debug
-                self.error(output)
-                raise RuntimeError("Compilation failed with code %d." % exit_code)
[…]

### 2017-09-06 cis_interface/drivers/ModelDriver.py

@@ -1,83 +1,0 @@
-#
-# This should not be used directly by modelers
-#
-import os
-import subprocess
-import sys
-from pprint import pformat
-from Driver import Driver
-import time
-
-
-def preexec():  # pragma: no cover
-    # Don't forward signals - used to ignore signals
-    os.setpgrp()
-
-
-class ModelDriver(Driver):
-    r"""Base class form Model drivers.
-
-    Args:
-        name (str): Driver name.
-        args (str or list): Argument(s) for running the model on the command
-            line.
-        \*\*kwargs: Additional keyword arguments are passed to parent class's
-            __init__ method.
-
-    Attributes (in additon to parent class's):
-        args (list): Argument(s) for running the model on the command line.
-        process (:class:`subprocess.Popen`): Process used to run the model.
-        env (dict): Dictionary of environment variables.
-
-    """
-
-    def __init__(self, name, args, **kwargs):
-        super(ModelDriver, self).__init__(name, **kwargs)
-        self.debug(str(args))
-        if isinstance(args, str):
-            self.args = [args]
-        else:
-            self.args = args
-        self.process = None
-        # This gets added to before run with the channel args
-        self.env = os.environ.copy()
-
-    def run(self):
-        r"""Run the model on a new process, receiving output from."""
-        self.debug(':run %s from %s with cwd %s and env %s',
-                   self.args, os.getcwd(), self.workingDir, pformat(self.env))
-        with self.lock:
-            try:
-                self.process = subprocess.Popen(
-                    ['stdbuf', '-o0'] + self.args, bufsize=0,
-                    stdin=subprocess.PIPE, stderr=subprocess.STDOUT,
-                    stdout=subprocess.PIPE,
-                    env=self.env, cwd=self.workingDir, preexec_fn=preexec)
-            except:  # pragma: debug
-                self.exception('(%s): Exception starting in %s with wd %s',
-                               self.args, os.getcwd, self.workingDir)
-                return
-        
-        # Re-direct output
-        while True:
-            time.sleep(0)
-            with self.lock:
-                if self.process:
-                    line = self.process.stdout.readline()
-                else:
-                    return
-            if not line:
-                break
-            sys.stdout.write(line)
-            sys.stdout.flush()
-            
-        self.debug(':run: done')
-
-    def terminate(self):
-        r"""Terminate the process running the model."""
-        self.debug(':terminate()')
-        with self.lock:
-            if self.process:
-                self.debug(':terminate(): terminate process')
-                self.process.terminate()
-                self.process = None

### 2017-09-06 cis_interface/drivers/RMQInputDriver.py

@@ -1,84 +1,0 @@
-"""Module for receiving input from a RabbitMQ server."""
-import requests
-from pprint import pformat
-from RMQDriver import RMQDriver
-from IODriver import IODriver
-
-
-class RMQInputDriver(RMQDriver, IODriver):
-    r"""Driver for receiving input from a RabbitMQ server.
-
-    Args:
-        name (str): The name of the local message queue that the driver should
-            connect to.
-        args (str): The name of the RabbitMQ message queue that the driver
-            should connect to.
-        \*\*kwargs: Additional keyword arguments are passed to parent class's
-            __init__ method.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-
-    def __init__(self, name, args, **kwargs):
-        super(RMQInputDriver, self).__init__(
-            name, suffix="_IN", queue=args, **kwargs)
-        self.debug()
-
-    def printStatus(self):
-        r"""Print the driver status."""
-        self.debug('::printStatus')
-        super(RMQInputDriver, self).printStatus()
-        url = 'http://%s:%s/api/%s/%s/%s' % (
-            self.server, 15672, 'queues', '%2f', self.queue)
-        res = requests.get(url, auth=(self.user, self.passwd))
-        jdata = res.json()
-        qdata = jdata.get('message_stats', '')
-        self.error(": server info: %s", pformat(qdata))
-
-    def start_communication(self):
-        r"""Begin consuming messages and add the callback for cancelling
-        consumption."""
-        self.debug('::start_communication')
-        # one at a time, don't stuff the Qs
-        self.channel.basic_qos(prefetch_count=1)
-        self._consumer_tag = self.channel.basic_consume(
-            self.on_message, queue=self.queue)
-
-    def on_message(self, ch, method, props, body):
-        r"""Action to perform when a message is received. Send it to the
-        local queue and acknowledge the message."""
-        self.debug('::on_message: received message # %s from %s',
-                   method.delivery_tag, props.app_id)
-        with self.lock:
-            if self._closing:
-                return
-        self.ipc_send(body)
-        with self.lock:
-            if self._closing:
-                return
-            ch.basic_ack(delivery_tag=method.delivery_tag)
-
-    def stop_communication(self, **kwargs):
-        r"""Stop sending/receiving messages. Only RMQInputDriver should
-        explicitly delete the queue."""
-        with self.lock:
-            self._closing = True
-            if self.channel:
[…]

### 2017-09-06 cis_interface/drivers/RMQOutputDriver.py

@@ -1,95 +1,0 @@
-"""Module for sending output to a RabbitMQ server."""
-from RMQDriver import RMQDriver
-from IODriver import IODriver
-
-
-class RMQOutputDriver(RMQDriver, IODriver):
-    r"""Driver for sending output to a RabbitMQ server.
-
-    Args:
-        name (str): The name of the local message queue that the driver should
-            connect to.
-        args (str): The name of the RabbitMQ message queue that the driver
-            connect to.
-        \*\*kwargs: Additional keyword arguments are passed to parent class's
-            __init__ method.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-    
-    def __init__(self, name, args, **kwargs):
-        super(RMQOutputDriver, self).__init__(
-            name, suffix="_OUT", queue=args, **kwargs)
-        self.debug()
-
-    def printStatus(self):
-        r"""Print the driver status."""
-        self.debug('::printStatus():')
-        super(RMQOutputDriver, self).printStatus()
-        msg = '%-30s' % ('RMQOutputDriver(' + self.name + '):')
-        with self.lock:
-            if self._q_obj is not None:
-                msg += '%-30s' % (str(self._q_obj.method.message_count) +
-                                  ' in RMQ server queue')
-            else:
-                msg += '%-30s' % 'queue not found'
-        print(msg)
-
-    def start_communication(self):
-        r"""Start publishing messages from the local queue."""
-        self.debug('::start_communication')
-        self.publish_message()
-    
-    def publish_message(self):
-        r"""Continue receiving messages from the local queue and passing them
-        to the RabbitMQ server until the queue is closed."""
-        with self.lock:
-            if self._closing:  # pragma: debug
-                return
-        while True:
-            self.debug("::publish_message(): IPC recv")
-            data = self.ipc_recv()
-            if data is None:
-                self.debug("::publish_message(): queue closed!")
-                break
-            elif len(data) == 0:
-                self.debug("::publish_message(): no data, reschedule")
-                self.connection.add_timeout(self.sleeptime,
-                                            self.publish_message)
-                break
-            self.debug("::publish_message(): IPC recv got %d bytes", len(data))
-            self.debug("::publish_message(): send %d bytes to AMQP", len(data))
-            with self.lock:
-                if self._closing:  # pragma: debug
-                    return
-                self.channel.basic_publish(
-                    exchange=self.exchange, routing_key=self.queue,
[…]

### 2017-09-06 cis_interface/drivers/RMQServerDriver.py

@@ -1,95 +1,0 @@
-import pika
-from RMQDriver import RMQDriver
-from RPCDriver import RPCDriver
-
-
-_new_client_msg = "PSI_NEW_CLIENT"
-_end_client_msg = "PSI_END_CLIENT"
-
-
-class RMQServerDriver(RMQDriver, RPCDriver):
-    r"""Class for handling server side RPC type RabbitMQ communication.
-
-    Args:
-        name (str): The name of the local IPC message queues that the driver
-            should use.
-        args (str, optional): The name of the RabbitMQ message queue that the
-            driver should connect to. Defaults to name + '_SERVER' if not set.
-        \*\*kwargs: Additional keyword arguments are passed to parent class's
-            __init__ method.
-
-    Attributes (in addition to the parent class's):
-        clients (set): The unique clients subscribed to this server.
-    
-    """
-
-    def __init__(self, name, args=None, **kwargs):
-        if args is None:
-            args = name + '_SERVER'
-        super(RMQServerDriver, self).__init__(name, queue=args, **kwargs)
-        self.debug()
-        self.clients = set([])
-
-    @property
-    def n_clients(self):
-        r"""int: The number of clients that are submitting requests to the
-        server."""
-        return len(self.clients)
-
-    def on_queue_declareok(self, method_frame):
-        r"""Actions to perform once the queue is succesfully declared."""
-        self.debug('::Declaring the server request queue.')
-        self.purge_queue()
-        super(RMQServerDriver, self).on_queue_declareok(method_frame)
-
-    def start_communication(self, no_ack=False):
-        r"""Start consuming messages from the queue."""
-        self.debug('::start_consuming')
-        self.channel.basic_qos(prefetch_count=1)
-        self.consumer_tag = self.channel.basic_consume(self.on_message,
-                                                       queue=self.queue)
-
-    def on_message(self, ch, method, props, body):
-        r"""Actions to perform when a message is received."""
-        with self.lock:
-            if self._closing:
-                return
-        # TODO: handle possibility of message larger than AMQP server memory
-        if body == _new_client_msg:
-            self.debug('::New client (%s)' % props.reply_to)
-            self.clients.add(props.reply_to)
-        elif body == _end_client_msg:
-            self.debug('::Client signed off (%s)' % props.reply_to)
-            self.clients.remove(props.reply_to)
-            if self.n_clients == 0:
-                self.debug('::All clients have signed off. Stopping.')
-                self.stop()
-        else:
-            self.clients.add(props.reply_to)
[…]

### 2017-09-06 cis_interface/drivers/RPCDriver.py

@@ -1,181 +1,0 @@
-from Driver import Driver
-from IODriver import IODriver
-
-
-class RPCDriver(Driver):
-    r"""Base class for any driver that requires to access to input & output
-    queues for RPC type functionality.
-
-    Args:
-        name (str): The name of the message queue set that the driver should
-            connect to. (name + "_IN" and name + "_OUT")
-        \*\*kwargs: Additional keyword arguments are passed to parent class's
-            __init__ method.
-
-    Attributes (in addition to parent class's):
-        iipc (:class:`cis_interface.drivers.IODriver`): Driver for the input
-            message queue.
-        oipc (:class:`cis_interface.drivers.IODriver`): Driver for the output
-            message ueue.
-
-    """
-    def __init__(self, name, args=None, **kwargs):
-        super(RPCDriver, self).__init__(name, **kwargs)
-        self.debug()
-        self.iipc = IODriver(name, '_IN', **kwargs)
-        self.oipc = IODriver(name, '_OUT', **kwargs)
-
-    @property
-    def env(self):
-        r"""dict: Environment variables."""
-        out = {}
-        out.update(self.iipc.env)
-        out.update(self.oipc.env)
-        return out
-
-    def run(self):
-        r"""Run the input/output queue drivers."""
-        super(RPCDriver, self).run()
-        self.debug('.run()')
-        self.iipc.start()
-        self.oipc.start()
-        self.debug('.run() done')
-
-    def graceful_stop(self):
-        r"""Allow the IPC queues to terminate gracefully."""
-        self.debug('.graceful_stop()')
-        self.iipc.graceful_stop()
-        self.oipc.graceful_stop()
-        super(RPCDriver, self).graceful_stop()
-        self.debug('.graceful_stop() done')
-
-    def terminate(self):
-        r"""Terminate input/output queue drivers."""
-        self.debug('.terminate()')
-        self.iipc.terminate()
-        self.oipc.terminate()
-        super(RPCDriver, self).terminate()
-        self.debug('.terminate() done')
-
-    def printStatus(self):
-        r"""Print information on the status of the driver."""
-        super(RPCDriver, self).printStatus()
-        self.iipc.printStatus(beg_msg='RPC Input Driver:')
-        self.oipc.printStatus(beg_msg='RPC Ouput Driver:')
-
-    def recv_wait(self, use_output=False, timeout=0.0):
-        r"""Receive a message smaller than maxMsgSize. This method will wait
[…]

### 2017-09-06 cis_interface/drivers/__init__.py

@@ -1,24 +1,0 @@
-r"""IO and Model drivers."""
-
-# import Driver
-
-# # Model drivers
-# import ModelDriver
-# import PythonModelDriver
-# import GCCModelDriver
-# import MatlabModelDriver
-
-# # IO drivers
-# import IODriver
-# import RMQConnection
-# import RMQInputDriver
-# import RMQOutputDriver
-# import FileInputDriver
-# import FileOutputDriver
-# import AsciiFileInputDriver
-# import AsciiFileOutputDriver
-# import AsciiTableInputDriver
-# import AsciiTableOutputDriver
-
-
-__all__ = []

### 2017-09-06 cis_interface/drivers/tests/test_AsciiFileInputDriver.py

@@ -1,60 +1,0 @@
-import nose.tools as nt
-import test_FileInputDriver as parent
-
-
-class TestAsciiFileInputParam(parent.TestFileInputParam):
-    r"""Test parameters for AsciiFileInputDriver.
-
-    Attributes (in addition to the parent class):
-        -
-
-    """
-
-    def __init__(self):
-        super(TestAsciiFileInputParam, self).__init__()
-        self.driver = 'AsciiFileInputDriver'
-        self.attr_list += ['file_kwargs', 'file']
-        self.args = self.args
-
-
-class TestAsciiFileInputDriverNoStart(TestAsciiFileInputParam,
-                                      parent.TestFileInputDriverNoStart):
-    r"""Test runner for AsciiFileInputDriver without start.
-
-    Attributes (in addition to the parent class):
-        -
-
-    """
-    pass
-
-    
-class TestAsciiFileInputDriver(TestAsciiFileInputParam,
-                               parent.TestFileInputDriver):
-    r"""Test runner for AsciiFileInputDriver.
-
-    Attributes (in addition to the parent class):
-        -
-
-    """
-
-    def assert_before_stop(self):
-        r"""Assertions to make before stopping the driver instance."""
-        super(parent.TestFileInputDriver, self).assert_before_stop()
-        while self.instance.n_msg == 0:
-            self.instance.sleep()
-        iline = 0
-        while True:
-            data = self.instance.ipc_recv()
-            if (data is None) or (data == self.instance.eof_msg):
-                break
-            if len(data) > 0:
-                while self.file_lines[iline].startswith('#'):
-                    iline += 1  # pragma: no cover
-                nt.assert_equal(data, self.file_lines[iline])
-                iline += 1
-        nt.assert_equal(len(self.file_lines), iline)
-        
-    def assert_after_terminate(self):
-        r"""Assertions to make after stopping the driver instance."""
-        super(TestAsciiFileInputDriver, self).assert_after_terminate()
-        assert(not self.instance.file.is_open)

### 2017-09-06 cis_interface/drivers/tests/test_FileOutputDriver.py

@@ -1,75 +1,0 @@
-import os
-import nose.tools as nt
-import test_IODriver as parent
-
-
-class TestFileOutputParam(parent.TestIOParam):
-    r"""Test parameters for FileOutputDriver.
-
-    Attributes (in addition to parent class's):
-        filepath (str): Full path to test file.
-
-    """
-
-    def __init__(self):
-        super(TestFileOutputParam, self).__init__()
-        self.driver = 'FileOutputDriver'
-        self.filepath = os.path.abspath('ascii_input.txt')
-        self.args = self.filepath
-        self.attr_list += ['args', 'fd', 'lock']
-        
-
-class TestFileOutputDriverNoStart(TestFileOutputParam,
-                                  parent.TestIODriverNoStart):
-    r"""Test runner for FileOutputDriver without start.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-    pass
-
-
-class TestFileOutputDriver(TestFileOutputParam, parent.TestIODriver):
-    r"""Test runner for FileOutputDriver.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-
-    def setup(self):
-        r"""Create a driver instance and start the driver."""
-        super(TestFileOutputDriver, self).setup()
-        self.instance.ipc_send(self.file_contents)
-
-    def teardown(self):
-        r"""Remove the instance, stoppping it."""
-        super(TestFileOutputDriver, self).teardown()
-        if os.path.isfile(self.filepath):
-            os.remove(self.filepath)
-
-    def assert_after_stop(self):
-        r"""Assertions to make after stopping the driver instance."""
-        super(TestFileOutputDriver, self).assert_after_stop()
-        assert(os.path.isfile(self.filepath))
-        with open(self.filepath, 'rb') as fd:
-            data = fd.read()
-        nt.assert_equal(data, self.file_contents)
-
-    def assert_after_terminate(self):
-        r"""Assertions to make after terminating the driver instance."""
-        super(TestFileOutputDriver, self).assert_after_terminate()
-        assert(self.instance.fd is None)
-
-    def test_send_recv(self):
-        r"""Test sending/receiving small message."""
-        pass
-        # self.instance.close_file()
-        # super(TestFileOutputDriver, self).test_send_recv()
-
-    def test_send_recv_nolimit(self):
-        r"""Test sending/receiving large message."""
-        pass
-        # self.instance.close_file()
-        # super(TestFileOutputDriver, self).test_send_recv_nolimit()

### 2017-09-06 cis_interface/drivers/tests/test_GCCModelDriver.py

@@ -1,46 +1,0 @@
-import os
-from cis_interface.tests import scripts
-import test_ModelDriver as parent
-from cis_interface.drivers.GCCModelDriver import _incl_interface
-
-
-class TestGCCModelParam(parent.TestModelParam):
-    r"""Test parameters for GCCModelDriver.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-
-    def __init__(self):
-        super(TestGCCModelParam, self).__init__()
-        self.driver = 'GCCModelDriver'
-        self.args = [scripts['c'], '1', '-I' + _incl_interface]
-        self.attr_list += ['compiled']
-        
-
-class TestGCCModelDriverNoStart(TestGCCModelParam,
-                                parent.TestModelDriverNoStart):
-    r"""Test runner for GCCModelDriver without start.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-    pass
-
-
-class TestGCCModelDriver(TestGCCModelParam, parent.TestModelDriver):
-    r"""Test runner for GCCModelDriver.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-
-    def teardown(self):
-        r"""Remove the instance, stoppping it."""
-        super(TestGCCModelDriver, self).teardown()
-        fexec = os.path.splitext(self.args[0])[0] + '.out'
-        if os.path.isfile(fexec):
-            os.remove(fexec)

### 2017-09-06 cis_interface/drivers/tests/test_MatInputDriver.py

@@ -1,53 +1,0 @@
-import os
-import numpy as np
-from scipy.io import savemat
-from cis_interface.backwards import pickle
-import test_IODriver as parent
-
-
-class TestMatInputDriver(parent.TestIODriver):
-    r"""Test runner for MatInputDriver.
-
-    Attributes (in addition to parent class's):
-        filepath (str): Full path to test file.
-    
-    """
-
-    def __init__(self):
-        super(TestMatInputDriver, self).__init__()
-        self.driver = 'MatInputDriver'
-        self.filepath = os.path.abspath('mat_input.mat')
-        self.args = self.filepath
-
-    def setup(self):
-        r"""Create a driver instance and start the driver."""
-        with open(self.filepath, 'wb') as fd:
-            savemat(fd, self.data_dict)
-        super(TestMatInputDriver, self).setup()
-
-    def teardown(self):
-        r"""Remove the instance, stoppping it."""
-        super(TestMatInputDriver, self).teardown()
-        if os.path.isfile(self.filepath):
-            os.remove(self.filepath)
-
-    def assert_before_stop(self):
-        r"""Assertions to make before stopping the driver instance."""
-        super(TestMatInputDriver, self).assert_before_stop()
-        msg_recv = self.instance.recv_wait_nolimit()
-        dat_recv = pickle.loads(msg_recv)
-        for k in self.data_dict.keys():
-            np.testing.assert_array_equal(dat_recv[k], self.data_dict[k])
-
-    def assert_after_terminate(self):
-        r"""Assertions to make after stopping the driver instance."""
-        super(TestMatInputDriver, self).assert_after_terminate()
-        assert(self.instance.fd is None)
-
-    def test_send_recv(self):
-        r"""Test sending/receiving small message."""
-        pass
-
-    def test_send_recv_nolimit(self):
-        r"""Test sending/receiving large message."""
-        pass

### 2017-09-06 cis_interface/drivers/tests/test_MatOutputDriver.py

@@ -1,72 +1,0 @@
-import os
-import numpy as np
-from scipy.io import loadmat
-import test_IODriver as parent
-
-
-class TestMatOutputParam(parent.TestIOParam):
-    r"""Test parameters for MatOutputDriver.
-
-    Attributes (in addition to parent class's):
-        filepath (str): Full path to test file.
-    
-    """
-
-    def __init__(self):
-        super(TestMatOutputParam, self).__init__()
-        self.driver = 'MatOutputDriver'
-        self.filepath = os.path.abspath('mat_input.mat')
-        self.args = self.filepath
-        
-
-class TestMatOutputDriverNoStart(TestMatOutputParam,
-                                 parent.TestIODriverNoStart):
-    r"""Test runner for MatOutputDriver.
-
-    Attributes (in addition to parent class's):
-        -
-    
-    """
-    pass
-
-
-class TestMatOutputDriver(TestMatOutputParam, parent.TestIODriver):
-    r"""Test runner for MatOutputDriver.
-
-    Attributes (in addition to parent class's):
-        -
-    
-    """
-
-    def setup(self):
-        r"""Create a driver instance and start the driver."""
-        super(TestMatOutputDriver, self).setup()
-        self.instance.ipc_send_nolimit(self.pickled_data)
-
-    def teardown(self):
-        r"""Remove the instance, stoppping it."""
-        super(TestMatOutputDriver, self).teardown()
-        if os.path.isfile(self.filepath):
-            os.remove(self.filepath)
-
-    def assert_after_stop(self):
-        r"""Assertions to make after stopping the driver instance."""
-        super(TestMatOutputDriver, self).assert_after_stop()
-        assert(os.path.isfile(self.filepath))
-        with open(self.filepath, 'rb') as fd:
-            dat_recv = loadmat(fd)
-        for k in self.data_dict.keys():
-            np.testing.assert_array_equal(dat_recv[k], self.data_dict[k])
-
-    def assert_after_terminate(self):
-        r"""Assertions to make after stopping the driver instance."""
-        super(TestMatOutputDriver, self).assert_after_terminate()
-        assert(self.instance.fd is None)
-
-    def test_send_recv(self):
-        r"""Test sending/receiving small message."""
-        pass
-
-    def test_send_recv_nolimit(self):
-        r"""Test sending/receiving large message."""
-        pass

### 2017-09-06 cis_interface/drivers/tests/test_MatlabModelDriver.py

@@ -1,55 +1,0 @@
-import os
-from cis_interface.tests import scripts
-import test_ModelDriver as parent
-from cis_interface import runner
-from cis_interface.examples import yamls as ex_yamls
-
-
-_session_fname = os.path.join(os.getcwd(), 'nt_screen_session.txt')
-
-
-def test_multiple():
-    r"""Test that creates multiple matlab drivers."""
-    os.environ['FIB_ITERATIONS'] = '3'
-    os.environ['FIB_SERVER_SLEEP_SECONDS'] = '1'
-    cr = runner.get_runner(ex_yamls['rpcfib_matlab'])
-    cr.run()
-
-
-class TestMatlabModelDriver(parent.TestModelDriver,
-                            parent.TestModelDriverNoStart):
-    r"""Test runner for MatlabModelDriver.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-
-    def __init__(self):
-        super(TestMatlabModelDriver, self).__init__()
-        self.driver = "MatlabModelDriver"
-        self.args = [scripts["matlab"], "test", 1]
-        self.attr_list += ['started_matlab', 'mlengine']
-
-    def test_a(self):
-        r"""Dummy test to start matlab."""
-        if self.instance.screen_session is None:  # pragma: debug
-            print("Matlab was not started by this test. Close any " +
-                  "existing Matlab sessions to test creation/removal.")
-        else:
-            with open(_session_fname, 'w') as f:
-                f.write(self.instance.screen_session)
-            self.instance.screen_session = None
-            self.instance.started_matlab = False
-
-    def test_z(self):
-        r"""Dummy test to stop matlab."""
-        if os.path.isfile(_session_fname):
-            with open(_session_fname, 'r') as f:
-                session = f.read()
-            os.remove(_session_fname)
-            self.instance.screen_session = session
-            self.instance.started_matlab = True
-        else:  # pragma: debug
-            print("Skipping removal of Matlab session as the test did " +
-                  "not create it.")

### 2017-09-06 cis_interface/drivers/tests/test_ModelDriver.py

@@ -1,39 +1,0 @@
-import test_Driver as parent
-
-
-class TestModelParam(parent.TestParam):
-    r"""Test parameters for basic ModelDriver class.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-    
-    def __init__(self):
-        super(TestModelParam, self).__init__()
-        self.driver = 'ModelDriver'
-        self.args = ['sleep', '0.1']
-        self.attr_list += ['args', 'process', 'env']
-        
-
-class TestModelDriverNoStart(TestModelParam, parent.TestDriverNoStart):
-    r"""Test runner for basic ModelDriver class.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-    pass
-
-
-class TestModelDriver(TestModelParam, parent.TestDriver):
-    r"""Test runner for basic ModelDriver class.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-    
-    def run_before_stop(self):
-        r"""Commands to run while the instance is running."""
-        self.instance.wait()

### 2017-09-06 cis_interface/drivers/tests/test_PythonModelDriver.py

@@ -1,37 +1,0 @@
-from cis_interface.tests import scripts
-import test_ModelDriver as parent
-
-
-class TestPythonModelParam(parent.TestModelParam):
-    r"""Test parameters for PythonModelDriver.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-
-    def __init__(self):
-        super(TestPythonModelParam, self).__init__()
-        self.driver = "PythonModelDriver"
-        self.args = scripts["python"]
-
-
-class TestPythonModelDriver(TestPythonModelParam, parent.TestModelDriver):
-    r"""Test runner for PythonModelDriver.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-    pass
-
-
-class TestPythonModelDriverNoStart(TestPythonModelParam,
-                                   parent.TestModelDriverNoStart):
-    r"""Test runner for PythonModelDriver without start.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-    pass

### 2017-09-06 cis_interface/drivers/tests/test_RMQClientDriver.py

@@ -1,93 +1,0 @@
-import os
-import pika
-import nose.tools as nt
-import test_RMQDriver as parent1
-from test_IODriver import IOInfo
-from cis_interface.drivers.RMQClientDriver import (
-    _new_client_msg, _end_client_msg)
-from cis_interface import runner
-from cis_interface.examples import yamls as ex_yamls
-
-
-def test_yaml():
-    r"""Test Server/Client setup using runner."""
-    os.environ['FIB_ITERATIONS'] = '3'
-    os.environ['FIB_SERVER_SLEEP_SECONDS'] = '1'
-    cr = runner.get_runner(ex_yamls['rpcfib_python'])
-    cr.run()
-
-
-class TestRMQClientParam(parent1.TestRMQParam, IOInfo):
-    r"""Test parameters for RMQClientDriver class.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-
-    def __init__(self):
-        super(TestRMQClientParam, self).__init__()
-        self.driver = 'RMQClientDriver'
-        self.args = None
-        self.attr_list += ['request_queue', 'response', 'corr_id',
-                           '_deliveries', '_acked', '_nacked',
-                           '_message_number']
-        self._temp_queue = 'TestRMQClientDriver_SERVER'
-        if getattr(self, 'channel', None):
-            self.channel.queue_purge(queue=self.temp_queue)
-            
-
-class TestRMQClientDriverNoStart(TestRMQClientParam,
-                                 parent1.TestRMQDriverNoStart):
-    r"""Test class for RMQClientDriver class without start.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-    pass
-
-
-class TestRMQClientDriver(TestRMQClientParam, parent1.TestRMQDriver):
-    r"""Test class for RMQClientDriver class.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-
-    def setup(self):
-        r"""Recover new client message on start-up."""
-        super(TestRMQClientDriver, self).setup()
-        nt.assert_equal(self.instance.request_queue, self.temp_queue)
-        # New client message
-        method_frame, props, rmq_msg = self.temp_basic_get()
-        self.channel.basic_ack(delivery_tag=method_frame.delivery_tag)
-        # print('received setup', rmq_msg)
-        nt.assert_equal(rmq_msg, _new_client_msg)
-        
-    def teardown(self):
-        r"""Recover end client message on teardown."""
[…]

### 2017-09-06 cis_interface/drivers/tests/test_RMQDriver.py

@@ -1,132 +1,0 @@
-import os
-import pika
-import nose.tools as nt
-import test_Driver as parent
-
-
-class TestRMQParam(parent.TestParam):
-    r"""Test parameters for RMQDriver class.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-
-    def __init__(self):
-        super(TestRMQParam, self).__init__()
-        self.driver = 'RMQDriver'
-        self.args = None
-        self.attr_list += ['user', 'server', 'passwd', 'exchange',
-                           'connection', 'queue', 'channel',
-                           'routing_key', 'consumer_tag',
-                           '_opening', '_closing', 'times_connected']
-        self._temp_queue = ''
-        self._connection = None
-        self._channel = None
-        self.inst_kwargs['user'] = os.environ.get('PSI_MSG_USER', None)
-
-        
-class TestRMQDriverNoStart(TestRMQParam, parent.TestDriverNoStart):
-    r"""Test class for RMQDriver class without start.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-    pass
-
-
-class TestRMQDriver(TestRMQParam, parent.TestDriver):
-    r"""Test class for RMQDriver class.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-
-    def teardown(self):
-        r"""Make sure the queue is empty before closing the driver."""
-        self.instance.purge_queue()
-        super(TestRMQDriver, self).teardown()
-
-    def declare_temp_queue(self):
-        r"""Create a blocking connection, channel, and queue for testing."""
-        self._connection = pika.BlockingConnection(
-            self.instance.connection_parameters)
-        self._channel = self._connection.channel()
-        out = self._channel.queue_declare(
-            queue=self._temp_queue,
-            auto_delete=True)
-        self._temp_queue = out.method.queue
-        self._channel.queue_bind(
-            queue=self._temp_queue,
-            exchange=self.instance.exchange,
-            routing_key=self._temp_queue)
-
-    @property
-    def temp_queue(self):
-        r"""Temporary queue for responses etc."""
-        if not self._temp_queue:
-            self.declare_temp_queue()
-        return self._temp_queue
-
-    @property
-    def connection(self):
-        r"""Temporary connection."""
-        if self._connection is None:  # pragma: debug
-            self.declare_temp_queue()
-        return self._connection
-        
-    @property
-    def channel(self):
-        r"""Temporary channel."""
-        if self._channel is None:
-            self.declare_temp_queue()
[…]

### 2017-09-06 cis_interface/drivers/tests/test_RMQInputDriver.py

@@ -1,51 +1,0 @@
-import nose.tools as nt
-import test_RMQDriver as parent1
-from test_IODriver import IOInfo
-
-
-class TestRMQInputParam(parent1.TestRMQParam, IOInfo):
-    r"""Test parameters for RMQInputDriver.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-
-    def __init__(self):
-        super(TestRMQInputParam, self).__init__()
-        self.driver = 'RMQInputDriver'
-        self.args = 'test'
-
-        
-class TestRMQInputDriverNoStart(TestRMQInputParam,
-                                parent1.TestRMQDriverNoStart):
-    r"""Test runner for RMQInputDriver without start.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-    pass
-
-
-class TestRMQInputDriver(TestRMQInputParam, parent1.TestRMQDriver):
-    r"""Test runner for RMQInputDriver.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-
-    def test_RMQ_recv(self):
-        r"""Receive a small message from AMQP server."""
-        self.instance.rmq_send(self.msg_short)
-        self.instance.sleep(0.1)
-        msg_recv = self.instance.recv_wait(timeout=3)
-        nt.assert_equal(msg_recv, self.msg_short)
-
-    def test_RMQ_recv_nolimit(self):
-        r"""Receive a large message from AMQP server."""
-        self.instance.rmq_send_nolimit(self.msg_long)
-        self.instance.sleep(0.1)
-        msg_recv = self.instance.recv_wait_nolimit(timeout=3)
-        nt.assert_equal(msg_recv, self.msg_long)

### 2017-09-06 cis_interface/drivers/tests/test_RMQServerDriver.py

@@ -1,80 +1,0 @@
-import pika
-import time
-import nose.tools as nt
-import test_RMQDriver as parent1
-from test_IODriver import IOInfo
-from cis_interface.drivers.RMQServerDriver import (
-    _new_client_msg, _end_client_msg)
-
-
-class TestRMQServerParam(parent1.TestRMQParam, IOInfo):
-    r"""Test parameters for RMQServerDriver class.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-
-    def __init__(self):
-        super(TestRMQServerParam, self).__init__()
-        self.driver = 'RMQServerDriver'
-        self.args = None
-        self.attr_list += ['clients']
-        
-
-class TestRMQServerDriverNoStart(TestRMQServerParam,
-                                 parent1.TestRMQDriverNoStart):
-    r"""Test class for RMQServerDriver class without start.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-    pass
-
-
-class TestRMQServerDriver(TestRMQServerParam, parent1.TestRMQDriver):
-    r"""Test class for RMQServerDriver class.
-
-    Attributes (in addition to parent class's):
-        -
-
-    """
-
-    def test_client_count(self):
-        r"""Test to ensure client count is correct."""
-        nt.assert_equal(self.instance.n_clients, 0)
-        # Send new client message
-        with self.instance.lock:
-            self.instance.channel.basic_publish(
-                exchange=self.instance.exchange,
-                routing_key=self.instance.queue,
-                properties=pika.BasicProperties(reply_to=self.temp_queue),
-                body=_new_client_msg)
-        time.sleep(0.1)
-        nt.assert_equal(self.instance.n_clients, 1)
-        # Send end client message
-        with self.instance.lock:
-            self.instance.channel.basic_publish(
-                exchange=self.instance.exchange,
-                routing_key=self.instance.queue,
-                properties=pika.BasicProperties(reply_to=self.temp_queue),
-                body=_end_client_msg)
-        time.sleep(0.1)
-        nt.assert_equal(self.instance.n_clients, 0)
-
-    def test_msg(self):
-        r"""Test routing of a message through the IPC & RMQ queues."""
-        # Send message to RMQ input & receive from IPC input
-        with self.instance.lock:
-            self.instance.channel.basic_publish(
-                exchange=self.instance.exchange,
-                routing_key=self.instance.queue,
[…]

### 2017-09-06 cis_interface/examples/SaM-multi/SaM-integrated.sh

@@ -1,5 +1,0 @@
-#!/bin/bash
-[ -d Output ] || mkdir Output
-# ----------------Your Commands------------------- #
-
-cisrun integrated.yml > ./Output/SaM_log.txt

### 2017-09-06 cis_interface/examples/SaM-multi/src/SaM.py

@@ -1,33 +1,0 @@
-#!/usr/bin/python
-import sys
-import signal
-from cis_interface.interface.PsiInterface import PsiInput, PsiOutput
-
-
-def signal_handler(signal, frame):
-    print('You pressed Ctrl+C!')
-    sys.exit(0)
-
-    
-if __name__ == '__main__':
-    signal.signal(signal.SIGINT, signal_handler)
-
-    print('Hello from Python!')
-    in1 = PsiInput('pyinput1')
-    in2 = PsiInput('pystatic')
-    out1 = PsiOutput('pyoutput')
-
-    adata = in1.recv()[1]
-    print("Python received {} from pyinput1".format(adata))
-    bdata = in2.recv()[1]
-    print("Python received {} from pystatic".format(bdata))
-
-    a = int(adata)
-    b = int(bdata)
-    sum = a + b
-
-    outdata = str(sum)
-    out1.send(outdata)
-    print("Python sent {} to pyoutput".format(outdata))
-
-    print('Goodbye from Python!')

### 2017-09-06 cis_interface/examples/SaM/SaMQ_GCC.sh

@@ -1,8 +1,0 @@
-#!/bin/bash
-
-# make output dir if it's not there
-[ -d Output ] || mkdir Output
-
-# ----------------Your Commands------------------- #
-
-cisrun SaMQ_GCC.yml > ./Output/SaM_log.txt

### 2017-09-06 cis_interface/examples/SaM/SaM_GCC.sh

@@ -1,9 +1,0 @@
-#!/bin/bash
-
-# make output dir if it's not there
-[ -d Output ] || mkdir Output
-
-# ----------------Your Commands------------------- #
-
-
-cisrun SaM_GCC.yml > ./Output/SaM_log.txt

### 2017-09-06 cis_interface/examples/SaM/src/SaM.c

@@ -1,22 +1,0 @@
-// Author Venkatraman Srinivasan
-#include <stdio.h>
-#include <stdlib.h>
-
-#include "PsiInterface.h"
-
-int main(int argc,char *argv[]){
-    const uint BSIZE = 8192; // the max
-    char buf1[BSIZE];
-    char buf2[BSIZE];
-    char outbuf[BSIZE];
-
-    PsiInput Input = psi_input("input1");
-    PsiInput Static = psi_input("static");
-    PsiOutput Output = psi_output("output");
-
-    psi_recv(Input, buf1, BSIZE);
-    psi_recv(Static, buf2, BSIZE);
-    sprintf(outbuf,"Sum = %d", atoi(buf1) + atoi(buf2));
-    psi_send(Output, outbuf, strlen(outbuf));
-    return 0;
-}

### 2017-09-06 cis_interface/examples/SaM/src/SaM.m

@@ -1,13 +1,0 @@
-PsiInterface = py.importlib.import_module('cis_interface.interface.PsiInterface');
-
-in = PsiInterface.PsiInput('input1');
-st = PsiInterface.PsiInput('static');
-out = PsiInterface.PsiOutput('output');
-
-Input = in.recv()
-Static = st.recv()
-Output = str2num(char(Input))+str2num(char(Static))
-sOutput = sprintf('Sum = %d',Output);
-out.send(sOutput);
-
-exit();

### 2017-09-06 cis_interface/examples/SaM/src/SaMQ.c

@@ -1,26 +1,0 @@
-// Author Venkatraman Srinivasan
-#include <stdio.h>
-#include <stdlib.h>
-
-#include "PsiInterface.h"
-
-int main(int argc,char *argv[]){
-    const uint BSIZE = 8192; // the max
-    char buf1[BSIZE];
-    char buf2[BSIZE];
-    char outbuf[BSIZE];
-
-    printf("start");
-    PsiInput Static = psi_input("static");
-    PsiInput Qin = psi_input("qin");
-    PsiOutput Qout = psi_output("qout");
-    PsiOutput Output = psi_output("output");
-
-    
-    psi_recv(Static, buf1, BSIZE);
-    psi_send(Qout, buf1, strlen(buf1));
-    psi_recv(Qin, buf2, BSIZE);
-    sprintf(outbuf,"Sum = %d", atoi(buf1) + atoi(buf2));
-    psi_send(Output, outbuf, strlen(outbuf));
-    return 0;
-}

### 2017-09-06 cis_interface/examples/__init__.py

@@ -1,40 +1,0 @@
-"""Tools for accessing examples from python."""
-import os
-
-
-yaml_list = [
-    ('ascii_io_python', 'ascii_io', 'ascii_io_Python'),
-    ('ascii_io_matlab', 'ascii_io', 'ascii_io_Matlab'),
-    ('ascii_io_gcc', 'ascii_io', 'ascii_io_GCC'),
-    ('ascii_io_all', 'ascii_io', 'ascii_io_all'),
-    ('sam_python', 'SaM', 'SaM_Python'),
-    ('sam_matlab', 'SaM', 'SaM_Matlab'),
-    ('sam_gcc', 'SaM', 'SaM_GCC'),
-    ('sam_all', 'SaM-multi', 'integrated'),
-    ('hello_python', 'python', 'hellopython'),
-    ('hello_matlab', 'matlab', 'helloMatlab'),
-    ('hello_gcc', ('cpp', 'hello'), 'hello_c.yml'),
-    ('rpcfib_python', ('python', 'rpcFib'),
-     ('rpcFibCli', 'rpcFibCliPar', 'rpcFibSrv')),
-    ('rpcfib_matlab', ('matlab', 'rpcFib'),
-     ('rpcFibCli', 'rpcFibCliPar', 'rpcFibSrv')),
-    ('rpcfib_gcc', ('cpp', 'rpcFib'),
-     ('rpcFibCli', 'rpcFibCliPar', 'rpcFibSrv'))
-]
-yamls = {}
-for k, d, f in yaml_list:
-    if isinstance(d, str):
-        dtup = [d]
-    else:
-        dtup = list(d)
-    alldir = os.path.join(os.path.dirname(__file__), *dtup)
-    if isinstance(f, str):
-        yamls[k] = os.path.join(alldir, f + '.yml')
-    else:
-        yamls[k] = []
-        for iyml in f:
-            yamls[k].append(
-                os.path.join(alldir, iyml + '.yml'))
-
-
-__all__ = ['yamls']

### 2017-09-06 cis_interface/examples/ascii_io/ascii_io.sh

@@ -1,37 +1,0 @@
-#!/bin/bash
-
-python make_input.py
-
-export PSI_DEBUG="INFO"
-export PSI_NAMESPACE="AsciiIO"
-
-# make output dir if it's not there
-[ -d Output ] || mkdir Output
-
-yaml= 
-
-# ----------------Your Commands------------------- #
-case $1 in
-    "" | -a | --all )
-	echo "Running Python, Matlab, C integration"
-	yaml='ascii_io_all.yml'
-	;;
-    -p | --python )
-	echo "Running Python"
-	yaml='ascii_io_Python.yml'
-	;;
-    -m | --matlab )
-	echo "Running Matlab"
-	yaml='ascii_io_Matlab.yml'
-	;;
-    -c | --gcc )
-	echo "Running C"
-	yaml='ascii_io_GCC.yml'
-	;;
-    * )
-	echo "Running ", $1
-	yaml=$1
-	;;
-esac
-
-cisrun $yaml

### 2017-09-06 cis_interface/examples/ascii_io/src/ascii_io.c

@@ -1,132 +1,0 @@
-// Author Venkatraman Srinivasan
-#include <stdio.h>
-#include <stdlib.h>
-
-#include "PsiInterface.h"
-
-PsiAsciiFileInput get_input_file(int type) {
-  const char *name;
-  if (type == 0)
-    name = "Input/inputC_file.txt";
-  else
-    name = "inputC_file";
-  PsiAsciiFileInput out = psi_ascii_file_input(name, type);
-  printf("%s\n", name);
-  return out;
-}
-
-PsiAsciiFileOutput get_output_file(int type) {
-  const char *name;
-  if (type == 0)
-    name = "Output/outputC_file.txt";
-  else
-    name = "outputC_file";
-  PsiAsciiFileOutput out = psi_ascii_file_output(name, type);
-  return out;
-}
-
-PsiAsciiTableInput get_input_table(int type) {
-  const char *name;
-  if (type == 0)
-    name = "Input/inputC_table.txt";
-  else
-    name = "inputC_table";
-  PsiAsciiTableInput out = psi_ascii_table_input(name, type);
-  return out;
-}
-
-PsiAsciiTableOutput get_output_table(int type) {
-  const char *name;
-  if (type == 0)
-    name = "Output/outputC_table.txt";
-  else
-    name = "outputC_table";
-  PsiAsciiTableOutput out = psi_ascii_table_output(name, type, "%5s\t%ld\t%f\n");
-  return out;
-}
-
-PsiAsciiTableInput get_input_array(int type) {
-  const char *name;
-  if (type == 0)
-    name = "Input/inputC_array.txt";
-  else
-    name = "inputC_array";
-  PsiAsciiTableInput out = psi_ascii_table_input(name, type);
-  return out;
-}
-
-PsiAsciiTableOutput get_output_array(int type) {
-  const char *name;
-  if (type == 0)
-    name = "Output/outputC_array.txt";
-  else
-    name = "outputC_array";
-  PsiAsciiTableOutput out = psi_ascii_table_output(name, type, "%5s\t%ld\t%f\n");
-  return out;
-}
-
-int main(int argc,char *argv[]){
-  char line[LINE_SIZE_MAX];
-  const uint BSIZE = 8192; // the max
-  char name[BSIZE];
-  int number;
-  double value;
-  int ret;
-  char *name_arr;
-  long *number_arr;
-  double *value_arr;
-
[…]

### 2017-09-06 cis_interface/examples/cpp/hello/hello.sh

@@ -1,6 +1,0 @@
-#!/bin/bash
-
-export PSI_NAMESPACE="hello_c"
-
-cisrun hello_c.yml
-

### 2017-09-06 cis_interface/examples/cpp/hello/hello_c.c

@@ -1,45 +1,0 @@
-
-#include "PsiInterface.h"
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <signal.h>
-
-
-/*
-* Read file, send to queue, read back from queue, store to file
-*/
-
-void sig_handler(int signo) {
-    printf("received signal %d\n", signo);
-}
-
-
-int main(int argc, char *argv[]) {
-    const int bufsz=8192;
-    char buf[bufsz];
-
-    signal(SIGINT, sig_handler);
-
-    printf("hello from C\n");
-
-    /* Matching with the the model yaml */
-    PsiInput in = psi_input("input"); 
-    PsiOutput out = psi_output("output");
-    printf("hello_c: created I/Os, %d, %d\n", in, out);
-
-    int ret;
-    ret = psi_recv(in, buf, bufsz);
-    if (ret < 0)
-        perror("psi_recv");
-    printf("hello_c received %d bytes: %s\n", ret, buf );
-    sleep(5);
-    ret = psi_send(out, buf, ret);
-    if (ret < 0)
-        perror("psi_send:");
-    printf("hello_c: send returns %d\n", ret);
-
-    printf("bye\n");
-    
-}
-

### 2017-09-06 cis_interface/examples/cpp/maxMsg/maxMsgCli.c

@@ -1,54 +1,0 @@
-
-#include "PsiInterface.h"
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <signal.h>
-
-void rand_str(char *dest, size_t length) {
-    char charset[] = "0123456789"
-                     "abcdefghijklmnopqrstuvwxyz"
-                     "ABCDEFGHIJKLMNOPQRSTUVWXYZ";
-
-    while (length-- > 0) {
-        size_t index = (double) rand() / RAND_MAX * (sizeof charset - 1);
-        *dest++ = charset[index];
-    }
-    *dest = '\0';
-}
-
-
-int main(int argc, char *argv[]) {
-    printf("maxMsgCli: hello psiMaxMsgSize is %d\n", PSI_MSG_MAX);
-
-    // create a max message, send/recv and verify
-    psiRpc_t rpc = psiRpcClient("cli_output", "%s", "cli_input", "%s");
-    
-    int fib;
-
-    char output[PSI_MSG_MAX];
-    char input[PSI_MSG_MAX];
-    char verify[PSI_MSG_MAX];
-
-    rand_str(output, PSI_MSG_MAX-1);
-    output[PSI_MSG_MAX] == '\0';
-    
-    // save a copy of the string
-    memcpy(verify, output, PSI_MSG_MAX);
-
-    if (rpcCall(rpc, output, &input)) {
-            printf("RPC error");
-            exit(-1);
-    }
-    if (memcmp(output, input, PSI_MSG_MAX)) {
-        printf("input/output do not match");
-        exit(-1);
-    } else {
-        printf("maxCliMsg: CONFIRM\n");
-    }
-
-    // all done, say goodbye
-    printf("maxMsgCli: bye\n");
-    
-}
-

### 2017-09-06 cis_interface/examples/cpp/rpcFib/rpcFibCli.c

@@ -1,46 +1,0 @@
-
-#include "PsiInterface.h"
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <signal.h>
-
-
-int main(int argc, char *argv[]) {
-    char *notset = "NOT SET";
-    char *ns = getenv("PSI_NAMESPACE");
-    if (ns == NULL) ns = notset;
-    char *rank = getenv("PSI_RANK");
-    if (rank == NULL) rank = notset;
-    char *hostname = getenv("PSI_HOST");
-    if (hostname == NULL) hostname = notset;
-
-   
-    printf("fibcli(C): hello, on host %s with PSI_NAMESPACE %s and PSI_RANK %s\n", hostname, ns, rank);
-    int iterations = atoi(argv[1]);
-    printf("running %d iterations\n", iterations);
-
-    // send fibbonacci via RPC
-    // create the rpc object to make the calls
-    psiRpc_t rpc = psiRpcClient("cli_fib", "%d", "cli_fib", "%d %d");
-    
-    // a place for the return value
-    int fib = -1;
-    int fibNo = -1;
-
-    // make a bunch of rpc's and display the results
-    for (int i = 1; i <= iterations; i++) {
-        printf("fibcli(C): fib(->%-2d) ::: ", i);
-        if (rpcCall(rpc, i, &fibNo, &fib) < 0) {
-            printf("RPC error, exit");
-            exit(-1);
-        }
-        printf("fib(%2d<-) = %-2d<-\n", fibNo, fib);
-    }
-
-    // all done, say goodbye
-    printf("fibcli(C) says bye\n");
-    exit(0);
-    
-}
-

### 2017-09-06 cis_interface/examples/cpp/rpcFib/rpcFibCliPar.c

@@ -1,53 +1,0 @@
-
-#include "PsiInterface.h"
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <signal.h>
-
-
-int main(int argc, char *argv[]) {
-    char *notset = "NOT SET";
-    char *ns = getenv("PSI_NAMESPACE");
-    if (ns == NULL) ns = notset;
-    char *rank = getenv("PSI_RANK");
-    if (rank == NULL) rank = notset;
-    char *host = getenv("PSI_HOST");
-    if (host == NULL) host = notset;
-
-    printf("Pfibcli(C): hello, system %s, PSI_NAMESPACE %s, PSI_RANK %s\n", host, ns, rank);
-    int iterations = atoi(argv[1]);
-    printf("running %d iterations\n", iterations);
-    
-
-    // send fibbonacci via RPC
-    // create the rpc object to make the calls
-    psiRpc_t rpc = psiRpcClient("cli_par_fib", "%d", "cli_par_fib", "%d %d");
-    
-    // a place for the return value
-    int fib = -1;
-    int fibNo = -1;
-
-    // make a bunch of rpc's and display the results
-    for (int i = 1; i <= iterations; i++) {
-      printf("Pfibcli(C): fib(->%2d) ::: ", i);
-      if (rpcSend(rpc, i)) {
-	printf("RPC send error");
-	exit(-1);
-      }
-    }
-    printf("\n");
-    for (int i = 1; i <= iterations; i++) {
-      if (rpcRecv(rpc, &fibNo, &fib)) {
-	printf("RPC recv error");
-	exit(-1);
-      }
-      printf("Pfibcli(C) : fib(%2d<-) = %-2d<-\n", fibNo, fib);
-    }
-
-    // all done, say goodbye
-    printf("Pfibcli(C) says bye\n");
-    exit(0);
-    
-}
-

### 2017-09-06 cis_interface/examples/matlab/rpcFib/rpcFibCliPar.m

@@ -1,37 +1,0 @@
-
-function rpcFibCliPar(iterations)
-  PsiInterface = py.importlib.import_module('cis_interface.interface.PsiInterface');
-  iterations = str2num(iterations);
-
-  fprintf('fibcliPar(M): hello, %d iterations', iterations);
-
-  rpc = PsiInterface.PsiRpc('cli_par_fib', '%d', 'cli_par_fib', '%d %d');
-
-  for i = 1:iterations
-    fprintf('fibcli(M):  fib(->%-2d) ::: ', i);
-    ret = rpc.rpcSend(i);
-    if (~ret)
-      disp('send failure');
-      exit(-1);
-    end
-  end
-
-  for i = 1:iterations
-    fprintf('fibcli(M):  fib(->%-2d) ::: ', i);
-    ret = rpc.rpcRecv();
-    if (~ret{1}) 
-      disp('end of input');
-      exit(-1)
-      break;
-    end
-    ret = ret{2}
-    disp(sprintf('fib(%2d<-) = %-2d<-', ret{1}, ret{2}));
-  end
-
-  disp('fibcliPar(M) ends');
-
-  exit(0);
-end
-
-
-

### 2017-09-06 cis_interface/examples/matlab/rpcFib/rpcFibSrv.m

@@ -1,42 +1,0 @@
-
-function rpcFibSrv(sleeptime)
-  PsiInterface = py.importlib.import_module('cis_interface.interface.PsiInterface');
-  sleeptime = str2num(sleeptime);
-  
-  disp('rpcFibSrv starts')
-
-  rpc = PsiInterface.PsiRpc('srv_fib', '%d %d', 'srv_fib', '%d');
-
-  while 1
-    input = rpc.rpcRecv();
-    if (~input{1})
-      D = sprintf('rpcFibSrv(M): end of input');
-      disp(D);
-      exit(0);
-    end
-    input = input{2};
-    D = sprintf('rpcFibSrv(M): got input %d', input{1});
-    disp(D);
-
-    pprev = 0;
-    prev = 1;
-    result = 1;
-    n = 1;
-    while n < input{1}
-      result = prev + pprev;
-      pprev = prev;
-      prev = result;
-      n = n+1;
-    end
-    pause(sleeptime);
-    D = sprintf('rpcFibSrv(M): sending result %d %d', input{1}, result);
-    disp(D);
-    rpc.rpcSend(input{1}, result);
-  end
-
-  disp('rpcFibSrv: done')
-  exit(0);
-end
-
-
-

### 2017-09-06 cis_interface/examples/python/rpcFib/rpcFibSrv.py

@@ -1,55 +1,0 @@
-from __future__ import print_function
-import os
-import sys
-import time
-import logging
-from logging import debug
-from cis_interface.interface.PsiInterface import PsiRpc
-import socket
-
-
-def fibServer(sleeptime):
-    print('fibsrv(P): hello on system {} with sleep {}'.format(
-        socket.gethostname(), sleeptime))
-
-    rpc = PsiRpc("srv_fib", "%d %d", "srv_fib", "%d")
-
-    while True:
-        print('fibsrv(P): input<- ', end='')
-        retval, rpc_in = rpc.rpcRecv()
-        if not retval:
-            print('rpcFibSrv: end of input')
-            break
-
-        #
-        # compute fibonacci(input)
-        #
-        pprev = 0
-        prev = 1
-        result = 1
-        fib_no = 1
-        arg = rpc_in[0]
-        while fib_no < arg:
-            result = prev + pprev
-            pprev = prev
-            prev = result
-            fib_no = fib_no + 1
-        print(' ::: ->(%2d %2d)' % (arg, result))
-
-        time.sleep(float(sleeptime))
-        rpc.rpcSend(arg, result)
-
-    debug('fibsrv(P): bye')
-    sys.exit(0)
-
-    
-if __name__ == '__main__':
-    LOGLEVEL = logging.NOTSET
-    if 'PSI_CLIENT_DEBUG' in os.environ:
-        LOGLEVEL = getattr(logging, os.environ['PSI_CLIENT_DEBUG'])
-    if 'RMQ_DrEBUG' in os.environ:
-        RMQLOGLEVEL = getattr(logging, os.environ['RMQ_DEBUG'])
-    logging.basicConfig(level=LOGLEVEL, stream=sys.stdout,
-                        format=sys.argv[0].split('/')[-1] + ': %(message)s')
-    print('psirun', sys.argv)
-    fibServer(sys.argv[1])

### 2017-09-06 cis_interface/interface/PsiInterface.h

@@ -1,812 +1,0 @@
-
-#include <fcntl.h>           /* For O_* constants */
-#include <sys/stat.h>        /* For mode constants */
-#include <sys/msg.h>
-#include <sys/types.h>
-#include <sys/sem.h>
-#include <sys/shm.h>
-#include <string.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <stdarg.h>
-#include <unistd.h>
-#include <errno.h>
-#include <../dataio/AsciiFile.h>
-#include <../dataio/AsciiTable.h>
-
-
-#define PSI_MSG_MAX 1024*64
-#define PSI_MSG_EOF "EOF!!!"
-
-typedef struct msgbuf_t {
-  long mtype;       
-  char data[PSI_MSG_MAX];
-} msgbuf_t;
-
-
-
-
-//
-// basic IO and formatted IO support
-//
-// usage:
-//      1. one-time: create input and output channels - holding in varibales
-//      2. Prepare:  Output: Format data in buffer, input: allocat buffer
-//      3. Call send/receive routines:  
-//          ret = psi_send(output_channel, buf, length_of_buf)
-//      or
-//          ret = psi_recv(input_channel, buf*, length_of_buf)
-
-// logging support: - this is compile-time:   if PSI_DEBUG changes recompile for the new level
-
-static inline
-void psiLog(const char* fmt, ...) {
-  va_list ap;
-  va_start(ap, fmt);
-  fprintf(stdout, "%d:", getpid());   
-  vfprintf(stdout, fmt, ap);
-  fprintf(stdout, "\n");
-  va_end(ap);
-}
-#ifdef PSI_DEBUG
-  #if PSI_DEBUG == INFO
-    #define info psiLog
-    #define debug while (0) psiLog
-  #elif PSI_DEBUG == DEBUG
-    #define debug psiLog
-    #define info psiLog
-  #else
-    #define debug while (0) psiLog
-    #define info while (0) psiLog
-  #endif
-#else
-  #define debug while (0) psiLog
-  #define info while (0) psiLog
-#endif
-
-typedef struct PsiInput {
-  int _handle;
-  const char *_name;
-  char *_fmt;
-} PsiInput;
-
-typedef struct PsiOutput {
-  int _handle ;
-  const char *_name;
-  char *_fmt;
-} PsiOutput;
-
-
[…]

### 2017-09-06 cis_interface/interface/PsiInterface.py

@@ -1,644 +1,0 @@
-from logging import debug
-import os
-import time
-from sysv_ipc import MessageQueue
-from cis_interface.backwards import pickle
-from cis_interface.interface.scanf import scanf
-from cis_interface.dataio.AsciiFile import AsciiFile
-from cis_interface.dataio.AsciiTable import AsciiTable
-
-
-PSI_MSG_MAX = 1024 * 64
-PSI_MSG_EOF = "EOF!!!"
-
-
-class PsiInput(object):
-    r"""Class for handling input from a message queue.
-
-    Args:
-        name (str): The name of the message queue. Combined with the
-            suffix '_INT', it should match an environment variable containing
-            a message queue key.
-        
-    Attributes:
-        name (str): The name of the message queue.
-        qname (str): The name of the message queue combined with the suffix
-            '_IN'.
-        q (:class:`sysv_ipc.MessageQueue`): Message queue.
-        
-    """
-    name = None
-    qName = None
-    q = None
-    
-    def __init__(self, name):
-        self.name = name
-        self.qName = name + '_IN'
-        self.q = None
-        debug("PsiInput(%s):", name)
-        if self.qName not in os.environ:
-            raise Exception('PsiInterface cant see %s in env.' % self.qName)
-            # print('ERROR:  PsiInterface cant see ' + name + ' in env')
-            # exit(-1)
-        qid = os.environ.get(self.qName, '')
-        qid = int(qid)
-        debug("PsiInput(%s): qid %s", self.name, qid)
-        self.q = MessageQueue(qid, max_message_size=PSI_MSG_MAX)
-        clidebug = os.environ.get('PSI_CLIENT_DEBUG', False)
-        self.sleeptime = 0.25
-        if clidebug:
-            self.sleeptime = 2.0
-
-    def recv(self):
-        r"""Receive a message smaller than PSI_MSG_MAX. The process will
-        sleep until there is a message in the queue to receive.
-
-        Returns:
-            tuple (bool, str): The success or failure of receiving a message
-                and the message received.
-
-        """
-        payload = (False, '')
-        debug("PsiInput(%s).recv()", self.name)
-        try:
-            while self.q.current_messages == 0:
-                debug("PsiInput(%s): recv() - no data, sleep", self.name)
-                time.sleep(self.sleeptime)
-            debug("PsiInput(%s).recv(): message ready, read it", self.name)
-            data, _ = self.q.receive()  # ignore ident
[…]

### 2017-09-06 cis_interface/interface/__init__.py

@@ -1,7 +1,0 @@
-r"""Routines for interfacing with framework from Python/C/C++/Matlab."""
-import PsiInterface
-from PsiInterface import (
-    PsiInput, PsiOutput, PsiRpc)
-
-
-__all__ = ['PsiInterface', 'PsiInput', 'PsiOutput', 'PsiRpc']

### 2017-09-06 cis_interface/interface/tests/test_PsiInterface.py

@@ -1,490 +1,0 @@
-import os
-import numpy as np
-import nose.tools as nt
-from cis_interface.interface import PsiInterface, PSI_MSG_EOF
-from cis_interface.drivers import (IODriver, RPCDriver,
-                                   AsciiFileInputDriver,
-                                   AsciiTableInputDriver)
-from cis_interface.drivers.tests.test_IODriver import IOInfo
-from cis_interface.backwards import pickle
-
-
-class TestPsiInput(IOInfo):
-    r"""Test basic input to python."""
-    def __init__(self):
-        super(TestPsiInput, self).__init__()
-        self.name = 'test'
-        self.driver = None
-        self.instance = None
-
-    def setup(self):
-        r"""Start driver and instance."""
-        self.driver = IODriver.IODriver(self.name, '_IN')
-        self.driver.start()
-        os.environ.update(self.driver.env)
-        self.instance = PsiInterface.PsiInput(self.name)
-
-    def teardown(self):
-        r"""Stop the driver."""
-        self.driver.stop()
-
-    def test_init(self):
-        r"""Test error on init."""
-        nt.assert_raises(Exception, PsiInterface.PsiInput, 'error')
-
-    def test_recv(self):
-        r"""Test receiving small message."""
-        # self.driver.ipc_send(self.msg_short)
-        self.driver.sched_task(0.01, self.driver.ipc_send,
-                               args=[self.msg_short])
-        msg_flag, msg_recv = self.instance.recv()
-        assert(msg_flag)
-        nt.assert_equal(msg_recv, self.msg_short)
-
-    def test_recv_nolimit(self):
-        r"""Test receiving large message."""
-        self.driver.ipc_send_nolimit(self.msg_long)
-        msg_flag, msg_recv = self.instance.recv_nolimit()
-        assert(msg_flag)
-        nt.assert_equal(msg_recv, self.msg_long)
-
-
-class TestPsiOutput(IOInfo):
-    r"""Test basic output to python."""
-    def __init__(self):
-        super(TestPsiOutput, self).__init__()
-        self.name = 'test'
-        self.driver = None
-        self.instance = None
-
-    def setup(self):
-        r"""Start driver and instance."""
-        self.driver = IODriver.IODriver(self.name, '_OUT')
-        self.driver.start()
-        os.environ.update(self.driver.env)
[…]

### 2017-09-06 cis_interface/tests/__init__.py

@@ -1,27 +1,0 @@
-"""Testing things."""
-import os
-
-# Test data
-data_dir = os.path.join(os.path.dirname(__file__), 'data')
-data_list = [
-    ('txt', 'ascii_file.txt'),
-    ('table', 'ascii_table.txt')]
-data = {k: os.path.join(data_dir, v) for k, v in data_list}
-
-# Test scripts
-script_dir = os.path.join(os.path.dirname(__file__), 'scripts')
-script_list = [
-    ('c', 'gcc_model.c'),
-    ('matlab', 'matlab_model.m'),
-    ('python', 'python_model.py')]
-scripts = {k: os.path.join(script_dir, v) for k, v in script_list}
-    
-# Test yamls
-yaml_dir = os.path.join(os.path.dirname(__file__), 'yamls')
-yaml_list = [
-    ('c', 'gcc_model.yml'),
-    ('matlab', 'matlab_model.yml'),
-    ('python', 'python_model.yml')]
-yamls = {k: os.path.join(yaml_dir, v) for k, v in yaml_list}
-
-__all__ = ['data', 'scripts', 'yamls']

### 2017-09-06 cis_interface/tests/scripts/python_io.py

@@ -1,23 +1,0 @@
-import logging
-from logging import debug
-import os
-import sys
-from PsiInterface import PsiInput, PsiOutput
-
-
-def runhello():
-    debug('hello pythonfiles from %s', os.getcwd())
-    inf = PsiInput('inFile')
-    outf = PsiOutput('outFile')
-    fdata = inf.recv()
-    debug('got input %s', fdata)
-    print fdata
-    outf.send(fdata)
-    debug('sent output')
-    debug('bye')
-
-
-if __name__ == '__main__':
-    logging.basicConfig(level=logging.DEBUG,
-                        format=sys.argv[0].split('/')[-1] + ':%(message)s')
-    runhello()

### 2017-09-06 cis_interface/tests/test_runner.py

@@ -1,49 +1,0 @@
-import nose.tools as nt
-from cis_interface import runner
-from cis_interface.drivers import Driver, ModelDriver, IODriver
-from cis_interface.tests import scripts
-from cis_interface.examples import yamls
-
-
-def test_import_driver():
-    r"""Check a few drivers."""
-    drvs = [('Driver', Driver.Driver),
-            ('ModelDriver', ModelDriver.ModelDriver),
-            ('IODriver', IODriver.IODriver)]
-    for n, dans in drvs:
-        dres = runner.import_driver(n)
-        nt.assert_equal(dres, dans)
-
-
-def test_create_driver():
-    r"""Test driver creation w/ and w/o args."""
-    runner.create_driver('Driver', 'test_io_driver')
-    runner.create_driver('ModelDriver', 'test_model_driver',
-                         args=scripts['python'])
-
-
-def test_get_runner():
-    r"""Use get_runner to start a run."""
-    cr = runner.get_runner([yamls['ascii_io_python']])
-    cr.run()
-
-    
-def test_runner_terminate():
-    r"""Start a runner, then stop it early."""
-    cr = runner.get_runner([yamls['ascii_io_python']])
-    cr.loadDrivers()
-    cr.startDrivers()
-    cr.printStatus()
-    cr.terminate()
-
-
-class TestCisRunner(object):
-    r"""Tests of the CisRunner class."""
-    def __init__(self):
-        nt.assert_raises(IOError, runner.CisRunner,
-                         ['fake_yaml.yml'], 'test_psi_run')
-        self.runner = runner.CisRunner([yamls['ascii_io_python']],
-                                       'test_psi_run')
-
-    def test_parseModelYaml(self):
-        nt.assert_raises(IOError, self.runner.parseModelYaml, 'fake_yaml.yml')

### 2017-09-06 cis_interface/tests/test_tools.py

@@ -1,17 +1,0 @@
-from cis_interface import tools
-
-
-def test_ipcs():
-    tools.ipcs()
-
-    
-def test_ipc_queues():
-    tools.ipc_queues()
-
-
-def test_ipcrm():
-    tools.ipcrm()
-
-
-def test_ipcrm_queues():
-    tools.ipcrm_queues()

### 2017-09-06 cis_interface/tools.py

@@ -1,65 +1,0 @@
-"""This modules offers various tools."""
-from subprocess import Popen, PIPE
-
-
-def ipcs(options=[]):
-    r"""Get the output from running the ipcs command.
-
-    Args:
-        options (list): List of flags that should be used. Defaults to an empty
-            list.
-
-    Returns:
-        list: Captured output.
-
-    """
-    cmd = ' '.join(['ipcs'] + options)
-    p = Popen(cmd, stdin=PIPE, stdout=PIPE, stderr=PIPE, shell=True)
-    output, err = p.communicate()
-    exit_code = p.returncode
-    if exit_code != 0:  # pragma: debug
-        print(err.decode('utf-8'))
-        raise Exception("Error on spawned process. See output.")
-    return output.decode('utf-8')
-
-
-def ipc_queues():
-    r"""Get a list of active IPC queues.
-
-    Returns:
-       list: List of IPC queues.
-
-    """
-    skip_lines = [
-        '------ Message Queues --------',
-        'key        msqid      owner      perms      used-bytes   messages    ',
-        '']
-    out = ipcs(['-q']).split('\n')
-    qlist = []
-    for l in out:
-        if l not in skip_lines:
-            qlist.append(l)
-    return qlist
-
-
-def ipcrm(options=[]):
-    r"""Remove IPC constructs using the ipcrm command.
-
-    Args:
-        options (list): List of flags that should be used. Defaults to an empty
-            list.
-
-    """
-    cmd = ' '.join(['ipcrm'] + options)
-    p = Popen(cmd, stdin=PIPE, stdout=PIPE, stderr=PIPE, shell=True)
-    output, err = p.communicate()
-    exit_code = p.returncode
-    if exit_code != 0:  # pragma: debug
-        print(err.decode('utf-8'))
-        raise Exception("Error on spawned process. See output.")
-    print(output.decode('utf-8'))
-
-
-def ipcrm_queues():
-    r"""Delete existing IPC queues."""
-    ipcrm(['--all=msg'])

### 2018-10-01 papers/lang2018/ms.tex
Draft

\newif\iftwocol
\newif\ifplacefig
\newif\ifdraft
\newif\ifinclfig
\newif\ifieee

% For IEEE 
%\ieeetrue
%\twocoltrue
%\placefigtrue
%\inclfigtrue % false for word count (no figs)
%\draftfalse

% For Draft
\ieeefalse
\twocoltrue
\placefigtrue
\inclfigtrue
\drafttrue

% For ApJ preprint
%\ieeefalse
%\twocolfalse
%\placefigfalse
%\inclfigtrue
%\draftfalse

% For ArXiv
%\ieeefalse
%\twocoltrue
%\placefigtrue
%\inclfigtrue
%\draftfalse

%---------------------------------------------------------------------------------
% PREAMBLE
\iftwocol
	\ifieee
		\documentclass[journal]{IEEEtran}
	\else
		\documentclass[apj]{emulateapj}
	\fi
\else
	\ifieee
		\documentclass[peerreview]{IEEEtran}
	\else
		\documentclass[preprint]{aastex}
	\fi
\fi

\pdfoutput=1
\usepackage[normalem]{ulem}
\usepackage{color}
\usepackage{graphicx}
\usepackage{enumerate}
\ifdraft
	\ifieee
		\usepackage[numbers]{natbib}
		\bibliographystyle{ieeetr}
	\else
		\usepackage{natbib}
		\citestyle{apj}
		\bibliographystyle{apj}
		\bibpunct[,]{(}{)}{;}{a}{}{,}
	\fi
\else
	\ifieee
		\usepackage[numbers]{natbib}
		\bibliographystyle{ieeetr}
	\fi
\fi

\usepackage[bookmarks=false]{hyperref}

%\makeatletter

%---------------------------------------------------------------------------------
% CONTROL THINGS
\newcommand{\todo}[1]{{\color{red}{#1}}}
\newcommand{\addref}{{\color{red}(ref. needed)}}
\newcommand{\Msol}{\,M$_{\odot}$}
\newcommand{\TB}{\,Tbytes}
\newcommand{\cis}{{\tt cis\_interface}{}}

%---------------------------------------------------------------------------------
% MANUSCRIPT INFO
\begin{document}

\shorttitle{{\cis}}

\title{{\cis}: A Python Package for Integrating Computational Models Across Languages and Scales}

\author{Meagan~Lang\altaffilmark{1}}

\altaffiltext{1}{National Center for Supercomputing Application, 
University of Illinois, Urbana-Champaign, IL email: {\tt langmm@illinois.edu}}

%---------------------------------------------------------------------------------
% ABSTRACT (145 words)
\begin{abstract}
Thousands of computational models have been created within the plant biology community within the past two decades that have the potential to be combined into complex integration networks capable of capturing more complex biological processes than possible as isolated models. However, the technological barriers introduced by differences in language and data formats as slowed this progress. We present \href{https://github.com/cropsinsilico/cis_interface}{\cis}, a Python package for running integration networks with connections between models across languages and scales. {\cis} coordinates parallel execution of models in Python, C, C++, and Matlab and handles communication in a number of data formats common to computational plant modeling. {\cis} is designed to be user-friendly, including visual tools for composing the integration networks. 
\end{abstract}

%\begin{IEEEkeywords}
%Plant modeling, Modeling, Communication
%\end{IEEEkeywords}

%\IEEEpeerreviewmaketitle

\maketitle
%---------------------------------------------------------------------------------
%---------------------------------------------------------------------------------
% INTRO (183 words)
\section{Introduction}\label{S:intro}
%
%\IEEEPARstart{C}{omputational} 
Computational plant biologists have produced a wealth of computational models for different biological processes governing plant growth and development, covering scales from atomic to global. Although usually developed independently to address questions specific to an organ, species, or growth process, computational plant models can also have applications beyond their original scope. For example, a photosynthesis model developed for maize can also be adapted to work for other species like soy \todo{cite Kavya and Yu}. Computational models for different scales, organs, or processes also have the potential for being combined to probe more complex biological mechanisms. With advances in compute power, it should be possible to link biologically related models to create integration networks capable of capturing the response of entire plants, fields, or regions. However, independently developed computational models often have compatibility issues resulting from differences in programming language or data format that make such collaborations difficult.

We present an Open Source Python package, {\cis}, for creating and running integrations networks from existing computational models written in Python, Matlab, C and C++. Although developed for the purpose of integrating plant biology models, {\cis} can be applied to any situation requiring coordination between models in the supported languages. \S\ref{S:background} provides background information on existing efforts for cross-language model integration, \S\ref{S:methods} describes the methods used by the {\cis} package to integrate models, \S\ref{S:results} presents several tests demonstrating the performance of message passing between integrated models, and \S\ref{S:discuss} summarizes the features of the {\cis} package, describes a few use cases, and outlines areas for future development.

%---------------------------------------------------------------------------------
%---------------------------------------------------------------------------------
% BACKGROUND
\section{Background}\label{S:background}
%
Computational plant models are usually written with a very specific research question in mind. The resulting programs are highly tuned for one purpose and written specifically for scientists within an isolated group of collaborators. Scientists will generally write programs in the language they are most familiar with and, given that there is no consensus on one programming language as ideal for all scientific applications, the language of computational models will often vary between research groups, if not between members of the groups themselves. The same can be said of data formats which can be highly tailored to the target question and may or may not include metadata like the data types, fields, or units. To integrate two biological models, the following questions must be addressed:
%
\begin{enumerate}
	\item Orchestration: How will models be executed?
	\item Communication: How will information be passed from one model to the next across languages?
	\item Translation: How will data output by one model be translated into a format understood by the next model?
	%\item Computational Efficiency: Will the integration make good use of computational resources?
	%\item Human Efficiency: How much hands-on work will the integration take and can products be reused in the future?
\end{enumerate}

These three questions can be addressed through manual integration on the command line or using scripting (e.g. running model A, converting output from A into input expected by model B, running model B). However, manual integration is 1) computational inefficient when many iterations are necessary, such as during parameter searches or steady state convergence tests, 2) unique to every integration, requiring the production of new scripts, and 3) complex when more than $>3$ models are being integrated. Within both computational biology and the larger scientific computing community, groups have developed around different solutions to the problem of connecting models.

%---------------------------------------------------------------------------------
% DOMAIN SPECIFIC LANGUAGE
\subsection{Domain Specific Language (DSL)}\label{SS:dsl}
%
One solution is to agree upon a language. The questions of orchestration, communication, and translation become trivial when models are all written in the same language. Beyond selecting a single programming language, many communities have  developed dedicated domain specific languages (DSLs) for model representation. For example, the Systems Biology Markup Language \citep[SBML][]{Hucka2003} is an XML-based format for representing models of biological processes. Scientists can compose their models using SBML markup and then run their model using software designed to parse SBML files. Other DSLs for biological models include LPy \citep{Boudon2012}, CellML \citep{Cuellar2003}, and BioPAX \citep{Demir2010}. Dedicated model DSLs improve the reusability of models written in the DSL and have the advantage of often implicitly handling data transformation. However, they do nothing for existing models that are not in the DSL or cannot be expressed within the constraints of the DSL. In addition, learning a new DSL, in addition to a programming language, can be daunting for new researchers.

%---------------------------------------------------------------------------------
% FRAMEWORKS
\subsection{Workflow/Data Flow/Framework}\label{SS:framework}
%
Another approach is a workflow/data flow/framework tools for models that are written in the same programming language or can be wrapped. For example, there are many generic tools in Python for coordinating the execution of tasks in parallel or serial \citep[e.g.][]{babuji18, celery, luigi}. While these tools are powerful for organizing complex work flows, it is the user's onus to make sure that the components they connect are compatible and handle an sort of data transformation that might be necessary (e.g. unit conversion or field selection).

There are also domain specific frameworks for connecting biological models. For example, OpenAlea \citep{Pradal2015} allows users to compose networks from different components including plant models, analysis tools, and visualizations. Domain specific frameworks are more intuitive and ultimately more flexible than DSLs in the types of operations they allow models to perform since they have access to the full power of a programming language. However, they are stricter in several respects: 1) the model must be written in (or exposable to), the language of the framework and 2) the model must be written in a way that is aware of the framework and/or the format of models with which it will interact. Like DSLs, frameworks also provided limited support for models written without them in mind.

{\cis} interface is an example of a framework that overcomes these issues by exposing light-weight interfaces in the language of the model (See \S\ref{SSS:interface}) that permit messages to be passed between the model processes as they run in parallel (See \S\ref{SS:communication}). As a result, models writers only need knowledge of the language in which their model is written.

%---------------------------------------------------------------------------------
% MESSAGE PASSING TOOLS
%\subsection{Message Passing Tools}\label{SS:messages}
%
%There are many existing software solutions for passing messages between processes. Implementations of the Message Passing Interface (MPI) standard, such as MPICH \citep{Gropp1996} and OpenMPI \citep{Gabriel2004}, allow communication between instances of a program launched on parallel processes with distributed memory. MPI implementations are robust, fast, and scale well, however, they are optimized for processing high volumes of structured data on supercomputers and are not as well suited to the case of dynamic integration networks composed of diverse processes operating on different timescales.

%---------------------------------------------------------------------------------
%---------------------------------------------------------------------------------
% METHODS
\section{Methods}\label{S:methods}

{\cis} was designed to address the questions from \S\ref{S:background} while being:

\begin{itemize}
	\item Easy to use. Require as little modification to the model source code as possible and only in the language of the model itself.
	\item Efficient. Allow models to run in parallel with asynchronous communication that dosen't block when a message is sent.
	\item Flexible. Provide the same interface to the user, regardless of the communication mechanism being used or the platform the model is being executed on.
\end{itemize}

{\cis} currently support models written in Python, Matlab, C, and C++ with 
additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}.

%---------------------------------------------------------------------------------
% ORCHESTRATION
\subsection{Orchestration}\label{SS:orchestration}
%
%At execution, {\cis} launches models on new processes based on information contained in a YAML specification file (See \S\ref{SSS:yaml} below) and establishes

%---------------------------------------------------------------------------------
% YAML
\subsubsection{YAML Specification}\label{SSS:yaml}
%
Users specify information about models and integration networks via declarative 
YAML files \citep{Ben-Kiki2009}. The YAML file format was selected because it is human readable 
and there are many existing tools for parsing YAML formats in many different 
languages. The declarative format allows user to specify exactly what they 
what to do, without describing how it should be done. While the information 
about models and integration networks can be contained in a single YAML 
file, the information can naturally be split between two files, one containing 
information about the models and one containing the connections comprising the 
integration network. This separation is advantageous because the model YAML 
can be re-used, unchanged, in conjunction with other integration networks.

Model YAMLs include information about the location of the 
model source code, the language the model is written in, how the model should 
be run, and any input or output variables including their type and units. 
Integration networks are specified via YAML files declaring the connections 
between models. Connections are declared by pairing an output variable from 
one model with the input variable of another model. The {\cis} 
command line interface (CLI) sets up the necessary communication mechanisms to 
then direct data from one model to the next in the specified pattern. Models 
can have as many input and/or output variables as is desired.
Connections between models are specified by references to the 
input/output variables associated with each model. This format allows users the 
flexibility to create complex integration networks.

%---------------------------------------------------------------------------------
% LANGUAGE SUPPORT
%\subsubsection{Language Support}\label{SSS:language}
%
%{\cis} currently support models written in Python, Matlab, C, and C++ with 
%additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}. 
%Support for model languages is achieved through a language specific driver and API.

%---------------------------------------------------------------------------------
% MODEL DRIVERS
\subsubsection{Model Drivers}\label{SS:model_driver}
%
Model drivers handle model execution, monitoring, and compilation if necessary. 
While every model is executed on a new process, how the model is handled depends 
on the language it is written in. Models written in 
interpreted languages (Python and Matlab) are executed on the command line 
with the interpreter. In the case of Matlab, where there is significant overhead 
associated with starting the Matlab interpreter, a Matlab shared engine is used 
to execute the model. To speed up the execution, the shared engine can be started 
in advance and then reused.

For the compiled languages (C and C++) there are a few options. The user can 
compile the model themselves, provided they link against the appropriate 
{\cis} header library. Alternatively, users can provide the location of 
the model source code and let {\cis} handle the compilation, including 
linking against the appropriate {\cis} header library. {\cis} 
also has support for compiling models using Make \citep{Stallman2004} and 
CMake \citep{Martin2006} for 
models that already have a Makefile or CMakeLists.txt. To use these drivers, 
lines are added to the recipe in order to allow linking against {\cis}.

%---------------------------------------------------------------------------------
% COMMUNICATION
\subsection{Communication}\label{SS:communication}
%
Communication within {\cis} was designed to be flexible in terms of the 
languages, platforms, and data types. To accomplish this, {\cis} leverages 
several tools different tools for communication which are applied as needed. The 
particular mechanism used by {\cis} for communication is hidden from the 
user, who will always use the same interface in the language of their model 
(See \S\ref{SSS:interface}).

%---------------------------------------------------------------------------------
% ASYNCHRONOUS MESSAGE MASSING
\subsubsection{Asynchronous Message Passing}\label{SSS:async}
%
Regardless of the specific communication mechanism, the same asynchronous
communication strategy is used. Models do not block on sending messages to 
output channels; the model is free to continue working on its task while 
a separate thread waits for the message to be routed and received. Similarly, 
a thread continuously pings input channels, moving received messages into a 
a buffer queue so that they are ready and waiting for the receiving model 
when it asks for input. As a result, models in complex integration networks 
can work in parallel, improving the overall efficiency with which computational 
resources are used. Figure \ref{fig:async} describes the general flow of messages.
%
\ifinclfig
 	\begin{figure}[htbp]
	\begin{center}
	\includegraphics[width=\columnwidth,keepaspectratio]{./images/io_drivers.png}
	\caption{Diagram of how messages are passed asynchronously using input/output drivers and an intermediate channel.}
	\label{fig:async}
	\end{center}
	\end{figure}
\fi
%
\begin{itemize}
	\item A model sends a message in the form of a native data object via a language 
specific API to one of the output channels declared in the model YAML. The output channel interface encodes the message and sends it.
	\item An output connection driver (written in Python) runs in a separate thread, listening to the model output channel. When the model sends a message, the model side connection driver checks that the message is in the expected format and then forwards it to an intermediate channel. The intermediate channel may seem unnecessary, but it is used as a buffer for future support on distributed architectures (e.g. if one model is running on a remote machine). In these cases, the intermediate channel will connect to a RabbitMQ broker using security credentials.
	\item An input connection driver (also written in Python) runs in a separate thread, listening to the intermediate channel. When a message is received, it is then forwarded to the input channel of the receiving model as specified in the integration network YAML.
	\item The receiving model receives the message.
\end{itemize}

In additional to communication between two models, users can also specify that a 
model should receive/send input/output from/to a file. This is specified in 
the integration network YAML and does not impact the way the model will 
receive/send messages.

%---------------------------------------------------------------------------------
% SYSTEM V IPC
\subsubsection{System V IPC Queues}\label{SSS:ipc}
%
The first communication mechanism used by {\cis} was System V interprocess 
communication (IPC) message queues \citep{Rusling1999} on Posix (Linux and Mac OSX) systems. 
IPC message queues allow messages to be passed between models running on separate 
processes on the same machine. While IPC message queues are light weight, fast 
(See \S\ref{S:results}), and are part of Posix operating systems, they do not work in all 
situations. Sys V IPC queues are not natively supported by Windows operating systems 
and do not allow communication between remote processes. In addition, IPC queues also 
have relatively low default message size limits on Mac OSX systems (2048 bytes). 
While this can be handled by splitting large messages into multiple smaller messages 
(See \S\ref{SSS:large}), the time required to send a message increases with the number of 
message it must be broken into (See \S\ref{S:results}). As a result, Sys V IPC queues are considered by 
{\cis} to be a fallback on Posix systems if the necessary ZeroMQ libraries have 
not been installed.

%---------------------------------------------------------------------------------
% ZEROMQ
\subsubsection{ZeroMQ}\label{SSS:zmq}
%
The preferred communication mechanism used by {\cis} are ZeroMQ sockets 
\citep{Akgul2013}. ZeroMQ 
provides broker-less communication via a number of protocols and patterns with 
bindings in a wide variety of languages that can be installed on Posix and Windows 
operating systems. ZeroMQ was adopted by {\cis} in order to allow support 
on Windows and for future target languages (See \S\ref{SS:future}) that could not be 
accomplished using System V IPC queues. In addition, while ZeroMQ allows 
interprocess communication via IPC queues like System V IPC queues, ZeroMQ also 
supports protocols for distributed communication via an Internet Protocol (IP) 
network. While {\cis} does not currently support using these protocols 
for distributed integration networks, this one of the plans for future improvement.

%---------------------------------------------------------------------------------
% RABBITMQ
\subsubsection{RabbitMQ}\label{SSS:rmq}
%
While ZeroMQ provides broker-less communication means, {\cis} includes support 
for brokered communication via RabbitMQ \citep{RMQ}. {\cis} does not currently use 
RabbitMQ for communication unless explicitly specified by the user in their 
integration network YAML. RabbitMQ support was originally added to {\cis} 
as a supplement to System V IPC queues in the case of future support for distributed 
integration networks. In future development, RabbitMQ brokered communication will be 
used for establishing integration networks with remote models run a services 
(See \S\ref{SS:future}).

%---------------------------------------------------------------------------------
% SENDING/RECEIVING LARGE MESSAGES
\subsubsection{Sending/Receiving Large Messages}\label{SSS:large}
%
All of the communication tools leveraged by {\cis} have intrinsic limits 
on the allowed size for a single message. Some of these limits can be quite large 
($2^{20}$ for ZeroMQ and RabbitMQ), while others are very limiting 
(2048 on Mac OSX for Sys V IPC queues). Although messages consisting of a few scalars are unlikely to 
exceed these limits, biological inputs and outputs are often much more complex. 
For example, structural data represented as a 3D mesh can easily exceed these 
limits. To handle messages that are larger than the limit of the communication 
mechanism being used, {\cis} splits the message up into multiple smaller 
messages. In addition, for large messages, {\cis} creates new, temporary 
communication channels that are used exclusively for a single message and 
then destroyed. The address associated with the temporary channel is send in 
header information as a message on the main channel along with information about 
the message that will be sent through the temporary channel like size and type. 
Temporary channels are used for large messages to prevent mistakenly combining 
the pieces from two different large messages that were received at the same time 
such as in the case that a model is receiving input from two different models 
working in parallel.

%---------------------------------------------------------------------------------
% INTERFACE
\subsubsection{Interface}\label{SSS:interface}
%
{\cis} provides functions and classes for communication that are 
written in each of the supported languages. This allows users to program in the 
language(s) they are already familiar with. The Python interface provides 
communication classes for sending and receiving messages. The Matlab 
interface provides a simple wrapper class for the Python class, that exposes 
the appropriate methods and handles conversion between Python and Matlab 
data types. The C interface provides structures and functions for accessing 
communication channels and sending/receiving messages. The C++ interface 
provides classes that wrap the C structures with functions called as methods.

In addition to basic input and output, each interface also provides access 
to more complex data types and communication patterns.

%---------------------------------------------------------------------------------
% SERIALIZATION
%
\subsection{Transformation}\label{SS:transformation}
%
%\subsubsection{Data Formats}\label{SSS:dataformats}
%
Messages are passed as raw bytes. In order to understand the messages begin passed, parallel processes that communicate must agree upon the format used to do so. Without community standards, different models will often use very different data formats for their input and output. Differences between data formats can include, but are not limited to, type, precision, fields, or units. While some data formats are self-descriptive and include these types of information as meta data, this is not true of all data formats. To combat this, {\cis} requires models explicitly specify the format of input and output expected by a model in the model YAML. {\cis} can then handle a number of conversion between models without prompting as well as serialization/deserialization to the correct type in . Data formats currently supported by {\cis} include:
%
\begin{itemize}
	\item Scalars (e.g. integers, floats)
	\item Arrays
	\item Tabular (e.g. CSV or tab-delimited)
	\item Pandas data frames
	\item PLY
	\item OBJ
\end{itemize}

%\subsubsection{Units}\label{SSS:units}
%
In addition, {\cis} offers the option to specify units for scalars, arrays, tabular data, and pandas data frames. Units are tracked using the {\tt unyt} package \citep{Goldbaum2018}. If two models use different units (and both are specified), {\cis} will automatically perform the necessary conversions before passing data from one model to the next.

%---------------------------------------------------------------------------------
%---------------------------------------------------------------------------------
% RESULTS
\section{Results}\label{S:results}
%
In order to evaluate {\cis} performance tests were run on Linux, Mac OSX, and Windows for different communication mechanisms and language combinations. During each run, $N_{msg}$ messages of size $S_{msg}$ were sent from the source model (in one language) to the destination model (in another language) which then output the messages to file for verification. Performance tests were run using the perf package \citep{Stinner2018}. Each run was repeated 20 times to rule out fluctuations due to external loads on the test machines.

% COMMTYPE
\subsection{Communication Mechanism}
%
Figure \ref{fig:commtype} compares the execution times for the different communication mechanisms discussed in \S\ref{SS:communication} for sending different numbers/sizes of messages from one Python model to another Python model. The tests were run on a \todo{Dell?} tower with 10 quad-core Intel(R) Xeon(R) CPU E5-2650 v3 @ 2.30GHz processors running Ubuntu 14.04. The left panel of Figure \ref{fig:commtype} shows the total time required to run both models and send a varying number of 1000 bytes messages from one model to the other. The right panel of Figure \ref{fig:commtype} shows the total time required to run the models and send 5 messages of varying lengths from one to the other.
%
\ifinclfig
 	\begin{figure}[htbp]
	\begin{center}
	\includegraphics[width=\columnwidth,keepaspectratio]{./images/scaling_commtype.png}
	\caption{Comparison of communication time scaling across communication methods on Linux. Left: Scaling of execution time with number of 1000 byte messages sent. Right: Scaling of execution time with the size of the 5 messages sent.}
	\label{fig:commtype}
	\end{center}
	\end{figure}
\fi
%
Although Sys V IPC queues are slightly faster than ZeroMQ TCP sockets (\todo{0.012} s/msg vs. \todo{0.019} s/msg), IPC queues have a much smaller maximum size limit for messages. As discussed in \S\ref{SSS:large}, messages larger than this limit (2048 bytes) must be broken up into multiple messages. This compounds the time required to send these messages making IPC queues a poor choice for sending data over $10^5$ bytes. This limit does not kick in for ZeroMQ sockets until much later ($\sim10^6$ bytes).

% LANGUAGE
\subsection{Language}
%
Figure \ref{fig:language} compares the execution times for integrations with communications between models in different combinations of languages. The tests were run on \todo{the Linux machine from above} using ZeroMQ communication.
%
\ifinclfig
 	\begin{figure}[htbp]
	\begin{center}
	\includegraphics[width=\columnwidth,keepaspectratio]{./images/scaling_language.png}
	\caption{Comparison of communication time scaling between languages on Linux.}
	\label{fig:language}
	\end{center}
	\end{figure}
\fi
%
For the most part, performances are not affected by the language of either model. Each language entails some overhead. For Python and Matlab, the overhead comes from starting the interpreter. For C and C++, the overhead comes from compiling the source code.
%
\todo{results for Matlab}

% PYTHON VERSION
\subsection{Python Version}
%
Figure \ref{fig:python} compares the execution times when using different versions of Python. The performance tests for both Python 2.7 and Python 3.5 were performed on the Linux machine from above using ZeroMQ communication.
%
\todo{figure and results}

% PLATFORM
\subsection{Platform}
%
Figure \ref{fig:platform} compares the execution times on different platforms. The performance tests reported for Linux were run on the machine from above. The tests for Mac OS X were run on a 2015 MacBook Pro with a 2.9 GHz Intel Core i5 processor running macOS High Sierra 10.13.4. The tests for Windows were run on the same machine from a dual boot of Windows 10. All platform dependent performance tests use ZeroMQ communication.
%
\todo{figure and results}

%---------------------------------------------------------------------------------
%---------------------------------------------------------------------------------
% SUMMARY & DISCUSSION
\section{Summary \& Discussion}\label{S:discuss}

%---------------------------------------------------------------------------------
% CURRENT STATUS
\subsection{Current Status}\label{SS:current}
%
{\cis} is an open-source Python package for connecting computational models across programming languages in scales to form integration networks. Models are run in parallel with asynchronous message passing handled under-the-hood via threading and one of three communication mechanisms. {\cis} works on Linux, Mac OSX, and Windows operating systems with Python 2.7 and $3.4+$. {\cis} currently supports running models written in Python, C, C++, and Matlab with additional support for compiling models C/C++ using Make \citep{Stallman2004} or CMake \citep{Martin2006} and running models written in the LPy \citep{Boudon2012} DSL.

{\cis} has already been used to integrate several plant models within the Crops in Silico organization with ongoing work to add additional models to the integration network. Figure \ref{fig:network} shows the progress so far. Each square is a model and lines between the models represent connections where information is exchanged. Solid lines are connections we have already implemented, dashed lines are connections we are currently working on, and dotted lines are connections planned for the future. Results from one integration can be found in \todo{cite Kavya/Yu}.
%
\ifinclfig
 	\begin{figure}[htbp]
	\begin{center}
	\includegraphics[width=\columnwidth,keepaspectratio]{./images/CiS-BioCro-Planned.png}
	\caption{Progress towards full Crops in Silico integration network.}
	\label{fig:network}
	\end{center}
	\end{figure}
\fi
%
\todo{more here? teaser results?}

%---------------------------------------------------------------------------------
% FUTURE WORK
\subsection{Current/Future Improvements}\label{SS:future}
%
{\cis} is being actively developed to expand the number of models 
that can be used in integration networks and the complexity of integration 
networks that can be handled. Several improvements are already in progress/planned for {\cis}.

% LANGUAGE SUPPORT
\subsubsection{Language Support}\label{SS:langsupport}
%
The biggest barrier to running new models is language support. Plant models 
are written in many different languages. While the current language support 
covers many, additional languages must be added to unlock the full potential. 
Surveys of the plant modeling community have helped identify the core languages 
that models have been written in. Based on these results, we plan to expand 
{\cis} to support models written in R, Fortran, Java, and the SBML \citep{Hucka2003} DSL.
We also plan to add support for running Matlab models using Octave \citep{Eaton2002}. Matlab 
models require a Matlab license to run. Given that plant modelers do not all 
use Matlab, it cannot be assumed that everyone will have access to a Matlab 
license. Octave is open source and provides much of the same functionality 
as Matlab and can run many Matlab codes. Support for Octave will improve 
modelers ability to collaborate without worrying about access to a Matlab 
license.

% DISTRIBUTED SYSTEMS
\subsubsection{Distributed Systems}\label{SS:distributed}
%
High performance computing (HPC) and cloud compute resources are powerful 
tools in the current computing ecosystem that could be used for running 
complex integration networks. To this end, we plan to expand {\cis} 
support for running integration networks on distributed compute resources. 
{\cis} already uses communication tools that can be adapted for use 
in a distributed pattern. We will leverage tools like the libsubmit package from the 
Parsl project \citep{babuji18} for 
automating the submission process to HPC and compute resources. In addition, 
we will add tools for running models as a service and using RabbitMQ to 
permit access to these models within integration networks.

% CONTROL FLOW
\subsubsection{Control Flow}\label{SS:control}
%
Currently, modelers must explicitly specify how a model should process input 
within the model code including things like which input variables are static 
and read in once versus which variables change and should be looped over. 
We plan on adding options to {\cis} for dynamically generating model code 
based a user provided function call and list of static/variable input channels. 
This will allow better reusability of model code such as during parameter studies.

% DATA AGGREGATIOIN
\subsubsection{Data Aggregation}\label{SS:dataagg}
%
Much of the original version of {\cis} centered around the use of 
tabular input data. While tabular data is used heavily by plant models, it 
presents several barriers for constructing integration networks. Tabular input 
data only works when one model outputs the same columns that are expected as 
input by another model. However, this is unlikely to be the case if two models 
are developed independently. Future improvements to {\cis} will allow 
tabular input to models to be composed by aggregating output data from more than 
one model and/or file.

% JSON DATA TYPE SPECIFICATION
\subsubsection{JSON Data Type Specification}\label{SS:json}
%
While {\cis} supports serialization of several data types/formats, we 
would like to make serialization as flexible as possible. To this end, 
{\cis} will be adapted to read JSON schema \citep{jsonschema} for user defined types that 
can be used to automatically create the appropriate data structures alongside
methods for parsing and serializing them.

% GUI
\subsubsection{Graphical User Interface}\label{SS:gui}
%
In an effort to make {\cis} as user friendly as possible, we have also begun work on a graphical user interface (GUI) for entering model information, composing integration networks from an existing palette of models, and displaying basic output from an integration run. The current prototype of the GUI can only handle model ingestion and network composition; model execution must be done locally. However, our ultimate goal with the GUI is to provide a service where users can register their model, compose and run integration networks on dedicated cloud compute resources, and view output in real time all from the web browser. \todo{Mike and Craig as authors/acknowledgements? Provide \href{https://hackathon.cis.ndslabs.org/}{link?}}

%---------------------------------------------------------------------------------
% CODE
\section*{Code}\label{S:code}
The {\cis} package is available publicly on \href{https://github.com/cropsinsilico/cis_interface}{Github} and can be 
installed using {\tt pip} or {\tt conda}. To ensure code health, {\cis} boasts 100\% code coverage with automated testing on Linux, Mac OSX, and Windows for Python version 2.7, 3.4, 3.5, 3.6, and 3.7 via continuous integration with TravisCI \citep{travisci} and Appveyor \citep{appveyor}. Full documentation for {\cis} can be found \href{https://cropsinsilico.github.io/cis_interface/}{here} including step-by-step directions from the tutorial conducted during the \href{https://cropsinsilico.github.io/cis_interface/hackathon2018/index.html}{2018 Crops in Silico Hackathon}.

%---------------------------------------------------------------------------------
% ACKNOWLEDGMENTS
%\section*{Acknowledgment}
%\acknowledgments
%The authors would like to thank ...

%---------------------------------------------------------------------------------
% REFERENCES
\todo{Fix code citations}
\ifdraft
	\bibliography{CiS}
\else
	\input{ms-bbl.tex}
\fi

%---------------------------------------------------------------------------------
% FIGURES
\ifplacefig
\else
% put figures here if figures go after
\fi

%---------------------------------------------------------------------------------
% COMMENTS
%\newpage
%\todo{To do:
%\begin{itemize}
%\item Improve conclusion
%\item Non-uniform test data set?
%\item Performance of runs on cosmology dataset?
%\item Move figures to the end for submission
%\item iPython notebook for running tests
%\end{itemize}
% }

\end{document}

### 2018-10-08 papers/lang2018/ms.tex
Update figures

@@
 %
-\todo{results for Matlab}
+\todo{results for Matlab \& discussion}
 
@@
 %
-\todo{figure and results}
+\ifinclfig
+ 	\begin{figure}[htbp]
+	\begin{center}
+	\includegraphics[width=\columnwidth,keepaspectratio]{./images/scaling_python.png}
+	\caption{Comparison of communication time scaling between Python versions on Linux.}
+	\label{fig:python}
+	\end{center}
+	\end{figure}
+\fi
+%
+\todo{discussion}
 
@@
 %
-\todo{figure and results}
+\ifinclfig
+ 	\begin{figure}[htbp]
+	\begin{center}
+	\includegraphics[width=\columnwidth,keepaspectratio]{./images/scaling_platform.png}
+	\caption{Comparison of communication time scaling between operating systems.}
+	\label{fig:platform}
+	\end{center}
+	\end{figure}
+\fi
+%
+\todo{Windows results \& discussion}
 

### 2018-10-08 papers/lang2018/ms.tex
Edits to text and begin adding tables

@@
 \begin{abstract}
-Thousands of computational models have been created within the plant biology community within the past two decades that have the potential to be combined into complex integration networks capable of capturing more complex biological processes than possible as isolated models.
-However, the technological barriers introduced by differences in language and data formats as slowed this progress.
+Thousands of computational models have been created within the plant biology community in the past two decades that have the potential to be combined into complex integration networks capable of capturing more complex biological processes than possible with isolated models.
+However, the technological barriers introduced by differences in language and data formats has slowed this progress.
 We present \href{https://github.com/cropsinsilico/cis_interface}{\cis}, a Python package for running integration networks with connections between models across languages and scales.
-{\cis} coordinates parallel execution of models in Python, C, C++, and Matlab and handles communication in a number of data formats common to computational plant modeling.
-{\cis} is designed to be user-friendly, including visual tools for composing the integration networks.
+{\cis} coordinates parallel execution of models in Python, C, C++, and Matlab on Linux, Mac OSX, and Windows operating systems, and handles communication in a number of data formats common to computational plant modeling.
+{\cis} is designed to be user-friendly, with visual tools developed in parallel for composing the integration networks.
 \end{abstract}
@@
 %\IEEEPARstart{C}{omputational} 
-Computational plant biologists have produced a wealth of computational models for different biological processes governing plant growth and development, covering scales from atomic to global.
+Computational plant biologists have produced a wealth of computational models to describe the biological processes governing plant growth and development, covering scales from atomic to global.
 Although usually developed independently to address questions specific to an organ, species, or growth process, computational plant models can also have applications beyond their original scope.
-For example, a photosynthesis model developed for maize can also be adapted to work for other species like soy \todo{cite Kavya and Yu}.
-Computational models for different scales, organs, or processes also have the potential for being combined to probe more complex biological mechanisms.
-With advances in compute power, it should be possible to link biologically related models to create integration networks capable of capturing the response of entire plants, fields, or regions.
+Many biological processes are the same across different species of plants, allowing models developed for one species to be adapted for another by modifying input parameters.
+In addition, computational models for different scales, organs, or processes can also be combined, or integrated, to probe more complex biological mechanisms.
+By integrating a metabolic flux model with a genetic regulatory model for carbon uptake, \citet{integration_prep} was able to identify gene candidates for regulating photosynthesis under elevated CO$_{2}$.
+With advances in compute power, it should be possible to link biologically related models to create complex integration networks capable of capturing the response of entire plants, fields, or regions.
 However, independently developed computational models often have compatibility issues resulting from differences in programming language or data format that make such collaborations difficult.
 
-We present an Open Source Python package, {\cis}, for creating and running integrations networks from existing computational models written in Python, Matlab, C and C++.
+We present an Open Source Python package, {\cis}, for creating and running integration networks by connecting existing computational models written in Python, Matlab, C and C++.
 Although developed for the purpose of integrating plant biology models, {\cis} can be applied to any situation requiring coordination between models in the supported languages.
@@
 Scientists will generally write programs in the language they are most familiar with and, given that there is no consensus on one programming language as ideal for all scientific applications, the language of computational models will often vary between research groups, if not between members of the groups themselves.
-The same can be said of data formats which can be highly tailored to the target question and may or may not include metadata like the data types, fields, or units.
+The same can be said of data formats which can be highly tailored to the target question and may or may not include metadata like the data types, field names, or units.
 To integrate two biological models, the following questions must be addressed:
@@
 %
-Another approach is a workflow/data flow/framework tools for models that are written in the same programming language or can be wrapped.
+Another approach is a workflow/data flow/framework tool for models that are written in the same programming language or can be wrapped.
 For example, there are many generic tools in Python for coordinating the execution of tasks in parallel or serial \citep[e.g.][]{babuji18, celery, luigi}.
-While these tools are powerful for organizing complex work flows, it is the user's onus to make sure that the components they connect are compatible and handle an sort of data transformation that might be necessary (e.g.
+While these tools are powerful for organizing complex work flows, it is the user's onus to make sure that the components they connect are compatible and handle any data transformation that might be necessary (e.g.
 unit conversion or field selection).
@@
 There are also domain specific frameworks for connecting biological models.
-For example, OpenAlea \citep{Pradal2015} allows users to compose networks from different components including plant models, analysis tools, and visualizations.
+For example, OpenAlea \citep{Pradal2015} allows users to compose networks from different components including plant models, analysis tools, and visualization tools.
 Domain specific frameworks are more intuitive and ultimately more flexible than DSLs in the types of operations they allow models to perform since they have access to the full power of a programming language.
@@
 	\item Easy to use. Require as little modification to the model source code as possible and only in the language of the model itself.
-	\item Efficient. Allow models to run in parallel with asynchronous communication that dosen't block when a message is sent.
+	\item Efficient. Allow models to run in parallel with asynchronous communication that dosn't block when a message is sent.
 	\item Flexible. Provide the same interface to the user, regardless of the communication mechanism being used or the platform the model is being executed on.
@@
 %
+{\cis} is executed via a command line interface (CLI), {\tt cisrun} with YAML specification files (\S\ref{SSS:yaml}) as input.
+Based on the information contained in the YAML specification files, {\cis} establishes a network of asynchronous communication channels (\S\ref{SS:communication}), and launches the models on new processes (\S\ref{SS:model_driver}).
 %At execution, {\cis} launches models on new processes based on information contained in a YAML specification file (See \S\ref{SSS:yaml} below) and establishes
@@
 YAML files \citep{Ben-Kiki2009}. The YAML file format was selected because it is human readable 
-and there are many existing tools for parsing YAML formats in many different 
+and there are many existing tools for parsing YAML formats in different programming 
 languages. The declarative format allows user to specify exactly what they 
-what to do, without describing how it should be done. While the information 
+want to do, without describing how it should be done. While the information 
 about models and integration networks can be contained in a single YAML 
-file, the information can naturally be split between two files, one containing 
-information about the models and one containing the connections comprising the 
+file, the information can naturally be split between two or more files, one (or more) containing 
+information about the model(s) and one containing the connections comprising the 
 integration network. This separation is advantageous because the model YAML 
@@
 model source code, the language the model is written in, how the model should 
-be run, and any input or output variables including their type and units. 
+be run, and any input or output variables including their data type and units. 
 Integration networks are specified via YAML files declaring the connections 
@@
 one model with the input variable of another model. The {\cis} 
-command line interface (CLI) sets up the necessary communication mechanisms to 
+CLI sets up the necessary communication mechanisms to 
 then direct data from one model to the next in the specified pattern. Models 
@@
 with the interpreter. In the case of Matlab, where there is significant overhead 
-associated with starting the Matlab interpreter, a Matlab shared engine is used 
+associated with starting the Matlab interpreter (See \S\ref{SS:results_language}), a Matlab shared engine is used 
 to execute the model. To speed up the execution, the shared engine can be started 
@@
 lines are added to the recipe in order to allow linking against {\cis}.
+
+Once model are running, each model driver uses a thread to monitor the progress of the model and report back model output (e.g.
+log messages printed to stdout) and any status changes to the master process.
+If a model issues any errors, the master process will shut down any model processes that are still running.
+If a model completes without any errors, the master process will cleanup any input/output channels that are no longer required.
+The master process will only complete once an error is encountered or all model processes have completed.
 
@@
 a thread continuously pings input channels, moving received messages into a 
-a buffer queue so that they are ready and waiting for the receiving model 
+buffer queue so that they are ready and waiting for the receiving model 
 when it asks for input. As a result, models in complex integration networks 
@@
 %
-\begin{itemize}
+\begin{enumerate}
 	\item A model sends a message in the form of a native data object via a language 
@@
 	\item An output connection driver (written in Python) runs in a separate thread, listening to the model output channel.
-When the model sends a message, the model side connection driver checks that the message is in the expected format and then forwards it to an intermediate channel.
+When the model sends a message, the output connection driver checks that the message is in the expected format and then forwards it to an intermediate channel.
 The intermediate channel may seem unnecessary, but it is used as a buffer for future support on distributed architectures (e.g.
@@
 When a message is received, it is then forwarded to the input channel of the receiving model as specified in the integration network YAML.
-	\item The receiving model receives the message.
-\end{itemize}
+	\item The receiving model receives the message in the form of an analogous native data object.
+\end{enumerate}
 
-In additional to communication between two models, users can also specify that a 
+In addition to communication between two models, users can also specify that a 
 model should receive/send input/output from/to a file. This is specified in 
 the integration network YAML and does not impact the way the model will 
-receive/send messages.
+receive/send messages. In this way, users can test their model integration in 
+isolation with input/output form/to files rather than another model.
 
@@
 processes on the same machine. While IPC message queues are light weight, fast 
-(See \S\ref{S:results}), and are part of Posix operating systems, they do not work in all 
+(See \S\ref{SS:results_commtype}), and are part of Posix operating systems, they do not work in all 
 situations. Sys V IPC queues are not natively supported by Windows operating systems 
 and do not allow communication between remote processes. In addition, IPC queues also 
-have relatively low default message size limits on Mac OSX systems (2048 bytes). 
-While this can be handled by splitting large messages into multiple smaller messages 
-(See \S\ref{SSS:large}), the time required to send a message increases with the number of 
-message it must be broken into (See \S\ref{S:results}). As a result, Sys V IPC queues are considered by 
-{\cis} to be a fallback on Posix systems if the necessary ZeroMQ libraries have 
+have relatively low default message size limits on Mac OSX systems (2048 bytes). Once 
+the queue is full or if the message is larger than the limit, any process attempting to send an additional message will block until a 
+sufficient number of messages has been removed from the queue to accommodate the new message. 
+For messages larger than the limit, the sending process will block indefinitely. 
+This can be handled by splitting large messages into multiple smaller messages 
+(See \S\ref{SSS:large}), however, the time required to send a message increases with the number of 
+message it must be broken into (See \S\ref{SS:results_commtype}). These limits make sending large messages 
+relatively inefficient when compared with other communications mechanisms. 
+As a result, Sys V IPC queues are used by 
+{\cis} as a fallback on Posix systems if the necessary ZeroMQ libraries have 
 not been installed.
@@
 network. While {\cis} does not currently support using these protocols 
-for distributed integration networks, this one of the plans for future improvement.
+for distributed integration networks, this is part of the plans for future improvement.
 
@@
 communication channels that are used exclusively for a single message and 
-then destroyed. The address associated with the temporary channel is send in 
+then destroyed. The address associated with the temporary channel is sent in 
 header information as a message on the main channel along with information about 
-the message that will be sent through the temporary channel like size and type. 
+the message that will be sent through the temporary channel like size and data type. 
 Temporary channels are used for large messages to prevent mistakenly combining 
@@
 %
-{\cis} provides functions and classes for communication that are 
-written in each of the supported languages. This allows users to program in the 
-language(s) they are already familiar with. The Python interface provides 
+{\cis} provides interface functions/classes for communication that are 
+written in each of the supported languages. Language specific interfaces 
+allow users to program in the 
+language(s) with which they are already familiar. The Python interface provides 
 communication classes for sending and receiving messages. The Matlab 
@@
 While some data formats are self-descriptive and include these types of information as meta data, this is not true of all data formats.
-To combat this, {\cis} requires models explicitly specify the format of input and output expected by a model in the model YAML.
-{\cis} can then handle a number of conversion between models without prompting as well as serialization/deserialization to the correct type in .
+To combat this, {\cis} requires models to explicitly specify the format of input and output expected by a model in the model YAML.
+{\cis} can then handle a number of conversion between models without prompting as well as serialization/deserialization to the correct type in each of the supported languages.
 Data formats currently supported by {\cis} include:
@@
 %
-In order to evaluate {\cis} performance tests were run on Linux, Mac OSX, and Windows for different communication mechanisms and language combinations.
+In order to evaluate {\cis}, performance tests were run on machines with Linux, Mac OSX, and Windows operating systems for different communication mechanisms, Python versions, and language combinations.
 During each run, $N_{msg}$ messages of size $S_{msg}$ were sent from the source model (in one language) to the destination model (in another language) which then output the messages to file for verification.
 Performance tests were run using the perf package \citep{Stinner2018}.
-Each run was repeated 20 times to rule out fluctuations due to external loads on the test machines.
+Each run was repeated 10 times to rule out fluctuations due to external loads on the test machines.
 
 % COMMTYPE
-\subsection{Communication Mechanism}
+\subsection{Communication Mechanism}\label{SS:results_commtype}
 %
@@
 	\caption{Comparison of communication time scaling across communication methods on Linux.
+Lines show the average run time for each test with the standard deviation shown in the shaded region.
 Left: Scaling of execution time with number of 1000 byte messages sent.
@@
 %
-Although Sys V IPC queues are slightly faster than ZeroMQ TCP sockets (\todo{0.012} s/msg vs.
-\todo{0.019} s/msg), IPC queues have a much smaller maximum size limit for messages.
-As discussed in \S\ref{SSS:large}, messages larger than this limit (2048 bytes) must be broken up into multiple messages.
-This compounds the time required to send these messages making IPC queues a poor choice for sending data over $10^5$ bytes.
-This limit does not kick in for ZeroMQ sockets until much later ($\sim10^6$ bytes).
+
+Table \ref{tab:commtype} shows how the two communication mechanisms compare as determined by performing a linear fit to the scaling of execution time with message count from Figure \ref{fig:commtype}.
+Time per message (slope of Figure \ref{fig:commtype}) is the average amount of time required to send a single message containing 1000 bytes and constrains how quickly messages can be passed between the models.
+Overhead (intercept of Figure \ref{fig:commtype}) is the amount of execution time that would be required if no messages were passed between the models and includes the time required to set up communication mechanisms, start the model drivers, and clean up the integration network.
+%
+\begin{table}[htbp]
+%\caption{The time required per message and overhead for startup/teardown of integration for different communication mechanisms as determined by a linear fit to the scaling of execution time with message count.
+}
+\begin{center}
+\begin{tabular}{|c|c|c|}
+\hline
+Mechanism	& Time per Message (s) 	& Overhead (s) 	\\\hline
+IPC			& 0.010				& 1.60			\\
+ZMQ 		& 0.012				& 1.73			\\\hline
+\end{tabular}
+\end{center}
+\caption{Effect of communication mechanism on performance.}
+\label{tab:commtype}
+\end{table}%
+%
+Although Sys V IPC queues are slightly faster than ZeroMQ TCP sockets in both time per message and overhead for messages smaller than the limit, IPC queues have a much smaller maximum size limit for messages.
+As discussed in \S\ref{SSS:large}, messages larger than this limit (2048 bytes) must be broken up into multiple messages, compounding the time required to send these messages and making IPC queues a poor choice for sending data over $\sim10^5$ bytes.
+This limit is much larger for ZeroMQ sockets ($\sim10^6$ bytes).
+However, for messages less than the limit, both ZMQ and IPC communication has no measurable dependence on message size.
 
 % LANGUAGE
-\subsection{Language}
+\subsection{Language}\label{SS:results_language}
 %
-Figure \ref{fig:language} compares the execution times for integrations with communications between models in different combinations of languages.
-The tests were run on \todo{the Linux machine from above} using ZeroMQ communication.
+Figure \ref{fig:language} compares the execution times for integrations with communications between models in different combinations of languages and Table \ref{tab:language} reports the time per message and overhead.
+The runs for Figure \ref{fig:language} and Table \ref{tab:language} were run on the Linux machine from above while those for Figure \ref{fig:language_matlab} and Table \ref{tab:language_matlab} were run on a 2015 MacBook Pro with a 2.9 GHz Intel Core i5 processor running macOS High Sierra 10.13.4 that had Matlab R2017a installed.
+In both cases, ZeroMQ communication was used.
 %
@@
 	\includegraphics[width=\columnwidth,keepaspectratio]{./images/scaling_language.png}
-	\caption{Comparison of communication time scaling between languages on Linux.}
+	\caption{Comparison of communication time scaling between languages on Linux. The same fiducial message size (1000 bytes) and count (5) was used as in Figure \ref{fig:commtype}.}
 	\label{fig:language}
@@
 %
-For the most part, performances are not affected by the language of either model.
-Each language entails some overhead.
-For Python and Matlab, the overhead comes from starting the interpreter.
-For C and C++, the overhead comes from compiling the source code.
+\begin{table}[htbp]
+\begin{center}
+\begin{tabular}{|c|c|c|c|}
+\hline
+Source 	& Destination 	& Time per Message (s) 	& Overhead (s) 	\\\hline
+C 				& C					& 0.11				& 1.36			\\
+C 				& C++				& 0.12				& 1.42			\\
+C 				& Python				& 0.12				& 1.99			\\
+\hline%
+C++ 				& C					& 0.12				& 1.42			\\
+C++ 				& C++				& 0.11				& 1.49			\\
+C++ 				& Python				& 0.12				& 2.05			\\
+\hline%
+Python			& C					& 0.12				& 2.15			\\
+Python			& C++				& 0.12				& 2.21			\\
+Python			& Python				& 0.12				& 1.73			\\
+\hline
+\end{tabular}
+\end{center}
+\caption{Effect of model language on performance (Linux).}
+\label{tab:language}
+\end{table}%
 %
-\todo{results for Matlab \& discussion}
+For the most part, the time required per message is not affected by the language of either model.
+There is a very small increase in the time per message when there is a Python model involved, but this is much lower than the standard deviation resulting from background activity on the test machine and so is in conclusive.
+
+The largest difference between the different test cases is in in the overhead.
+Every model driver entails some overhead.
+For interpreted languages (e.g.
+Python and Matlab), the majority of the overhead comes from starting the interpreter.
+For Matlab, this is particularly time consuming.
+Although this is only a one-time cost required at the start of an integration, {\cis} does offer ways to alleviate this if multiple runs are necessary by starting a Matlab engine prior to the first integration that can be used by subsequent runs \todo{(results that include existing Matlab engine)}.
+
+For compiled languages (C and C++), the overhead comes from compiling the source code which is done in serial.
+The time required for compilation will depend on the compiler used and the complexity of the model code.
+For command line compiled models, {\cis} forces the model to be recompiled every time to ensure any changes to the source code are propagated.
+However, for models that use Make and CMake, overhead due to compile time can be reduced by using existing builds.
+
+Interestingly, the recorded overheads were not symmetric or additive.
+For example, the Python-to-C integrations required more overhead than the C-to-Python integrations.
+This is because the models are started in parallel, but do not require the same amount of time to start.
+In a Python-to-C integration, the C model must wait for the Python model to finish its much slower startup and begin sending message, while in a C-to-Python integration, the C model can immediately begin sending messages that will be received by the Python model once it finishes starting up.
+In addition, the Python-to-Python integrations required less time than any of the other integrations because, although Python models take longer to start up, the start up is \todo{explanation?}
+%
+\todo{figure/table for Matlab \& discussion}
 
 % PYTHON VERSION
-\subsection{Python Version}
+\subsection{Python Version}\label{SS:results_python}
 %
@@
 % PLATFORM
-\subsection{Platform}
+\subsection{Operating System}\label{SS:results_platform}
 %
-Figure \ref{fig:platform} compares the execution times on different platforms.
+Figure \ref{fig:platform} compares the execution times on different operating systems.
 The performance tests reported for Linux were run on the machine from above.
@@
 %---------------------------------------------------------------------------------
+% TODO
+\section*{TODO}
+\todo{%
+\begin{itemize}
+	\item Fix code citations to include URLs
+	\item Increase size of font on figures or increase size of figures (depends on iSP format)
+	\item Get real in-prep citation from Kavya
+	\item Confirm sentence about Yu/Kavya integration for accuracy
+	\item Get grant numbers/titles for acknowledgments
+	\item Ask Matt what the hardware is for DXL
+	\item Add plots that include Matlab results when they finish
+	\item Add plots that include Windows results when they finish
+\end{itemize}
+}
+
+%---------------------------------------------------------------------------------
 % ACKNOWLEDGMENTS
@@
 % REFERENCES
-\todo{Fix code citations}
 \ifdraft

### 2018-10-08 cis_interface/tests/test_timing.py
Include tests of per_message plots

@@ -72,9 +72,9 @@ def test_plot_scalings():
     kwargs = dict(msg_size=[1], msg_size0=1, msg_count=[1], msg_count0=1)
     timing.plot_scalings(compare='commtype', **kwargs)
     timing.plot_scalings(compare='language', **kwargs)
-    timing.plot_scalings(compare='platform',
+    timing.plot_scalings(compare='platform', per_message=True,
                          compare_values=[platform._platform], **kwargs)
-    timing.plot_scalings(compare='python',
+    timing.plot_scalings(compare='python', per_message=True,
                          compare_values=[backwards._python_version], **kwargs)
 
 

### 2018-10-08 cis_interface/timing.py
Print summary of fit in table

@@ -105,7 +105,7 @@ def perf_func(loops, timer, nmsg, msg_size):
         while not flag:
             try:
                 t0 = perf.perf_counter()
-                timer.run(run_uuid)
+                tI = timer.run(run_uuid)
                 t1 = perf.perf_counter()
                 tdif = t1 - t0
                 timer.after_run(run_uuid, tdif)
@@ -168,6 +168,9 @@ class TimedRun(CisTestBase, tools.CisClass):
             tests.
 
     """
+
+    _use_mine = False
+
     def __init__(self, lang_src, lang_dst, name='timed_pipe',
                  scalings_file=None, perf_file=None,
                  comm_type=None, platform=None, python_ver=None, **kwargs):
@@ -229,6 +232,39 @@ class TimedRun(CisTestBase, tools.CisClass):
                                             self.lang_src, self.lang_dst,
                                             nmsg, msg_size)
         return out
+
+    @property
+    def max_msg_size(self):
+        r"""int: Largest size of message that can be sent without being split."""
+        return tools.get_CIS_MSG_MAX(comm_type=self.comm_type)
+
+    @property
+    def default_msg_size(self):
+        r"""list: Default message sizes for scaling tests."""
+        if self.comm_type.startswith('IPC'):
+            msg_size = [1, 1e2, 1e3, 1e4, 1e5]
+        else:
+            msg_size = [1, 1e2, 1e3, 1e4, 1e5, 1e6, 1e7]
+        return msg_size
+
+    @property
+    def default_msg_count(self):
+        r"""list: Default message count for scaling tests."""
+        return [5, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100]
+
+    @property
+    def time_per_byte(self):
+        return self.fit_scaling_size()[0]
+
+    @property
+    def time_per_message(self):
+        r"""float: Time required to send a single message of 1000 bytes."""
+        return self.fit_scaling_count()[0]
+
+    @property
+    def startup_time(self):
+        r"""float: Time required to set up communications and start models."""
+        return self.fit_scaling_count()[1]
 
     @property
     def description_prefix(self):
@@ -407,21 +443,33 @@ class TimedRun(CisTestBase, tools.CisClass):
         self.reset_default_comm()
         del self.entries[run_uuid], self.fyaml[run_uuid], self.foutput[run_uuid]
 
-    def run(self, run_uuid):
+    def run(self, run_uuid, time_func=None):
[…]

### 2018-10-09 papers/lang2018/ms.tex
Updates to results

@@
 %
-For the most part, the time required per message is not affected by the language of either model.
+For the most part, the time required per message is not affected by the language of either model for smaller messages.
 There is a very small increase in the time per message when there is a Python model involved, but this is much lower than the standard deviation resulting from background activity on the test machine and so is in conclusive.
+Language does play a role for messages larger than the maximum ($2^{20}$ bytes), when messages must be broken into smaller pieces.
+In particular, integrations that include a C or C++ receiving model scale more strongly than for those that have a Python receiving model because, in addition to the input/output connection drivers, the Python interface itself is asynchronous while the C/C++ interface is not.
+This means that while Python receiving models have concurrent operations to receive, backlog, and confirm messages, C/C++ models will block on each receive until the input connection driver completes the confirmation handshake.
+In addition, because C/C++ models are not continuously receiving and backlogging messages, the channel will become saturated.
+Once it reaches its maximum, the input connection driver can no longer pass along messages and must wait for the model to receive the next message.
+One remedy that is being explored for future releases, is to make the C/C++ interface asynchronous as well.
+However, this is a low priority as only models with very rapid and high volume input (such as the test models) are affected.
 
@@
 For example, the Python-to-C integrations required more overhead than the C-to-Python integrations.
-This is because the models are started in parallel, but do not require the same amount of time to start.
+This is because the models are started in parallel, but do not require the same amount of time to start and being execution.
+C models, while requiring compilation, start much faster than Python models.
 In a Python-to-C integration, the C model must wait for the Python model to finish its much slower startup and begin sending message, while in a C-to-Python integration, the C model can immediately begin sending messages that will be received by the Python model once it finishes starting up.
-In addition, the Python-to-Python integrations required less time than any of the other integrations because, although Python models take longer to start up, the start up is \todo{explanation?}
+In addition, the Python-to-Python integrations required less time than any of the other integrations that included Python because, although Python models take longer to start up, the start up is done in parallel while the compilation for the C and C++ models is done in serial.
+This source of overhead could be ameliorated by compiling models in parallel or in advanced, but is usually not significant enough to effect overall performance.
 %
@@
 %
-Figure \ref{fig:python} compares the execution times when using different versions of Python.
-The performance tests for both Python 2.7 and Python 3.5 were performed on the Linux machine from above using ZeroMQ communication.
+Figure \ref{fig:python} and Table \ref{tab:python} compare the execution times when using different versions of Python.
+The performance tests for both Python 2.7 and Python 3.5 were performed on the Linux machine from above using ZeroMQ communication and two different integration (Python-to-Python and C-to-C).
 %
@@
 	\includegraphics[width=\columnwidth,keepaspectratio]{./images/scaling_python.png}
-	\caption{Comparison of communication time scaling between Python versions on Linux.}
+	\caption{Comparison of communication time scaling between Python versions on Linux. The solid lines are for the Python-to-Python integrations and the dotted lines are for the C-to-C integration.}
 	\label{fig:python}
@@
 %
-\todo{discussion}
+\begin{table}[htbp]
+\begin{center}
+\begin{tabular}{|c|c|c|c|}
+\hline
+Python Version	& Language	& Time per Message (s) 	& Overhead (s) 	\\\hline
+2.7			& Python		& 0.012				& 1.73			\\
+2.7			& C			& 0.011				& 1.36			\\
+3.5 			& Python		& 0.013				& 2.62			\\
+3.5			& C			& 0.011				& 1.36			\\\hline
+\end{tabular}
+\end{center}
+\caption{Effect of Python version on performance.}
+\label{tab:python}
+\end{table}          
+%
+There is $\sim1$\,s additional overhead with Python 3.5 compared to Python 2.7 due to the increased startup time for the Python 3.5 interpreter.
+This overhead is only present for integrations that include Python models.
+For comparison, there is no difference in the time per message and overhead between the different Python versions for the C-to-C integration (dashed lines in Figure \ref{fig:python}).
+There is a slight difference in time per message between the two versions for Python models that is especially apparent in the scaling of execution time with message size for large messages ($>10^6$ bytes).
+\todo{discussion of large message scaling}
 
@@
 %
-Figure \ref{fig:platform} compares the execution times on different operating systems.
+Figure \ref{fig:platform} and Table \ref{tab:platform} compare the execution times on different operating systems.
 The performance tests reported for Linux were run on the machine from above.
@@
 The tests for Windows were run on the same machine from a dual boot of Windows 10.
-All platform dependent performance tests use ZeroMQ communication.
+All platform dependent performance tests use ZeroMQ communication and the Python-to-Python integration.
 %
@@
 \fi
+%
+\begin{table}[htbp]
+\begin{center}
+\begin{tabular}{|c|c|c|}
+\hline
+Operating System	& Time per Message (s) 	& Overhead (s) 	\\\hline
+Linux			& \todo{?}				& \todo{?}			\\
+Mac OSX			& \todo{?}				& \todo{?}			\\
+Windows			& \todo{?}				& \todo{?}			\\\hline
+\end{tabular}
+\end{center}
+\caption{Effect of operating system on performance.}
+\label{tab:platform}
+\end{table}%
 %
@@
 Models are run in parallel with asynchronous message passing handled under-the-hood via threading and one of three communication mechanisms.
-{\cis} works on Linux, Mac OSX, and Windows operating systems with Python 2.7 and $3.4+$.
+{\cis} works on Linux, Mac OSX, and Windows operating systems with Python 2.7, 3.4, 3.5, 3.6 and 3.7.
 {\cis} currently supports running models written in Python, C, C++, and Matlab with additional support for compiling models C/C++ using Make \citep{Stallman2004} or CMake \citep{Martin2006} and running models written in the LPy \citep{Boudon2012} DSL.

### 2018-10-11 papers/lang2018/ms.tex
Add Matlab results

@@
 % For Draft
-\ieeefalse
+\ieeetrue
 \twocoltrue
@@
 
+\ifieee
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd{\@makecaption}
+  {\scshape}
+  {}
+  {}
+  {}
+\makeatletter
+\patchcmd{\@makecaption}
+  {\\}
+  {.\ }
+  {}
+  {}
+\def\tablename{Table}
+\makeatother
+\else
 %\makeatletter
+\fi
 
@@
 
+\ifieee
+\else
 	\shorttitle{{\cis}}
+\fi
 
@@
 
+\ifieee
+	\author{Meagan~Lang%
+	\thanks{The authors are with the National Center for Supercomputing Application, University of Illinois, Urbana-Champaign, IL 61801 USA (e-mail: langmm@illinois.edu)}}
+\else
 	\author{Meagan~Lang\altaffilmark{1}}
 
-\altaffiltext{1}{National Center for Supercomputing Application, 
-University of Illinois, Urbana-Champaign, IL email: {\tt langmm@illinois.edu}}
+	\altaffiltext{1}{National Center for Supercomputing Application, University of Illinois, Urbana-Champaign, IL email: {\tt langmm@illinois.edu}}
+\fi
+
+\ifieee
+	\maketitle
+\fi
 
@@
 We present \href{https://github.com/cropsinsilico/cis_interface}{\cis}, a Python package for running integration networks with connections between models across languages and scales.
-{\cis} coordinates parallel execution of models in Python, C, C++, and Matlab on Linux, Mac OSX, and Windows operating systems, and handles communication in a number of data formats common to computational plant modeling.
+{\cis} coordinates parallel execution of models in Python, C, C++, and Matlab on Linux, Mac OS X, and Windows operating systems, and handles communication in a number of data formats common to computational plant modeling.
 {\cis} is designed to be user-friendly, with visual tools developed in parallel for composing the integration networks.
@@
 
-%\begin{IEEEkeywords}
-%Plant modeling, Modeling, Communication
-%\end{IEEEkeywords}
+\ifieee
+	\begin{IEEEkeywords}
+		Plant modeling, Modeling, Communication
+	\end{IEEEkeywords}
+	\IEEEpeerreviewmaketitle
+\else
+	\maketitle
+\fi
 
-%\IEEEpeerreviewmaketitle
-
-\maketitle
 %---------------------------------------------------------------------------------
@@
 In addition, computational models for different scales, organs, or processes can also be combined, or integrated, to probe more complex biological mechanisms.
-By integrating a metabolic flux model with a genetic regulatory model for carbon uptake, \citet{integration_prep} was able to identify gene candidates for regulating photosynthesis under elevated CO$_{2}$.
+By integrating a metabolic flux model with a genetic regulatory model for carbon uptake,
+%
+\ifieee
+	\citep{integration_prep} 
+\else
+	\citet{integration_prep} 
+\fi
+%
+was able to identify gene candidates for regulating photosynthesis under elevated CO$_{2}$.
 With advances in compute power, it should be possible to link biologically related models to create complex integration networks capable of capturing the response of entire plants, fields, or regions.
@@
 The first communication mechanism used by {\cis} was System V interprocess 
-communication (IPC) message queues \citep{Rusling1999} on Posix (Linux and Mac OSX) systems. 
+communication (IPC) message queues \citep{Rusling1999} on Posix (Linux and Mac OS X) systems. 
 IPC message queues allow messages to be passed between models running on separate 
@@
 and do not allow communication between remote processes. In addition, IPC queues also 
-have relatively low default message size limits on Mac OSX systems (2048 bytes). Once 
+have relatively low default message size limits on Mac OS X systems (2048 bytes). Once 
 the queue is full or if the message is larger than the limit, any process attempting to send an additional message will block until a 
@@
 ($2^{20}$ for ZeroMQ and RabbitMQ), while others are very limiting 
-(2048 on Mac OSX for Sys V IPC queues). Although messages consisting of a few scalars are unlikely to 
+(2048 on Mac OS X for Sys V IPC queues). Although messages consisting of a few scalars are unlikely to 
 exceed these limits, biological inputs and outputs are often much more complex. 
@@
 %
-In order to evaluate {\cis}, performance tests were run on machines with Linux, Mac OSX, and Windows operating systems for different communication mechanisms, Python versions, and language combinations.
+In order to evaluate {\cis}, performance tests were run on machines with Linux, Mac OS X, and Windows operating systems for different communication mechanisms, Python versions, and language combinations.
 During each run, $N_{msg}$ messages of size $S_{msg}$ were sent from the source model (in one language) to the destination model (in another language) which then output the messages to file for verification.
@@
 Figure \ref{fig:commtype} compares the execution times for the different communication mechanisms discussed in \S\ref{SS:communication} for sending different numbers/sizes of messages from one Python model to another Python model.
-The tests were run on a \todo{Dell?} tower with 10 quad-core Intel(R) Xeon(R) CPU E5-2650 v3 @ 2.30GHz processors running Ubuntu 14.04.
+The tests were run on a Dell tower with Dual Intel(R) Xeon(R) CPU E5-2650 v3 @ 2.30GHz processors running Ubuntu 14.04.
 The left panel of Figure \ref{fig:commtype} shows the total time required to run both models and send a varying number of 1000 bytes messages from one model to the other.
@@
 The runs for Figure \ref{fig:language} and Table \ref{tab:language} were run on the Linux machine from above while those for Figure \ref{fig:language_matlab} and Table \ref{tab:language_matlab} were run on a 2015 MacBook Pro with a 2.9 GHz Intel Core i5 processor running macOS High Sierra 10.13.4 that had Matlab R2017a installed.
-In both cases, ZeroMQ communication was used.
+In both cases, ZeroMQ communication was used with Python 2.7.
 %
@@
 	\includegraphics[width=\columnwidth,keepaspectratio]{./images/scaling_language.png}
-	\caption{Comparison of communication time scaling between languages on Linux. The same fiducial message size (1000 bytes) and count (5) was used as in Figure \ref{fig:commtype}.}
+	\caption{Comparison of communication time scaling between languages on Linux without Matlab. The same fiducial message size (1000 bytes) and count (5) was used as in Figure \ref{fig:commtype}.}
 	\label{fig:language}
@@
 Source 	& Destination 	& Time per Message (s) 	& Overhead (s) 	\\\hline
-C 				& C					& 0.11				& 1.36			\\
-C 				& C++				& 0.12				& 1.42			\\
-C 				& Python				& 0.12				& 1.99			\\
+C 				& C					& 0.011				& 1.36			\\
+C 				& C++				& 0.012				& 1.42			\\
+C 				& Python				& 0.012				& 1.99			\\
 \hline%
-C++ 				& C					& 0.12				& 1.42			\\
-C++ 				& C++				& 0.11				& 1.49			\\
-C++ 				& Python				& 0.12				& 2.05			\\
+C++ 				& C					& 0.012				& 1.42			\\
+C++ 				& C++				& 0.011				& 1.49			\\
+C++ 				& Python				& 0.012				& 2.05			\\
 \hline%
-Python			& C					& 0.12				& 2.15			\\
-Python			& C++				& 0.12				& 2.21			\\
-Python			& Python				& 0.12				& 1.73			\\
+Python			& C					& 0.012				& 2.15			\\
+Python			& C++				& 0.012				& 2.21			\\
+Python			& Python				& 0.012				& 1.73			\\
 \hline
@@
 %
-For the most part, the time required per message is not affected by the language of either model for smaller messages.
+\ifinclfig
+ 	\begin{figure}[htbp]
+	\begin{center}
+	\includegraphics[width=\columnwidth,keepaspectratio]{./images/scaling_language_matlab.png}
+	\caption{Comparison of communication time scaling between languages on Mac OS X including Matlab.}
+	\label{fig:language_matlab}
+	\end{center}
+	\end{figure}
+\fi
+%
+\begin{table}[htbp]
+\begin{center}
+\begin{tabular}{|c|c|c|c|}
+\hline
+Source 	& Destination 	& Time per Message (s) 	& Overhead (s) 	\\\hline
+C 				& C					& 0.015				& 1.11			\\
+C 				& C++				& 0.016				& 1.15			\\
+C 				& Python				& 0.019				& 2.86			\\
+C				& Matlab				& 0.505				& 12.98			\\
+\hline%
+C++ 				& C					& 0.019				& 1.10			\\
+C++ 				& C++				& 0.016				& 1.25			\\
+C++ 				& Python				& 0.016				& 3.10			\\
+C++ 				& Matlab				& 0.505				& 13.45			\\
+\hline%
+Python			& C					& 0.019				& 3.17			\\
+Python			& C++				& 0.019				& 3.26			\\
+Python			& Python				& 0.019				& 2.86			\\
+Python			& Matlab				& 0.500				& 11.61			\\
+\hline%
+Matlab			& C					& 0.144				& 11.82			\\
+Matlab			& C++				& 0.142				& 14.79			\\
+Matlab			& Python				& 0.139				& 12.23			\\
+Matlab			& Matlab				& 0.559				& 19.28			\\
+\hline%
+Matlab (started)		& Matlab (started)		& 0.537				& 1.25			\\
+\hline
+\end{tabular}
+\end{center}
+\caption{Effect of model language on performance (Mac OS X) with Matlab.}
+\label{tab:language_matlab}
+\end{table}%
+%
+
+Except for Matlab models, the time required per message is not affected by the language of either model for smaller messages.
 There is a very small increase in the time per message when there is a Python model involved, but this is much lower than the standard deviation resulting from background activity on the test machine and so is in conclusive.
-Language does play a role for messages larger than the maximum ($2^{20}$ bytes), when messages must be broken into smaller pieces.
+Matlab models have a much higher time per message than any other language.
+This results from the way the Matlab interface was implemented, by calling the Python interface.
+The time required per message is largest when the receiving model is written in Matlab because the receiving models both receives messages from the other model and send output to a file, resulting in twice the number of calls to the wrapped interface.
+The performance of the Matlab interface may be improved in the future by either implementing it in Matlab directly or by wrapping the C routines.
+
+Language also plays a role for messages larger than the maximum ($2^{20}$ bytes), when messages must be broken into smaller pieces.
 In particular, integrations that include a C or C++ receiving model scale more strongly than for those that have a Python receiving model because, in addition to the input/output connection drivers, the Python interface itself is asynchronous while the C/C++ interface is not.
@@
 Python and Matlab), the majority of the overhead comes from starting the interpreter.
-For Matlab, this is particularly time consuming.
-Although this is only a one-time cost required at the start of an integration, {\cis} does offer ways to alleviate this if multiple runs are necessary by starting a Matlab engine prior to the first integration that can be used by subsequent runs \todo{(results that include existing Matlab engine)}.
+For Matlab, this is particularly time consuming ($>10$ s).
+Although this is only a one-time cost required at the start of an integration, {\cis} does offer ways to alleviate this if multiple runs are necessary by starting a Matlab engine prior to the first integration that can be used by subsequent runs.
+When Matlab is started in advance (the dashed-dotted orange line in Figure \ref{fig:language_matlab}), the overhead for Matlab models drops to 1.25\,s, less than Python (2.86\,s) but slightly more than C (1.11\,s).
 
@@
 This source of overhead could be ameliorated by compiling models in parallel or in advanced, but is usually not significant enough to effect overall performance.
-%
-\todo{figure/table for Matlab \& discussion}
 
@@
 For comparison, there is no difference in the time per message and overhead between the different Python versions for the C-to-C integration (dashed lines in Figure \ref{fig:python}).
-There is a slight difference in time per message between the two versions for Python models that is especially apparent in the scaling of execution time with message size for large messages ($>10^6$ bytes).
-\todo{discussion of large message scaling}
+There is a slight difference in time per message between the two versions for all models that shows up in the scaling of execution time with message size for large messages ($>10^6$ bytes).
+Differences at large message sizes between the two versions arrises from the use of Python drivers to transport messages between models and is more pronounced in integrations with Python models due to the increased amount of Python code.
 
@@
 Linux			& \todo{?}				& \todo{?}			\\
-Mac OSX			& \todo{?}				& \todo{?}			\\
+Mac OS X			& \todo{?}				& \todo{?}			\\
 Windows			& \todo{?}				& \todo{?}			\\\hline
@@
 %
-{\cis} is an open-source Python package for connecting computational models across programming languages in scales to form integration networks.
+{\cis} is an open-source Python package for connecting computational models across programming languages and scales to form integration networks.
 Models are run in parallel with asynchronous message passing handled under-the-hood via threading and one of three communication mechanisms.
-{\cis} works on Linux, Mac OSX, and Windows operating systems with Python 2.7, 3.4, 3.5, 3.6 and 3.7.
+{\cis} works on Linux, Mac OS X, and Windows operating systems with Python 2.7, 3.4, 3.5, 3.6 and 3.7.
 {\cis} currently supports running models written in Python, C, C++, and Matlab with additional support for compiling models C/C++ using Make \citep{Stallman2004} or CMake \citep{Martin2006} and running models written in the LPy \citep{Boudon2012} DSL.
@@
 Each square is a model and lines between the models represent connections where information is exchanged.
-Solid lines are connections we have already implemented, dashed lines are connections we are currently working on, and dotted lines are connections planned for the future.
-Results from one integration can be found in \todo{cite Kavya/Yu}.
+Solid lines are connections that have already implemented, dashed lines are connections currently being implemented, and dotted lines are connections planned for the future.
+Results from one integration, a metabolic model and genetic response to CO$_2$ model, can be found in
+%
+\ifieee
+	\citep{integration_prep}. 
+\else
+	\citet{integration_prep}. 
+\fi
 %
@@
 that can be used in integration networks and the complexity of integration 
-networks that can be handled. Several improvements are already in progress/planned for {\cis}.
+networks that can be executed. Several improvements are already in progress/planned for {\cis}.
 
@@
 are written in many different languages. While the current language support 
-covers many, additional languages must be added to unlock the full potential. 
+covers many of the most popular languages among plant biologists, 
+additional languages must be added to unlock the full set of potential integrations. 
 Surveys of the plant modeling community have helped identify the core languages 
@@
 {\cis} to support models written in R, Fortran, Java, and the SBML \citep{Hucka2003} DSL.
+
 We also plan to add support for running Matlab models using Octave \citep{Eaton2002}. Matlab 
@@
 within the model code including things like which input variables are static 
-and read in once versus which variables change and should be looped over. 
+and received once versus which variables change and should be looped over. 
 We plan on adding options to {\cis} for dynamically generating model code 
@@
 However, our ultimate goal with the GUI is to provide a service where users can register their model, compose and run integration networks on dedicated cloud compute resources, and view output in real time all from the web browser.
-\todo{Mike and Craig as authors/acknowledgements?
-Provide \href{https://hackathon.cis.ndslabs.org/}{link?}}
+%\todo{Provide \href{https://hackathon.cis.ndslabs.org/}{link?}}
 
@@
 installed using {\tt pip} or {\tt conda}.
-To ensure code health, {\cis} boasts 100\% code coverage with automated testing on Linux, Mac OSX, and Windows for Python version 2.7, 3.4, 3.5, 3.6, and 3.7 via continuous integration with TravisCI \citep{travisci} and Appveyor \citep{appveyor}.
+To ensure code health, {\cis} boasts 100\% code coverage with automated testing on Linux, Mac OS X, and Windows for Python version 2.7, 3.4, 3.5, 3.6, and 3.7 via continuous integration with TravisCI \citep{travisci} and Appveyor \citep{appveyor}.
 Full documentation for {\cis} can be found \href{https://cropsinsilico.github.io/cis_interface/}{here} including step-by-step directions from the tutorial conducted during the \href{https://cropsinsilico.github.io/cis_interface/hackathon2018/index.html}{2018 Crops in Silico Hackathon}.
@@
 % TODO
+\todo{%
 \section*{TODO}
-\todo{%
 \begin{itemize}
@@
 	\item Get grant numbers/titles for acknowledgments
-	\item Ask Matt what the hardware is for DXL
-	\item Add plots that include Matlab results when they finish
+	\item Ask Matt about author list
 	\item Add plots that include Windows results when they finish
+	\item Include statistics in tables on scaling at large message size?
+	\item Re-run Matlab-to-Matlab, $nmsg=5$, $msg\_size=1000$
 \end{itemize}
@@
 % ACKNOWLEDGMENTS
-%\section*{Acknowledgment}
-%\acknowledgments
-%The authors would like to thank ...
+\section*{Acknowledgments}
+\ifieee
+\else
+\acknowledgments
+\fi
+%
+This work was supported by funding from the Foundation for Food and Agriculture Research (FFAR), Institute for Sustainability, Energy, and Environment (iSEE), and the National Center for Supercomputing Applications (NCSA).
+The authors would like to thank Mike Lambert and Craig Willis for their ongoing work on a graphical user interface for visually composing and running integration networks for {\cis} as well as David Raila for his work on the original integration code that {\cis} was based on.
 

### 2018-10-11 cis_interface/timing.py
Change color for matlab with startup

@@ -1082,7 +1082,7 @@ def plot_scalings(compare='commtype', compare_values=None,
             for i in range(nml):
                 ml_sessions.append(MatlabModelDriver.start_matlab())
             label += ' (Existing ML Session)'
-            plot_kws['alpha'] = 0.5
+            plot_kws['color'] = 'orange'
         axs, fit = TimedRun.class_plot(name=test_name, axs=axs, label=label,
                                        yscale=yscale, plot_kws=plot_kws, **kws)
         fits[label] = fit

### 2018-10-11 papers/lang2018/ms.tex
Edits from Matt plus new figures

@@
 % For Draft
-\ieeetrue
-\twocoltrue
+\ieeefalse
+\twocolfalse
 \placefigtrue
@@
 was able to identify gene candidates for regulating photosynthesis under elevated CO$_{2}$.
-With advances in compute power, it should be possible to link biologically related models to create complex integration networks capable of capturing the response of entire plants, fields, or regions.
+With advances in computational power, it should be possible to link biologically-related models to create complex integration networks capable of capturing the response of entire plants, fields, or regions.
 However, independently developed computational models often have compatibility issues resulting from differences in programming language or data format that make such collaborations difficult.
@@
 running model A, converting output from A into input expected by model B, running model B).
-However, manual integration is 1) computational inefficient when many iterations are necessary, such as during parameter searches or steady state convergence tests, 2) unique to every integration, requiring the production of new scripts, and 3) complex when more than $>3$ models are being integrated.
+However, manual integration is 1) computationally inefficient when many iterations are necessary, such as during parameter searches or steady state convergence tests, 2) unique to every integration, requiring the production of new scripts, and 3) complex when more than $>3$ models are being integrated.
 Within both computational biology and the larger scientific computing community, groups have developed around different solutions to the problem of connecting models.
@@
 \begin{itemize}
-	\item Easy to use. Require as little modification to the model source code as possible and only in the language of the model itself.
-	\item Efficient. Allow models to run in parallel with asynchronous communication that dosn't block when a message is sent.
-	\item Flexible. Provide the same interface to the user, regardless of the communication mechanism being used or the platform the model is being executed on.
+	\item {\bf Easy to use.} Require as little modification to the model source code as possible and only in the language of the model itself.
+	\item {\bf Efficient.} Allow models to run in parallel with asynchronous communication that dosn't block when a message is sent.
+	\item {\bf Flexible.} Provide the same interface to the user, regardless of the communication mechanism being used or the platform the model is being executed on.
 \end{itemize}
@@
 {\cis} currently support models written in Python, Matlab, C, and C++ with 
-additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}.
+additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}. Support for additional languages is planned for future development (See \S\ref{SS:langsupport}).
 
@@
 be run, and any input or output variables including their data type and units. 
-Integration networks are specified via YAML files declaring the connections 
+Integration networks are specified by declaring the connections 
 between models. Connections are declared by pairing an output variable from 
@@
 languages, platforms, and data types. To accomplish this, {\cis} leverages 
-several tools different tools for communication which are applied as needed. The 
+three tools different tools for communication, 
+System V IPC Queues (\S\ref{SSS:ipc}), ZeroMQ (\S\ref{SSS:zmq}), and RabbitMQ (\S\ref{SSS:rmq}), 
+which are applied as needed. The 
 particular mechanism used by {\cis} for communication is hidden from the 
@@
 \begin{enumerate}
-	\item A model sends a message in the form of a native data object via a language 
-specific API to one of the output channels declared in the model YAML. The output channel interface encodes the message and sends it.
-	\item An output connection driver (written in Python) runs in a separate thread, listening to the model output channel.
+	\item A model sends a message in the form of a native data object via a language-specific API to one of the output channels declared in the model YAML.
+The output channel interface encodes the message and sends it.
+	\item An output connection driver (written in Python) runs in a separate thread on the master process, listening to the model output channel.
 When the model sends a message, the output connection driver checks that the message is in the expected format and then forwards it to an intermediate channel.
-The intermediate channel may seem unnecessary, but it is used as a buffer for future support on distributed architectures (e.g.
+The intermediate channel is used as a buffer for support on distributed architectures (e.g.
 if one model is running on a remote machine).
@@
 When a message is received, it is then forwarded to the input channel of the receiving model as specified in the integration network YAML.
-	\item The receiving model receives the message in the form of an analogous native data object.
+	\item The receiving model receives the message in the form of an analogous native data object. Interface receive calls can be either blocking or non-blocking, but are blocking by default.
 \end{enumerate}
@@
 network. While {\cis} does not currently support using these protocols 
-for distributed integration networks, this is part of the plans for future improvement.
+for distributed integration networks, this is a future development plan that has been prepared for.
 
@@
 %
-While ZeroMQ provides broker-less communication means, {\cis} includes support 
-for brokered communication via RabbitMQ \citep{RMQ}. {\cis} does not currently use 
-RabbitMQ for communication unless explicitly specified by the user in their 
-integration network YAML. RabbitMQ support was originally added to {\cis} 
-as a supplement to System V IPC queues in the case of future support for distributed 
-integration networks. In future development, RabbitMQ brokered communication will be 
-used for establishing integration networks with remote models run a services 
-(See \S\ref{SS:future}).
+While broker-less communication like ZeroMQ is light weight and fast, it is not as fault 
+tolerant as brokered messaging systems that confirm message delivery and can resend 
+dropped messages. Resilience to dropped messages, while not as necessary for integrations 
+running entirely on a local machine, will be more important for integrations running on 
+distributed resources with less reliable connections. As a result, {\cis} includes support 
+for brokered communication via RabbitMQ \citep{RMQ} that will be used during future 
+development to allow integrations to run on distributed resources or include remote models 
+run as services (\S\ref{SS:distributed}). Due to the slower message speed, 
+{\cis} does not currently use RabbitMQ for communication unless explicitly specified by the user in their 
+integration network YAML.
 
@@
 on the allowed size for a single message. Some of these limits can be quite large 
-($2^{20}$ for ZeroMQ and RabbitMQ), while others are very limiting 
-(2048 on Mac OS X for Sys V IPC queues). Although messages consisting of a few scalars are unlikely to 
+($2^{20}$ bytes for ZeroMQ and RabbitMQ), while others are very limiting 
+(2048 bytes on Mac OS X for Sys V IPC queues). Although messages consisting of a few scalars are unlikely to 
 exceed these limits, biological inputs and outputs are often much more complex. 
@@
 then destroyed. The address associated with the temporary channel is sent in 
-header information as a message on the main channel along with information about 
+header information as a message on the main channel along with metadata about 
 the message that will be sent through the temporary channel like size and data type. 
@@
 To combat this, {\cis} requires models to explicitly specify the format of input and output expected by a model in the model YAML.
-{\cis} can then handle a number of conversion between models without prompting as well as serialization/deserialization to the correct type in each of the supported languages.
+{\cis} can then handle a number of conversions between models without prompting as well as serialization/deserialization to the correct type in each of the supported languages.
 Data formats currently supported by {\cis} include:
@@
 	\item Arrays
-	\item Tabular (e.g. CSV or tab-delimited)
+	\item Text-encoded tables (e.g. CSV or tab-delimited)
 	\item Pandas data frames
@@
 Performance tests were run using the perf package \citep{Stinner2018}.
-Each run was repeated 10 times to rule out fluctuations due to external loads on the test machines.
+Each run was repeated 10 times to mitigate fluctuations due to external loads on the test machines.
 
@@
 Operating System	& Time per Message (s) 	& Overhead (s) 	\\\hline
-Linux			& \todo{?}				& \todo{?}			\\
-Mac OS X			& \todo{?}				& \todo{?}			\\
+Linux			& 0.012				& 1.73			\\
+Mac OS X			& 0.019				& 2.86			\\
 Windows			& \todo{?}				& \todo{?}			\\\hline
@@
 %
+Both the time per message and overhead changed between operating systems with the Linux machine performing the most consistently and fastest.
+Because the Linux test machine was very different from the test machine for both Mac OS X and Windows, it cannot be determined that operating system alone contributed to the observed differences.
+The same can be said of the Mac OS X and Windows tests given that operating systems are often optimized for their intended hardware.
+However, the tests do show that {\cis} performed consistently on each of the supported operating systems.
 \todo{Windows results \& discussion}
@@
 %
-\todo{more here? teaser results?}
+%\todo{more here? teaser results?}
 
@@
 % TODO
-\todo{%
-\section*{TODO}
-\begin{itemize}
-	\item Fix code citations to include URLs
-	\item Increase size of font on figures or increase size of figures (depends on iSP format)
-	\item Get real in-prep citation from Kavya
-	\item Confirm sentence about Yu/Kavya integration for accuracy
-	\item Get grant numbers/titles for acknowledgments
-	\item Ask Matt about author list
-	\item Add plots that include Windows results when they finish
-	\item Include statistics in tables on scaling at large message size?
-	\item Re-run Matlab-to-Matlab, $nmsg=5$, $msg\_size=1000$
-\end{itemize}
-}
+%\todo{%
+%\section*{TODO}
+%\begin{itemize}
+%	\item Increase size of font on figures or increase size of figures (depends on iSP format)
+%	\item Get real in-prep citation from Kavya
+%	\item Confirm sentence about Yu/Kavya integration for accuracy
+%	\item Get grant numbers/titles for acknowledgments
+%	\item Ask Matt about author list
+%	\item Add plots that include Windows results when they finish
+%	\item Include statistics in tables on scaling at large message size?
+%	\item Re-run Matlab-to-Matlab, $nmsg=5$, $msg\_size=1000$
+%	\item Re-run Python-to-Matlab, $nmsg=50$, $msg\_size=1000$
+%\end{itemize}
+%}
 

### 2018-10-11 cis_interface/timing.py
Added data point for IPC scaling to show log

@@ -266,7 +266,7 @@ class TimedRun(CisTestBase, tools.CisClass):
     def default_msg_size(self):
         r"""list: Default message sizes for scaling tests."""
         if self.comm_type.startswith('IPC'):
-            msg_size = [1, 1e2, 1e3, 1e4, 1e5]
+            msg_size = [1, 1e2, 1e3, 1e4, 5e4, 1e5]
         else:
             msg_size = [1, 1e2, 1e3, 1e4, 1e5, 1e6, 5e6, 1e7]
         return msg_size

### 2018-10-11 cis_interface/timing.py
Update plots to use different line styles when possible for printing

@@ -997,8 +997,8 @@ def plot_scalings(compare='commtype', compare_values=None,
     if compare == 'commtype':
         color_var = 'comm_type'
         color_map = {'ZMQComm': 'b', 'IPCComm': 'r', 'RMQComm': 'g'}
-        style_var = None
-        style_map = None
+        style_var = 'comm_type'
+        style_map = {'ZMQComm': '-', 'IPCComm': '--', 'RMQComm': ':'}
         var_list = compare_values
         var_kws = [{color_var: k} for k in var_list]
         kws2label = lambda x: x['comm_type'].split('Comm')[0]  # noqa: E731

### 2018-10-11 papers/lang2018/ms.tex
Add moore grant to acknowledgements

@@
 %
-This work was supported by funding from the Foundation for Food and Agriculture Research (FFAR), Institute for Sustainability, Energy, and Environment (iSEE), and the National Center for Supercomputing Applications (NCSA).
+This work was supported by funding from the Foundation for Food and Agriculture Research (FFAR), Institute for Sustainability, Energy, and Environment (iSEE), the National Center for Supercomputing Applications (NCSA), and the Gordon and Betty Moore Foundation's Data-Driven Discovery Initiative through Grant GBMF4561 to Matthew Turk.
+%
 The authors would like to thank Mike Lambert and Craig Willis for their ongoing work on a graphical user interface for visually composing and running integration networks for {\cis} as well as David Raila for his work on the original integration code that {\cis} was based on.

### 2018-10-12 papers/lang2018/ms.tex
Minor edits responding to Steve's comments and update to include Windows resutls

@@
 {\cis} coordinates parallel execution of models in Python, C, C++, and Matlab on Linux, Mac OS X, and Windows operating systems, and handles communication in a number of data formats common to computational plant modeling.
-{\cis} is designed to be user-friendly, with visual tools developed in parallel for composing the integration networks.
+{\cis} is designed to be user-friendly, with visual tools developed in parallel for composing the integration networks and can be accessed at {\tt https://github.com/cropsinsilico/cis\_interface}.
 \end{abstract}
@@
 %\IEEEPARstart{C}{omputational} 
-Computational plant biologists have produced a wealth of computational models to describe the biological processes governing plant growth and development, covering scales from atomic to global.
+Plant biologists have produced a wealth of computational models to describe the biological processes governing plant growth and development, covering scales from atomistic to global.
 Although usually developed independently to address questions specific to an organ, species, or growth process, computational plant models can also have applications beyond their original scope.
@@
 We present an Open Source Python package, {\cis}, for creating and running integration networks by connecting existing computational models written in Python, Matlab, C and C++.
+{\cis} is available on Github\footnote{{\tt https://github.com/cropsinsilico/cis\_interface}}, and can be installed from PyPI\footnote{{\tt https://pypi.org/project/cis\_interface/}} via {\tt pip install {\cis}} or conda-forge\footnote{{\tt https://github.com/conda-forge/cis\_interface-feedstock}} via {\tt conda -c conda-forge install {\cis}}.
 Although developed for the purpose of integrating plant biology models, {\cis} can be applied to any situation requiring coordination between models in the supported languages.
@@
 The resulting programs are highly tuned for one purpose and written specifically for scientists within an isolated group of collaborators.
-Scientists will generally write programs in the language they are most familiar with and, given that there is no consensus on one programming language as ideal for all scientific applications, the language of computational models will often vary between research groups, if not between members of the groups themselves.
+Scientists will generally write programs in the language they are most familiar with and, given that there is no consensus on one programming language as ideal for all scientific applications, the language will often vary between research groups, if not between members within the groups themselves.
 The same can be said of data formats which can be highly tailored to the target question and may or may not include metadata like the data types, field names, or units.
@@
 Dedicated model DSLs improve the reusability of models written in the DSL and have the advantage of often implicitly handling data transformation.
-However, they do nothing for existing models that are not in the DSL or cannot be expressed within the constraints of the DSL.
-In addition, learning a new DSL, in addition to a programming language, can be daunting for new researchers.
+However, they provide no help for existing models that are not in the DSL or those that cannot be expressed within the constraints of the DSL.
+In addition, learning a new DSL, in addition to a programming language, can be daunting.
 
@@
 For example, there are many generic tools in Python for coordinating the execution of tasks in parallel or serial \citep[e.g.][]{babuji18, celery, luigi}.
-While these tools are powerful for organizing complex work flows, it is the user's onus to make sure that the components they connect are compatible and handle any data transformation that might be necessary (e.g.
-unit conversion or field selection).
+While these tools are powerful for organizing complex work flows, it is the user's onus to make sure that the components they connect are compatible and handle any data transformation that might be necessary, e.g.
+unit conversion or field selection.
 
@@
 However, they are stricter in several respects: 1) the model must be written in (or exposable to), the language of the framework and 2) the model must be written in a way that is aware of the framework and/or the format of models with which it will interact.
-Like DSLs, frameworks also provided limited support for models written without them in mind.
+Like DSLs, frameworks also provided limited support for models written without the framework in mind.
 
-{\cis} interface is an example of a framework that overcomes these issues by exposing light-weight interfaces in the language of the model (See \S\ref{SSS:interface}) that permit messages to be passed between the model processes as they run in parallel (See \S\ref{SS:communication}).
+{\cis} interface is an example of a framework that overcomes these issues by exposing simple and easily accessible interfaces in the language of the model (See \S\ref{SSS:interface}) that permit messages to be passed between the model processes as they run in parallel (See \S\ref{SS:communication}).
 As a result, models writers only need knowledge of the language in which their model is written.
@@
 	\item {\bf Easy to use.} Require as little modification to the model source code as possible and only in the language of the model itself.
-	\item {\bf Efficient.} Allow models to run in parallel with asynchronous communication that dosn't block when a message is sent.
+	\item {\bf Efficient.} Allow models to run in parallel with asynchronous communication that doesn't block model execution when a message is sent, but has yet to be received.
 	\item {\bf Flexible.} Provide the same interface to the user, regardless of the communication mechanism being used or the platform the model is being executed on.
@@
 %
-{\cis} is executed via a command line interface (CLI), {\tt cisrun} with YAML specification files (\S\ref{SSS:yaml}) as input.
-Based on the information contained in the YAML specification files, {\cis} establishes a network of asynchronous communication channels (\S\ref{SS:communication}), and launches the models on new processes (\S\ref{SS:model_driver}).
+{\cis} is executed via a command line interface (CLI), {\tt cisrun} with specification files (\S\ref{SSS:yaml}) as input.
+Based on the information contained in the specification files, {\cis} establishes a network of asynchronous communication channels (\S\ref{SS:communication}), and launches the models on new processes (\S\ref{SS:model_driver}).
 %At execution, {\cis} launches models on new processes based on information contained in a YAML specification file (See \S\ref{SSS:yaml} below) and establishes
@@
 and there are many existing tools for parsing YAML formats in different programming 
-languages. The declarative format allows user to specify exactly what they 
+languages \citep[e.g.][]{Simonov2006,pyyaml,jsyaml}. The declarative format allows user to specify exactly what they 
 want to do, without describing how it should be done. While the information 
@@
 if one model is running on a remote machine).
-In these cases, the intermediate channel will connect to a RabbitMQ broker using security credentials.
+In these cases, the intermediate channel will connect to a RabbitMQ broker using security credentials (See \S\ref{SSS:rmq}).
 	\item An input connection driver (also written in Python) runs in a separate thread, listening to the intermediate channel.
@@
 \begin{itemize}
-	\item Scalars (e.g. integers, floats)
+	\item Scalars (e.g. integers, decimals)
 	\item Arrays
 	\item Text-encoded tables (e.g. CSV or tab-delimited)
-	\item Pandas data frames
-	\item PLY
-	\item OBJ
+	\item Pandas data frames \citep{pandas}
+	\item 3D geometry structures (PLY \citep{ply} \& OBJ \citep{obj})
 \end{itemize}
@@
 In addition, {\cis} offers the option to specify units for scalars, arrays, tabular data, and pandas data frames.
-Units are tracked using the {\tt unyt} package \citep{Goldbaum2018}.
+Units are tracked using the {\tt unyt} package \citep{Goldbaum2018}, which allows physical units to be associated with scalars and arrays.
 If two models use different units (and both are specified), {\cis} will automatically perform the necessary conversions before passing data from one model to the next.
@@
 Mac OS X			& 0.019				& 2.86			\\
-Windows			& \todo{?}				& \todo{?}			\\\hline
+Windows			& 0.029				& 3.73			\\\hline
 \end{tabular}
@@
 %
-Both the time per message and overhead changed between operating systems with the Linux machine performing the most consistently and fastest.
+Both the time per message and overhead changed between operating systems with the Linux machine performing the most consistently and fastest and Windows offering the slowest speeds.
 Because the Linux test machine was very different from the test machine for both Mac OS X and Windows, it cannot be determined that operating system alone contributed to the observed differences.
-The same can be said of the Mac OS X and Windows tests given that operating systems are often optimized for their intended hardware.
+The same can be said of comparing the Mac OS X and Windows tests given that operating systems are often optimized for their intended hardware.
 However, the tests do show that {\cis} performed consistently on each of the supported operating systems.
-\todo{Windows results \& discussion}
 
@@
 Each square is a model and lines between the models represent connections where information is exchanged.
-Solid lines are connections that have already implemented, dashed lines are connections currently being implemented, and dotted lines are connections planned for the future.
+Solid lines are connections that have already been implemented, dashed lines are connections currently being implemented, and dotted lines are connections planned for the future.
 Results from one integration, a metabolic model and genetic response to CO$_2$ model, can be found in

### 2018-10-23 papers/lang2018/ms.tex
Updated figures and tables for Python-to-Matlab & Matlab-to-Matlab timings

@@
+\newif\ifcode
+
+\codefalse
+
 \newif\iftwocol
@@
 \usepackage{enumerate}
+\usepackage{listings}
 \ifdraft
@@
 {\cis} coordinates parallel execution of models in Python, C, C++, and Matlab on Linux, Mac OS X, and Windows operating systems, and handles communication in a number of data formats common to computational plant modeling.
-{\cis} is designed to be user-friendly, with visual tools developed in parallel for composing the integration networks and can be accessed at {\tt https://github.com/cropsinsilico/cis\_interface}.
+{\cis} is designed to be user-friendly and can be accessed at {\tt https://github.com/cropsinsilico/cis\_interface}.
 \end{abstract}
@@
 Many biological processes are the same across different species of plants, allowing models developed for one species to be adapted for another by modifying input parameters.
-In addition, computational models for different scales, organs, or processes can also be combined, or integrated, to probe more complex biological mechanisms.
-By integrating a metabolic flux model with a genetic regulatory model for carbon uptake,
-%
-\ifieee
-	\citep{integration_prep} 
-\else
-	\citet{integration_prep} 
-\fi
-%
-was able to identify gene candidates for regulating photosynthesis under elevated CO$_{2}$.
+In addition, computational models for different scales, organs, or processes are often related via biological dependencies.
+%can also be combined, or integrated, to probe more complex biological mechanisms.
+%By integrating a metabolic flux model with a genetic regulatory model for carbon uptake, 
+%%
+%\ifieee
+%	\citep{integration_prep} 
+%\else
+%	\citet{integration_prep} 
+%\fi
+%%
+%was able to identify gene candidates for regulating photosynthesis under elevated CO$_{2}$. 
 With advances in computational power, it should be possible to link biologically-related models to create complex integration networks capable of capturing the response of entire plants, fields, or regions.
@@
 
-We present an Open Source Python package, {\cis}, for creating and running integration networks by connecting existing computational models written in Python, Matlab, C and C++.
-{\cis} is available on Github\footnote{{\tt https://github.com/cropsinsilico/cis\_interface}}, and can be installed from PyPI\footnote{{\tt https://pypi.org/project/cis\_interface/}} via {\tt pip install {\cis}} or conda-forge\footnote{{\tt https://github.com/conda-forge/cis\_interface-feedstock}} via {\tt conda -c conda-forge install {\cis}}.
+Consider the task of integrating a root growth model with a shoot growth model. The root growth model is written in C and can be expressed as
+\begin{equation}
+R_{t+1} = R_{t} \times r_{r} + R_{t}.
+\end{equation}
+The inputs to the root growth model are
+\begin{itemize}
+	\item $R_{t}$: the root mass at time $t$
+	\item $r_{r}$: the relative root growth rate
+\end{itemize}
+and the output of the root growth model is $R_{t+1}$, the root mass at the next time step. 
+%
+The shoot growth model is written in Python and can be expressed as
+\begin{equation}
+S_{t+1} = S_{t} \times r_{s} + S_{t} - (R_{t+1} - R_{t}).
+\end{equation}
+The inputs to the shoot model are 
+\begin{itemize}
+	\item $S_{t}$: the shoot mass at time $t$
+	\item $r_{s}$: the relative shoot growth rate
+	\item $R_{t}$: the root mass at time $t$ 
+	\item $R_{t+1}$: the root mass at time $t+1$ 
+\end{itemize}
+and the output of the shoot growth model is $S_{t+1}$, the shoot mass at the next time step.
+%
+While it is possible to run both models in isolation if all of the appropriate input variables are provided, the value of $R_{t+1}$ calculated by the root growth model could be used as input to the shoot growth model.
+However, because the two models are written in different languages, the two models cannot be directly integrated.
+% without translating one of the models into the language of the other.
+Although translating one or both of these simple toy models is relatively straight forward, actual plant models are much more complex.
+In the case of real plant models, model translation is time consuming and results in unnecessary replication of model algorithms.
+
+We present an Open Source Python package, {\cis}, for creating and running integration networks by connecting existing computational models written in different programming language.
+{\cis} is available on Github\footnote{{\tt https://github.com/cropsinsilico/cis\_interface}} with full documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/}} and can be installed from PyPI\footnote{{\tt https://pypi.org/project/cis\_interface/}} via {\tt pip install {\cis}} or conda-forge\footnote{{\tt https://github.com/conda-forge/cis\_interface-feedstock}} via {\tt conda -c conda-forge install {\cis}}.
 Although developed for the purpose of integrating plant biology models, {\cis} can be applied to any situation requiring coordination between models in the supported languages.
@@
 The same can be said of data formats which can be highly tailored to the target question and may or may not include metadata like the data types, field names, or units.
-To integrate two biological models, the following questions must be addressed:
+To integrate two biological models like the root and shoot models from \S\ref{S:intro}, the following questions must be addressed:
 %
@@
 	\item Translation: How will data output by one model be translated into a format understood by the next model?
-	%\item Computational Efficiency: Will the integration make good use of computational resources?
-	%\item Human Efficiency: How much hands-on work will the integration take and can products be reused in the future?
 \end{enumerate}
 
-These three questions can be addressed through manual integration on the command line or using scripting (e.g.
-running model A, converting output from A into input expected by model B, running model B).
+These three questions can be addressed through manual integration (e.g.
+running the root model, converting the output root mass into the units expected by the shoot model, and then running the shoot model).
 However, manual integration is 1) computationally inefficient when many iterations are necessary, such as during parameter searches or steady state convergence tests, 2) unique to every integration, requiring the production of new scripts, and 3) complex when more than $>3$ models are being integrated.
@@
 %
-Another approach is a workflow/data flow/framework tool for models that are written in the same programming language or can be wrapped.
+Another approach is a workflow/data flow/framework tool for models that are written in the same programming language or can be wrapped using an intermediary \citep[e.g.
+Cython, ][]{Behnel2011a}.
 For example, there are many generic tools in Python for coordinating the execution of tasks in parallel or serial \citep[e.g.][]{babuji18, celery, luigi}.
@@
 {\cis} is executed via a command line interface (CLI), {\tt cisrun} with specification files (\S\ref{SSS:yaml}) as input.
-Based on the information contained in the specification files, {\cis} establishes a network of asynchronous communication channels (\S\ref{SS:communication}), and launches the models on new processes (\S\ref{SS:model_driver}).
+Based on the information contained in the specification files, {\cis} establishes a network of asynchronous communication channels (\S\ref{SS:communication}), and launches the models on new processes (\S\ref{SS:model_execution}).
 %At execution, {\cis} launches models on new processes based on information contained in a YAML specification file (See \S\ref{SSS:yaml} below) and establishes
@@
 % YAML
-\subsubsection{YAML Specification}\label{SSS:yaml}
+\subsubsection{Specification Files}\label{SSS:yaml}
 %
@@
 be run, and any input or output variables including their data type and units. 
+%
+\ifcode
+For example, Listings \ref{lst:rootYAML} and \ref{lst:shootYAML} are the model YAMLs for the example root and shoot growth models from \S\ref{S:intro}.
+These model YAMLs could be used in any integration without being modified because the model input and output variable will not change.
+%
+\lstinputlisting[caption={Model YAML for root model},label={lst:rootYAML}]{../../cis_interface/examples/paper/root.yml}
+%
+\lstinputlisting[caption={Model YAML for shoot model},label={lst:shootYAML}]{../../cis_interface/examples/paper/shoot.yml}
+%
+\fi
+%
 Integration networks are specified by declaring the connections 
-between models. Connections are declared by pairing an output variable from 
+between models and connections are declared by pairing an output variable from 
 one model with the input variable of another model. The {\cis} 
@@
 then direct data from one model to the next in the specified pattern. Models 
-can have as many input and/or output variables as is desired.
-Connections between models are specified by references to the 
-input/output variables associated with each model. This format allows users the 
-flexibility to create complex integration networks.
+can have as many input and/or output variables as is desired and connections between models are specified by references to the 
+input/output variables associated with each model. In addition, input and output variables can also be connected to files.
+%
+\ifcode
+For example, in the connection YAML for integrating the root and shoot models (Listing \ref{lst:connYAML}), the first two connections link the {\tt root\_input} and {\tt shoot\_input} input channels from the root and shoot model YAML's to input files.
+The third connection links the {\tt next\_root\_mass} output variable from the root model YAML (Listing \ref{lst:rootYAML}) to the {\tt next\_root\_mass} input variable from the shoot model YAML (Listing \ref{lst:shootYAML}).
+The forth connection links the {\tt next\_shoot\_mass} to an output file.
+%
+\lstinputlisting[caption={Connection YAML for integration of root and shoot models},label={lst:connYAML}]{../../cis_interface/examples/paper/paper.yml}
+%
+\fi
+%
+This format allows users the 
+flexibility to create complex integration networks and test models in isolation before running the entire integration network.
 
 %---------------------------------------------------------------------------------
-% LANGUAGE SUPPORT
-%\subsubsection{Language Support}\label{SSS:language}
+% MODEL EXECUTION
+\subsubsection{Model Execution}\label{SS:model_execution}
 %
-%{\cis} currently support models written in Python, Matlab, C, and C++ with 
-%additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}. 
-%Support for model languages is achieved through a language specific driver and API.
+{\cis} launches each model in an integration on its own process, allowing models to complete independent operations in parallel and complete tasks more quickly.
+For example, the root and shoot growth models from \S\ref{S:intro} require 2.42\,s and 4.06\,s respectively to run for 100 time steps in isolation.
+If these two models were manually integrated in serial via the command line, the integration would require a total of 6.48\,s.
+However, the same integration requires only 4.52\,s using {\cis}, a speed-up of 1.43.
+The exact speedup provided by parallel integration using {\cis} depends on how independent the models are and how equally distributed the work is between the models.
+If the models are fully independent, the speedup is limited by the time required to execute the slowest model.
+In the example, this would be the shoot growth model which required almost twice the time to execute as the root model.
+If the models are dependent on one-another, the models will have to wait for messages to be passed and the speedup will not be a great.
 
-%---------------------------------------------------------------------------------
-% MODEL DRIVERS
-\subsubsection{Model Drivers}\label{SS:model_driver}
-%
-Model drivers handle model execution, monitoring, and compilation if necessary. 
 While every model is executed on a new process, how the model is handled depends 
@@
 interpreted languages (Python and Matlab) are executed on the command line 
-with the interpreter. In the case of Matlab, where there is significant overhead 
-associated with starting the Matlab interpreter (See \S\ref{SS:results_language}), a Matlab shared engine is used 
-to execute the model. To speed up the execution, the shared engine can be started 
-in advance and then reused.
+with the interpreter. In the case of Matlab, where a significant amount of time is required to start 
+the Matlab interpreter (See \S\ref{SS:results_language}), Matlab shared engines are used 
+to execute Matlab models. Matlab shared engines are Matlab instances that other processes can 
+submit Matlab code to for execution. While shared engines also require the same amount of time 
+to start as a standard instance of the Matlab interpreter, they can be started in advance and then 
+reused.
 
@@
 the model source code and let {\cis} handle the compilation, including 
-linking against the appropriate {\cis} header library. {\cis} 
+linking against the appropriate {\cis} interface library. {\cis} 
 also has support for compiling models using Make \citep{Stallman2004} and 
 CMake \citep{Martin2006} for 
-models that already have a Makefile or CMakeLists.txt. To use these drivers, 
+models that already have a Makefile or CMakeLists.txt. To use these tools to compile a model, 
 lines are added to the recipe in order to allow linking against {\cis}.
 
-Once model are running, each model driver uses a thread to monitor the progress of the model and report back model output (e.g.
-log messages printed to stdout) and any status changes to the master process.
-If a model issues any errors, the master process will shut down any model processes that are still running.
-If a model completes without any errors, the master process will cleanup any input/output channels that are no longer required.
+Once the models are running, {\cis} uses threads on the master process to monitor the progress of the model and report back model output (e.g.
+log messages printed to stdout or stderr) and any status changes.
+If a model issues any errors, the master process will shut down any model processes that are still running and close any connections, discarding any unprocessed messages.
+If a model completes without any errors, the master process will cleanup any connections that are no longer required after waiting for all messages to be processed.
 The master process will only complete once an error is encountered or all model processes have completed.
@@
 %
-Communication within {\cis} was designed to be flexible in terms of the 
-languages, platforms, and data types. To accomplish this, {\cis} leverages 
+In order for integrations to be possible, models must be able to send and receive information to and from other models written in different languages via some communication mechanism. 
+Communication within {\cis} integration networks was designed to be flexible in terms of the 
+languages, platforms, and data types available. To accomplish this, {\cis} leverages 
 three tools different tools for communication, 
 System V IPC Queues (\S\ref{SSS:ipc}), ZeroMQ (\S\ref{SSS:zmq}), and RabbitMQ (\S\ref{SSS:rmq}), 
-which are applied as needed. The 
-particular mechanism used by {\cis} for communication is hidden from the 
-user, who will always use the same interface in the language of their model 
-(See \S\ref{SSS:interface}).
+which each have their own strengths and weaknesses. The 
+particular communication mechanism used by {\cis} is determined by the platform, available libraries, and integration strategy.
+However, regardless of the communication mechanisms, the user will always use the same interface in the language of their model (See \S\ref{SSS:interface}).
+
+%---------------------------------------------------------------------------------
+% SYSTEM V IPC
+\subsubsection{System V IPC Queues}\label{SSS:ipc}
+%
+The first communication mechanism used by {\cis} was System V interprocess 
+communication (IPC) message queues \citep{Rusling1999} on Posix (Linux and Mac OS X) systems. 
+IPC message queues allow messages to be passed between models running on separate 
+processes on the same machine. While IPC message queues are simple, fast 
+(See \S\ref{SS:results_commtype}), and are built into most Posix operating systems, they do not work in all 
+situations. IPC queues are not natively supported by Windows operating systems 
+and do not allow communication between remote processes. In addition, IPC queues also 
+have relatively low default message size limits on Mac OS X systems (2048 bytes). Once 
+the queue is full or if the message is larger than the limit, any process attempting to send an additional message will block until a 
+sufficient number of messages has been removed from the queue to accommodate the new message. 
+For messages larger than the limit, the sending process will block indefinitely. 
+This can be handled by splitting large messages into multiple smaller messages 
+(See \S\ref{SSS:large}), however, the time required to send a message increases with the number of 
+message it must be broken into (See \S\ref{SS:results_commtype}). These limits make sending large messages 
+relatively inefficient when compared with other communications mechanisms. 
+As a result, IPC queues are used by 
+{\cis} as a fallback on Posix systems if none of the other supported communication libraries have 
+not been installed.
+
+%---------------------------------------------------------------------------------
+% ZEROMQ
+\subsubsection{ZeroMQ}\label{SSS:zmq}
+%
+The preferred communication mechanism used by {\cis} are ZeroMQ sockets 
+\citep[ZMQ][]{Akgul2013}. ZMQ 
+provides broker-less communication via a number of protocols and patterns with 
+bindings in a wide variety of languages that can be installed on Posix and Windows 
+operating systems. ZMQ was adopted by {\cis} in order to allow support 
+on Windows and for future target languages (See \S\ref{SS:future}) that could not be 
+accomplished using IPC queues. In addition, while ZMQ allows 
+interprocess communication via IPC, ZMQ also 
+supports protocols for distributed communication via an Internet Protocol (IP) 
+network. While {\cis} does not currently support using these protocols 
+for distributed integration networks, this is a future development plan that has been prepared for.
+
+%---------------------------------------------------------------------------------
+% RABBITMQ
+\subsubsection{RabbitMQ}\label{SSS:rmq}
+%
+While broker-less communication like ZMQ is light weight and fast, it is not as fault 
+tolerant as brokered messaging systems that confirm message delivery and can resend 
+dropped messages. Resilience to dropped messages, while not as necessary for integrations 
+running entirely on a local machine, will be more important for integrations running on 
+distributed resources with less reliable connections. As a result, {\cis} includes support 
+for brokered communication via RabbitMQ \citep[RMQ][]{RMQ} that will be used during future 
+development to allow integrations to run on distributed resources or include remote models 
+run as services (\S\ref{SS:distributed}). Due to the slower message speed, 
+{\cis} does not currently use RMQ for communication unless explicitly specified by the user in their 
+integration network YAML.
 
@@
 %
-Regardless of the specific communication mechanism, the same asynchronous
-communication strategy is used. Models do not block on sending messages to 
+The same asynchronous communication strategy is used with each of the communication mechanisms supported by {\cis}. 
+Models do not block on sending messages to 
 output channels; the model is free to continue working on its task while 
-a separate thread waits for the message to be routed and received. Similarly, 
-a thread continuously pings input channels, moving received messages into a 
+an \emph{output driver} waits for the message to be routed and received on a separate master process thread. Similarly, 
+an \emph{input driver} continuously pings input channels on another thread, moving received messages into a 
 buffer queue so that they are ready and waiting for the receiving model 
-when it asks for input. As a result, models in complex integration networks 
-can work in parallel, improving the overall efficiency with which computational 
-resources are used. Figure \ref{fig:async} describes the general flow of messages.
+when it asks for input. These drivers work together to move message along from one model to the next like a conveyor belt. 
+As a result, models in complex integration networks are not affected by the rate at which dependent model consume their output and the speedup offered by running the models in parallel is improved.
+Figure \ref{fig:async} describes the general flow of messages.
 %
@@
 if one model is running on a remote machine).
-In these cases, the intermediate channel will connect to a RabbitMQ broker using security credentials (See \S\ref{SSS:rmq}).
+In these cases, the intermediate channel will connect to a RMQ broker using security credentials.
 	\item An input connection driver (also written in Python) runs in a separate thread, listening to the intermediate channel.
@@
 
-In addition to communication between two models, users can also specify that a 
-model should receive/send input/output from/to a file. This is specified in 
-the integration network YAML and does not impact the way the model will 
-receive/send messages. In this way, users can test their model integration in 
-isolation with input/output form/to files rather than another model.
-
-%---------------------------------------------------------------------------------
-% SYSTEM V IPC
-\subsubsection{System V IPC Queues}\label{SSS:ipc}
-%
-The first communication mechanism used by {\cis} was System V interprocess 
-communication (IPC) message queues \citep{Rusling1999} on Posix (Linux and Mac OS X) systems. 
-IPC message queues allow messages to be passed between models running on separate 
-processes on the same machine. While IPC message queues are light weight, fast 
-(See \S\ref{SS:results_commtype}), and are part of Posix operating systems, they do not work in all 
-situations. Sys V IPC queues are not natively supported by Windows operating systems 
-and do not allow communication between remote processes. In addition, IPC queues also 
-have relatively low default message size limits on Mac OS X systems (2048 bytes). Once 
-the queue is full or if the message is larger than the limit, any process attempting to send an additional message will block until a 
-sufficient number of messages has been removed from the queue to accommodate the new message. 
-For messages larger than the limit, the sending process will block indefinitely. 
-This can be handled by splitting large messages into multiple smaller messages 
-(See \S\ref{SSS:large}), however, the time required to send a message increases with the number of 
-message it must be broken into (See \S\ref{SS:results_commtype}). These limits make sending large messages 
-relatively inefficient when compared with other communications mechanisms. 
-As a result, Sys V IPC queues are used by 
-{\cis} as a fallback on Posix systems if the necessary ZeroMQ libraries have 
-not been installed.
-
-%---------------------------------------------------------------------------------
-% ZEROMQ
-\subsubsection{ZeroMQ}\label{SSS:zmq}
-%
-The preferred communication mechanism used by {\cis} are ZeroMQ sockets 
-\citep{Akgul2013}. ZeroMQ 
-provides broker-less communication via a number of protocols and patterns with 
-bindings in a wide variety of languages that can be installed on Posix and Windows 
-operating systems. ZeroMQ was adopted by {\cis} in order to allow support 
-on Windows and for future target languages (See \S\ref{SS:future}) that could not be 
-accomplished using System V IPC queues. In addition, while ZeroMQ allows 
-interprocess communication via IPC queues like System V IPC queues, ZeroMQ also 
-supports protocols for distributed communication via an Internet Protocol (IP) 
-network. While {\cis} does not currently support using these protocols 
-for distributed integration networks, this is a future development plan that has been prepared for.
-
-%---------------------------------------------------------------------------------
-% RABBITMQ
-\subsubsection{RabbitMQ}\label{SSS:rmq}
-%
-While broker-less communication like ZeroMQ is light weight and fast, it is not as fault 
-tolerant as brokered messaging systems that confirm message delivery and can resend 
-dropped messages. Resilience to dropped messages, while not as necessary for integrations 
-running entirely on a local machine, will be more important for integrations running on 
-distributed resources with less reliable connections. As a result, {\cis} includes support 
-for brokered communication via RabbitMQ \citep{RMQ} that will be used during future 
-development to allow integrations to run on distributed resources or include remote models 
-run as services (\S\ref{SS:distributed}). Due to the slower message speed, 
-{\cis} does not currently use RabbitMQ for communication unless explicitly specified by the user in their 
-integration network YAML.
+%In addition to communication between two models, users can also specify that a 
+%model should receive/send input/output from/to a file. This is specified in 
+%the integration network YAML and does not impact the way the model will 
+%receive/send messages. In this way, users can test their model integration in 
+%isolation with input/output form/to files rather than another model.
 
@@
 on the allowed size for a single message. Some of these limits can be quite large 
-($2^{20}$ bytes for ZeroMQ and RabbitMQ), while others are very limiting 
-(2048 bytes on Mac OS X for Sys V IPC queues). Although messages consisting of a few scalars are unlikely to 
+($2^{20}$ bytes for ZMQ and RMQ), while others are very limiting 
+(2048 bytes on Mac OS X for IPC queues). Although messages consisting of a few scalars are unlikely to 
 exceed these limits, biological inputs and outputs are often much more complex. 
@@
 % RESULTS
-\section{Results}\label{S:results}
+\section{Results \& Discussion}\label{S:results}
 %
 In order to evaluate {\cis}, performance tests were run on machines with Linux, Mac OS X, and Windows operating systems for different communication mechanisms, Python versions, and language combinations.
-During each run, $N_{msg}$ messages of size $S_{msg}$ were sent from the source model (in one language) to the destination model (in another language) which then output the messages to file for verification.
+During each run, $N_{msg}$ messages of size $S_{msg}$ were sent from a source model (in one language) to a destination model (in another language) which then sends the messages to an output file for verification.
+These are test models that do not perform any operations outside of passing the messages in order to accurately gauge the performance of the {\cis} machinery on its own.
 Performance tests were run using the perf package \citep{Stinner2018}.
-Each run was repeated 10 times to mitigate fluctuations due to external loads on the test machines.
+Each run was repeated 10 times with warm-up runs in between to mitigate fluctuations due to external loads on the test machines.
 
@@
 Time per message (slope of Figure \ref{fig:commtype}) is the average amount of time required to send a single message containing 1000 bytes and constrains how quickly messages can be passed between the models.
-Overhead (intercept of Figure \ref{fig:commtype}) is the amount of execution time that would be required if no messages were passed between the models and includes the time required to set up communication mechanisms, start the model drivers, and clean up the integration network.
+Overhead (intercept of Figure \ref{fig:commtype}) is the amount of execution time that would be required if no messages were passed between the models and includes the time required to set up communication mechanisms, start the models, and clean up the integration network.
 %
@@
 %
-Although Sys V IPC queues are slightly faster than ZeroMQ TCP sockets in both time per message and overhead for messages smaller than the limit, IPC queues have a much smaller maximum size limit for messages.
+Although IPC queues are slightly faster than ZMQ TCP sockets in both time per message and overhead for messages smaller than the limit, IPC queues have a much smaller maximum size limit for messages.
 As discussed in \S\ref{SSS:large}, messages larger than this limit (2048 bytes) must be broken up into multiple messages, compounding the time required to send these messages and making IPC queues a poor choice for sending data over $\sim10^5$ bytes.
-This limit is much larger for ZeroMQ sockets ($\sim10^6$ bytes).
+This limit is much larger for ZMQ sockets ($\sim10^6$ bytes).
 However, for messages less than the limit, both ZMQ and IPC communication has no measurable dependence on message size.
@@
 The runs for Figure \ref{fig:language} and Table \ref{tab:language} were run on the Linux machine from above while those for Figure \ref{fig:language_matlab} and Table \ref{tab:language_matlab} were run on a 2015 MacBook Pro with a 2.9 GHz Intel Core i5 processor running macOS High Sierra 10.13.4 that had Matlab R2017a installed.
-In both cases, ZeroMQ communication was used with Python 2.7.
+In both cases, ZMQ communication was used with Python 2.7.
 %
@@
 Python			& Python				& 0.019				& 2.86			\\
-Python			& Matlab				& 0.500				& 11.61			\\
+Python			& Matlab				& 0.500				& 11.31			\\
 \hline%
@@
 Matlab			& Python				& 0.139				& 12.23			\\
-Matlab			& Matlab				& 0.559				& 19.28			\\
+Matlab			& Matlab				& 0.526				& 21.71			\\
 \hline%
@@
 The largest difference between the different test cases is in in the overhead.
-Every model driver entails some overhead.
+Every model language entails some overhead.
 For interpreted languages (e.g.
@@
 Figure \ref{fig:python} and Table \ref{tab:python} compare the execution times when using different versions of Python.
-The performance tests for both Python 2.7 and Python 3.5 were performed on the Linux machine from above using ZeroMQ communication and two different integration (Python-to-Python and C-to-C).
+The performance tests for both Python 2.7 and Python 3.5 were performed on the Linux machine from above using ZMQ communication and two different integration (Python-to-Python and C-to-C).
 %
@@
 There is a slight difference in time per message between the two versions for all models that shows up in the scaling of execution time with message size for large messages ($>10^6$ bytes).
-Differences at large message sizes between the two versions arrises from the use of Python drivers to transport messages between models and is more pronounced in integrations with Python models due to the increased amount of Python code.
+Differences at large message sizes between the two versions arrises from the use of Python classes to transport messages between models and is more pronounced in integrations with Python models due to the increased amount of Python code.
 
@@
 The tests for Windows were run on the same machine from a dual boot of Windows 10.
-All platform dependent performance tests use ZeroMQ communication and the Python-to-Python integration.
+All platform dependent performance tests use ZMQ communication and the Python-to-Python integration.
 %
@@
 % SUMMARY & DISCUSSION
-\section{Summary \& Discussion}\label{S:discuss}
+\section{Summary}\label{S:discuss}
 
@@
 Figure \ref{fig:network} shows the progress so far.
-Each square is a model and lines between the models represent connections where information is exchanged.
+%
+\ifinclfig
+ 	\begin{figure}[htbp]
+	\begin{center}
+	\includegraphics[width=\columnwidth,keepaspectratio]{./images/CiS-Languages.png}
+	\caption{Progress towards full Crops in Silico integration network.
+Each square represents a model and the color of the square indicates the language in which the model is written.
+Arrows between models represent connections being made between the models using {\cis}.
+Solid lines are connections that have already been made, dashed lines are in-progress connections, and dotted lines are planned connections.}
+	\label{fig:network}
+	\end{center}
+	\end{figure}
+\fi
+%
+Each square is a model, color indicates the programming language that the model is written in, and lines between the models represent connections where information is exchanged.
 Solid lines are connections that have already been implemented, dashed lines are connections currently being implemented, and dotted lines are connections planned for the future.
-Results from one integration, a metabolic model and genetic response to CO$_2$ model, can be found in
+When complete, the integration network will include 9 models of molecular, genetic, cellular, organ, and plant level processes in 4 different programming languages, produced by $>10$ researchers at 4 different institutions.
+Although not yet complete, partial integrations from this network are already yielding new insights.
+For example, by integrating a metabolic flux model  with gene expression data from soybean (\emph{Glycine max}) plants grown under ambient (380\,$\mu$mol/mol) and elevated (550\,$\mu$mol/mol) carbon dioxide (CO2),
 %
@@
 %
-\ifinclfig
- 	\begin{figure}[htbp]
-	\begin{center}
-	\includegraphics[width=\columnwidth,keepaspectratio]{./images/CiS-BioCro-Planned.png}
-	\caption{Progress towards full Crops in Silico integration network.}
-	\label{fig:network}
-	\end{center}
-	\end{figure}
-\fi
-%
-%\todo{more here? teaser results?}
+simulated the increased carbon assimilation rate observed in field studies \citep{Bernacchi2005}.
+The integrated modeling also identified gene candidates predicted to regulate enzymes involved in the light and dark reactions of photosynthesis under elevated CO$_2$.
 
@@
 automating the submission process to HPC and compute resources. In addition, 
-we will add tools for running models as a service and using RabbitMQ to 
+we will add tools for running models as a service and using RMQ to 
 permit access to these models within integration networks.
@@
 %
-This work was supported by funding from the Foundation for Food and Agriculture Research (FFAR), Institute for Sustainability, Energy, and Environment (iSEE), the National Center for Supercomputing Applications (NCSA), and the Gordon and Betty Moore Foundation's Data-Driven Discovery Initiative through Grant GBMF4561 to Matthew Turk.
+This work was supported by funding from the Foundation for Food and Agriculture Research (FFAR) through Grant 515760 and Cost Share C2252 with the Institute for Sustainability, Energy, and Environment (iSEE) and the National Center for Supercomputing Applications (NCSA) at the University of Illinois at Urbana-Champaign (UIUC), as well as the Gordon and Betty Moore Foundation's Data-Driven Discovery Initiative through Grant GBMF4561 to Matthew Turk.
 %

### 2018-10-23 papers/lang2018/ms.tex
Minor edit

@@
 	\begin{IEEEkeywords}
-		Plant modeling, Modeling, Communication
+		Plant Modeling, Modeling, Communication, Computational Framework, Model Integration, Parallel Processing
 	\end{IEEEkeywords}
@@
 Scientists will generally write programs in the language they are most familiar with and, given that there is no consensus on one programming language as ideal for all scientific applications, the language will often vary between research groups, if not between members within the groups themselves.
-The same can be said of data formats which can be highly tailored to the target question and may or may not include metadata like the data types, field names, or units.
+The same can be said of data formats which can be highly tailored to the target question and may or may not include metadata such as the data types, field names, or units.
 To integrate two biological models like the root and shoot models from \S\ref{S:intro}, the following questions must be addressed:

### 2018-10-24 papers/lang2018/ms.tex
Add reference to FLAME

@@
 Scientists can compose their models using SBML markup and then run their model using software designed to parse SBML files.
-Other DSLs for biological models include LPy \citep{Boudon2012}, CellML \citep{Cuellar2003}, and BioPAX \citep{Demir2010}.
+Other DSLs for biological models include LPy \citep{Boudon2012}, CellML \citep{Cuellar2003}, FLAME \citep{Coakley2012}, and BioPAX \citep{Demir2010}.
 Dedicated model DSLs improve the reusability of models written in the DSL and have the advantage of often implicitly handling data transformation.
-However, they provide no help for existing models that are not in the DSL or those that cannot be expressed within the constraints of the DSL.
-In addition, learning a new DSL, in addition to a programming language, can be daunting.
+However, they provide no help for existing models that are not written in the DSL or those that cannot be expressed within the constraints of the DSL.
+Furthermore, learning a new DSL, in addition to a programming language, can be daunting.
 

### 2018-11-04 cis_interface/command_line.py
Replace OSX in images and paper content to MacOS

@@ -95,12 +95,12 @@ def cistime_paper():
     _lang_list_nomatlab.remove('matlab')
     timing.plot_scalings(compare='platform', python_ver='2.7')
     # All plots on Linux, no matlab
-    timing.plot_scalings(compare='commtype', platform='Linux', python_ver='2.7')
+    timing.plot_scalings(compare='comm_type', platform='Linux', python_ver='2.7')
     timing.plot_scalings(compare='python_ver', platform='Linux')
     timing.plot_scalings(compare='language', platform='Linux', python_ver='2.7',
                          compare_values=_lang_list_nomatlab)
-    # Language comparision on OSX, with matlab
-    timing.plot_scalings(compare='language', platform='OSX', python_ver='2.7',
+    # Language comparision on MacOS, with matlab
+    timing.plot_scalings(compare='language', platform='MacOS', python_ver='2.7',
                          compare_values=_lang_list)
 
 

### 2018-11-04 papers/lang2018/ms.tex
Replace OSX in images and paper content to MacOS

@@
 We present \href{https://github.com/cropsinsilico/cis_interface}{\cis}, a Python package for running integration networks with connections between models across languages and scales.
-{\cis} coordinates parallel execution of models in Python, C, C++, and Matlab on Linux, Mac OS X, and Windows operating systems, and handles communication in a number of data formats common to computational plant modeling.
+{\cis} coordinates parallel execution of models in Python, C, C++, and Matlab on Linux, Mac OS, and Windows operating systems, and handles communication in a number of data formats common to computational plant modeling.
 {\cis} is designed to be user-friendly and can be accessed at {\tt https://github.com/cropsinsilico/cis\_interface}.

### 2018-11-04 papers/lang2018/prep.sh
Add missing paper files and ignore latex temp files

@@ -1,0 +1,6 @@
+cp /Users/langmm/Library/texmf/bibtex/bib/CiS.bib /Users/langmm/cis_interface/papers/lang2018/
+cp /Users/langmm/cis_interface/images/compare_timed_pipe_commtype_python_python_Linux_27_average.png /Users/langmm/cis_interface/papers/lang2018/images/scaling_commtype.png
+cp /Users/langmm/cis_interface/images/compare_timed_pipe_language_ZMQComm_Linux_27_average.png /Users/langmm/cis_interface/papers/lang2018/images/scaling_language.png
+cp /Users/langmm/cis_interface/images/compare_timed_pipe_language_ZMQComm_MacOS_27_average.png /Users/langmm/cis_interface/papers/lang2018/images/scaling_language_matlab.png
+cp /Users/langmm/cis_interface/images/compare_timed_pipe_platform_ZMQComm_python_python_27_average.png /Users/langmm/cis_interface/papers/lang2018/images/scaling_platform.png
+cp /Users/langmm/cis_interface/images/compare_timed_pipe_python_ZMQComm_python_python_Linux_average.png /Users/langmm/cis_interface/papers/lang2018/images/scaling_python.png

### 2019-01-01 papers/lang2018/ms.tex
Response to reviewer 2 and addition of example in response to reviewer 1

@@
 	\else
-		\documentclass[preprint]{aastex}
+		\documentclass[preprint]{aastex6}
 	\fi
@@
 \usepackage{listings}
+\usepackage{amsmath}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+%\usepackage[noend]{algpseudocode}
 \ifdraft
@@
 \newcommand{\TB}{\,Tbytes}
-\newcommand{\cis}{{\tt cis\_interface}{}}
+\newcommand{\pkg}{{\tt \todo{cis\_interface}}{}}
+\newcommand{\pkgrun}{{\tt \todo{cisrun}}}
+\newcommand{\A}{Appendix{ }}
+
+%---------------------------------------------------------------------------------
+% SOURCE CODE
+\newcommand{\rootsrc}{\lstinputlisting[language=C,label={lst:src:root},caption={root.h: Source code for root model in C.}]{../../cis_interface/examples/root_to_shoot/src/root.h}}
+\newcommand{\shootsrc}{\lstinputlisting[language=Python,label={lst:src:shoot},caption={shoot.py: Source code for shoot model in Python.}]{../../cis_interface/examples/root_to_shoot/src/shoot.py}}
+
+%---------------------------------------------------------------------------------
+% WRAPPER CODE
+\newcommand{\rootwrap}{\lstinputlisting[language=C,label={lst:wrap:root},caption={root\_wrapper.c: Wrapper source code for root model in C.}]{../../cis_interface/examples/root_to_shoot/src/root_wrapper.c}}
+\newcommand{\shootwrap}{\lstinputlisting[language=Python,label={lst:wrap:shoot},caption={shoot\_wrapper.py: Wrapper source code for shoot model in Python}]{../../cis_interface/examples/root_to_shoot/src/shoot_wrapper.py}}
+
+%---------------------------------------------------------------------------------
+% YAML FILES
+\newcommand{\rootyml}{\lstinputlisting[label={lst:yml:root},caption={root.yml: Model YAML specification file for root model.}]{../../cis_interface/examples/root_to_shoot/root.yml}}
+\newcommand{\shootyml}{\lstinputlisting[label={lst:yml:shoot},caption={shoot.yml: Model YAML specification file for shoot model.}]{../../cis_interface/examples/root_to_shoot/shoot.yml}}
+\newcommand{\rootymlfiles}{\lstinputlisting[label={lst:yml:root_files},caption={root\_files.yml: Connection YAML specification file for running the root model in isolation.}]{../../cis_interface/examples/root_to_shoot/root_files.yml}}
+\newcommand{\shootymlfiles}{\lstinputlisting[label={lst:yml:shoot_files},caption={shoot\_files.yml: Model YAML specification file for running the shoot model in isolation.}]{../../cis_interface/examples/root_to_shoot/shoot_files.yml}}
+\newcommand{\intyml}{\lstinputlisting[label={lst:yml:int},caption={root\_to\_shoot.yml: Connection YAML specification file for the root/shoot integration.}]{../../cis_interface/examples/root_to_shoot/root_to_shoot.yml}}
+
+%---------------------------------------------------------------------------------
+% CODE STYLE
+\usepackage{color}
+
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{1,1,1}
+%\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{backcolour},   
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\footnotesize,
+    breakatwhitespace=false,         
+    breaklines=true,                 
+    captionpos=b,                    
+    keepspaces=true,                 
+    numbers=left,                    
+    numbersep=5pt,                  
+    showspaces=false,                
+    showstringspaces=false,
+    showtabs=false,                  
+    tabsize=2
+}
+
+\lstset{style=mystyle}
 
@@
 \else
-	\shorttitle{{\cis}}
-\fi
-
-\title{{\cis}: A Python Package for Integrating Computational Models Across Languages and Scales}
+	\shorttitle{{\pkg}}
+\fi
+
+\title{{\pkg}: A Python Package for Integrating Computational Models Across Languages and Scales}
 
@@
 However, the technological barriers introduced by differences in language and data formats has slowed this progress.
-We present \href{https://github.com/cropsinsilico/cis_interface}{\cis}, a Python package for running integration networks with connections between models across languages and scales.
-{\cis} coordinates parallel execution of models in Python, C, C++, and Matlab on Linux, Mac OS, and Windows operating systems, and handles communication in a number of data formats common to computational plant modeling.
-{\cis} is designed to be user-friendly and can be accessed at {\tt https://github.com/cropsinsilico/cis\_interface}.
+We present \href{https://github.com/cropsinsilico/cis_interface}{\pkg}, a Python package for running integration networks with connections between models across languages and scales.
+{\pkg} coordinates parallel execution of models in Python, C, C++, and Matlab on Linux, Mac OS, and Windows operating systems, and handles communication in a number of data formats common to computational plant modeling.
+{\pkg} is designed to be user-friendly and can be accessed at {\tt https://github.com/cropsinsilico/cis\_interface}.
 \end{abstract}
@@
 
-Consider the task of integrating a root growth model with a shoot growth model. The root growth model is written in C and can be expressed as
+Consider the task of integrating a root growth model with a shoot growth model. The root growth model is written in C (see Listing \ref{lst:src:root}) and can be expressed as
 \begin{equation}
-R_{t+1} = R_{t} \times r_{r} + R_{t}.
+R_{t+1} = R_{t} \times r_{r} \times dt + R_{t}.
 \end{equation}
@@
 	\item $r_{r}$: the relative root growth rate
+	\item $dt$: the time step
 \end{itemize}
@@
 %
-The shoot growth model is written in Python and can be expressed as
+The shoot growth model is written in Python (see Listing \ref{lst:src:shoot}) and can be expressed as
 \begin{equation}
-S_{t+1} = S_{t} \times r_{s} + S_{t} - (R_{t+1} - R_{t}).
+S_{t+1} = S_{t} \times r_{s} \times dt + S_{t} - (R_{t+1} - R_{t}).
 \end{equation}
@@
 	\item $r_{s}$: the relative shoot growth rate
+	\item $dt$: the time step
 	\item $R_{t}$: the root mass at time $t$ 
@@
 However, because the two models are written in different languages, the two models cannot be directly integrated.
-% without translating one of the models into the language of the other.
-Although translating one or both of these simple toy models is relatively straight forward, actual plant models are much more complex.
-In the case of real plant models, model translation is time consuming and results in unnecessary replication of model algorithms.
-
-We present an Open Source Python package, {\cis}, for creating and running integration networks by connecting existing computational models written in different programming language.
-{\cis} is available on Github\footnote{{\tt https://github.com/cropsinsilico/cis\_interface}} with full documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/}} and can be installed from PyPI\footnote{{\tt https://pypi.org/project/cis\_interface/}} via {\tt pip install {\cis}} or conda-forge\footnote{{\tt https://github.com/conda-forge/cis\_interface-feedstock}} via {\tt conda -c conda-forge install {\cis}}.
-Although developed for the purpose of integrating plant biology models, {\cis} can be applied to any situation requiring coordination between models in the supported languages.
-\S\ref{S:background} provides background information on existing efforts for cross-language model integration, \S\ref{S:methods} describes the methods used by the {\cis} package to integrate models, \S\ref{S:results} presents several tests demonstrating the performance of message passing between integrated models, and \S\ref{S:discuss} summarizes the features of the {\cis} package, describes a few use cases, and outlines areas for future development.
+To integrate the models a scientist must either "manually" integrate them by running one model and then the other or translate one of the models into the language of the other so that it can be called directly.
+For a tightly coupled set of models such as these where one model depends directly on the result from the other at the current time step, manual integration is very inefficient for more than a handful of time steps even when done using a script.
+Although translating one or both of these simple toy models would be relatively straight forward, actual plant models are much more complex and could contain thousands of lines of code that would be time consuming to translate and result in unnecessary replication of model algorithms.
+
+We present an Open Source Python package, {\pkg}, for creating and running integration networks by connecting existing computational models written in different programming languages.
+{\pkg} was developed as part of the Crops \emph{in Silico} \citep{Marshall-Colon2017} initiative to build a complete crop \emph{in silico} from the level of the genes to the level of the field.
+{\pkg} is available on Github\footnote{{\tt https://github.com/cropsinsilico/cis\_interface}} with full documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/}} and can be installed from PyPI\footnote{{\tt https://pypi.org/project/cis\_interface/}} via {\tt pip install {\pkg}} or conda-forge\footnote{{\tt https://github.com/conda-forge/cis\_interface-feedstock}} via {\tt conda -c conda-forge install {\pkg}}.
+Although developed for the purpose of integrating plant biology models, {\pkg} can be applied to any situation requiring coordination between models in the supported languages.
+\S\ref{S:background} provides background information on existing efforts for cross-language model integration, \S\ref{S:methods} describes the methods used by the {\pkg} package to integrate models, \S\ref{S:results} presents several tests demonstrating the performance of message passing between integrated models, and \S\ref{S:discuss} summarizes the features of the {\pkg} package, describes a few use cases, and outlines areas of ongoing and future development.
 
@@
 %
-Computational plant models are usually written with a very specific research question in mind.
-The resulting programs are highly tuned for one purpose and written specifically for scientists within an isolated group of collaborators.
-Scientists will generally write programs in the language they are most familiar with and, given that there is no consensus on one programming language as ideal for all scientific applications, the language will often vary between research groups, if not between members within the groups themselves.
-The same can be said of data formats which can be highly tailored to the target question and may or may not include metadata such as the data types, field names, or units.
-To integrate two biological models like the root and shoot models from \S\ref{S:intro}, the following questions must be addressed:
+Computational plant models are usually written with a very specific research question in mind \citep[e.g.
+describing a specific metabolic pathway in C4 photosynthesis,][]{Wang2014}.
+The resulting model codes are often highly tuned for this one purpose and are written specifically for scientists within an isolated group of collaborators.
+As a result, it is unlikely that models produced by two independent groups will be directly compatible in terms of language, data format, or units \citep{Marshall-Colon2017}.
+
+There are models written in Matlab \citep{ZHU2013, Wang2014}, Python \citep{Pradal2009}, C++ \citep{Merks2011, Postma2017}, Microsoft Excel \citep{Sharkey2016}, VisualBasic \citep{Humphries1995, HALL2013}, R \citep{WANG2015}, Java \citep{Song2013, Kappas2013}, Fortran \citep{Goudriaan1994}, and several domain specific languages \citep[e.g.
+SBML,][]{Hucka2003}.
+The variety of data formats used by these codes is just as diverse.
+While some specific subfields have settled on standards for things like microarray gene expression data \citep[e.g.
+MINiML,][]{GEO2017}, many models use unique data formats that may or may not include metadata such as the data types, field names, or units.
+%For example, OpenAlea \citep{Pradal2008} defines its own Multiscale Tree Graph (MTG) tree structure for describing plant topology at different scales that is an extension to the MTG format used by AMAPmod \citep{Godin1998}.
+For example, while LPy \citep{Boudon2012} and Houdini FX \citep{HoudiniFX} both allow users to specify plant structures via L-systems, they differ slightly in their syntax and so are not directly compatible.
+Or, as another example, 3D canopy structures can be generated in any unit and standard 3D geometry formats like Ply \citep{ply} and Obj \citep{obj} do not include units in their metadata.
+As a result, if an LPy model produces a canopy in Ply format with units of centimeters, the photosynthetic photon flux density (PPFD) output by a ray tracer like fastTracer \citep{Song2013} (which expects a geometry in units of meters) would be incorrect.
+
+As a result, integrating two biological models like the root and shoot models from \S\ref{S:intro} requires that the following questions be addressed first:
 %
@@
 running the root model, converting the output root mass into the units expected by the shoot model, and then running the shoot model).
-However, manual integration is 1) computationally inefficient when many iterations are necessary, such as during parameter searches or steady state convergence tests, 2) unique to every integration, requiring the production of new scripts, and 3) complex when more than $>3$ models are being integrated.
+However, manual integration is 1) computationally inefficient when many iterations are necessary, such as during parameter searches or steady state convergence tests, 2) unique to every integration, requiring the production of new scripts, and 3) complex when more than three models are being integrated.
 Within both computational biology and the larger scientific computing community, groups have developed around different solutions to the problem of connecting models.
+These solutions can generally be grouped into two categories: domain specific languages and frameworks.
 
@@
 
-{\cis} interface is an example of a framework that overcomes these issues by exposing simple and easily accessible interfaces in the language of the model (See \S\ref{SSS:interface}) that permit messages to be passed between the model processes as they run in parallel (See \S\ref{SS:communication}).
+{\pkg} is an example of a framework that overcomes these issues by exposing simple and easily accessible interfaces in the language of the model (See \S\ref{SSS:interface}) that permit messages to be passed between the model processes as they run in parallel (See \S\ref{SS:communication}).
 As a result, models writers only need knowledge of the language in which their model is written.
@@
 
-{\cis} was designed to address the questions from \S\ref{S:background} while being:
+{\pkg} was designed to address the questions from \S\ref{S:background} while being:
 
@@
 
-{\cis} currently support models written in Python, Matlab, C, and C++ with 
+{\pkg} currently support models written in Python, Matlab, C, and C++ with 
 additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}. Support for additional languages is planned for future development (See \S\ref{SS:langsupport}).
@@
 %
-{\cis} is executed via a command line interface (CLI), {\tt cisrun} with specification files (\S\ref{SSS:yaml}) as input.
-Based on the information contained in the specification files, {\cis} establishes a network of asynchronous communication channels (\S\ref{SS:communication}), and launches the models on new processes (\S\ref{SS:model_execution}).
-%At execution, {\cis} launches models on new processes based on information contained in a YAML specification file (See \S\ref{SSS:yaml} below) and establishes
+{\pkg} is executed via a command line interface (CLI), {\pkgrun} with specification files (\S\ref{SSS:yaml}) as input.
+Based on the information contained in the specification files, {\pkg} establishes a network of asynchronous communication channels (\S\ref{SS:communication}), and launches the models on new processes (\S\ref{SS:model_execution}).
+%At execution, {\pkg} launches models on new processes based on information contained in a YAML specification file (See \S\ref{SSS:yaml} below) and establishes
 
@@
 %
-\ifcode
-For example, Listings \ref{lst:rootYAML} and \ref{lst:shootYAML} are the model YAMLs for the example root and shoot growth models from \S\ref{S:intro}.
-These model YAMLs could be used in any integration without being modified because the model input and output variable will not change.
-%
-\lstinputlisting[caption={Model YAML for root model},label={lst:rootYAML}]{../../cis_interface/examples/paper/root.yml}
-%
-\lstinputlisting[caption={Model YAML for shoot model},label={lst:shootYAML}]{../../cis_interface/examples/paper/shoot.yml}
-%
-\fi
+%For example, Listings \ref{lst:yml:root} and \ref{lst:yml:shoot} from \A\ref{A:yml} are the model YAMLs for the example root and shoot growth models from \S\ref{S:intro}.
+These model YAMLs could be used in any integration without being modified because the model inputs and outputs will not change.
+%
+%\rootyml
+%
+%\shootyml
 %
@@
 between models and connections are declared by pairing an output variable from 
-one model with the input variable of another model. The {\cis} 
+one model with the input variable of another model. The {\pkg} 
 CLI sets up the necessary communication mechanisms to 
@@
 %
-\ifcode
-For example, in the connection YAML for integrating the root and shoot models (Listing \ref{lst:connYAML}), the first two connections link the {\tt root\_input} and {\tt shoot\_input} input channels from the root and shoot model YAML's to input files.
-The third connection links the {\tt next\_root\_mass} output variable from the root model YAML (Listing \ref{lst:rootYAML}) to the {\tt next\_root\_mass} input variable from the shoot model YAML (Listing \ref{lst:shootYAML}).
+%For example, in the connection YAML for integrating the root and shoot models (Listing \ref{lst:yml:int} from \A\ref{A:yml}), the first two connections link the {\tt root\_input} and {\tt shoot\_input} input channels from the root and shoot model YAML's to input files.
+The third connection links the {\tt next\_root\_mass} output variable from the root model YAML (Listing \ref{lst:yml:root}) to the {\tt next\_root\_mass} input variable from the shoot model YAML (Listing \ref{lst:yml:shoot}).
 The forth connection links the {\tt next\_shoot\_mass} to an output file.
 %
-\lstinputlisting[caption={Connection YAML for integration of root and shoot models},label={lst:connYAML}]{../../cis_interface/examples/paper/paper.yml}
-%
-\fi
+%\intyml
 %
@@
 %
-{\cis} launches each model in an integration on its own process, allowing models to complete independent operations in parallel and complete tasks more quickly.
+{\pkg} launches each model in an integration on its own process, allowing models to complete independent operations in parallel and complete tasks more quickly.
 For example, the root and shoot growth models from \S\ref{S:intro} require 2.42\,s and 4.06\,s respectively to run for 100 time steps in isolation.
 If these two models were manually integrated in serial via the command line, the integration would require a total of 6.48\,s.
-However, the same integration requires only 4.52\,s using {\cis}, a speed-up of 1.43.
-The exact speedup provided by parallel integration using {\cis} depends on how independent the models are and how equally distributed the work is between the models.
+However, the same integration requires only 4.52\,s using {\pkg}, a speed-up of 1.43.
+The exact speedup provided by parallel integration using {\pkg} depends on how independent the models are and how equally distributed the work is between the models.
 If the models are fully independent, the speedup is limited by the time required to execute the slowest model.
 In the example, this would be the shoot growth model which required almost twice the time to execute as the root model.
-If the models are dependent on one-another, the models will have to wait for messages to be passed and the speedup will not be a great.
+If the models are dependent on one-another, the models will have to wait for messages to be passed and the speedup will not be as great.
 
@@
 compile the model themselves, provided they link against the appropriate 
-{\cis} header library. Alternatively, users can provide the location of 
-the model source code and let {\cis} handle the compilation, including 
-linking against the appropriate {\cis} interface library. {\cis} 
+{\pkg} header library. Alternatively, users can provide the location of 
+the model source code and let {\pkg} handle the compilation, including 
+linking against the appropriate {\pkg} interface library. {\pkg} 
 also has support for compiling models using Make \citep{Stallman2004} and 
@@
 models that already have a Makefile or CMakeLists.txt. To use these tools to compile a model, 
-lines are added to the recipe in order to allow linking against {\cis}.
-
-Once the models are running, {\cis} uses threads on the master process to monitor the progress of the model and report back model output (e.g.
+lines are added to the recipe in order to allow linking against {\pkg}.
+
+Once the models are running, {\pkg} uses threads on the master process to monitor the progress of the model and report back model output (e.g.
 log messages printed to stdout or stderr) and any status changes.
@@
 In order for integrations to be possible, models must be able to send and receive information to and from other models written in different languages via some communication mechanism. 
-Communication within {\cis} integration networks was designed to be flexible in terms of the 
-languages, platforms, and data types available. To accomplish this, {\cis} leverages 
-three tools different tools for communication, 
+Communication within {\pkg} integration networks was designed to be flexible in terms of the 
+languages, platforms, and data types available. To accomplish this, {\pkg} leverages 
+three different tools for communication, 
 System V IPC Queues (\S\ref{SSS:ipc}), ZeroMQ (\S\ref{SSS:zmq}), and RabbitMQ (\S\ref{SSS:rmq}), 
 which each have their own strengths and weaknesses. The 
-particular communication mechanism used by {\cis} is determined by the platform, available libraries, and integration strategy.
+particular communication mechanism used by {\pkg} is determined by the platform, available libraries, and integration strategy.
 However, regardless of the communication mechanisms, the user will always use the same interface in the language of their model (See \S\ref{SSS:interface}).
@@
 %
-The first communication mechanism used by {\cis} was System V interprocess 
+The first communication mechanism used by {\pkg} was System V interprocess 
 communication (IPC) message queues \citep{Rusling1999} on Posix (Linux and Mac OS X) systems. 
@@
 As a result, IPC queues are used by 
-{\cis} as a fallback on Posix systems if none of the other supported communication libraries have 
+{\pkg} as a fallback on Posix systems if none of the other supported communication libraries have 
 not been installed.
@@
 %
-The preferred communication mechanism used by {\cis} are ZeroMQ sockets 
+The preferred communication mechanism used by {\pkg} are ZeroMQ sockets 
 \citep[ZMQ][]{Akgul2013}. ZMQ 
@@
 bindings in a wide variety of languages that can be installed on Posix and Windows 
-operating systems. ZMQ was adopted by {\cis} in order to allow support 
+operating systems. ZMQ was adopted by {\pkg} in order to allow support 
 on Windows and for future target languages (See \S\ref{SS:future}) that could not be 
@@
 supports protocols for distributed communication via an Internet Protocol (IP) 
-network. While {\cis} does not currently support using these protocols 
+network. While {\pkg} does not currently support using these protocols 
 for distributed integration networks, this is a future development plan that has been prepared for.
@@
 running entirely on a local machine, will be more important for integrations running on 
-distributed resources with less reliable connections. As a result, {\cis} includes support 
+distributed resources with less reliable connections. As a result, {\pkg} includes support 
 for brokered communication via RabbitMQ \citep[RMQ][]{RMQ} that will be used during future 
@@
 run as services (\S\ref{SS:distributed}). Due to the slower message speed, 
-{\cis} does not currently use RMQ for communication unless explicitly specified by the user in their 
+{\pkg} does not currently use RMQ for communication unless explicitly specified by the user in their 
 integration network YAML.
@@
 %
-The same asynchronous communication strategy is used with each of the communication mechanisms supported by {\cis}. 
+The same asynchronous communication strategy is used with each of the communication mechanisms supported by {\pkg}. 
 Models do not block on sending messages to 
@@
 %
-All of the communication tools leveraged by {\cis} have intrinsic limits 
+All of the communication tools leveraged by {\pkg} have intrinsic limits 
 on the allowed size for a single message. Some of these limits can be quite large 
@@
 limits. To handle messages that are larger than the limit of the communication 
-mechanism being used, {\cis} splits the message up into multiple smaller 
-messages. In addition, for large messages, {\cis} creates new, temporary 
+mechanism being used, {\pkg} splits the message up into multiple smaller 
+messages. In addition, for large messages, {\pkg} creates new, temporary 
 communication channels that are used exclusively for a single message and 
@@
 %
-{\cis} provides interface functions/classes for communication that are 
+{\pkg} provides interface functions/classes for communication that are 
 written in each of the supported languages. Language specific interfaces 
@@
 While some data formats are self-descriptive and include these types of information as meta data, this is not true of all data formats.
-To combat this, {\cis} requires models to explicitly specify the format of input and output expected by a model in the model YAML.
-{\cis} can then handle a number of conversions between models without prompting as well as serialization/deserialization to the correct type in each of the supported languages.
-Data formats currently supported by {\cis} include:
+To combat this, {\pkg} requires models to explicitly specify the format of input and output expected by a model in the model YAML.
+{\pkg} can then handle a number of conversions between models without prompting as well as serialization/deserialization to the correct type in each of the supported languages.
+Data formats currently supported by {\pkg} include:
 %
@@
 %
-In addition, {\cis} offers the option to specify units for scalars, arrays, tabular data, and pandas data frames.
+In addition, {\pkg} offers the option to specify units for scalars, arrays, tabular data, and pandas data frames.
 Units are tracked using the {\tt unyt} package \citep{Goldbaum2018}, which allows physical units to be associated with scalars and arrays.
-If two models use different units (and both are specified), {\cis} will automatically perform the necessary conversions before passing data from one model to the next.
+If two models use different units (and both are specified), {\pkg} will automatically perform the necessary conversions before passing data from one model to the next.
+
+%---------------------------------------------------------------------------------
+%---------------------------------------------------------------------------------
+% WORKED EXAMPLE
+\section{Worked Example}\label{S:example}
+%
+In order to illustrate how {\pkg} is intended to be used by model writers, the following walks through the integration of the two examples models from \S\ref{S:intro}.
+All of the source code and YAML files discussed in this section are available in the {\pkg} GitHub repository and will be included with future releases on PyPI and conda-forge.
+This example assumes that the user starts with the following existing model source code:
+%
+\rootsrc
+%
+\shootsrc
+%
+The root source code in Listing \ref{lst:src:root} is a standalone C header library containing a single function {\tt calc\_root\_mass} that calculates and returns the root mass at the next time step (a double precision floating point number) from input of the relative root growth rate ({\tt r\_r}, double), the time step ({\tt dt}, double), and the root mass as the current time step ({\tt R\_t}, double).
+%
+The shoot source code in Listing \ref{lst:src:shoot} is a standalone Python module containing a single function {\tt calc\_shoot\_mass} that calculates and returns the shoot mass at the next time step from input of the relative shoot growth rate ({\tt r\_s}), the time step ({\tt dt}, double), the shoot mass at the current time step ({\tt S\_t}), the root mass at the current time step ({\tt R\_t}), and the root mass at the next time step ({\tt R\_tp1}).
+
+Both of the example models used here start out in the form of the function.
+If possible, model writers should try to pose or wrap their model as a function prior to starting the integration process.
+This format makes the integration process easier and allows for the model to be used in a larger number of integration patterns.
+
+%---------------------------------------------------------------------------------
+% MODEL YAML
+\subsection{Model YAMLs}\label{SS:model_yaml}
+%
+The first step in the integration process is to create model YAML files describing the models including the location of the source code and the input/output variables.
+This step only has to be completed once per model and the model YAML can then be reused to include the model in any integration.
+The model YAMLs for the example root and shoot models are shown in Listings \ref{lst:yml:root} and \ref{lst:yml:shoot} respectively.
+%
+\rootyml
+%
+\shootyml
+%
+Each model yaml must include, at minimum, a name that is unique within the set of models being integrated, the language that the model is written in, and the location of the model source code (this can be a list/sequence if the model is split between multiple files).
+If the path to the source files is not absolute, as in the case of Listings \ref{lst:yml:root} and \ref{lst:yml:shoot}, the path is interpreted as being relative to the directory containing the model YAML.
+In this case, the source code for each model is a wrapper (described in \S\ref{SS:wrapper}) that calls the model code such that no modification needs to be done to the original model code in order to integrate it.
+
+In additional to the basic required fields, each model should have an entry for each of its input/output variables in the appropriate {\tt "inputs"} or {\tt "outputs"} section.
+At a minimum, input/output entries should include a name that will be used to identify a communication channel connected to the model.
+It is also highly recommended that the units of each variable also be specified (if applicable) so that {\pkg} can handle any necessary conversions.
+
+There are other optional model and input/output keywords for more advanced cases (e.g.
+compilation flags or data types), but they are beyond the scope of this limited introduction.
+Additional examples and descriptions of these options can be found in the documentation \footnote{{\tt https://cropsinsilico.github.io/cis\_interface/yaml.html}}.
+
+It should also be noted that it is possible to pass multiple variables as a single input/output.
+However, unless the variables will \emph{always} be coupled in the model, it is advised that every variable be specified separately for clarity and to allow flexibility in the way other models can integrate with it.
+In addition, if brevity is a concern because a model has a large number of input variables, there are features currently under active development (see \S\ref{SS:datagg}) that will allow fields to be aggregated within connection YAMLs and simplify the send and receive calls in such cases.
+
+%---------------------------------------------------------------------------------
+% WRAPPER
+\subsection{Wrapper}\label{SS:wrapper}
+%
+Once the model YAMLs are complete, the next step is to write a wrapper for each model that will make the necessary calls to the {\pkg} API to set up channels and send/receive messages.
+Writing the wrapper is the most involved step in any integration as it requires the most thought about how one model should interact with others, but this wrapper is written in the same language as the model and so should be a comfortable process for the model author since it is a language they are already familiar with.
+There is work in progress to add features which will allow {\pkg} to perform this step automatically based on YAML options for simple cases such as single loops or if statements (see \S\ref{SS:control}), but these will not cover all potential cases and it is instructive to walk through the process and understand how the wrapper acts as an intermediary between {\pkg} and the actual model function.
+
+\subsubsection{Integration Pattern}
+%
+To write the wrapper, a integration pattern must first be selected.
+For the root/shoot model integration, the obvious pattern is to loop over time steps, outputting the evolution of the root and shoot masses over time.
+Although this is the pattern adopted for this example, you could imagine another pattern in which the loop is performed over the growth rates in a parameter sweep that would require a slightly different wrapper.
+
+Once an integration pattern is selected, the wrappers can be flushed out in pseudocode independently by assuming that all input is received from a file and all output is sent to a file (i.e.
+the models are independent).
+The pseudocode for the root and shoot model wrappers is shown in Algorithms \ref{alg:root} and \ref{alg:shoot} respectively.
+%
+\begin{algorithm}[H]
+    \caption{Root wrapper}
+    \label{alg:root}
+    \begin{algorithmic}[1] 
+    \State \textbf{Receive} $r_r, R_t$
+    \State \textbf{Send} $R_{t}$ \Comment{So that there is a complete record.}
+    \While{true}
+        \If {$dt$ available}
+        		\State \textbf{Receive} $dt$
+	\Else
+		\State \textbf{Break}
+	\EndIf
+    	\State $R_{t+1} \gets {\tt calc\_root\_mass}(r_r, dt, R_t)$
+	\State $R_t \gets R_{t+1}$
+	\State \textbf{Send} $R_{t}$
+    \EndWhile
+    \end{algorithmic}
+\end{algorithm}
+%
+\begin{algorithm}[H]
+    \caption{Shoot wrapper}
+    \label{alg:shoot}
+    \begin{algorithmic}[1] 
+    \State \textbf{Receive} $r_s, S_t, R_t$
+    \State \textbf{Send} $S_{t}$ \Comment{So that there is a complete record.}
+    \While{true}
+        \If {$dt$ available}
+        		\State \textbf{Receive} $dt$
+	\Else
+		\State \textbf{Break}
+	\EndIf
+	\State \textbf{Receive} $R_{t+1}$
+	\State $S_{t+1} \gets {\tt calc\_shoot\_mass}(r_s, dt, S_t, R_t, R_{t+1})$
+	\State $S_t \gets S_{t+1}$
+	\State $R_t \gets R_{t+1}$
+	\State \textbf{Send} $S_{t}$
+    \EndWhile
+    \end{algorithmic}
+\end{algorithm}
+%
+Both models follow a similar pattern.
+They first receive static variables that will not change over the course of the run and then send the initial mass to output so that the output record is a complete history of the mass.
+Then both entire a while loop that is only broken when there is not a new time step available.
+When a new time step is available, it is received along with the next root mass in the case of the shoot model.
+Next, both model wrappers make the call to the actual model function.
+Finally, the models advance the time step by setting the masses to those calculated for the next time step and output the calculated mass.
+
+\subsubsection{Wrapper Code}
+%
+With some translation into the appropriate language and the addition of calls to the appropriate {\pkg} model interface to establish input and output channels, the code for the root and shoot model wrappers can then be written as Listings \ref{lst:wrap:root} and \ref{lst:wrap:shoot} respectively.
+In the syntax of the target language, the wrapper should be executable.
+For Python/Matlab this might be a script, while for C/C++ this should be compilable as an executable with a {\tt main} function.
+%
+\rootwrap
+%
+\shootwrap
+%
+Prior to any calls, both model wrappers must first locate the necessary {\pkg} and model code that they will call on via directives.
+In the C root model wrapper, this takes the form of {\tt \#include} statements on lines 3-6.
+In the Python shoot model wrapper, this takes the form of {\tt import} statements.
+{\pkg} takes care of the paths at compile/runtime so that the appropriate API library can be located while it is expected that the model source code is in the same directory as the wrapper and automatically discovered.
+
+The first step in each model wrapper is to connect to the appropriate channels via the {\pkg} interface.
+This step occurs on lines 14-17 in the root model wrapper (Listing \ref{lst:wrap:root}) and lines 8-13 in the shoot model wrapper (Listing \ref{lst:wrap:shoot}).
+Regardless of the language, each has a similar call signature.
+Inputs require a single input that is a string specifying the name of the model input from the model YAML that the returned object should access.
+Outputs are similar in that their first argument is a string specifying the name of the model output from the mode YAML that the returned object should access.
+In addition, outputs can also be provided with a format string that tells {\pkg} interface what data type to expect how the output should be formatted if it is sent to a file.
+Additional information about how these strings are processed can be found in the ``C-Style Format Strings" section of the documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/c\_format\_strings.html}}.
+
+The next step in each model wrapper is to receive the static input variables (those that are not being looped over).
+This step occurs on lines 19-33 in the root model wrapper (Listing \ref{lst:wrap:root}) and lines 15-34 in the shoot model wrapper (Listing \ref{lst:wrap:shoot}).
+Next, each model wrapper sends the initial mass to the output so that the output will contain the entire mass history.
+This occurs on lines 58-63 in the root model wrapper and lines 36-39 in the shoot model wrapper.
+The majority of these lines are devoted to error handling and print statements.
+The syntax for sending/receiving messages differs slightly from language to language, but each will return a flag indicating if the send/receive was successful or not.
+If an input channel is still open, a receive call will block until the channel is closed or a message is received.
+If the input channel is closed (either because of an error or because it was closed by the source), the flag will indicate this and the received message should not be used.
+If an output channel is open, a send call will return immediately while a worker thread handles asynchronous completion of the send request.
+If an output channel is closed, a send call will return immediately with a flag indicating failure.
+Additional information about the interface calls for sending and receiving can be found in the ``Model Interface" section of the documentation\footnote{{\tt \todo{link}}}.
+
+Once the static variables have been received and the initial masses have been sent to output, both model wrappers enter their while loop.
+The flag returned by the receive call for the new time step is then used to decide if the loop should be broken (lines 46-52 in the root model wrapper and lines 45-51 in the shoot model wrapper).
+In the case of the shoot model wrapper, a failure to also receive the root mass for the next time step results in an error as it is required that there be a new root mass for each new time step (lines 53-59).
+Finally, both model wrappers make calls to the appropriate model functions, send the result to the output, and advance the masses to the next time step.
+
+While it is valid to include these calls in the model code itself, this is not advised for several reasons.
+First, reproducibility is important.
+If there is a model that has been used to produce past scientific results, it is important that the model is preserved (in the form it was in during production of the results) so that other scientists can reproduce that result if need be.
+The use of a wrapper allows the original model code to be preserved while still adding the necessary API calls for integration via {\pkg}.
+Second, although the original model code could be duplicated and then altered to preserve the original code, this duplication results in redundancy and complicates the process of incorporating changes that may occur in the original model code due to bug fixes or added features.
+Third, it is possible that a model may be used in several different integration patterns.
+If the model code is altered directly, it will be difficult to add new integration patterns and there will again be redundant copies of the model.
+
+%---------------------------------------------------------------------------------
+% CONNECTION YAMLS
+\subsection{Connection YAMLS}
+%
+Once the wrappers are complete, it is time to write the connection YAML. Connection YAMLs contain information about how inputs and outputs should be connected between models and files.
+
+\subsubsection{Isolated Models}
+%
+Before connecting the models to each other, it is useful to first write a connection YAML that connects all of a model's inputs and outputs to files so that it can be run (and tested) in isolation.
+In addition, much of the connection YAML for an isolated model can be reused in connection YAMLs for the model in an integration.
+The connection YAMLs for running the root and shoot models in isolation with file input/output are show in Listings \ref{lst:yml:root_files} and \ref{lst:yml:shoot_files} respectively.
+%
+\rootymlfiles
+%
+\shootymlfiles
+%
+Each model's connection YAML contains one item under {\tt connections} for each input/output listed in the model's model YAML.
+Here they have been grouped by input and output, but they can be in any order.
+Every model input/output must have a connection for an integration to be valid.
+If there were a model input/output without a connection, whether to a file or a model, an error would be raised at run time.
+
+Every connection has, at a minimum, an input and and output.
+Connection inputs can be model outputs, files, or a set of multiple model outputs or files.
+Connection outputs can be model inputs, files, or a set of multiple model inputs or files.
+An error will be raised if a connection just connects two files.
+In the connection YAMLs shown in Listings \ref{lst:yml:root_files} and \ref{lst:yml:shoot_files}, all of the connections include a file as the input or output.
+Therefore, these connection YAMLs are closed systems.
+For connections including files, there is also the option of specifying a {\tt filetype} that will determine how {\pkg} reads the file.
+All of the files used in this example have {\tt filetype} of {\tt table}, indicating that the files are ASCII tables with some number of columns and rows, assumed to be tab-delimited by default.
+By default, {\pkg} will read a table row by row, splitting each row up into its constituent column elements.
+The output connections in Listings \ref{lst:yml:root_files} and \ref{lst:yml:shoot_files} also include a {\tt field\_names} entry.
+This instructs {\pkg} to add the designated field names as a header line in the output table that is produced.
+
+There are many file formats that {\pkg} supports which have additional YAML options.
+Information about these file formats and their YAML options can be found in the ``Connection Options" subsection of the ``YAML Files" section of the documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/yaml.html\#connection-options}}.
+
+\subsubsection{Integrated Models}
+%
+The connection YAML for integrating the two models is shown in Listing \ref{lst:yml:int} and should look very similar to the connection YAMLs for running the models in isolation from Listings \ref{lst:yml:root_files} and \ref{lst:yml:shoot_files}.
+%
+\intyml
+%
+The connections on lines 3-11 in Listing \ref{lst:yml:int} are identical to lines 3-11 in Listing \ref{lst:yml:root_files}.
+Similarly, lines 18-29 and 32-35 in Listing \ref{lst:yml:int} are identical to lines 3-14 and 20-23 respectively in Listing \ref{lst:yml:shoot_files}.
+This duplication results because there is only one connection between the two models.
+All of the remaining model inputs and outputs are still connected to files.
+As a result, the only new connection in Listing \ref{lst:yml:int} is on lines 14-15.
+This connection between the two models occupies the {\tt next\_root\_mass} output from the root model and the {\tt next\_root\_mass} input to the shoot model, thereby eliminating the need for the output connection on lines 14-17 of Listing \ref{lst:yml:root_files} and the input connection on lines 15-17 of Listing \ref{lst:yml:shoot_files}.
+Although the model output and input being connected in this example have the same name ({\tt next\_root\_mass}), this is not a requirement.
+
+A careful observer would note that there is a discrepancy between the units of these two models.
+For one, both models are receiving their time steps from the same file, but, as designated in their model YAMLs, the root model expects its time steps to have units of hours and the shoot model expects its time steps to have units of days.
+In addition, the root model outputs masses in units of grams, while the shoot model expects the root masses to have units of kilograms.
+However, there is no need to do any conversions within the models themselves.
+Because the units are specified in the model YAMLs and the headers of the input tables, {\pkg} is able to perform the appropriate conversions during the asynchronous transfer of data from one model to the next using the {\tt unyt} package \citep{Goldbaum2018}.
+
+In addition to units transformations, {\pkg} is also capable of handling basic transformations between compatible data types (e.g.
+int to float or obj to ply) and inferring data types from sent/received messages in dynamically typed programming languages (i.e.
+Python \& Matlab).
+For more advanced transformations, users can instruct {\pkg} to use any arbitrary Python function to transform data by passing the module import path as a connection field.
+
+%---------------------------------------------------------------------------------
+% RUNNING THE INTEGRATION
+\subsection{Running the Integration}
+%
+Integrations are run by calling the command line utility {\pkgrun}. {\pkgrun} takes as input one or more paths to YAML files describing the integration. These files can be passed in any order.
+
+\subsubsection{Isolated Models}
+%
+To run the models in isolation, the user would pass {\pkgrun} the model YAML and the connection YAML specifying connections to files. For the root model this would be 
+%
+\begin{quote}
+{\tt {\pkgrun} root.yml root\_files.yml} 
+\end{quote}
+%
+and for the shoot model this would be 
+%
+\begin{quote}
+{\tt {\pkgrun} shoot.yml shoot\_files.yml}. 
+\end{quote}
+
+\subsubsection{Integrated Models}
+%
+To run the integrated models, the suer would pass {\pkgrun} both model YAMLs and the connection YAML specifying the complete integration. For the example, this would be 
+\begin{quote}
+{\tt {\pkgrun} root.yml shoot.yml root\_to\_shoot.yml}
+\end{quote}
+and the output to {\tt stdout} would include interleaved messages from both models as well as log messages from {\pkg}.
+
+%---------------------------------------------------------------------------------
+% ADVANCED TOPICS
+\subsection{Advanced Integration Topics}
+%
+The example integration used here was simplified in several respects which we would like to address for those who would use the package for more complex integrations.
+\begin{description}
+\item[Model Complexity] Both of the example models has a limited and homogenous set of scalar inputs.
+More realistic models are unlikely to have data limited to such cases.
+{\pkg} also has support for sending more complex data types like those discussed in \S\ref{SS:transformation}.
+In addition, work is underway to add generic support for any data type that can be expressed as a JSON object (see \S\ref{SS:json}).
+Examples using these different data types can be found in the ``Formatted I/O" section of the documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/formatted\_io.html}}.
+\item[Time Step Synchronization] The example models used here both took the same time step as input.
+This is an unlikely case in actual models, particularly for those that are capturing processes at different physical scales (e.g.
+cellular vs.
+field).
+Without strict constraints on the problem formulation and model relationships, it is not possible for {\pkg} to determine how two time steps should be reconciled.
+Therefore, the current version of {\pkg} requires the user to handle this part of the integration.
+There are plans for new features which allow automated creation and integration of models that can be described symbolically as Ordinary Differential Equations (ODEs, see \S\ref{SS:ODE}).
+For two coupled ODE models, {\pkg} may be able to determine the correct time step for synchronization, but, as an incorrect time step could drastically impact the results of an integration, it will still be recommend that users play an active roll in determining or evaluating the correct time step for a given integration.
+\item[Intermediate Output] In the integrated example presented, the output from the root model is passed to the shoot model without being output to a file.
+However, in real integrations, it is likely that users would want to know the value of this output as well.
+In order to output to both a model and a file, users must be able to direction connections to multiple channels (i.e.
+a model input and a file for output).
+While the current version of {\pkg} has limited support for ``forking" connections, full support for this feature is under way as part of the data aggregation feature (see \S\ref{SS:dataagg}) and will be a part of the version 1.0 release.
+\end{description}
 
@@
 %
-In order to evaluate {\cis}, performance tests were run on machines with Linux, Mac OS X, and Windows operating systems for different communication mechanisms, Python versions, and language combinations.
+In order to evaluate {\pkg}, performance tests were run on machines with Linux, Mac OS X, and Windows operating systems for different communication mechanisms, Python versions, and language combinations.
 During each run, $N_{msg}$ messages of size $S_{msg}$ were sent from a source model (in one language) to a destination model (in another language) which then sends the messages to an output file for verification.
-These are test models that do not perform any operations outside of passing the messages in order to accurately gauge the performance of the {\cis} machinery on its own.
+These are test models that do not perform any operations outside of passing the messages in order to accurately gauge the performance of the {\pkg} machinery on its own.
 Performance tests were run using the perf package \citep{Stinner2018}.
@@
 For Matlab, this is particularly time consuming ($>10$ s).
-Although this is only a one-time cost required at the start of an integration, {\cis} does offer ways to alleviate this if multiple runs are necessary by starting a Matlab engine prior to the first integration that can be used by subsequent runs.
+Although this is only a one-time cost required at the start of an integration, {\pkg} does offer ways to alleviate this if multiple runs are necessary by starting a Matlab engine prior to the first integration that can be used by subsequent runs.
 When Matlab is started in advance (the dashed-dotted orange line in Figure \ref{fig:language_matlab}), the overhead for Matlab models drops to 1.25\,s, less than Python (2.86\,s) but slightly more than C (1.11\,s).
@@
 The time required for compilation will depend on the compiler used and the complexity of the model code.
-For command line compiled models, {\cis} forces the model to be recompiled every time to ensure any changes to the source code are propagated.
+For command line compiled models, {\pkg} forces the model to be recompiled every time to ensure any changes to the source code are propagated.
 However, for models that use Make and CMake, overhead due to compile time can be reduced by using existing builds.
@@
 The same can be said of comparing the Mac OS X and Windows tests given that operating systems are often optimized for their intended hardware.
-However, the tests do show that {\cis} performed consistently on each of the supported operating systems.
+However, the tests do show that {\pkg} performed consistently on each of the supported operating systems.
+
+%---------------------------------------------------------------------------------
+% SPEEDUP
+\subsection{Speedup}\label{SS:speedup}
+%
+\todo{Write this!}
 
@@
 %
-{\cis} is an open-source Python package for connecting computational models across programming languages and scales to form integration networks.
+{\pkg} is an open-source Python package for connecting computational models across programming languages and scales to form integration networks.
 Models are run in parallel with asynchronous message passing handled under-the-hood via threading and one of three communication mechanisms.
-{\cis} works on Linux, Mac OS X, and Windows operating systems with Python 2.7, 3.4, 3.5, 3.6 and 3.7.
-{\cis} currently supports running models written in Python, C, C++, and Matlab with additional support for compiling models C/C++ using Make \citep{Stallman2004} or CMake \citep{Martin2006} and running models written in the LPy \citep{Boudon2012} DSL.
-
-{\cis} has already been used to integrate several plant models within the Crops in Silico organization with ongoing work to add additional models to the integration network.
+{\pkg} works on Linux, Mac OS X, and Windows operating systems with Python 2.7, 3.4, 3.5, 3.6 and 3.7.
+{\pkg} currently supports running models written in Python, C, C++, and Matlab with additional support for compiling models C/C++ using Make \citep{Stallman2004} or CMake \citep{Martin2006} and running models written in the LPy \citep{Boudon2012} DSL.
+
+%is unique as a framework in several respects:
+%\begin{itemize}
+%	\item Asynchronous communication via 3 different communication mechanisms. 
+%	\item Create and coordination of asynchrnous communication networks that 
+%\end{itemize}
+
+Integration via {\pkg} requires minimal modification of the existing model source code yet greatly increases the model's potential by allowing it to be reused in combination with other models.
+{\pkg} has already been used to integrate several plant models within the Crops in Silico organization \citep{Marshall-Colon2017} with ongoing work to add additional models to the integration network.
 Figure \ref{fig:network} shows the progress so far.
@@
 Each square represents a model and the color of the square indicates the language in which the model is written.
-Arrows between models represent connections being made between the models using {\cis}.
+Arrows between models represent connections being made between the models using {\pkg}.
 Solid lines are connections that have already been made, dashed lines are in-progress connections, and dotted lines are planned connections.}
@@
 When complete, the integration network will include 9 models of molecular, genetic, cellular, organ, and plant level processes in 4 different programming languages, produced by $>10$ researchers at 4 different institutions.
+
 Although not yet complete, partial integrations from this network are already yielding new insights.
-For example, by integrating a metabolic flux model  with gene expression data from soybean (\emph{Glycine max}) plants grown under ambient (380\,$\mu$mol/mol) and elevated (550\,$\mu$mol/mol) carbon dioxide (CO2),
+For example, by integrating a metabolic flux model  with gene expression data from soybean (\emph{Glycine max}) plants grown under ambient (380\,$\mu$mol/mol) and elevated (550\,$\mu$mol/mol) carbon dioxide (CO$_2$),
 %
 \ifieee
-	\citep{integration_prep}. 
+	\citep{integration_prep} 
 \else
-	\citet{integration_prep}. 
+	\citet{integration_prep} 
 \fi
@@
 
+In addition to allowing scientists to reuse existing models in new integrations, {\pkg} has the potential to significantly increase a model's accessibility.
+{\pkg} could be used a method of running models without knowledge of the model's implementation.
+At publication, model code could be released along with a wrapper and YAML specification file for use with {\pkg}.
+These additional materials would be particularly powerful when considered alongside the interactive features being developed for use with {\pkg} (see \S\ref{SS:gui} and \S\ref{SS:function}).
+
 %---------------------------------------------------------------------------------
@@
 %
-{\cis} is being actively developed to expand the number of models 
+{\pkg} is being actively developed to expand the number of models 
 that can be used in integration networks and the complexity of integration 
-networks that can be executed. Several improvements are already in progress/planned for {\cis}.
+networks that can be executed. Several improvements are already in progress/planned for {\pkg}.
 
@@
 that models have been written in. Based on these results, we plan to expand 
-{\cis} to support models written in R, Fortran, Java, and the SBML \citep{Hucka2003} DSL.
+{\pkg} to support models written in R, Fortran, Java, and the SBML \citep{Hucka2003} DSL.
 
@@
 tools in the current computing ecosystem that could be used for running 
-complex integration networks. To this end, we plan to expand {\cis} 
+complex integration networks. To this end, we plan to expand {\pkg} 
 support for running integration networks on distributed compute resources. 
-{\cis} already uses communication tools that can be adapted for use 
+{\pkg} already uses communication tools that can be adapted for use 
 in a distributed pattern. We will leverage tools like the libsubmit package from the 
@@
 and received once versus which variables change and should be looped over. 
-We plan on adding options to {\cis} for dynamically generating model code 
+We plan on adding options to {\pkg} for dynamically generating model code 
 based a user provided function call and list of static/variable input channels. 
 This will allow better reusability of model code such as during parameter studies.
+Control flow will be part of the {\pkg} 1.0 release.
+
+% ODE
+\subsubsection{Symbolic ODE Models}\label{SS:ODE}
+%
+Many biological models can be posed as a series of coupled ordinary differential equations \citep[ODEs, ][]{WANG2015}.
+The equations describing such models can be easily formulated into symbolic representations \citep[e.g.
+SymPy][]{Meurer2017} with well defined inputs and outputs that can be solved by existing tools designed specifically for that task \citep[e.g.
+scipy.integrate.ode][]{Jones2001, Hairer1993}.
+Using these tools, we will add features to {\pkg} that allow ODE models to be constructed directly from symbolic representation provided in YAML specification files.
 
@@
 %
-Much of the original version of {\cis} centered around the use of 
+Much of the original version of {\pkg} centered around the use of 
 tabular input data. While tabular data is used heavily by plant models, it 
@@
 input by another model. However, this is unlikely to be the case if two models 
-are developed independently. Future improvements to {\cis} will allow 
+are developed independently. Future improvements to {\pkg} will allow 
 tabular input to models to be composed by aggregating output data from more than 
-one model and/or file.
+one model and/or file. Data aggregation will be part of the {\pkg} 1.0 release.
 
@@
 %
-While {\cis} supports serialization of several data types/formats, we 
+While {\pkg} supports serialization of several data types/formats, we 
 would like to make serialization as flexible as possible. To this end, 
-{\cis} will be adapted to read JSON schema \citep{jsonschema} for user defined types that 
+{\pkg} will be adapted to read JSON schema \citep{jsonschema} for user defined types that 
 can be used to automatically create the appropriate data structures alongside
-methods for parsing and serializing them.
+methods for parsing and serializing them. JSON serialization will be part of the {\pkg} 1.0 release.
 
@@
 %
-In an effort to make {\cis} as user friendly as possible, we have also begun work on a graphical user interface (GUI) for entering model information, composing integration networks from an existing palette of models, and displaying basic output from an integration run.
+In an effort to make {\pkg} as user friendly as possible, we have also begun work on a graphical user interface (GUI) for entering model information, composing integration networks from an existing palette of models, and displaying basic output from an integration run.
 The current prototype of the GUI can only handle model ingestion and network composition; model execution must be done locally.
 However, our ultimate goal with the GUI is to provide a service where users can register their model, compose and run integration networks on dedicated cloud compute resources, and view output in real time all from the web browser.
+This GUI for model integration and execution will also greatly increase the accessibility of models as standalone scientific products.
+Once complete, users will be able to use the GUI to graphically run models and adjust parameters without any knowledge of the underlying algorithms, programming language, and/or data formats.
+Such a workflow could be a powerful tool for both the peer review process and collaboration.
 %\todo{Provide \href{https://hackathon.cis.ndslabs.org/}{link?}}
+
+% NOTEBOOK
+\subsubsection{Models as Functions}\label{SS:function}
+%
+In addition to a dedicated GUI, tools are also being developed for using {\pkg} to call integrated models as Python functions, allowing users to interactively explore and visualize a model.
+This feature is particularly powerful when combined with Jupyter notebooks \citep{Kluyver2016}, which allow code to be presented alongside text and/or images describing what the code does.
+Model developers could publish a Jupyter notebook exposing their model as a function via {\pkg} that can be used by those unfamiliar with modeling or the model's particular programming language to explore the model's behavior.
+This workflow would be particularly well suited to examining models during the peer review process and allowing on-the-fly parameter adjustment/estimation by experimentalists.
+Widgets can make these notebooks even more user friendly, allowing code to be replaced with graphical interfaces (sliders, toggles, etc.) for modifying model parameters and allowing even those unfamiliar with Python to explore the model.
+In addition, tools like Binder, BinderHub, and mybinder.org \citep{Jupyter2018} make it possible for those interested in a model to launch such notebooks stored in public repositories from their browser in custom environments deployed on cloud resources without the need to install {\pkg} or any of the model code on their machine.
+They can then interact with the model using either widgets or the full set of Python libraries.
+Model execution as Python functions will be part of the {\pkg} 1.0 release.
 
@@
 \section*{Code}\label{S:code}
-The {\cis} package is available publicly on \href{https://github.com/cropsinsilico/cis_interface}{Github} and can be 
+The {\pkg} package is available publicly on \href{https://github.com/cropsinsilico/cis_interface}{Github} and can be 
 installed using {\tt pip} or {\tt conda}.
-To ensure code health, {\cis} boasts 100\% code coverage with automated testing on Linux, Mac OS X, and Windows for Python version 2.7, 3.4, 3.5, 3.6, and 3.7 via continuous integration with TravisCI \citep{travisci} and Appveyor \citep{appveyor}.
-Full documentation for {\cis} can be found \href{https://cropsinsilico.github.io/cis_interface/}{here} including step-by-step directions from the tutorial conducted during the \href{https://cropsinsilico.github.io/cis_interface/hackathon2018/index.html}{2018 Crops in Silico Hackathon}.
+To ensure code health, {\pkg} boasts 100\% code coverage with automated testing on Linux, Mac OS X, and Windows for Python version 2.7, 3.4, 3.5, 3.6, and 3.7 via continuous integration with TravisCI \citep{travisci} and Appveyor \citep{appveyor}.
+Full documentation for {\pkg} can be found \href{https://cropsinsilico.github.io/cis_interface/}{here} including step-by-step directions from the tutorial conducted during the \href{https://cropsinsilico.github.io/cis_interface/hackathon2018/index.html}{2018 Crops in Silico Hackathon}.
 
@@
 %
-The authors would like to thank Mike Lambert and Craig Willis for their ongoing work on a graphical user interface for visually composing and running integration networks for {\cis} as well as David Raila for his work on the original integration code that {\cis} was based on.
+The authors would like to thank Mike Lambert and Craig Willis for their ongoing work on a graphical user interface for visually composing and running integration networks for {\pkg} as well as David Raila for his work on the original integration code that {\pkg} was based on.
 
@@
 % FIGURES
-\ifplacefig
-\else
-% put figures here if figures go after
-\fi
+%\ifplacefig
+%\else
+%% put figures here if figures go after
+%\fi
+
+%%---------------------------------------------------------------------------------
+%%---------------------------------------------------------------------------------
+%% APPENDICES
+%\appendix
+%
+%%---------------------------------------------------------------------------------
+%% SOURCE CODE
+%\section{Example Source Code}\label{A:src}
+%
+%\subsection{Root Model}\label{A:src:root}
+%\rootsrc
+%
+%\subsection{Shoot Model}\label{A:src:shoot}
+%\shootsrc
+%
+%%---------------------------------------------------------------------------------
+%% YAML FILES
+%\section{Example YAML Specification Files}\label{A:yml}
+%
+%\subsection{Root Model}\label{A:yml:root}
+%\rootyml
+%
+%\subsection{Shoot Model}\label{A:yml:shoot}
+%\shootyml
+%
+%\subsection{Integrated Models}\label{A:yml:integration}
+%\intyml
+%
+%%---------------------------------------------------------------------------------
+%% WRAPPER CODE
+%\section{Wrapper Source Code}\label{A:wrap}
+%
+%\subsection{Root Model}\label{A:wrap:root}
+%\rootwrap
+%
+%\subsection{Shoot Model}\label{A:wrap:shoot}
+%\shootwrap
 
@@
 % COMMENTS
-%\newpage
-%\todo{To do:
-%\begin{itemize}
-%\item Improve conclusion
-%\item Non-uniform test data set?
-%\item Performance of runs on cosmology dataset?
-%\item Move figures to the end for submission
-%\item iPython notebook for running tests
-%\end{itemize}
-% }
+\newpage
+\todo{To do:
+\begin{itemize}
+\item Replace links with those referencing new package name
+\item Forking example
+\item Transformation example
+\item Add JSON to documentation
+\item Add forking to documentation
+\item Add transformation to documentation
+\item Add schema to documentation
+\item Regenerate documentation
+\end{itemize}
+}
 

### 2019-01-01 cis_interface/examples/__init__.py

@@ -23,7 +23,7 @@ ex_dict = {'gs_lesson1': ('python', 'matlab', 'c', 'cpp'),
            'maxMsg': ('python', 'matlab', 'c', 'cpp', 'all', 'all_nomatlab'),
            'timed_pipe': ('python', 'matlab', 'c', 'cpp'),
            'fakeplant': ('python', 'matlab', 'c', 'cpp', 'all', 'all_nomatlab'),
-           'paper': ('python', 'c', 'all', 'all_nomatlab')}
+           'root_to_shoot': ('python', 'c', 'all', 'all_nomatlab')}
 ext_map = {'python': '.py',
            'matlab': '.m',
            'c': '.c',
@@ -83,9 +83,9 @@ for k, lang in ex_dict.items():
             yml_names = ['server_python.yml',
                          'client_%s.yml' % ilang]
             src_names = ['server.py', 'client%s' % ext_map[ilang]]
-        elif k == 'paper':
+        elif k == 'root_to_shoot':
             if ilang.startswith('all'):
-                yml_names = ['root.yml', 'shoot.yml', 'paper.yml']
+                yml_names = ['root.yml', 'shoot.yml', 'root_to_shoot.yml']
                 src_names = ['root.c', 'shoot.py']
             elif ilang == 'python':
                 yml_names = ['shoot.yml', 'shoot_files.yml']

### 2019-01-01 cis_interface/examples/paper/src/root.c

@@ -1,49 +1,0 @@
-#include <stdio.h>
-#include <stdlib.h>
-#include "CisInterface.h"
-
-
-double calc_root_mass(double r_r, double R_t) {
-  return R_t + (R_t * r_r);
-}
-
-
-int main(int argc, char *argv[]) {
-  int i, ret, nstep;
-  double r_r, R_t, R_tp1;
-
-  cisInput_t RootInput = cisInput("root_input");
-  cisOutput_t NextRootMass = cisOutputFmt("next_root_mass", "%lf\n");
-
-  // Receive number of timesteps, root growth rate, and initial root mass
-  ret = cisRecv(RootInput, &nstep, &r_r, &R_t);
-  if (ret < 0) {
-    printf("root: Error receiving root input.\n");
-    return -1;
-  }
-  printf("received nsteps = %d\n", nstep);
-
-  // Send initial root mass
-  ret = cisSend(NextRootMass, R_t);
-  if (ret < 0) {
-    printf("root: Error sending initial root mass.\n");
-    return -1;
-  }
-
-  // Loop over timesteps, outputing root masses
-  for (i = 0; i < nstep; i++) {
-
-    R_tp1 = calc_root_mass(r_r, R_t);
-    ret = cisSend(NextRootMass, R_tp1);
-    if (ret < 0) {
-      printf("root: Error sending root mass for timestep %d.\n", i+1);
-      return -1;
-    }
-
-    R_t = R_tp1;
-
-  }
-
-  return 0;
-
-}

### 2019-01-01 cis_interface/examples/paper/src/shoot.py

@@ -1,65 +1,0 @@
-import sys
-from cis_interface.interface import CisInput, CisOutput
-
-
-def calc_shoot_mass(r_s, S_t, R_t, R_tp1):
-    r"""Calculate the shoot mass.
-
-    Args:
-        r_s (float): Relative shoot growth rate.
-        S_t (float): Previous shoot mass.
-        R_t (float): Previous root mass.
-        R_tp1 (float): Root mass at the next timestep.
-
-    Returns:
-        float: Shoot mass at the next timestep.
-
-    """
-    return (S_t * r_s) + S_t - (R_tp1 - R_t)
-
-
-if __name__ == '__main__':
-
-    ShootInput = CisInput('shoot_input')
-    NextRootMass = CisInput('next_root_mass')
-    NextShootMass = CisOutput('next_shoot_mass', '%lf\n')
-
-    # Receive number of timesteps, shoot growth rate, and initial shoot mass
-    flag, input = ShootInput.recv()
-    if not flag:
-        print('shoot: Error receiving shoot growth rate.')
-        sys.exit(-1)
-    nstep, r_s, S_t = input
-
-    # Send initial shoot mass
-    flag = NextShootMass.send(S_t)
-    if not flag:
-        print('shoot: Error sending initial shoot mass.')
-        sys.exit(-1)
-
-    # Receive inital root mass
-    flag, input = NextRootMass.recv()
-    if not flag:
-        print('shoot: Error receiving initial root mass.')
-        sys.exit(-1)
-    R_t = input[0]
-
-    # Loop over timesteps, outputing shoot masses
-    for i in range(nstep):
-        flag, input = NextRootMass.recv()
-        if not flag:
-            print('shoot: Error receiving root mass for timestep %d.' % (i + 1))
-            sys.exit(-1)
-        R_tp1 = input[0]
-
-        S_tp1 = calc_shoot_mass(r_s, S_t, R_t, R_tp1)
-
-        flag = NextShootMass.send(S_tp1)
-        if not flag:
-            print('shoot: Error sending shoot amss for timestep %d.' % (i + 1))
-            sys.exit(-1)
-
-        S_t = S_tp1
-        R_t = R_tp1
-
-    sys.exit(0)

### 2019-01-01 cis_interface/examples/root_to_shoot/root_to_shoot.sh

@@ -12,6 +12,6 @@ case $1 in
 	;;
     * )
 	echo "Running Integration"
-	cisrun root.yml shoot.yml paper.yml
+	cisrun root.yml shoot.yml root_to_shoot.yml
 	;;
 esac

### 2019-01-01 cis_interface/examples/root_to_shoot/src/root.h

@@ -1,0 +1,3 @@
+double calc_root_mass(double r_r, double dt, double R_t) {
+  return R_t + (R_t * r_r * dt);
+};

### 2019-01-01 cis_interface/examples/root_to_shoot/src/shoot.py

@@ -1,0 +1,15 @@
+def calc_shoot_mass(r_s, dt, S_t, R_t, R_tp1):
+    r"""Calculate the shoot mass.
+
+    Args:
+        r_s (float): Relative shoot growth rate.
+        dt (float): The time step.
+        S_t (float): Previous shoot mass.
+        R_t (float): Previous root mass.
+        R_tp1 (float): Root mass at the next timestep.
+
+    Returns:
+        float: Shoot mass at the next timestep.
+
+    """
+    return (S_t * r_s * dt) + S_t - (R_tp1 - R_t)

### 2019-01-01 cis_interface/examples/root_to_shoot/src/shoot_wrapper.py

@@ -1,0 +1,73 @@
+# Import Python interface
+from cis_interface.interface import CisInput, CisOutput
+# Import Python module containing model calculation
+from shoot import calc_shoot_mass
+
+
+# Create input/output channels
+ShootGrowthRate = CisInput('shoot_growth_rate')
+InitShootMass = CisInput('init_shoot_mass')
+InitRootMass = CisInput('init_root_mass_shoot')
+TimeStep = CisInput('shoot_time_step')
+NextRootMass = CisInput('next_root_mass')
+NextShootMass = CisOutput('next_shoot_mass', '%lf\n')
+
+# Receive shoot growth rate
+flag, input = ShootGrowthRate.recv()
+if not flag:
+    raise RuntimeError('shoot: Error receiving shoot growth rate.')
+r_s = input[0]
+print('shoot: Received shoot growth rate: %f' % r_s)
+
+# Receive initial shoot mass
+flag, input = InitShootMass.recv()
+if not flag:
+    raise RuntimeError('shoot: Error receiving initial shootmass.')
+S_t = input[0]
+print('shoot: Received initial shoot mass: %f' % S_t)
+
+# Receive inital root mass
+flag, input = InitRootMass.recv()
+if not flag:
+    raise RuntimeError('shoot: Error receiving initial root mass.')
+R_t = input[0]
+print('shoot: Received initial root mass: %f' % R_t)
+
+# Send initial shoot mass
+flag = NextShootMass.send(S_t)
+if not flag:
+    raise RuntimeError('shoot: Error sending initial shoot mass.')
+
+# Keep advancing until there arn't any new input times
+i = 0
+while True:
+    
+    # Receive the time step
+    flag, input = TimeStep.recv()
+    if not flag:
+        print('shoot: No more time steps.')
+        break
+    dt = input[0]
+    print('shoot: Received next time step: %f' % dt)
+
+    # Receive the next root mass
+    flag, input = NextRootMass.recv()
+    if not flag:
+        # This raises an error because there must be a root mass for each time step
+        raise RuntimeError('shoot: Error receiving root mass for timestep %d.' % (i + 1))
+    R_tp1 = input[0]
+    print('shoot: Received next root mass: %f' % R_tp1)
+
+    # Calculate shoot mass
+    S_tp1 = calc_shoot_mass(r_s, dt, S_t, R_t, R_tp1)
[…]

### 2019-01-01 cis_interface/examples/tests/test_root_to_shoot.py

@@ -2,12 +2,12 @@ import os
 from cis_interface.examples.tests import TestExample
 
 
-class TestExamplePaper(TestExample):
-    r"""Test the Paper example."""
+class TestExampleRootToShoot(TestExample):
+    r"""Test the Root to Shoot example."""
 
     def __init__(self, *args, **kwargs):
-        super(TestExamplePaper, self).__init__(*args, **kwargs)
-        self._name = 'paper'
+        super(TestExampleRootToShoot, self).__init__(*args, **kwargs)
+        self._name = 'root_to_shoot'
 
     @property
     def output_dir(self):
@@ -34,4 +34,4 @@ class TestExamplePaper(TestExample):
         if self.output_dir is not None:
             if not os.path.isdir(self.output_dir):
                 os.mkdir(self.output_dir)
-        super(TestExamplePaper, self).run_example()
+        super(TestExampleRootToShoot, self).run_example()

### 2019-01-01 papers/lang2018/referee_comments.tex
Response to reviewer 2 and addition of example in response to reviewer 1

\documentclass[11pt]{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\title{Reviewer(s)' Comments to Author}
%\author{The Author}
%\date{}                                           % Activate to display a given date or no date

\newcommand{\pkg}{{\tt cis\_interface}{}}
\usepackage{enumitem,amssymb}
\newlist{todolist}{itemize}{2}
\setlist[todolist]{label=$\square$}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
\hspace{-2.5pt}}
\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}

\begin{document}
\maketitle
%\section{}
%\subsection{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reviewer: 1}

\subsection{Comments to the Author}

This is timely and a really nice piece of work. This Cis-interface has the potential to become a major tool to support the future development of the plant systems model integration. I have a few specific comments on this and would like the author to address. 
\begin{todolist}
\item First, it would be great to have a more clear description of the novelty of the pipeline developed. Is it mainly a compilation of existing tools, in that case, this needs to be clearly indicated. If not, then a clear statement in this might be great. 
\item[\done] Secondly, the work will benefit a lot from a more explicit description of a user case. For example, clearly write out the exact commands needed to finish one round of code integration. 
\item Thirdly, I would be keen to show some performance data of integration of different modules. No data are provided on this at this stage. 
\item Fourthly, the main purpose of this paper is to describe a platform which can help plant modelers of different research areas and using programming languages work together more effectively and easily, it would be great if author can explain the advantage of this new software from the perspective of users. 
\item Fifthly,  many of the software engineering methodologies described in the paper are widely used in the field of software engineering. It would be better to provide some more concrete description of how this particular tool help users.  
\item Finally, it would be helpful to provide a flow chart and pseudo-code of the core algorithms used to link different modules.
\end{todolist}

\subsection{Response from the Author}

We would like to thank the reviewer for their comments as we believe they helped make the paper much stronger and to better represent the work being described. We will address the reviewer's comments in the order they were raised:
\begin{enumerate}
\item ?
\item We whole heartedly agree with the reviewer that the manuscript greatly benefits from a concrete use case. As such, we have added a new section ``Worked Example" after the methods section that walks through, step-by-step, integration of the example models presented in the introduction. The code from the example is available on the paper branch of the {\pkg} GitHub repository and will be published as part of the documentation during the 1.0 release.
\item We believe that the reviewer is referring to performance data on integration of specific biological \emph{models}, but would appreciate some clarification if this is not the case. While we would like to show concrete example of the performance of some integrations done with {\pkg}, many of the integration are still in the pre-publication stage. As the code for the models is not yet public, these performance tests would not be informative to those without access. In addition, the performance increase to parallelism is highly dependent on how coupled the two models are. If two model are integrated as in the case of the added root/shoot integration example as they are, the shoot model will always be waiting for the root model to complete its calculation and there not be a speedup over scripted integration of the two models via an intermediary file. However, the shoot model had other calculations to perform before it needed the data from the root model, a speedup would be possible and completely dependent on the ratio of independent calculations to dependent calculations and the time required for model communication. The speedup calculation is even more complex when more than two models are involved. To reflect this complex relationship, we have added a subsection to the end results section entitled ``Speedup" that discusses this and present a equation for calculating the speedup for an integration of two models given the appropriate parameters.
\item ?
\item ?
\item ?
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reviewer: 2}

\subsection{Comments to the Author}
This is a very well written manuscript that is accessible to people from different backgrounds. Few minor comments:
\begin{todolist}
\item[\done] Background: Pg 4, L8-18: The first paragraph contains broad, bold statements with no citations as examples. To move away from a largely anecdotal paragraph, examples from the literature should be given. It would also help to cite the Crops in silico perspective article in Frontiers (https://www.frontiersin.org/articles/10.3389/fpls.2017.00786/full).
\item[\done] P4, L35-38: give examples here too. What solutions?
\item[\done] P6 line 17: a to "as"
\item[\done] P7 line 9: remove extra "tools"
\item[\done] In the conclusions, it would be helpful to have a few lines suggesting uses or applications of the Cis interface. For example, researchers could use cis{\_}interface as a user-friendly way to run model code either from the published literature or when reviewing manuscripts. Likewise, experimentalists can update parameter files based on in-house experimental data and then run the models, even if they have no prior expertise in modeling.
\end{todolist}

\subsection{Response from the Author}

We would like to thank the reviewer for their comments as we the manuscript is stronger as a result. We have responded to the reviewer's comments in the order they were posed.
\begin{enumerate}
\item We agree with the reviewer that the paragraph in question was vague and have added several references to examples from the literature and including the Crops in silico perspective article.
\item The sentence in question "Within both computational biology and the larger scientific computing community, groups have developed around different solutions to the problem of connecting models." was the last sentence in the introductory part of the background section and was meant to refer to the solutions described form the literature in the following subsections. We have clarified this by explicitly listing the solutions that are described in those sections.
\item We have added several sentences to the conclusions discussing the use cases for {\pkg} including as a tool for integrating models and as a tool for executing models without knowledge of the model's programming language. While we agree with the reviewer that {\pkg} could be used for peer review and collaboration by exposing models via a standard interface, we believe this application will be most powerful when used with the model integration GUI (which is separate form the {\pkg} package itself) or by calling models as functions from Jupyter notebooks (which is a feature that was added after submitting the manuscript and will be released in verison 1.0 of {\pkg}). As such, we have added a subsection for "Models as Functions" in the "Current/Future Improvements" section and discuss this application further there and in the "Graphical User Interface" subsection.
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reviewer: 3}

\subsection{Comments to the Author}
This manuscript reports on efforts to create a universal interface toolbox for integrating computational models across languages and scales. This topic is without doubt one of the key topics in further development and application of existing and to-be-developed plant models. I appreciate the authors? effort in developing an open source Python package cis{\_}interface, which can be installed on 3 different operating systems and support integration of models written in 4 different programming languages. I think the paper has great potential but I do have some comments:
\begin{todolist}
\item The author had a lengthy discussion on comparison of communication time cost for model integration with different communication mechanisms, programming languages, Python versions and operating systems, actually this is also the main result of current manuscript. However, without comparison with existed tools, I think this part should be greatly condensed. I believe readers will be more interested in how does cis{\_}interface work for combining different types of biological models into a complex network. These models may have different time step length (e.g. ODE vs constant step length) and different types and scales of input and output which need complex data conversion (e.g. total leaf nitrogen content to different protein levels). Moreover, how to perform integration of input for one model from different models? For example, canopy model only gives 3D mesh data with light level on each facet, however, metabolic model needs both light level and protein levels (calculated from another model) for each facet, how to correctly integrate the protein data and the light data?
\item I expect more detailed biological modeling examples to be provided, from which the readers are able to know how to use the cis{\_}interface and what?s the advantage of it. I also expect the author to suggest some general modeling principles from following which the newly developed model can easily be integrated with other models. 
\end{todolist}

\subsubsection{Minor comments}
\begin{todolist}
\item[\done] Page 7 line 9: remove redundant ?tools?
\end{todolist}

\subsection{Response from the Author}

\end{document}

### 2019-01-01 cis_interface/examples/paper/paper.sh

@@ -1,17 +1,0 @@
-#!/bin/bash
-
-# ----------------Your Commands------------------- #
-case $1 in
-    --root )
-	echo "Running Root Model"
-	cisrun root.yml root_files.yml
-	;;
-    --shoot )
-	echo "Running Shoot Model"
-	cisrun shoot.yml shoot_files.yml
-	;;
-    * )
-	echo "Running Integration"
-	cisrun root.yml shoot.yml paper.yml
-	;;
-esac

### 2019-01-01 cis_interface/examples/paper/src/shoot.py

@@ -1,65 +1,0 @@
-import sys
-from cis_interface.interface import CisInput, CisOutput
-
-
-def calc_shoot_mass(r_s, S_t, R_t, R_tp1):
-    r"""Calculate the shoot mass.
-
-    Args:
-        r_s (float): Relative shoot growth rate.
-        S_t (float): Previous shoot mass.
-        R_t (float): Previous root mass.
-        R_tp1 (float): Root mass at the next timestep.
-
-    Returns:
-        float: Shoot mass at the next timestep.
-
-    """
-    return (S_t * r_s) + S_t - (R_tp1 - R_t)
-
-
-if __name__ == '__main__':
-
-    ShootInput = CisInput('shoot_input')
-    NextRootMass = CisInput('next_root_mass')
-    NextShootMass = CisOutput('next_shoot_mass', '%lf\n')
-
-    # Receive number of timesteps, shoot growth rate, and initial shoot mass
-    flag, input = ShootInput.recv()
-    if not flag:
-        print('shoot: Error receiving shoot growth rate.')
-        sys.exit(-1)
-    nstep, r_s, S_t = input
-
-    # Send initial shoot mass
-    flag = NextShootMass.send(S_t);
-    if not flag:
-        print('shoot: Error sending initial shoot mass.')
-        sys.exit(-1)
-
-    # Receive inital root mass
-    flag, input = NextRootMass.recv()
-    if not flag:
-        print('shoot: Error receiving initial root mass.')
-        sys.exit(-1)
-    R_t = input[0]
-
-    # Loop over timesteps, outputing shoot masses
-    for i in range(nstep):
-        flag, input = NextRootMass.recv()
-        if not flag:
-            print('shoot: Error receiving root mass for timestep %d.' % (i + 1))
-            sys.exit(-1)
-        R_tp1 = input[0]
-
-        S_tp1 = calc_shoot_mass(r_s, S_t, R_t, R_tp1)
-
-        flag = NextShootMass.send(S_tp1)
-        if not flag:
-            print('shoot: Error sending shoot amss for timestep %d.' % (i + 1))
-            sys.exit(-1)
-
-        S_t = S_tp1
-        R_t = R_tp1
-
-    sys.exit(0)

### 2019-01-01 cis_interface/examples/root_to_shoot/root_to_shoot.sh

@@ -2,32 +2,16 @@
 
 # ----------------Your Commands------------------- #
 case $1 in
-    --root-iso )
-	echo "Running Root Model w/o Interface"
-	gcc -o root_isolated src/root_isolated.c
-	# time for i in {1..100}; 
-	# do
-	time ./root_isolated ./Input/root_growth_rate.txt ./Input/init_root_mass.txt ./Input/timesteps.txt ./Output/root_output.txt
-	# done
-	rm root_isolated
-	;;
-    --shoot-iso )
-	echo "Running Shoot Model w/o Interface"
-	# time for i in {1..100}; 
-	# do
-	time python src/shoot_isolated.py ./Input/shoot_growth_rate.txt ./Input/init_shoot_mass.txt ./Input/timesteps.txt ./Input/root_output.txt ./Output/shoot_output.txt
-	# done
-	;;
     --root )
 	echo "Running Root Model"
-	time cisrun root.yml root_files.yml
+	cisrun root.yml root_files.yml
 	;;
     --shoot )
 	echo "Running Shoot Model"
-	time cisrun shoot.yml shoot_files.yml
+	cisrun shoot.yml shoot_files.yml
 	;;
     * )
 	echo "Running Integration"
-	time cisrun root.yml shoot.yml root_to_shoot.yml
+	cisrun root.yml shoot.yml root_to_shoot.yml
 	;;
 esac

### 2019-01-01 cis_interface/examples/tests/test_paper.py

@@ -1,37 +1,0 @@
-import os
-from cis_interface.examples.tests import TestExample
-
-
-class TestExamplePaper(TestExample):
-    r"""Test the Paper example."""
-
-    def __init__(self, *args, **kwargs):
-        super(TestExamplePaper, self).__init__(*args, **kwargs)
-        self._name = 'paper'
-
-    @property
-    def output_dir(self):
-        r"""str: Output directory."""
-        if self.yamldir is None:  # pragma: debug
-            return None
-        return os.path.join(self.yamldir, 'Output')
-
-    @property
-    def output_files(self):
-        r"""list: Output files for the run."""
-        outdir = self.output_dir
-        filelist = ['root_output.txt',
-                    'shoot_output.txt']
-        out = [os.path.join(outdir, f) for f in filelist]
-        return out
-
-    def check_results(self):
-        r"""This should be overridden with checks for the result."""
-        pass
-
-    def run_example(self):
-        r"""This runs an example in the correct language."""
-        if self.output_dir is not None:
-            if not os.path.isdir(self.output_dir):
-                os.mkdir(self.output_dir)
-        super(TestExamplePaper, self).run_example()

### 2019-01-03 papers/lang2018/ms.tex
Add comment to todo list
Added section on speedup

@@
 
+To accomplish these objectives, {\pkg} is built upon several well-established open-source software packages along with tools developed explicitly for {\pkg}.
 {\pkg} currently support models written in Python, Matlab, C, and C++ with
-additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}. Support for additional languages is planned for future development (See \S\ref{SS:langsupport}).
+additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}.
+Support for additional languages is planned for future development (See \S\ref{SS:langsupport}).
+\todo{Statement on novelty}
 
@@
 In the example, this would be the shoot growth model which required almost twice the time to execute as the root model.
-If the models are dependent on one-another, the models will have to wait for messages to be passed and the speedup will not be as great.
+If the models are dependent on one-another, the models will have to wait for messages to be passed and the speedup will not be as great (see \S\ref{S:speedup} for additional discussion of parallelism and the speedups achievable using {\pkg}).
+\todo{Fix these numbers for most recent version of the code}.
 
@@
 {\pkg} header library. Alternatively, users can provide the location of 
-the model source code and let {\pkg} handle the compilation, including 
+the model source code and let one of several {\pkg} drivers handle the compilation, including 
 linking against the appropriate {\pkg} interface library. {\pkg} 
@@
 The master process will only complete once an error is encountered or all model processes have completed.
+
+\todo{algorithm}
 
@@
 However, regardless of the communication mechanisms, the user will always use the same interface in the language of their model (See \S\ref{SSS:interface}).
+
+{\pkg} includes its own implementation of asynchronous communication via each of the communication mechanisms that are supported.
+While there are existing tools for asynchronous communication using each of these tools, using a tool developed specifically for {\pkg} allows for a more uniform treatment of the different communication mechanisms, greater control over how threads are managed (e.g.
+killing a thread when messages cannot be interpreted), and a more seamless coordination with other {\pkg} processes (e.g.
+enforcing dependencies on other model and connection threads).
 
@@
 
+\todo{algorithm}
+
 %In addition to communication between two models, users can also specify that a 
@@
 
+\todo{algorithm}
+
 %---------------------------------------------------------------------------------
@@
 If two models use different units (and both are specified), {\pkg} will automatically perform the necessary conversions before passing data from one model to the next.
+
+\todo{algorithm?}
 
@@
 \begin{description}
-\item[Model Complexity] Both of the example models has a limited and homogenous set of scalar inputs.
+\item[Model Input/Output Complexity] Both of the example models has a limited and homogenous set of scalar inputs.
 More realistic models are unlikely to have data limited to such cases.
@@
 Examples using these different data types can be found in the ``Formatted I/O" section of the documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/formatted\_io.html}}.
+\item[Model Networks] It is possible to build up complex integration networks using {\pkg} one connection at a time.
+While not show here, integrations with more than two models are constructed in much the same way, one connection at a time.
+Tools are currently under development for composing integration networks visually from a palette of models (see \S\ref{SS:gui}) that will assist in this step.
+With or without such tools, it is recommended that users start with integrating models in isolation with input/output from/to files, then being testing each individual connection in isolation, and then slowly add the connections together to form a complete network.
 \item[Time Step Synchronization] The example models used here both took the same time step as input.
@@
 % SPEEDUP
-\subsection{Speedup}\label{SS:speedup}
-%
-\todo{Write this!}
+\subsection{Speedup from Parallelism}\label{SS:speedup}
+%
+While efforts have been made to limit the overhead introduced by {\pkg} in the startup and communication steps of the integration, {\pkg} cannot outperform a direct integration of two models if they are tightly coupled and must run sequentially.
+For example, consider an alternate version of the shoot model from \S\ref{S:example} where the shoot model itself calculates the current root mass, sends the root mass to the root model, and then waits for the response.
+While the root model is performing the calculation, the shoot model will be idle, waiting for the response and, while the shoot model is calculating the root mass or handling the response, the root model will be idle, waiting for the next input.
+As a result, such a system will essentially execute in serial, incurring the overhead from using {\pkg} to run the models and coordinate communication without the advantage of parallel execution.
+Because native language variable passing will always be faster, a direct integration of the two models by translating one into the language of the other will always be faster than the {\pkg} integration in such cases.
+However, if the two models are not as tightly coupled and have calculations that can occur simultaneously, {\pkg} can offer a significant speedup via parallelism.
+
+Speedup, in the case of parallelism, is defined as the ratio of the serial execution time to the parallel execution time.
+%
+\begin{equation}
+S = \frac{T_{serial}}{T_{parallel}}
+\end{equation}
+%
+For the example presented above, the speedup offered by {\pkg} through parallelism can be parameterized in terms of the durations of calculations for each model (designated model A \& B for brevity) performed in the different phases of the process (See Figure \ref{fig:empty_parallel}).
+%
+\ifinclfig
+ 	\begin{figure}[htbp]
+	\begin{center}
+	\includegraphics[width=\columnwidth,keepaspectratio]{./images/Parallelism-Empty.png}
+	\caption{Diagram of phases in parallel execution of models A (top) and B(bottom).
+Each model performs some setup actions in the initial phase, enters a loop that contains two communication (first from model A to model B, then from model B to model A), and then performs some teardown actions in the final phase.}
+	\label{fig:empty_parallel}
+	\end{center}
+	\end{figure}
+\fi
+%
+During the initial phase ($t_{init,A}$ \& $t_{init,B}$), both models perform setup tasks before immediately entering the loop.
+On entering the loop at Phase 1 ($t_{1,A}$ \& $t_{1,B}$), both models perform calculations in preparation for the first communication where Model A sends a message to Model B.
+In Phase 2 ($t_{2,A}$ \& $t_{2,B}$), Model B processes the message it received and Model A performs actions in preparation for receiving a response during the second communication.
+In Phase 3 ($t_{3,A}$ \& $t_{3,B}$), Model A analyzes the response and Model B performs any additional calculations needed before the end of the current loop.
+Following completion of Phase 3 tasks, the models will immediately either enter another loop or enter the final phase ($t_{final,A}$ \& $t_{final,B}$) to complete any necessary cleanup and teardown actions.
+%
+If the models in the toy example were run in serial (without {\pkg}), the total execution time would be 
+%
+\begin{equation}
+T_{serial} = t_{init,A} + t_{init,B} + N_{loop}\left[\sum_{i=1}^{3}\left(t_{i,A} + t_{i,B}\right)\right] + t_{final, A} + t_{final, B},
+\end{equation}
+%
+where $N_{loop}$ is the number of loops executed and $t_{x,A}$ and $t_{x,B}$ are the execution times of each phase of the process for models A and B respectively.
+%
+In order to calculate the parallel execution time, we must consider the effect of asynchronous communication.
+For every asynchronous call, the receiving model is limited by the sending model, but the sending model is not limited by the receiving model.
+For a single asynchronous communication between two parallel processes, the execution time will be
+%
+\begin{eqnarray}
+%t_{wait} & = & \max(0, t_{before,sender} - t_{before,receiver}) \\
+T_{async} & = & \max\left\{\left[t_{before,send} + t_{after,send}\right], \left[\max(t_{before,send}, t_{before,recv}) + t_{after,recv}\right]\right\},
+\end{eqnarray}
+%
+where $t_{before,send}$ and $t_{after,send}$ are the time the sending model spends in the phases before and after sending the message respectively and $t_{before,recv}$ and $t_{after,recv}$ are the same times, but for the receiving model.
+The parallel execution time for the example which includes a loop around two asynchronous communications is then a recursive relationship where the execution time for one loop depends upon the relative execution times of the two models for the previous loop.
+This relationship can be written as a recursive formula
+%
+\begin{equation}
+T_{parallel} = T_{overhead} + \max(T_{A,N_{loop}}, T_{B,N_{loop}}),
+\end{equation}
+\begin{equation}
+\begin{cases}
+T_{A, 0} = t_{init,A} + t_{final,A} \\
+T_{A, n} = \max\left[(T_{A, n-1} + t_{comm} + t_{1,A} + t_{2,A} - t_{final,A}), (T_{B, n} - t_{3,B} - t_{final,B})\right] + t_{3,A} + t_{final,A}\\
+T_{B, 0} = t_{init,B} + t_{final,B}\\
+T_{B, n} = \max\left[(T_{B, n-1} + t_{comm} + t_{1,B} - t_{final,B}), (T_{A, n-1} + t_{1,A} - t_{final,A})\right] + t_{2,B} + t_{3,B} + t_{final,B}\\
+%3x + 5y + z \\ 7x ? 2y + 4z \\ -6x + 3y + 2z 
+\end{cases}
+\end{equation}
+%
+where $T_{overhead}$ is the overhead incurred by using {\pkg} to run the models and setup the communication network, $t_{comm}$ is the time required to complete the exchanges via each {\pkg} asynchronous communication, $n$ is the number of loops, and $T_{A,n}$ and $T_{B,n}$ are the times required by each model to complete their individual task in parallel for the given number of loops.
+While $T_{overhead}$ and $t_{comm}$ can be determined from the previous results sections for a given set of languages, the other variables are entirely dependent on the speed of the code occupying each phase.
+This relationship can also be written more intuitively as an algorithm (See Algorithm \ref{alg:tpara})
+%
+% tA = tinitA
+% tB = tinitB
+% iloop = 0
+%
+% WHILE (iloop < nloop)
+%	tA += t1A
+%	tB += t1B
+%	# FIRST LOCK
+% 	IF tB <= tA: B waits for A at Lock 1 and they are synchronized
+%		tB = tA
+%	tA += t2A
+%	tB += t2B
+%	# SECOND LOCK
+%	IF tA <= tB: A waits for B at Lock2 and they are synchronized
+%		tA = tB
+%	tA += t3A
+%	tB += t3B
+%	iloop++
+% tA += tfinalA
+% tB += tfinalB
+% Tpara = max(tA, tB)
+%
+\begin{algorithm}[H]
+    \caption{Parallel Execution Time}
+    \label{alg:tpara}
+    \begin{algorithmic}[1] 
+    \State $T_{A} \gets t_{init,A}$
+    \State $T_{B} \gets t_{init,B}$
+    \State $iloop \gets 0$
+    \While{$iloop <= N_{loop}$}
+    	\State $T_{A} += t_{1,A}$
+	\State $T_{B} += t_{1,B}$
+	\State $T_{B} = \max(T_{A}, T_{B} + t_{comm})$ \Comment{Model B blocking on receive.}
+	\State $T_{A} += t_{2,A}$
+	\State $T_{B} += t_{2,B}$
+	\State $T_{A} = \max(T_{A} + t_{comm}, T_{B})$ \Comment{Model A blocking on receive.}
+	\State $T_{A} += t_{3,A}$
+	\State $T_{B} += t_{3,B}$
+	\State $iloop += 1$
+    \EndWhile
+    \State $T_{A} += t_{final,A}$
+    \State $T_{B} += t_{final,B}$
+    \State $T_{parallel} = \max(T_{A}, T_{B})$
+    \end{algorithmic}
+\end{algorithm}
+%
+If the models are completely dependent on one another such that neither has any calculations to perform while the other is working, the situation would be that shown in Figure \ref{fig:serial}. 
+%
+\ifinclfig
+ 	\begin{figure}[htbp]
+	\begin{center}
+	\includegraphics[width=\columnwidth,keepaspectratio]{./images/Parallelism-Serial.png}
+	\caption{Diagram of parallelism for a fully coupled set of models.
+The shaded arrows represent active calculation by models A (top) and B(bottom).
+Absence of an arrow indicates idle time where the model process is not performing any calculations.}
+	\label{fig:serial}
+	\end{center}
+	\end{figure}
+\fi
+%
+Assuming that $t_{init,B} < (t_{init,A} + t_{1,A})$, this integration would have a parallel execution time and speedup of
+%
+\begin{eqnarray}
+T_{parallel} & = & T_{overhead} + N_{loop}(t_{comm} + t_{1,A} + t_{2,B} + t_{3,A}) + \max\left[(t_{init,A} + t_{final,A}), (t_{init,B} + t_{final,B})\right] \\
+S & = & \frac{t_{init,A} + t_{init,B} + N_{loop}(t_{1,A} + t_{2,B} + t_{3,A}) + t_{final, A} + t_{final, B}}{T_{overhead} + N_{loop}(t_{comm} + t_{1,A} + t_{2,B} + t_{3,A}) + \max\left[(t_{init,A} + t_{final,A}), (t_{init,B} + t_{final,B})\right]}
+\end{eqnarray} 
+%
+If the initial and final phases require an insignificant amount of time in comparison with the time spent in the loop, this will approach a speedup of 1, indicating that there was no change in the execution time.
+However, due to the required overhead ($T_{overhead}$) and communication times ($t_{comm}$) incurred by a {\pkg} integration, it is more likely that the speedup will be below 1 in such a case unless the calculations have a long enough duration to render these costs inconsequential.
+For example, if we consider two Python models ($t_{comm} = 0.02$\,s/msg and $T_{overhead} = 2.86$\,s) where $t_{1,A}$, $t_{2,B}$, and $t_{3,A}$ are all 1\,s with initial and final phases that are 0.1\,s for both models, the speedup will be 0.92 for 10 loops and 0.99 for 100 loops.
+
+If the models are perfectly matched in that each spends exactly the same amount of time in each phase ($t_{init,A}=t_{init,B}$, $t_{1,A}=t_{1,B}$, etc.) as in Figure \ref{fig:parallel},
+%
+\ifinclfig
+ 	\begin{figure}[htbp]
+	\begin{center}
+	\includegraphics[width=\columnwidth,keepaspectratio]{./images/Parallelism-FullyParallel.png}
+	\caption{Diagram of parallelism for a set of perfectly matched models.}
+	\label{fig:parallel}
+	\end{center}
+	\end{figure}
+\fi
+%
+the parallel execution time and speedup of the integration would be
+%
+\begin{eqnarray}
+T_{parallel} & = & T_{overhead} + t_{init,A} + N_{loop}\left(t_{comm} + \sum_{i=1}^{3}t_{i,A}\right) + t_{final, A} \\
+S & = & \frac{2\left[t_{init,A} + N_{loop}\left(\sum_{i=1}^{3}t_{i,A}\right) + t_{final, A}\right]}{T_{overhead} + t_{init,A} + N_{loop}\left(t_{comm} + \sum_{i=1}^{3}t_{i,A}\right) + t_{final, A}}.
+\end{eqnarray} 
+%
+If the runtime is sufficiently long such that contributions from $T_{overhead}$ and $t_{comm}$ become insignificant, the speedup will approach 2, the maximum for two processes.
+For example, if we consider two Python models ($t_{comm} = 0.02$\,s/msg and $T_{overhead} = 2.86$\,s) where both models have Phases 1-3 of 1\,s in duration and initial and final phases of 0.1\,s in duration, the speedup will be 1.82 for 10 loops and 1.97 for 100 loops.
+
+However, the situations shows in Figures \ref{fig:serial} and \ref{fig:parallel} are rarely the case for actual models.
+Figure \ref{fig:partial} shows a more realistic case in which the models have some overlapping operations, but are not perfectly aligned.
+%
+\ifinclfig
+ 	\begin{figure}[htbp]
+	\begin{center}
+	\includegraphics[width=\columnwidth,keepaspectratio]{./images/Parallelism-PartiallyParallel.png}
+	\caption{Diagram of partial parallelism for a set of imperfectly matched models.}
+	\label{fig:partial}
+	\end{center}
+	\end{figure}
+\fi
+%
+Speedups in such cases would be between 1 and 2.
+For example given the parameters shown in Table \ref{tab:partial} for two Python models ($t_{comm} = 0.02$\,s/msg and $T_{overhead} = 2.86$\,s), the speedup will be 1.38 for 10 loops and 1.49 for 100 loops.
+%
+\begin{table}[h]
+\caption{Speedup calculation example parameters for a set of imperfectly matched models.}
+\begin{center}
+\begin{tabular}{|l|c|c|}
+\hline
+Phase 	& Duration in Model A (s) & Duration in Model B (s) \\\hline
+initial		& 0.1 	& 0.2 \\
+1		& 1.0		& 0.3 \\
+2		& 1.2		& 2.0 \\
+3		& 0.1		& 0.05 \\
+final		& 0.1		& 0.1 \\
+\hline
+\end{tabular}
+\end{center}
+\label{tab:partial}
+\end{table}%
+%
+
+For more complex integration involving more than 2 models or more than 2 communications per loop, speedup calculations will be more complex.
+In general, the maximum possible speedup will be greater for integrations of models that are more independent, but wrappers can be adjusted to deliberately perform tasks during periods where a model would be waiting for a message in order to get greater performance boosts.
 
@@
 
+%---------------------------------------------------------------------------------
+% USE CASES AND LIMITATION
+%
+\subsection{Use Cases \& Limitations}\label{SS:uses}
+%
 In addition to allowing scientists to reuse existing models in new integrations, {\pkg} has the potential to significantly increase a model's accessibility.
@@
 These additional materials would be particularly powerful when considered alongside the interactive features being developed for use with {\pkg} (see \S\ref{SS:gui} and \S\ref{SS:function}).
+%
+\todo{Add discussion of limitations}
 
@@
 \item Regenerate documentation
+\item Fix matlab autodocs for interface
 \end{itemize}

### 2019-01-03 papers/lang2018/referee_comments.tex
Added section on speedup

@@
 \item[\done] Secondly, the work will benefit a lot from a more explicit description of a user case. For example, clearly write out the exact commands needed to finish one round of code integration. 
-\item Thirdly, I would be keen to show some performance data of integration of different modules. No data are provided on this at this stage. 
+\item[\done] Thirdly, I would be keen to show some performance data of integration of different modules. No data are provided on this at this stage. 
 \item Fourthly, the main purpose of this paper is to describe a platform which can help plant modelers of different research areas and using programming languages work together more effectively and easily, it would be great if author can explain the advantage of this new software from the perspective of users.
@@
 \item We believe that the reviewer is referring to performance data on integration of specific biological \emph{models}, but would appreciate some clarification if this is not the case.
-While we would like to show concrete example of the performance of some integrations done with {\pkg}, many of the integration are still in the pre-publication stage.
+While we would like to show concrete examples of the performance of some integrations done with {\pkg} using actual models, most of the models that {\pkg} has been used with are still in the pre-publication stage.
 As the code for the models is not yet public, these performance tests would not be informative to those without access.
-In addition, the performance increase to parallelism is highly dependent on how coupled the two models are.
-If two model are integrated as in the case of the added root/shoot integration example as they are, the shoot model will always be waiting for the root model to complete its calculation and there not be a speedup over scripted integration of the two models via an intermediary file.
-However, the shoot model had other calculations to perform before it needed the data from the root model, a speedup would be possible and completely dependent on the ratio of independent calculations to dependent calculations and the time required for model communication.
+In addition, the performance increase due to parallelism is highly dependent on how coupled the models are and how many models are part of the integration.
+If two models are integrated that are tightly coupled, one model will always be waiting for the other to complete its calculation and there not be a speedup over scripted integration of the two models via an intermediary file.
+However, if one model had other calculations to perform before it needed the data from the other model, a speedup would be possible and completely dependent on the ratio of independent calculations to dependent calculations and the time required for model communication.
 The speedup calculation is even more complex when more than two models are involved.
-To reflect this complex relationship, we have added a subsection to the end results section entitled ``Speedup" that discusses this and present a equation for calculating the speedup for an integration of two models given the appropriate parameters.
+To reflect this complex relationship, we have added a subsection to the end of the results section entitled ``Speedup from Parallels" that discusses this and present an equation for calculating the speedup for an integration of two models given the appropriate parameters.
 \item ?

### 2019-01-03 papers/lang2018/speedup.py

@@ -1,0 +1,100 @@
+_lookup_overhead = {
+    'c': 0.55,
+    'c++': 0.63,
+    'cpp': 0.63,
+    'matlab': 10.85,
+    'python': 1.43}
+
+
+def calculate_speedup(src_lang, dst_lang, nloops,
+                      tinitA, t1A, t2A, t3A, tfinalA,
+                      tinitB, t1B, t2B, t3B, tfinalB,
+                      t_comm=None, T_overhead=None):
+    r"""Function to calculate the speedup resulting from parallelism and
+    asynchronous communication between two models.
+
+    Args:
+        src_lang (str): Source language.
+        dst_lang (str): Destination language.
+        nloops (int): Number of loops in integration.
+        tinitA (float): Duration of Initial Phase in Model A (before loop).
+        t1A (float): Duration of 1st Phase in Model A (before sending).
+        t2A (float): Duration of 2nd Phase in Model A (between sending and receiving).
+        t3A (float): Duration of 3rd Phase in Model A (after receiving).
+        tfinalA (float): Duration of Final Phase in Model A (after loop).
+        tinitB (float): Duration of Initial Phase in Model B (before loop).
+        t1B (float): Duration of 1st Phase in Model B (before receiving).
+        t2B (float): Duration of 2nd Phase in Model B (between receiving and sending).
+        t3B (float): Duration of 3rd Phase in Model B (after sending).
+        tfinalB (float): Duration of Final Phase in Model B (after loop).
+        t_comm (float, optional): Time required to move a message from an output
+            of one model to the input of another. Defaults to a value based on a
+            lookup table for communication between different langauges.
+        T_overhead (float, optional): Time required to setup an integration and
+            start both models on their own processes. Defaults to a value based on
+            a lookup table for starting different langauges.
+
+    Returns:
+        float: Speedup.
+
+    """
+    # Set defaults
+    if t_comm is None:
+        if src_lang.lower() == 'matlab':
+            if dst_lang.lower() == 'matlab':
+                t_comm = 0.53
+            else:
+                t_comm = 0.14
+        elif dst_lang.lower() == 'matlab':
[…]

### 2019-01-03 papers/lang2018/ms.tex
Added algorithms for async comms, connection drivers, model drivers, and master thread

@@
 %\usepackage[noend]{algpseudocode}
+\usepackage{caption}
+\algblockdefx[Try]{Try}{EndTry}%
+	{\textbf{try}}%
+	{\textbf{end try}}
+\algcblockdefx[TryExcept]{Try}{Except}{EndTry}%
+	{\textbf{except}}%
+	{\textbf{end try}}
+\algblockdefx[class]{Class}{EndClass}%
+	[2]{\textbf{class} {\sc #1}(#2)}%
+	{\textbf{end class}}
+\algblockdefx[classmethod]{ClassMethod}{EndClassMethod}%
+	[2]{\textbf{method} {\sc #1}(#2)}%
+	{\textbf{end method}}
+
+%\DeclareCaptionFormat{algor}{%
+%  \hrulefill\par\offinterlineskip\vskip1pt%
+%    \textbf{#1#2}#3\offinterlineskip\hrulefill}
+%\DeclareCaptionStyle{algori}{singlelinecheck=off,format=algor,labelsep=space}
+%\captionsetup[algorithm]{style=algori}
+
 \ifdraft
@@
 
-\todo{algorithm}
+The algorithms for a generic model driver and the master thread can be found in Algorithms \ref{alg:model} \& \ref{alg:master} respectively.
+% MODEL DRIVER
+\begin{algorithm}[H]
+    \caption{Generic model driver class.
+This algorithm does not represent the actual implementation of any one of the languages supported by {\pkg}.
+Each of the language-specific drivers will have slightly different methods for compilation (turning the source code into an executable), launching the model process, and monitoring the model's status via output and error codes.
+However, this algorithm does provided the general behavior pattern upon which every model driver is based.}
+    \label{alg:model}
+    \begin{algorithmic}[1]
+    \Class{ModelDriver}{$name, args, yml$}
+    	\State $errors \gets []$
+    	\State $executable \gets$ \Call{Compile}{args}
+	\State $process \gets$ \Call{ModelProcess}{} on a new process
+	\State $thread \gets$ \Call{ModelThread}{} on a new thread
+	\item[]
+	%
+	\ClassMethod{ModelProcess}{}
+		\State \Call{Executable}{}
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{ModelThread}{}
+		\While{\Call{IsAlive}{{\sc ModelProcess}}}
+			\If{\Call{HasOutput}{{\sc ModelProcess}}}
+				\State \textbf{Print} \Call{GetOutput}{{\sc ModelProcess}}
+			\Else
+				\State \textbf{Sleep}
+			\EndIf
+		\EndWhile
+		\If{\Call{HasErrors}{{\sc ModelProcess}}}
+			\State $errors \gets$ \Call{GetErrors}{{\sc ModelProcess}}
+		\EndIf
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{Terminate}{}
+		\State \Call{Kill}{{\sc ModelProcess}}
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{Stop}{}
+		\While{not $errors$}
+			\State \textbf{Sleep}
+		\EndWhile
+		\State \Call{Terminate}{}
+	\EndClassMethod
+	\item[]
+    \EndClass
+    \end{algorithmic}
+\end{algorithm}
+
+%
+% MASTER THREAD
+%\begin{center}
+\begin{algorithm}[H]
+	\caption{Basic steps executed on the master thread to create, start, and monitor model \& connection drivers.
+This pseudocode is not an exhaustive reproduction of the entire implementation, but represents the broad strokes needed to understand the purpose and behavior of the master thread.
+The algorithm for the {\sc ModelDriver} and {\sc ConnectionDriver} can be found in Algorithms \ref{alg:model} \& \ref{alg:conn} respectively.}
+	\label{alg:master}
+	\begin{algorithmic}[1]
+	\State $models$ $\gets$ \textbf{Parse} YAML files
+	\State $inputChannels \gets Map()$
+	\State $outputChannels \gets Map()$
+	% CREATE DRIVERS
+	\item[] \Comment Create Drivers
+	\ForAll{$m in models$}
+		\ForAll{$i in m.inputs$}
+			\If{not $i.isFile$}
+				\State $inputAddress \gets i.address$
+			\Else
+				\State $inputAddress \gets `generate'$
+			\EndIf
+			\State $i.driver \gets$ \Call{ConnectionDriver}{$inputAddress, `generate', i$}
+			\State $inputChannels[i.name] = i$
+		\EndFor
+	\EndFor
+	\ForAll{$m in models$}
+	\algstore{alg:master}
+	\end{algorithmic}
+\end{algorithm}
+% PAGE BREAK
+\begin{algorithm}[H]
+	\caption{Continuation of Algorithm \ref{alg:master}}
+	\begin{algorithmic}[1]
+	\algrestore{alg:master}
+		\ForAll{$o \in m.outputs$}
+			\If{$o.isFile$}
+				\State $outputAddress \gets o.address$
+			\Else
+				\If{$o.args \in inpurChannels$}
+					\State $outputAddress \gets inputChannels[o.args]$
+				\Else
+					\State \Call{Error}{}
+				\EndIf
+			\EndIf
+			\State $o.driver \gets$ \Call{ConnectionDriver}{$`generate', outputAddress, o$}
+			\State $outputChannels[o.name] = o$
+		\EndFor
+	\EndFor
+	\ForAll{$m in models$}
+		\State $m.driver \gets$ \Call{ModelDriver}{$m.name, m.args, m$}
+	\EndFor
+	% START DRIVERS
+	\item[] \Comment Start Drivers	
+	\ForAll{$m in models$}
+		\ForAll{$i in m.inputs$}
+			\State $i.driver.$\Call{Start}{}
+		\EndFor
+	\EndFor
+	\ForAll{$m in models$}
+		\ForAll{$o \in m.outputs$}
+			\State $o.driver.$\Call{Start}{}
+		\EndFor
+	\EndFor
+	\ForAll{$m in models$}
+		\State $m.driver.$\Call{Start}{}
+	\EndFor
+	% MONITOR DRIVERS
+	\item[] \Comment Monitor Drivers
+	\State $errorFlag \gets false$
+	\While{not $errorFlag$}
+		\ForAll{$m in models$}
+			\If{$m.driver.errors$}
+				\State $errorFlag \gets true$
+				\State \textbf{break}
+			\EndIf
+		\EndFor
+	\EndWhile
+	% STOP DRIVER
+	\item[] \Comment Stop Drivers
+	\If{$errorFlag$}
+		\State $allErrors \gets []$
+		\ForAll{$m in models$}
+			\State $m.$\Call{Terminate}{}
+			\State \textbf{Append} $allErrors$, $m.errors$
+		\EndFor
+		\State \Call{Error}{$allErrors$}
+	\Else
+		\ForAll{$m in models$}
+			\State $m.$\Call{Stop}{}
+		\EndFor
+	\EndIf
+	\end{algorithmic}
+\end{algorithm}
+%\end{center}
+%
+\todo{make sure this is split correctly across pages}
 
@@
 As a result, models in complex integration networks are not affected by the rate at which dependent model consume their output and the speedup offered by running the models in parallel is improved.
-Figure \ref{fig:async} describes the general flow of messages.
+Figure \ref{fig:async} describes the general flow of messages, Algorithms \ref{alg:async_send} \& \ref{alg:async_recv} describe the asynchronous output \& input channel procedures, and Algorithm \ref{alg:conn} describes the procedure followed by input \& output connection drivers.
 %
@@
 
-\todo{algorithm}
+% SENDER
+\begin{algorithm}[H]
+    \caption{Asynchronous communication class for sending messages to an output channel.
+The procedures {\sc Open, Close, Send, \& IsOpen} are procedures that are implemented by external communication packages and different slightly from one communication mechanism to the next.}
+    \label{alg:async_send}
+    \begin{algorithmic}[1]
+    % SENDER
+    \Class{AsyncSender}{$address$}
+    	\State $backlog \gets []$
+	\State $x \gets$ \Call{Open}{$address$}
+	\State \textbf{Run} \Call{BacklogThread}{} on a new thread
+	\item[]
+	%
+	\ClassMethod{BacklogThread}{}
+		\While{\Call{IsOpen}{$x$}}
+    			\If{$length(backlog) > 0$}
+				\State $flag \gets$ \Call{Send}{$x$, $backlog[0]$}
+				\If{$flag$}
+					\State \textbf{Remove} $backlog[0]$
+				\Else
+					\State \Call{Close}{$x$}
+				\EndIf
+			\Else
+				\State \textbf{Sleep}
+			\EndIf
+    		\EndWhile
+    	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{SendBacklog}{$msg$}
+		\If{\Call{IsOpen}{$x$}}
+			\State \textbf{Append} $backlog$, $msg$
+		\EndIf
+		\State \textbf{Return} \Call{IsOpen}{$x$}
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{Close}{}
+		\State \Call{Close}{$x$}
+		\State $backlog \gets []$
+	\EndClassMethod
+	\item[]
+    \EndClass
+    \end{algorithmic}
+\end{algorithm}
+
+% RECEIVER
+\begin{algorithm}[H]
+    \caption{Asynchronous communication class for receiving messages from an input channel.
+The procedures {\sc Open, Close, Recv, IsOpen, \& IsMessageWaiting} are procedures that are implemented by external communication packages and different slightly from one communication mechanism to the next.}
+    \label{alg:async_recv}
+    \begin{algorithmic}[1]
+    \item[]
+    \Class{AsyncReceiver}{$address$}
+    	\State $backlog \gets []$
+	\State $x \gets$ \Call{Open}{$address$}
+	\State $thread \gets$ \Call{BacklogThread}{} on a new thread
+	\item[]
+	%
+	\ClassMethod{BacklogThread}{}
+		\While{\Call{IsOpen}{$x$}}
+    			\If{\Call{IsMessageWaiting}{$x$}}
+				\State $flag, msg \gets$ \Call{Recv}{$x$}
+				\If{$flag$}
+					\State \textbf{Append} $backlog$, $msg$
+				\Else
+					\State \Call{Close}{$x$}
+				\EndIf
+			\Else
+				\State \textbf{Sleep}
+			\EndIf
+		\EndWhile
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{RecvBacklog}{$msg$}
+		\If{$length(backlog) > 0$}
+			\State $msg \gets backlog[0]$
+			\State \textbf{Remove} $backlog[0]$
+			\State \textbf{Return} true, $msg$
+		\Else
+			\State \textbf{Return} \Call{IsOpen}{$x$}, ''
+		\EndIf
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{Close}{}
+		\State \Call{Close}{$x$}
+		\State $backlog \gets []$
+	\EndClassMethod
+	\item[]
+    \EndClass
+    \end{algorithmic}
+\end{algorithm}
+%
+
+%
+\begin{algorithm}[H]
+    \caption{Connection Driver}
+    \label{alg:conn}
+    \begin{algorithmic}[1]
+    \Class{ConnectionDriver}{$inputAddress, outputAddress, yml$}
+    	\State $input \gets$ \Call{AsyncReceiver}{$inputAddress$}
+    	\State $output \gets$ \Call{AsyncSender}{$outputAddress$}
+	\State $thread \gets$ \Call{ConnectionThread}{} on a new thread
+	\item[]
+	%
+	\ClassMethod{ConnectionThread}{}
+		\While{true}
+			\State $flag, msg \gets$ $input.$\Call{RecvBacklog}{}
+			\If{not $flag$} \State \textbf{break} \EndIf
+			\If{$msg$}
+				\State \textbf{Transform} $msg$ based on units and expected data format
+				\State $flag \gets$ $output.$\Call{SendBacklog}{$msg$}
+				\If{not $flag$} \State \textbf{break} \EndIf
+			\Else
+				\State \textbf{Sleep}
+			\EndIf
+		\EndWhile
+		%
+		\State $input.$\Call{CloseAsync}{}
+		\State $output.$\Call{CloseAsync}{}
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{Terminate}{}
+		\State $input.$\Call{CloseAsync}{}
+		\State $output.$\Call{CloseAsync}{}
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{Stop}{}
+		\While{\Call{IsOpen}{$input.x$} and \Call{IsOpen}{$output.x$}}
+			\State \textbf{Sleep}
+		\EndWhile
+		\State \Call{Terminate}{}
+	\EndClassMethod
+	\item[]
+    \EndClass
+    \end{algorithmic}
+\end{algorithm}
+
+\todo{discuss algorithm}
 
@@
 
+%\begin{algorithm}[H]
+%    \caption{Large Messages}
+%    \label{alg:large}
+%    \begin{algorithmic}[1] 
+%    \end{algorithmic}
+%\end{algorithm}
+
 \todo{algorithm}
@@
 If two models use different units (and both are specified), {\pkg} will automatically perform the necessary conversions before passing data from one model to the next.
-
-\todo{algorithm?}
 

### 2019-01-04 papers/lang2018/ms.tex
Added advantages/limitations section

@@
 	{\textbf{end method}}
+\usepackage[toc,page]{appendix}
 
@@
 \begin{abstract}
-Thousands of computational models have been created within the plant biology community in the past two decades that have the potential to be combined into complex integration networks capable of capturing more complex biological processes than possible with isolated models.
+Thousands of computational models have been created within the plant biology community and broader scientific communities in the past two decades that have the potential to be combined into complex integration networks capable of capturing more complex biological processes than possible with isolated models.
 However, the technological barriers introduced by differences in language and data formats has slowed this progress.
@@
 {\pkg} is designed to be user-friendly and can be accessed at {\tt https://github.com/cropsinsilico/cis\_interface}.
+Although originally developed for plant models, {\pkg} can be used to connect computational models from any domain.
 \end{abstract}
@@
 % DOMAIN SPECIFIC LANGUAGE
-\subsection{Domain Specific Language (DSL)}\label{SS:dsl}
+\subsection{Domain Specific Languages (DSLs)}\label{SS:dsl}
 %
@@
 However, they are stricter in several respects: 1) the model must be written in (or exposable to), the language of the framework and 2) the model must be written in a way that is aware of the framework and/or the format of models with which it will interact.
-Like DSLs, frameworks also provided limited support for models written without the framework in mind.
+Like DSLs, frameworks also often provided limited support for models written without the framework in mind.
 
@@
 
-{\pkg} was designed to address the questions from \S\ref{S:background} while being:
+{\pkg} was designed to transform models into building blocks that can be easily combined with other blocks to build complex structures, thus promoting model reuse and collaboration between scientists with varying degrees of overlapping expertise.
+{\pkg} does so by addressing the questions from \S\ref{S:background} while being:
 
@@
 
-To accomplish these objectives, {\pkg} is built upon several well-established open-source software packages along with tools developed explicitly for {\pkg}.
-{\pkg} currently support models written in Python, Matlab, C, and C++ with
-additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}.
+{\pkg} is built upon several well-established open-source software packages(see below) along with new tools developed explicitly for {\pkg} including utilities for running and monitoring models written in other languages from Python, dynamically creating and managing communication networks, data type conversions between languages, and asynchronous message passing with variable underlying communication mechanisms.
+{\pkg} currently support models written in Python, Matlab, C, and C++ with additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}.
 Support for additional languages is planned for future development (See \S\ref{SS:langsupport}).
-\todo{Statement on novelty}
 
@@
 {\pkg} is executed via a command line interface (CLI), {\pkgrun} with specification files (\S\ref{SSS:yaml}) as input.
-Based on the information contained in the specification files, {\pkg} establishes a network of asynchronous communication channels (\S\ref{SS:communication}), and launches the models on new processes (\S\ref{SS:model_execution}).
+Based on the information contained in the specification files, {\pkg} dynamically establishes a network of asynchronous communication channels (\S\ref{SS:communication}), and launches the models on new processes (\S\ref{SS:model_execution}).
+Although written in Python, {\pkg} was written such that users need not have any knowledge of the Python language and can interact with {\pkg} solely through the CLI.
 %At execution, {\pkg} launches models on new processes based on information contained in a YAML specification file (See \S\ref{SSS:yaml} below) and establishes
@@
 In the example, this would be the shoot growth model which required almost twice the time to execute as the root model.
-If the models are dependent on one-another, the models will have to wait for messages to be passed and the speedup will not be as great (see \S\ref{S:speedup} for additional discussion of parallelism and the speedups achievable using {\pkg}).
+If the models are dependent on one-another, the models will have to wait for messages to be passed and the speedup will not be as great (see \S\ref{SS:speedup} for additional discussion of parallelism and the speedups achievable using {\pkg}).
 \todo{Fix these numbers for most recent version of the code}.
@@
 While every model is executed on a new process, how the model is handled depends 
-on the language it is written in. Models written in 
+on the language it is written in.
+{\pkg} has a dedicated driver for each of the supported languages with utilities that allow {\pkg} to launch and monitor executables written in that language from Python.
+Models written in
 interpreted languages (Python and Matlab) are executed on the command line 
@@
 
-The algorithms for a generic model driver and the master thread can be found in Algorithms \ref{alg:model} \& \ref{alg:master} respectively.
-% MODEL DRIVER
-\begin{algorithm}[H]
-    \caption{Generic model driver class.
-This algorithm does not represent the actual implementation of any one of the languages supported by {\pkg}.
-Each of the language-specific drivers will have slightly different methods for compilation (turning the source code into an executable), launching the model process, and monitoring the model's status via output and error codes.
-However, this algorithm does provided the general behavior pattern upon which every model driver is based.}
-    \label{alg:model}
-    \begin{algorithmic}[1]
-    \Class{ModelDriver}{$name, args, yml$}
-    	\State $errors \gets []$
-    	\State $executable \gets$ \Call{Compile}{args}
-	\State $process \gets$ \Call{ModelProcess}{} on a new process
-	\State $thread \gets$ \Call{ModelThread}{} on a new thread
-	\item[]
-	%
-	\ClassMethod{ModelProcess}{}
-		\State \Call{Executable}{}
-	\EndClassMethod
-	\item[]
-	%
-	\ClassMethod{ModelThread}{}
-		\While{\Call{IsAlive}{{\sc ModelProcess}}}
-			\If{\Call{HasOutput}{{\sc ModelProcess}}}
-				\State \textbf{Print} \Call{GetOutput}{{\sc ModelProcess}}
-			\Else
-				\State \textbf{Sleep}
-			\EndIf
-		\EndWhile
-		\If{\Call{HasErrors}{{\sc ModelProcess}}}
-			\State $errors \gets$ \Call{GetErrors}{{\sc ModelProcess}}
-		\EndIf
-	\EndClassMethod
-	\item[]
-	%
-	\ClassMethod{Terminate}{}
-		\State \Call{Kill}{{\sc ModelProcess}}
-	\EndClassMethod
-	\item[]
-	%
-	\ClassMethod{Stop}{}
-		\While{not $errors$}
-			\State \textbf{Sleep}
-		\EndWhile
-		\State \Call{Terminate}{}
-	\EndClassMethod
-	\item[]
-    \EndClass
-    \end{algorithmic}
-\end{algorithm}
-
-%
-% MASTER THREAD
-%\begin{center}
-\begin{algorithm}[H]
-	\caption{Basic steps executed on the master thread to create, start, and monitor model \& connection drivers.
-This pseudocode is not an exhaustive reproduction of the entire implementation, but represents the broad strokes needed to understand the purpose and behavior of the master thread.
-The algorithm for the {\sc ModelDriver} and {\sc ConnectionDriver} can be found in Algorithms \ref{alg:model} \& \ref{alg:conn} respectively.}
-	\label{alg:master}
-	\begin{algorithmic}[1]
-	\State $models$ $\gets$ \textbf{Parse} YAML files
-	\State $inputChannels \gets Map()$
-	\State $outputChannels \gets Map()$
-	% CREATE DRIVERS
-	\item[] \Comment Create Drivers
-	\ForAll{$m in models$}
-		\ForAll{$i in m.inputs$}
-			\If{not $i.isFile$}
-				\State $inputAddress \gets i.address$
-			\Else
-				\State $inputAddress \gets `generate'$
-			\EndIf
-			\State $i.driver \gets$ \Call{ConnectionDriver}{$inputAddress, `generate', i$}
-			\State $inputChannels[i.name] = i$
-		\EndFor
-	\EndFor
-	\ForAll{$m in models$}
-	\algstore{alg:master}
-	\end{algorithmic}
-\end{algorithm}
-% PAGE BREAK
-\begin{algorithm}[H]
-	\caption{Continuation of Algorithm \ref{alg:master}}
-	\begin{algorithmic}[1]
-	\algrestore{alg:master}
-		\ForAll{$o \in m.outputs$}
-			\If{$o.isFile$}
-				\State $outputAddress \gets o.address$
-			\Else
-				\If{$o.args \in inpurChannels$}
-					\State $outputAddress \gets inputChannels[o.args]$
-				\Else
-					\State \Call{Error}{}
-				\EndIf
-			\EndIf
-			\State $o.driver \gets$ \Call{ConnectionDriver}{$`generate', outputAddress, o$}
-			\State $outputChannels[o.name] = o$
-		\EndFor
-	\EndFor
-	\ForAll{$m in models$}
-		\State $m.driver \gets$ \Call{ModelDriver}{$m.name, m.args, m$}
-	\EndFor
-	% START DRIVERS
-	\item[] \Comment Start Drivers	
-	\ForAll{$m in models$}
-		\ForAll{$i in m.inputs$}
-			\State $i.driver.$\Call{Start}{}
-		\EndFor
-	\EndFor
-	\ForAll{$m in models$}
-		\ForAll{$o \in m.outputs$}
-			\State $o.driver.$\Call{Start}{}
-		\EndFor
-	\EndFor
-	\ForAll{$m in models$}
-		\State $m.driver.$\Call{Start}{}
-	\EndFor
-	% MONITOR DRIVERS
-	\item[] \Comment Monitor Drivers
-	\State $errorFlag \gets false$
-	\While{not $errorFlag$}
-		\ForAll{$m in models$}
-			\If{$m.driver.errors$}
-				\State $errorFlag \gets true$
-				\State \textbf{break}
-			\EndIf
-		\EndFor
-	\EndWhile
-	% STOP DRIVER
-	\item[] \Comment Stop Drivers
-	\If{$errorFlag$}
-		\State $allErrors \gets []$
-		\ForAll{$m in models$}
-			\State $m.$\Call{Terminate}{}
-			\State \textbf{Append} $allErrors$, $m.errors$
-		\EndFor
-		\State \Call{Error}{$allErrors$}
-	\Else
-		\ForAll{$m in models$}
-			\State $m.$\Call{Stop}{}
-		\EndFor
-	\EndIf
-	\end{algorithmic}
-\end{algorithm}
-%\end{center}
-%
-\todo{make sure this is split correctly across pages}
+The algorithms for a generic model driver and the master thread can be found in Algorithms \ref{alg:model} \& \ref{alg:master} respectively in \A\ref{A:algorithms}.
 
@@
 {\pkg} includes its own implementation of asynchronous communication via each of the communication mechanisms that are supported.
-While there are existing tools for asynchronous communication using each of these tools, using a tool developed specifically for {\pkg} allows for a more uniform treatment of the different communication mechanisms, greater control over how threads are managed (e.g.
+While there are existing tools for asynchronous communication using each of these communication mechanisms, using a tool developed specifically for {\pkg} allows for a more uniform treatment of the different communication mechanisms, greater control over how threads are managed (e.g.
 killing a thread when messages cannot be interpreted), and a more seamless coordination with other {\pkg} processes (e.g.
@@
 
+In addition to using the default ZMQ libraries, {\pkg} also includes supporting routines for allowing received messages to be confirmed.
+Because ZMQ is broker-less, a socket has no way of knowing if the message it sent was successfully received.
+This lack of confirmation makes it harder to determine if there is was error somewhere along the network.
+{\pkg} overcomes this by generating two ZMQ sockets for every connection: one for sending the messages and one for confirming them.
+Each time a message is received, the receiving socket confirmation thread will send a confirmation including a unique ID for the received message and then wait for a reply to that confirmation.
+The sending socket confirmation thread will continuously check the confirmation sockets for messages indicating that sent messages were received.
+Once a confirmation message is received, it will send a reply to the receiving confirmation socket and record the ID of the message that was confirmed.
+This handshake operation ensures that all messages are accounted for so {\pkg} knows if a message is lost and where it was lost.
+
 %---------------------------------------------------------------------------------
@@
 As a result, models in complex integration networks are not affected by the rate at which dependent model consume their output and the speedup offered by running the models in parallel is improved.
-Figure \ref{fig:async} describes the general flow of messages, Algorithms \ref{alg:async_send} \& \ref{alg:async_recv} describe the asynchronous output \& input channel procedures, and Algorithm \ref{alg:conn} describes the procedure followed by input \& output connection drivers.
+Figure \ref{fig:async} describes the general flow of messages, Algorithms \ref{alg:async_send} \& \ref{alg:async_recv} in \A\ref{A:algorithms} describe the asynchronous output \& input channel procedures, and Algorithm \ref{alg:conn} describes the procedure followed by input \& output connection drivers.
 %
@@
 \end{enumerate}
-
-% SENDER
-\begin{algorithm}[H]
-    \caption{Asynchronous communication class for sending messages to an output channel.
-The procedures {\sc Open, Close, Send, \& IsOpen} are procedures that are implemented by external communication packages and different slightly from one communication mechanism to the next.}
-    \label{alg:async_send}
-    \begin{algorithmic}[1]
-    % SENDER
-    \Class{AsyncSender}{$address$}
-    	\State $backlog \gets []$
-	\State $x \gets$ \Call{Open}{$address$}
-	\State \textbf{Run} \Call{BacklogThread}{} on a new thread
-	\item[]
-	%
-	\ClassMethod{BacklogThread}{}
-		\While{\Call{IsOpen}{$x$}}
-    			\If{$length(backlog) > 0$}
-				\State $flag \gets$ \Call{Send}{$x$, $backlog[0]$}
-				\If{$flag$}
-					\State \textbf{Remove} $backlog[0]$
-				\Else
-					\State \Call{Close}{$x$}
-				\EndIf
-			\Else
-				\State \textbf{Sleep}
-			\EndIf
-    		\EndWhile
-    	\EndClassMethod
-	\item[]
-	%
-	\ClassMethod{SendBacklog}{$msg$}
-		\If{\Call{IsOpen}{$x$}}
-			\State \textbf{Append} $backlog$, $msg$
-		\EndIf
-		\State \textbf{Return} \Call{IsOpen}{$x$}
-	\EndClassMethod
-	\item[]
-	%
-	\ClassMethod{Close}{}
-		\State \Call{Close}{$x$}
-		\State $backlog \gets []$
-	\EndClassMethod
-	\item[]
-    \EndClass
-    \end{algorithmic}
-\end{algorithm}
-
-% RECEIVER
-\begin{algorithm}[H]
-    \caption{Asynchronous communication class for receiving messages from an input channel.
-The procedures {\sc Open, Close, Recv, IsOpen, \& IsMessageWaiting} are procedures that are implemented by external communication packages and different slightly from one communication mechanism to the next.}
-    \label{alg:async_recv}
-    \begin{algorithmic}[1]
-    \item[]
-    \Class{AsyncReceiver}{$address$}
-    	\State $backlog \gets []$
-	\State $x \gets$ \Call{Open}{$address$}
-	\State $thread \gets$ \Call{BacklogThread}{} on a new thread
-	\item[]
-	%
-	\ClassMethod{BacklogThread}{}
-		\While{\Call{IsOpen}{$x$}}
-    			\If{\Call{IsMessageWaiting}{$x$}}
-				\State $flag, msg \gets$ \Call{Recv}{$x$}
-				\If{$flag$}
-					\State \textbf{Append} $backlog$, $msg$
-				\Else
-					\State \Call{Close}{$x$}
-				\EndIf
-			\Else
-				\State \textbf{Sleep}
-			\EndIf
-		\EndWhile
-	\EndClassMethod
-	\item[]
-	%
-	\ClassMethod{RecvBacklog}{$msg$}
-		\If{$length(backlog) > 0$}
-			\State $msg \gets backlog[0]$
-			\State \textbf{Remove} $backlog[0]$
-			\State \textbf{Return} true, $msg$
-		\Else
-			\State \textbf{Return} \Call{IsOpen}{$x$}, ''
-		\EndIf
-	\EndClassMethod
-	\item[]
-	%
-	\ClassMethod{Close}{}
-		\State \Call{Close}{$x$}
-		\State $backlog \gets []$
-	\EndClassMethod
-	\item[]
-    \EndClass
-    \end{algorithmic}
-\end{algorithm}
-%
-
-%
-\begin{algorithm}[H]
-    \caption{Connection Driver}
-    \label{alg:conn}
-    \begin{algorithmic}[1]
-    \Class{ConnectionDriver}{$inputAddress, outputAddress, yml$}
-    	\State $input \gets$ \Call{AsyncReceiver}{$inputAddress$}
-    	\State $output \gets$ \Call{AsyncSender}{$outputAddress$}
-	\State $thread \gets$ \Call{ConnectionThread}{} on a new thread
-	\item[]
-	%
-	\ClassMethod{ConnectionThread}{}
-		\While{true}
-			\State $flag, msg \gets$ $input.$\Call{RecvBacklog}{}
-			\If{not $flag$} \State \textbf{break} \EndIf
-			\If{$msg$}
-				\State \textbf{Transform} $msg$ based on units and expected data format
-				\State $flag \gets$ $output.$\Call{SendBacklog}{$msg$}
-				\If{not $flag$} \State \textbf{break} \EndIf
-			\Else
-				\State \textbf{Sleep}
-			\EndIf
-		\EndWhile
-		%
-		\State $input.$\Call{CloseAsync}{}
-		\State $output.$\Call{CloseAsync}{}
-	\EndClassMethod
-	\item[]
-	%
-	\ClassMethod{Terminate}{}
-		\State $input.$\Call{CloseAsync}{}
-		\State $output.$\Call{CloseAsync}{}
-	\EndClassMethod
-	\item[]
-	%
-	\ClassMethod{Stop}{}
-		\While{\Call{IsOpen}{$input.x$} and \Call{IsOpen}{$output.x$}}
-			\State \textbf{Sleep}
-		\EndWhile
-		\State \Call{Terminate}{}
-	\EndClassMethod
-	\item[]
-    \EndClass
-    \end{algorithmic}
-\end{algorithm}
-
-\todo{discuss algorithm}
 
@@
 working in parallel.
-
-%\begin{algorithm}[H]
-%    \caption{Large Messages}
-%    \label{alg:large}
-%    \begin{algorithmic}[1] 
-%    \end{algorithmic}
-%\end{algorithm}
-
-\todo{algorithm}
+The procedures for creating and using temporary channels for sending large messages can be seen in the {\sc Send} and {\sc Recv} methods from Algorithms \ref{alg:async_send} \& \ref{alg:async_recv} respectively in \A\ref{A:algorithms}.
 
@@
 to more complex data types and communication patterns.
+Pseudocode describing the asynchronous communication classes and functions underlying the interfaces is provided by Algorithms \ref{alg:async_send} \& \ref{alg:async_recv} in \A\ref{A:algorithms}.
 
@@
 However, unless the variables will \emph{always} be coupled in the model, it is advised that every variable be specified separately for clarity and to allow flexibility in the way other models can integrate with it.
-In addition, if brevity is a concern because a model has a large number of input variables, there are features currently under active development (see \S\ref{SS:datagg}) that will allow fields to be aggregated within connection YAMLs and simplify the send and receive calls in such cases.
+In addition, if brevity is a concern because a model has a large number of input variables, there are features currently under active development (see \S\ref{SS:dataagg}) that will allow fields to be aggregated within connection YAMLs and simplify the send and receive calls in such cases.
 
@@
 
-Integration via {\pkg} requires minimal modification of the existing model source code yet greatly increases the model's potential by allowing it to be reused in combination with other models.
+{\pkg} is designed to make models as reusable as possible.
+Once integrated via {\pkg}, models can be combined with any other models that also have a {\pkg} wrapper and YAML specification file with minimal, if any, modification.
+This "plug-and-play" approach to models encourages model writers to make their codes public in a reusable form, generating citations and reuse of their work, while also growing a community of models and model writers capable of probing areas otherwise unreachable.
+%Generic workflow, parallelization, and communication tools can be used to expose/connect models, but without a standard for making connections, every new integration will require a great deal of hands on attention and new code.
+Domain specific frameworks like OpenAlea \citep{Pradal2008} are better for improving the reusability for some models in that they can impose a standard for how models interact with each other, however none of the existing frameworks offer the same flexibility as {\pkg} in terms of languages with native support and model form.
+%Integration via {\pkg} requires minimal modification of the existing model source code yet greatly increases the model's potential by allowing it to be reused in combination with other models. 
+%
 {\pkg} has already been used to integrate several plant models within the Crops in Silico organization \citep{Marshall-Colon2017} with ongoing work to add additional models to the integration network.
@@
 
-%---------------------------------------------------------------------------------
-% USE CASES AND LIMITATION
-%
-\subsection{Use Cases \& Limitations}\label{SS:uses}
-%
-In addition to allowing scientists to reuse existing models in new integrations, {\pkg} has the potential to significantly increase a model's accessibility.
-{\pkg} could be used a method of running models without knowledge of the model's implementation.
+In addition to allowing scientists to reuse existing models in new integrations and make new connections, {\pkg} has the potential to significantly increase a model's accessibility and promote reuse.
+{\pkg} could be used as a method of running models without knowledge of the model's implementation.
 At publication, model code could be released along with a wrapper and YAML specification file for use with {\pkg}.
 These additional materials would be particularly powerful when considered alongside the interactive features being developed for use with {\pkg} (see \S\ref{SS:gui} and \S\ref{SS:function}).
-%
-\todo{Add discussion of limitations}
+Potential collaborators could use {\pkg} to run and explore models without the need to be familiar with the inner workings of the model implementation or even implementation language.
+
+%---------------------------------------------------------------------------------
+% USE CASES AND LIMITATION
+%
+\subsection{Advantages \& Limitations}\label{SS:uses}
+%
+\subsubsection{Advantages}\label{SS:advantages}
+%
+{\pkg} offers several advantages over manual integrations or other existing tools.
+%
+\begin{description}
+	\item[Dynamic Communication Network] While there are many existing tools for communication that have implementations in multiple languages \citep[some of which {\pkg} uses, e.g.][ZMQ, RabbitMQ, OpenMPI]{ZMQ, RMQ, Gabriel2004}, these tools require that the communication pattern be well established such that it can be hard coded into the model (or model wrapper) itself (e.g.
+using specific ports or queues).
+The use of hard coded values greatly restricts the reusability of a model.
+For every integration a model is used in, the hard coded values for every connection must be updated between the models on either end so that they match.
+{\pkg} is unique in that it dynamically establishes and manages the networks of communication resources (e.g.
+queues, channels, sockets) that are required for integrations to run so that models do not need to have these hard coded value and can be reused in multiple integrations, unmodified.
+This process includes creating new resources, connecting to existing resources, passing the appropriate resource addresses to the model environment, error handling/propagation, and resource cleanup.
+	\item[Hands-off Parallelism \& Asynchronous Communication] In addition to inter-language communication, {\pkg} users benefit from speedups due to parallelism and concurrency without any additional work.
+From the specified model connections, {\pkg} manages the parallel execution of models and asynchronous message passing under-the-hood so that the user doesn't have to think about it.
+Existing tools for asynchronous communication between unmanaged parallel tasks \citep[e.g.
+Jack2,][]{Magoules2018} requires explicit definition of the communication pattern in advance, prohibiting dynamic communication networks as discussed above.
+Work/data flow tools like Parsl \citep{babuji18} allow dynamic execution of processes along a dependency chain in parallel, but usually require addition code, are limited to passing data as files or data object native to the tool's implementation language, and are ill suited to models with persistent internal states or intermediate outputs.
+{\pkg} is unique in that it offers users asynchronous communication and parallelism that is flexible enough to support any communication pattern, implicitly handles conversion between native data types in the supported languages, and does not require any additional time investment on the part of the user.
+	\item[No New Language] Unlike DSLs or language specific frameworks, {\pkg} doesn't require a model to be written in one specific language.
+In addition to expanding the possible number of model combinations, this eases the burden on model writers for whom learning an additional programming language or DSL might be a significant obstacle.
+{\pkg} is designed such that model writers only need to be familiar with the language that the model is written in to being with.
+Although {\pkg} is a Python package, users do not even need to be familiar with the Python language and can interact with {\pkg} solely through the command line.
+While users do need to be able to compose YAML specification files in addition to their model code, these files have a very limited syntax and will be GUI composable in future releases.
+\end{description}
+
+\subsubsection{Limitations}\label{SS:limitations}
+%
+While {\pkg} is a powerful tool for integrating scientific models and opening up new avenues for collaboration, it is not applicable to every scenario.
+%
+\begin{description}
+	\item[Single Language Integrations] For models written in the same language, it is usually much more computationally efficient to either integrate the models directly (i.e.
+by calling one from the other) or use a language specific tools for parallel execution and passing data \citep[e.g.
+Parsl in Python][]{babuji18} than to use {\pkg}.
+This performance advantage will be particularly pronounced for models that finish quickly or tightly coupled integrations where models make a large number of exchanges in rapid succession (e.g.
+a set of coupled ODEs).
+Although the constituent models will still benefit from integration via {\pkg} in regards to reusability, {\pkg} may not be the best choice in such cases if execution time is a concern.
+	\item[Easily Translatable Models] For relatively simple models ($<100$ lines of code), it may be more efficient for a scientist to translate the model into the language of the model(s) it will be integrated with.
+{\pkg} is still useful as a tool more making the model accessible and more likely to be reused, but the performance advantage of running codes in the same language may be of greater importance in certain situations as discussed above.
+	\item[Unsupported Languages] While {\pkg} currently supports models written in Python, Matlab, C, \& C++, models are written in many other languages.
+There are plans to expand support for {\pkg} to other languages (see \S\ref{SS:langsupport}), but in the meantime {\pkg} will not be particularly useful for models that are not written in one of the languages already supported.
+\end{description}
 
@@
 networks that can be executed. Several improvements are already in progress/planned for {\pkg}.
+\todo{expand?}
 
@@
 %% APPENDICES
+\newpage
 %\appendix
 %
+\begin{appendices}
+
+%---------------------------------------------------------------------------------
+% ALGORITHMS
+\section{Algorithms}\label{A:algorithms}
+
+% MODEL DRIVER
+\begin{algorithm}[H]
+    \caption{Generic model driver class.
+This algorithm does not represent the actual implementation of any one of the languages supported by {\pkg}.
+Each of the language-specific drivers will have slightly different methods for compilation (turning the source code into an executable), launching the model process, and monitoring the model's status via output and error codes.
+However, this algorithm does provided the general behavior pattern upon which every model driver is based.}
+    \label{alg:model}
+    \begin{algorithmic}[1]
+    \Class{ModelDriver}{$name, args, yml$}
+    	\State $errors \gets []$
+    	\State $executable \gets$ \Call{Compile}{args}
+	\State $process \gets$ \Call{ModelProcess}{} on a new process
+	\State $thread \gets$ \Call{ModelThread}{} on a new thread
+	\item[]
+	%
+	\ClassMethod{ModelProcess}{}
+		\State \Call{Executable}{}
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{ModelThread}{}
+		\While{\Call{IsAlive}{{\sc ModelProcess}}}
+			\If{\Call{HasOutput}{{\sc ModelProcess}}}
+				\State \textbf{Print} \Call{GetOutput}{{\sc ModelProcess}}
+			\Else
+				\State \textbf{Sleep}
+			\EndIf
+		\EndWhile
+		\If{\Call{HasErrors}{{\sc ModelProcess}}}
+			\State $errors \gets$ \Call{GetErrors}{{\sc ModelProcess}}
+		\EndIf
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{Terminate}{}
+		\State \Call{Kill}{{\sc ModelProcess}}
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{Stop}{}
+		\While{not $errors$}
+			\State \textbf{Sleep}
+		\EndWhile
+		\State \Call{Terminate}{}
+	\EndClassMethod
+	\item[]
+    \EndClass
+    \end{algorithmic}
+\end{algorithm}
+
+%
+% MASTER THREAD
+%\begin{center}
+\begin{algorithm}[H]
+	\caption{Basic steps executed on the master thread to create, start, and monitor model \& connection drivers.
+This pseudocode is not an exhaustive reproduction of the entire implementation, but represents the broad strokes needed to understand the purpose and behavior of the master thread.
+The algorithm for the {\sc ModelDriver} and {\sc ConnectionDriver} can be found in Algorithms \ref{alg:model} \& \ref{alg:conn} respectively.}
+	\label{alg:master}
+	\begin{algorithmic}[1]
+	\State $models$ $\gets$ \textbf{Parse} YAML files
+	\State $inputChannels \gets Map()$
+	\State $outputChannels \gets Map()$
+	% CREATE DRIVERS
+	\item[] \Comment Create Drivers
+	\ForAll{$m in models$}
+		\ForAll{$i in m.inputs$}
+			\If{not $i.isFile$}
+				\State $inputAddress \gets i.address$
+			\Else
+				\State $inputAddress \gets $`generate'
+			\EndIf
+			\State $i.driver \gets$ \Call{ConnectionDriver}{$inputAddress$, `generate', $i$}
+			\State $inputChannels[i.name] = i$
+	% PAGE BREAK
+	\algstore{alg:master}
+	\end{algorithmic}
+\end{algorithm}
+\addtocounter{algorithm}{-1}
+\begin{algorithm}[H]
+	\caption{Continued}
+	\begin{algorithmic}[1]
+	\algrestore{alg:master}
+	% PAGE BREAK
+		\EndFor
+	\EndFor
+	\ForAll{$m in models$}
+		\ForAll{$o \in m.outputs$}
+			\If{$o.isFile$}
+				\State $outputAddress \gets o.address$
+			\Else
+				\If{$o.args \in inpurChannels$}
+					\State $outputAddress \gets inputChannels[o.args]$
+				\Else
+					\State \Call{Error}{}
+				\EndIf
+			\EndIf
+			\State $o.driver \gets$ \Call{ConnectionDriver}{`generate', $outputAddress$, $o$}
+			\State $outputChannels[o.name] = o$
+		\EndFor
+	\EndFor
+	\ForAll{$m in models$}
+		\State $m.driver \gets$ \Call{ModelDriver}{$m.name$, $m.args$, $m$}
+	\EndFor
+	% START DRIVERS
+	\item[] \Comment Start Drivers	
+	\ForAll{$m in models$}
+		\ForAll{$i in m.inputs$}
+			\State $i.driver.$\Call{Start}{}
+		\EndFor
+	\EndFor
+	\ForAll{$m in models$}
+		\ForAll{$o \in m.outputs$}
+			\State $o.driver.$\Call{Start}{}
+		\EndFor
+	\EndFor
+	\ForAll{$m in models$}
+		\State $m.driver.$\Call{Start}{}
+	\EndFor
+	% MONITOR DRIVERS
+	\item[] \Comment Monitor Drivers
+	\State $errorFlag \gets false$
+	\While{not $errorFlag$}
+		\ForAll{$m in models$}
+			\If{$m.driver.errors$}
+				\State $errorFlag \gets true$
+				\State \textbf{break}
+			\EndIf
+		\EndFor
+	\EndWhile
+	% STOP DRIVER
+	\item[] \Comment Stop Drivers
+	\If{$errorFlag$}
+		\State $allErrors \gets []$
+		\ForAll{$m in models$}
+			\State $m.$\Call{Terminate}{}
+			\State \textbf{Append} $allErrors$, $m.errors$
+		\EndFor
+		\State \Call{Error}{$allErrors$}
+	\Else
+		\ForAll{$m in models$}
+			\State $m.$\Call{Stop}{}
+		\EndFor
+	\EndIf
+	\end{algorithmic}
+\end{algorithm}
+
+% SENDER
+\begin{algorithm}[H]
+    \caption{Asynchronous communication class for sending messages to an output channel.
+$x$ is a communication object implemented by an external communication package (one of the three supported by {\pkg}).
+The package must have at minimum methods {\sc NewAddress, Open, Close, Send, \& IsOpen} for interacting with the object $x$.
+{\sc Encode} is dependent on the message type and will use an extended version of JSON encoding in the future (see \S\ref{SS:json}).}
+    \label{alg:async_send}
+    \begin{algorithmic}[1]
+    % SENDER
+    \Class{AsyncSender}{$address$}
+    	\State $backlog \gets []$
+	\State $temp \gets []$
+	\If{$address==$`generate'}
+		\State $address \gets$ \Call{NewAddress}{}
+	\EndIf
+	\State $x \gets$ \Call{Open}{$address$}
+	\State \textbf{Run} \Call{BacklogLoop}{} on a new thread
+	\item[]
+	%
+	\ClassMethod{BacklogLoop}{}
+		\While{$x.$\Call{IsOpen}{}}
+    			\If{$length(backlog) > 0$}
+				\State $flag \gets$ $x.$\Call{Send}{$backlog[0]$}
+				\If{$flag$}
+					\State \textbf{Remove} $backlog[0]$
+				\Else
+					\State $x.$\Call{Close}{}
+				\EndIf
+			\Else
+				\State \textbf{Sleep}
+			\EndIf
+    		\EndWhile
+    	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{SendBacklog}{$msg$}
+		\If{$x.$\Call{IsOpen}{}}
+			\State \textbf{Append} $msg$ to the end of $backlog$
+		\EndIf
+		\State \textbf{Return} $x.$\Call{IsOpen}{}
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{SendMultipart}{$msg$}
+		\State $currPos \gets 0$
+		\While{$currPos < length(msg)$}
+			\State $ilength \gets \min(x.maxMsgSize, length(msg) - currPos)$
+			\State \Call{Send}{$msg[currPos:(currPos + ilength)]$}
+			\State $currPos += ilength$
+		\EndWhile
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{Send}{$msg$}
+		\State $header \gets Map()$ \Comment{Additional information will be added to the header based on communication mechanism and encoding.}
+		\State $totalMsg \gets $\Call{Encode}{$msg, header$}
+		\If{$length(totalMsg) > x.maxMsgSize$}
+			\State $newTemp \gets$ \Call{AsyncSender}{`generate'}
+			\State \textbf{Append} $newTemp$ to the end of $temp$
+			\State $header[$`workerAddress'$] \gets newTemp.address$
+			\State $totalMsg \gets $\Call{Encode}{$msg, header$}
+			\If{not \Call{SendBacklog}{$totalMsg[:x.maxMsgSize]$}}
+				\State \textbf{Return} false
+			\EndIf
+			\If{not $newTemp.$\Call{SendMultipart}{$totalMsg[x.maxMsgSize:]$}}
+				\State \textbf{Return} false
+			\EndIf
+			\State \textbf{Return} true
+		\Else
+			\State \textbf{Return} \Call{SendBacklog}{$totalMsg$}
+		\EndIf
+	\EndClassMethod
+	\item[]
+	% PAGE BREAK
+	\algstore{alg:master}
+	\end{algorithmic}
+\end{algorithm}
+\addtocounter{algorithm}{-1}
+\begin{algorithm}[H]
+	\caption{Continued}
+	\begin{algorithmic}[1]
+	\algrestore{alg:master}
+	% PAGE BREAK
+	%
+	\ClassMethod{Close}{$dontWaitForRecv=false$}
+		\If{$x.$\Call{IsOpen}{}}
+			\State \Call{Send}{$EOF$}
+		\EndIf
+		\If{not $dontWaitForRecv$}
+			\While{$x.$\Call{IsOpen}{} and $((length(backlog) > 0)$ or $x.$\Call{IsMessageWaiting}{}$)$}
+				\State \textbf{Sleep}
+			\EndWhile
+		\EndIf
+		\State $x.$\Call{Close}{}
+		\State $backlog \gets []$
+		\ForAll{$t \in temp$}
+			\State $t.$\Call{Close}{$dontWaitForRecv$}
+		\EndFor
+	\EndClassMethod
+	\item[]
+    \EndClass
+    \end{algorithmic}
+\end{algorithm}
+
+% RECEIVER
+\begin{algorithm}[H]
+    \caption{Asynchronous communication class for receiving messages from an input channel.
+$x$ is a communication object implemented by an external communication package (one of the three supported by {\pkg}).
+The package must have at minimum methods {\sc NewAddress, Open, Close, Recv, IsOpen, \& IsMessageWaiting} for interacting with the object $x$.
+{\sc Decode} is dependent on the message type and will use an extended version of JSON decoding in the future (see \S\ref{SS:json}).}
+    \label{alg:async_recv}
+    \begin{algorithmic}[1]
+    \item[]
+    \Class{AsyncReceiver}{$address$}
+    	\State $backlog \gets []$
+	\State $temp \gets []$
+	\If{$address==$`generate'}
+		\State $address \gets$ \Call{NewAddress}{}
+	\EndIf
+	\State $x \gets$ \Call{Open}{$address$}
+	\State $thread \gets$ \Call{BacklogLoop}{} on a new thread
+	\item[]
+	%
+	\ClassMethod{BacklogLoop}{}
+		\While{$x$.\Call{IsOpen}{}}
+    			\If{$x$.\Call{IsMessageWaiting}{}}
+				\State $flag, msg \gets$ $x$.\Call{Recv}{}
+				\If{$flag$}
+					\State \textbf{Append} $msg$ to the end of $backlog$
+				\Else
+					\State $x$.\Call{Close}{}
+				\EndIf
+			\Else
+				\State \textbf{Sleep}
+			\EndIf
+		\EndWhile
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{RecvBacklog}{$msg$}
+		\If{$length(backlog) > 0$}
+			\State $msg \gets backlog[0]$
+			\State \textbf{Remove} $backlog[0]$
+			\State \textbf{Return} true, $msg$
+		\Else
+			\State \textbf{Return} $x.$\Call{IsOpen}{}, `'
+		\EndIf
+	\EndClassMethod
+	\item[]
+	% PAGE BREAK
+	\algstore{alg:master}
+	\end{algorithmic}
+\end{algorithm}
+\addtocounter{algorithm}{-1}
+\begin{algorithm}[H]
+	\caption{Continued}
+	\begin{algorithmic}[1]
+	\algrestore{alg:master}
+	% PAGE BREAK
+	%
+	\ClassMethod{RecvMultipart}{$msgSize$}
+		\State $msg \gets$ `'
+		\While{$length(msg) < msgSize$}
+			\State $flag, msgPart \gets$ \Call{Recv}{}
+			\If{not $flag$}
+				\State \textbf{Return} $flag, msg$
+			\EndIf
+			\State $msg \gets msg + msgPart$
+		\EndWhile
+		\State \textbf{Return} $true, msg$
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{Recv}{}
+		\State $flag, msgStart \gets $\Call{RecvBacklog}{}
+		\If{not $flag$}
+			\State \textbf{Return} $flag$, `'
+		\EndIf
+		\State $header \gets $\Call{DecodeHeader}{$msgStart$}
+		\If{`workerAddress' $\in header$}
+			\State $newTemp \gets$ \Call{AsyncReceiver}{$header[$`workerAddress'$]$}
+			\State \textbf{Append} $newTemp$ to the end of $temp$
+			\State $flag, msgRemain \gets newTemp.$\Call{RecvMultipart}{}
+			\If{not $flag$}
+				\State \textbf{Return} $flag$, `'
+			\EndIf
+			\State $header, msg \gets $\Call{Decode}{$msgStart + msgRemain$}
+		\Else
+			\State $header, msg \gets $\Call{Decode}{$msgStart$}
+		\EndIf
+		\State \textbf{Return} $true, msg$
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{Close}{}
+		\State $x.$\Call{Close}{}
+		\State $backlog \gets []$
+		\ForAll{$t \in temp$}
+			\State $t.$\Call{Close}{}
+		\EndFor
+	\EndClassMethod
+	\item[]
+    \EndClass
+    \end{algorithmic}
+\end{algorithm}
+%
+
+% CONNECTION DRIVER
+\begin{algorithm}[H]
+    \caption{Connection Driver}
+    \label{alg:conn}
+    \begin{algorithmic}[1]
+    \Class{ConnectionDriver}{$inputAddress, outputAddress, yml$}
+    	\State $input \gets$ \Call{AsyncReceiver}{$inputAddress$}
+    	\State $output \gets$ \Call{AsyncSender}{$outputAddress$}
+	\State $thread \gets$ \Call{ConnectionLoop}{} on a new thread
+	\item[]
+	%
+	\ClassMethod{ConnectionLoop}{}
+		\While{true}
+			\State $flag, msg \gets$ $input.$\Call{RecvBacklog}{}
+			\If{not $flag$} \State \textbf{break} \EndIf
+			\If{$msg$}
+				\State \textbf{Transform} $msg$ based on units and expected data format
+				\State $flag \gets$ $output.$\Call{SendBacklog}{$msg$}
+				\If{not $flag$} \State \textbf{break} \EndIf
+			\Else
+				\State \textbf{Sleep}
+			\EndIf
+		\EndWhile
+		%
+		\State $input.$\Call{Close}{}
+		\State $output.$\Call{Close}{}
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{Terminate}{}
+		\State $input.$\Call{Close}{}
+		\State $output.$\Call{Close}{$dontWaitForRecv=true$}
+	\EndClassMethod
+	\item[]
+	%
+	\ClassMethod{Stop}{}
+		\While{$input.x.$\Call{IsOpen}{} and $output.x.$\Call{IsOpen}{}}
+			\State \textbf{Sleep}
+		\EndWhile
+		\State \Call{Terminate}{}
+	\EndClassMethod
+	\item[]
+    \EndClass
+    \end{algorithmic}
+\end{algorithm}
+
+\end{appendices}
+
 %%---------------------------------------------------------------------------------
@@
 \item Fix matlab autodocs for interface
+\item Check that final version of algorithms is split between pages correctly
 \end{itemize}

### 2019-01-05 papers/lang2018/ms.tex
Updated times and speedup for root/shoot example and added response to reviewer 3

@@
 Although developed for the purpose of integrating plant biology models, {\pkg} can be applied to any situation requiring coordination between models in the supported languages.
-\S\ref{S:background} provides background information on existing efforts for cross-language model integration, \S\ref{S:methods} describes the methods used by the {\pkg} package to integrate models, \S\ref{S:results} presents several tests demonstrating the performance of message passing between integrated models, and \S\ref{S:discuss} summarizes the features of the {\pkg} package, describes a few use cases, and outlines areas of ongoing and future development.
+\S\ref{S:background} provides background information on existing efforts for cross-language model integration, \S\ref{S:methods} describes the methods used by the {\pkg} package to integrate models, \S\ref{S:example} provides a worked example of integrating the two model presented above, \S\ref{S:results} presents several tests demonstrating the performance of message passing between integrated models, and \S\ref{S:discuss} describes a few use cases for {\pkg},
+summarizes the features and limitations of the {\pkg} package, and outlines areas of ongoing and future development.
 
@@
 
-{\pkg} is an example of a framework that overcomes these issues by exposing simple and easily accessible interfaces in the language of the model (See \S\ref{SSS:interface}) that permit messages to be passed between the model processes as they run in parallel (See \S\ref{SS:communication}).
+{\pkg} is an example of a framework that overcomes these issues by exposing simple and easily accessible interfaces in the languages of the models (See \S\ref{SSS:interface}) that permit messages to be passed between the model processes as they run in parallel (See \S\ref{SS:communication}).
 As a result, models writers only need knowledge of the language in which their model is written.
@@
 
-{\pkg} is built upon several well-established open-source software packages(see below) along with new tools developed explicitly for {\pkg} including utilities for running and monitoring models written in other languages from Python, dynamically creating and managing communication networks, data type conversions between languages, and asynchronous message passing with variable underlying communication mechanisms.
+{\pkg} is built upon several well-established open-source software packages (see below) along with new tools developed explicitly for {\pkg} including utilities for running and monitoring models written in other languages from Python, dynamically creating and managing communication networks, data type conversions between languages, and asynchronous message passing with variable underlying communication mechanisms.
 {\pkg} currently support models written in Python, Matlab, C, and C++ with additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}.
@@
 %
-Users specify information about models and integration networks via declarative 
-YAML files \citep{Ben-Kiki2009}. The YAML file format was selected because it is human readable 
-and there are many existing tools for parsing YAML formats in different programming 
-languages \citep[e.g.][]{Simonov2006,pyyaml,jsyaml}. The declarative format allows user to specify exactly what they 
-want to do, without describing how it should be done. While the information 
-about models and integration networks can be contained in a single YAML 
-file, the information can naturally be split between two or more files, one (or more) containing 
-information about the model(s) and one containing the connections comprising the 
-integration network. This separation is advantageous because the model YAML 
-can be re-used, unchanged, in conjunction with other integration networks.
-
-Model YAMLs include information about the location of the 
-model source code, the language the model is written in, how the model should 
-be run, and any input or output variables including their data type and units. 
+Users specify information about models and integration networks via declarative YAML files \citep{Ben-Kiki2009}.
+The YAML file format was selected because it is human readable and there are many existing tools for parsing YAML formats in different programming languages \citep[e.g.][]{Simonov2006,pyyaml,jsyaml}.
+The declarative format allows user to specify exactly what they want to do, without describing how it should be done.
+While the information about models and integration networks can be contained in a single YAML file, the information can naturally be split between two or more files, one (or more) containing information about the model(s) and one containing the connections comprising the integration network.
+This separation is advantageous because the model YAML can be re-used, unchanged, in conjunction with other integration networks.
+
+Model YAMLs include information about the location of the model source code, the language the model is written in, how the model should be run, and any input or output variables including their data type and units.
 %
@@
 %
-Integration networks are specified by declaring the connections 
-between models and connections are declared by pairing an output variable from 
-one model with the input variable of another model. The {\pkg} 
-CLI sets up the necessary communication mechanisms to 
-then direct data from one model to the next in the specified pattern. Models 
-can have as many input and/or output variables as is desired and connections between models are specified by references to the 
-input/output variables associated with each model. In addition, input and output variables can also be connected to files.
+Integration networks are specified by declaring the connections between models and connections are declared by pairing an output variable from one model with the input variable of another model.
+The {\pkg} CLI sets up the necessary communication mechanisms to then direct data from one model to the next in the specified pattern.
+Models can have as many input and/or output variables as is desired and connections between models are specified by references to the input/output variables associated with each model.
+In addition, input and output variables can also be connected to files.
 %
@@
 %
-This format allows users the 
-flexibility to create complex integration networks and test models in isolation before running the entire integration network.
+This format allows users the flexibility to create complex integration networks and test models in isolation before running the entire integration network as will be seen in \S\ref{S:example}.
 
@@
 {\pkg} launches each model in an integration on its own process, allowing models to complete independent operations in parallel and complete tasks more quickly.
-For example, the root and shoot growth models from \S\ref{S:intro} require 2.42\,s and 4.06\,s respectively to run for 100 time steps in isolation.
-If these two models were manually integrated in serial via the command line, the integration would require a total of 6.48\,s.
-However, the same integration requires only 4.52\,s using {\pkg}, a speed-up of 1.43.
+For example, the root and shoot growth models from \S\ref{S:intro} (described further in \S\ref{S:example}) require 10.28\,s and 10.40\,s respectively to run for 100 time steps in isolation with direct input/output from/to files in their native language.
+If these two models were manually integrated in serial via the command line, the integration would require a total of 20.68\,s.
+However, the same integration requires only 16.03\,s using {\pkg}, a speed-up of 1.29.
 The exact speedup provided by parallel integration using {\pkg} depends on how independent the models are and how equally distributed the work is between the models.
 If the models are fully independent, the speedup is limited by the time required to execute the slowest model.
-In the example, this would be the shoot growth model which required almost twice the time to execute as the root model.
-If the models are dependent on one-another, the models will have to wait for messages to be passed and the speedup will not be as great (see \S\ref{SS:speedup} for additional discussion of parallelism and the speedups achievable using {\pkg}).
-\todo{Fix these numbers for most recent version of the code}.
-
-While every model is executed on a new process, how the model is handled depends 
-on the language it is written in.
+If the models are dependent on one-another, the models may have to wait to receive messages and the speedup will not be as great.
+\S\ref{SS:speedup} provides additional information about the speedups achievable through {\pkg} parallel integration and how model structure influences the size of the performance boost.
+
+While every model is executed on a new process, how the model is handled depends on the language it is written in.
 {\pkg} has a dedicated driver for each of the supported languages with utilities that allow {\pkg} to launch and monitor executables written in that language from Python.
-Models written in
-interpreted languages (Python and Matlab) are executed on the command line 
-with the interpreter. In the case of Matlab, where a significant amount of time is required to start 
-the Matlab interpreter (See \S\ref{SS:results_language}), Matlab shared engines are used 
-to execute Matlab models. Matlab shared engines are Matlab instances that other processes can 
-submit Matlab code to for execution. While shared engines also require the same amount of time 
-to start as a standard instance of the Matlab interpreter, they can be started in advance and then 
-reused.
-
-For the compiled languages (C and C++) there are a few options. The user can 
-compile the model themselves, provided they link against the appropriate 
-{\pkg} header library. Alternatively, users can provide the location of 
-the model source code and let one of several {\pkg} drivers handle the compilation, including 
-linking against the appropriate {\pkg} interface library. {\pkg} 
-also has support for compiling models using Make \citep{Stallman2004} and 
-CMake \citep{Martin2006} for 
-models that already have a Makefile or CMakeLists.txt. To use these tools to compile a model, 
-lines are added to the recipe in order to allow linking against {\pkg}.
+Models written in interpreted languages (Python and Matlab) are executed on the command line via the interpreter.
+In the case of Matlab, where a significant amount of time is required to start the Matlab interpreter (See \S\ref{SS:results_language}), Matlab shared engines are used to execute Matlab models.
+Matlab shared engines are Matlab instances that other processes can submit Matlab code to for execution.
+While shared engines also require the same amount of time to start as a standard instance of the Matlab interpreter, they can be started in advance and then reused.
+
+For the compiled languages (C and C++) there are a few options.
+The user can compile the model themselves, provided they include the source code for the appropriate dependencies at compilation and link against the appropriate {\pkg} header library.
+{\pkg} provides several command line tools for determining the locations of the necessary libraries and any required compilation/linking flags.
+Alternatively, users can provide the location of the model source code and let one of several {\pkg} drivers handle the compilation, including linking against the appropriate {\pkg} interface library.
+{\pkg} also has support for compiling models using Make \citep{Stallman2004} and CMake \citep{Martin2006} for models that already have a Makefile or CMakeLists.txt.
+To use these tools to compile a model, lines are added to the recipe in order to allow linking against {\pkg}.
 
@@
 
-The algorithms for a generic model driver and the master thread can be found in Algorithms \ref{alg:model} \& \ref{alg:master} respectively in \A\ref{A:algorithms}.
+The algorithms for a generic model driver and the master thread are provided as Algorithms \ref{alg:model} \& \ref{alg:master} respectively in \A\ref{A:algorithms}.
 
@@
 In order for integrations to be possible, models must be able to send and receive information to and from other models written in different languages via some communication mechanism.
-Communication within {\pkg} integration networks was designed to be flexible in terms of the 
-languages, platforms, and data types available. To accomplish this, {\pkg} leverages 
-three different tools for communication, 
-System V IPC Queues (\S\ref{SSS:ipc}), ZeroMQ (\S\ref{SSS:zmq}), and RabbitMQ (\S\ref{SSS:rmq}), 
-which each have their own strengths and weaknesses. The 
-particular communication mechanism used by {\pkg} is determined by the platform, available libraries, and integration strategy.
+Communication within {\pkg} integration networks was designed to be flexible in terms of the languages, platforms, and data types available.
+To accomplish this, {\pkg} leverages three different tools for communication, System V IPC Queues (\S\ref{SSS:ipc}), ZeroMQ (\S\ref{SSS:zmq}), and RabbitMQ (\S\ref{SSS:rmq}), which each have their own strengths and weaknesses.
+The particular communication mechanism used by {\pkg} is determined by the platform, available libraries, and integration strategy.
 However, regardless of the communication mechanisms, the user will always use the same interface in the language of their model (See \S\ref{SSS:interface}).
@@
 %
-The first communication mechanism used by {\pkg} was System V interprocess 
-communication (IPC) message queues \citep{Rusling1999} on Posix (Linux and Mac OS X) systems. 
-IPC message queues allow messages to be passed between models running on separate 
-processes on the same machine. While IPC message queues are simple, fast 
-(See \S\ref{SS:results_commtype}), and are built into most Posix operating systems, they do not work in all 
-situations. IPC queues are not natively supported by Windows operating systems 
-and do not allow communication between remote processes. In addition, IPC queues also 
-have relatively low default message size limits on Mac OS X systems (2048 bytes). Once 
-the queue is full or if the message is larger than the limit, any process attempting to send an additional message will block until a 
-sufficient number of messages has been removed from the queue to accommodate the new message. 
+The first communication mechanism used by {\pkg} was System V interprocess communication (IPC) message queues \citep{Rusling1999} on Posix (Linux and Mac OS X) systems.
+IPC message queues allow messages to be passed between models running on separate processes on the same machine.
+While IPC message queues are simple, fast (See \S\ref{SS:results_commtype}), and are built into most Posix operating systems, they do not work in all situations.
+IPC queues are not natively supported by Windows operating systems and do not allow communication between remote processes.
+In addition, IPC queues also have relatively low default message size limits on Mac OS X systems (2048 bytes).
+Once the queue is full or if the message is larger than the limit, any process attempting to send an additional message will block until a sufficient number of messages has been removed from the queue to accommodate the new message.
 For messages larger than the limit, the sending process will block indefinitely.
-This can be handled by splitting large messages into multiple smaller messages 
-(See \S\ref{SSS:large}), however, the time required to send a message increases with the number of 
-message it must be broken into (See \S\ref{SS:results_commtype}). These limits make sending large messages 
-relatively inefficient when compared with other communications mechanisms. 
-As a result, IPC queues are used by 
-{\pkg} as a fallback on Posix systems if none of the other supported communication libraries have 
-not been installed.
+This can be handled by splitting large messages into multiple smaller messages (See \S\ref{SSS:large}), however, the time required to send a message increases with the number of message it must be broken into (See \S\ref{SS:results_commtype}).
+These limits make sending large messages relatively inefficient when compared with other communications mechanisms.
+As a result, IPC queues are used by {\pkg} as a fallback on Posix systems if none of the other supported communication libraries have not been installed.
 
@@
 %
-The preferred communication mechanism used by {\pkg} are ZeroMQ sockets 
-\citep[ZMQ][]{Akgul2013}. ZMQ 
-provides broker-less communication via a number of protocols and patterns with 
-bindings in a wide variety of languages that can be installed on Posix and Windows 
-operating systems. ZMQ was adopted by {\pkg} in order to allow support 
-on Windows and for future target languages (See \S\ref{SS:future}) that could not be 
-accomplished using IPC queues. In addition, while ZMQ allows 
-interprocess communication via IPC, ZMQ also 
-supports protocols for distributed communication via an Internet Protocol (IP) 
-network. While {\pkg} does not currently support using these protocols 
-for distributed integration networks, this is a future development plan that has been prepared for.
+The preferred communication mechanism used by {\pkg} are ZeroMQ sockets \citep[ZMQ][]{Akgul2013}.
+ZMQ provides broker-less communication via a number of protocols and patterns with bindings in a wide variety of languages that can be installed on Posix and Windows operating systems.
+ZMQ was adopted by {\pkg} in order to allow support on Windows and for future target languages (See \S\ref{SS:future}) that could not be accomplished using IPC queues.
+In addition, while ZMQ allows interprocess communication via IPC, ZMQ also supports protocols for distributed communication via an Internet Protocol (IP) network.
+While {\pkg} does not currently support using these protocols for distributed integration networks, this is an avenue of future development that has been prepared for.
 
@@
 This lack of confirmation makes it harder to determine if there is was error somewhere along the network.
-{\pkg} overcomes this by generating two ZMQ sockets for every connection: one for sending the messages and one for confirming them.
+{\pkg} overcomes this by generating two ZMQ sockets for every connection: one for passing the messages and one for confirming them.
 Each time a message is received, the receiving socket confirmation thread will send a confirmation including a unique ID for the received message and then wait for a reply to that confirmation.
@@
 %
-While broker-less communication like ZMQ is light weight and fast, it is not as fault 
-tolerant as brokered messaging systems that confirm message delivery and can resend 
-dropped messages. Resilience to dropped messages, while not as necessary for integrations 
-running entirely on a local machine, will be more important for integrations running on 
-distributed resources with less reliable connections. As a result, {\pkg} includes support 
-for brokered communication via RabbitMQ \citep[RMQ][]{RMQ} that will be used during future 
-development to allow integrations to run on distributed resources or include remote models 
-run as services (\S\ref{SS:distributed}). Due to the slower message speed, 
-{\pkg} does not currently use RMQ for communication unless explicitly specified by the user in their 
-integration network YAML.
+While broker-less communication like ZMQ is light weight and fast, it is not as fault tolerant as brokered messaging systems that confirm message delivery and can resend dropped messages.
+Resilience to dropped messages, while not as necessary for integrations running entirely on a local machine, will be more important for integrations running on distributed resources with less reliable connections.
+As a result, {\pkg} includes support for brokered communication via RabbitMQ \citep[RMQ][]{RMQ} that will be used during future development to allow integrations to run on distributed resources or include remote models run as services (\S\ref{SS:distributed}).
+Due to the slower message speed, {\pkg} does not currently use RMQ for communication unless explicitly specified by the user in their integration network YAML.
 
@@
 The same asynchronous communication strategy is used with each of the communication mechanisms supported by {\pkg}.
-Models do not block on sending messages to 
-output channels; the model is free to continue working on its task while 
-an \emph{output driver} waits for the message to be routed and received on a separate master process thread. Similarly, 
-an \emph{input driver} continuously pings input channels on another thread, moving received messages into a 
-buffer queue so that they are ready and waiting for the receiving model 
-when it asks for input. These drivers work together to move message along from one model to the next like a conveyor belt. 
+Models do not block on sending messages to output channels; the model is free to continue working on its task while an \emph{output driver} waits for the message to be routed and received on a separate master process thread.
+Similarly, an \emph{input driver} continuously pings input channels on another thread, moving received messages into a buffer queue so that they are ready and waiting for the receiving model when it asks for input.
+These drivers work together to move message along from one model to the next like a conveyor belt.
 As a result, models in complex integration networks are not affected by the rate at which dependent model consume their output and the speedup offered by running the models in parallel is improved.
@@
 %
-All of the communication tools leveraged by {\pkg} have intrinsic limits 
-on the allowed size for a single message. Some of these limits can be quite large 
-($2^{20}$ bytes for ZMQ and RMQ), while others are very limiting 
-(2048 bytes on Mac OS X for IPC queues). Although messages consisting of a few scalars are unlikely to 
-exceed these limits, biological inputs and outputs are often much more complex. 
-For example, structural data represented as a 3D mesh can easily exceed these 
-limits. To handle messages that are larger than the limit of the communication 
-mechanism being used, {\pkg} splits the message up into multiple smaller 
-messages. In addition, for large messages, {\pkg} creates new, temporary 
-communication channels that are used exclusively for a single message and 
-then destroyed. The address associated with the temporary channel is sent in 
-header information as a message on the main channel along with metadata about 
-the message that will be sent through the temporary channel like size and data type. 
-Temporary channels are used for large messages to prevent mistakenly combining 
-the pieces from two different large messages that were received at the same time 
-such as in the case that a model is receiving input from two different models 
-working in parallel.
+All of the communication tools leveraged by {\pkg} have intrinsic limits on the allowed size for a single message.
+Some of these limits can be quite large ($2^{20}$ bytes for ZMQ and RMQ), while others are very limiting (2048 bytes on Mac OS X for IPC queues).
+Although messages consisting of a few scalars are unlikely to exceed these limits, biological inputs and outputs are often much more complex.
+For example, structural data represented as a 3D mesh can easily exceed these limits.
+To handle messages that are larger than the limit of the communication mechanism being used, {\pkg} splits the message up into multiple smaller messages.
+In addition, for large messages, {\pkg} creates new, temporary communication channels that are used exclusively for a single message and then destroyed.
+The address associated with the temporary channel is sent in header information as a message on the main channel along with metadata about the message that will be sent through the temporary channel like size and data type.
+Temporary channels are used for large messages to prevent mistakenly combining the pieces from two different large messages that were received at the same time such as in the case that a model is receiving input from two different models working in parallel.
 The procedures for creating and using temporary channels for sending large messages can be seen in the {\sc Send} and {\sc Recv} methods from Algorithms \ref{alg:async_send} \& \ref{alg:async_recv} respectively in \A\ref{A:algorithms}.
@@
 %
-{\pkg} provides interface functions/classes for communication that are 
-written in each of the supported languages. Language specific interfaces 
-allow users to program in the 
-language(s) with which they are already familiar. The Python interface provides 
-communication classes for sending and receiving messages. The Matlab 
-interface provides a simple wrapper class for the Python class, that exposes 
-the appropriate methods and handles conversion between Python and Matlab 
-data types. The C interface provides structures and functions for accessing 
-communication channels and sending/receiving messages. The C++ interface 
-provides classes that wrap the C structures with functions called as methods.
-
-In addition to basic input and output, each interface also provides access 
-to more complex data types and communication patterns.
-Pseudocode describing the asynchronous communication classes and functions underlying the interfaces is provided by Algorithms \ref{alg:async_send} \& \ref{alg:async_recv} in \A\ref{A:algorithms}.
+{\pkg} provides interface functions/classes for communication that are written in each of the supported languages.
+Language specific interfaces allow users to program in the language(s) with which they are already familiar.
+Each language interface is an implementation of Algorithms \ref{alg:async_send} \& \ref{alg:async_recv} from \A\ref{A:algorithms} that provides users with send and receive calls for passing messages.
+The Python interface provides communication classes for sending and receiving messages.
+The Matlab interface provides a simple wrapper class for the Python class, that exposes the appropriate methods and handles conversion between Python and Matlab data types.
+The C interface provides structures and functions for accessing communication channels and sending/receiving messages.
+The C++ interface provides classes that wrap the C structures with functions called as methods.
+
+In addition to basic input and output, each interface also provides access to more complex data types and communication patterns that can be found in the {\pkg} documentation. 
+%Pseudocode describing the asynchronous communication classes and functions underlying the interfaces is provided by Algorithms \ref{alg:async_send} \& \ref{alg:async_recv} in \A\ref{A:algorithms}.
 
@@
 Differences between data formats can include, but are not limited to, type, precision, fields, or units.
-While some data formats are self-descriptive and include these types of information as meta data, this is not true of all data formats.
+While some data formats are self-descriptive and include these types of information as metadata, this is not true of all data formats.
 To combat this, {\pkg} requires models to explicitly specify the format of input and output expected by a model in the model YAML.
@@
 The shoot source code in Listing \ref{lst:src:shoot} is a standalone Python module containing a single function {\tt calc\_shoot\_mass} that calculates and returns the shoot mass at the next time step from input of the relative shoot growth rate ({\tt r\_s}), the time step ({\tt dt}, double), the shoot mass at the current time step ({\tt S\_t}), the root mass at the current time step ({\tt R\_t}), and the root mass at the next time step ({\tt R\_tp1}).
+%
+In addition to the actual calculation, both models include {\tt sleep} statements (lines 11 \& 18 in Listings \ref{lst:src:root} \& \ref{lst:src:shoot} respectively).
+These statements are meant to simulate a longer, more involved calculation representative of a more realistic model.
 
@@
 %
-For the example presented above, the speedup offered by {\pkg} through parallelism can be parameterized in terms of the durations of calculations for each model (designated model A \& B for brevity) performed in the different phases of the process (See Figure \ref{fig:empty_parallel}).
+For the example presented above of one model (model A) sending output to another (model B) and then waiting for a response, the speedup offered by {\pkg} through parallelism can be parameterized in terms of the durations of calculations for each model performed in the different phases of the process (See Figure \ref{fig:empty_parallel}).
 %
@@
 Jack2,][]{Magoules2018} requires explicit definition of the communication pattern in advance, prohibiting dynamic communication networks as discussed above.
-Work/data flow tools like Parsl \citep{babuji18} allow dynamic execution of processes along a dependency chain in parallel, but usually require addition code, are limited to passing data as files or data object native to the tool's implementation language, and are ill suited to models with persistent internal states or intermediate outputs.
-{\pkg} is unique in that it offers users asynchronous communication and parallelism that is flexible enough to support any communication pattern, implicitly handles conversion between native data types in the supported languages, and does not require any additional time investment on the part of the user.
+Work/data flow tools like Parsl \citep{babuji18} allow dynamic execution of processes along a dependency chain in parallel, but usually require addition code, are limited to passing data as files or data object native to the tool's implementation language, and are ill suited to models with persistent internal states or intermediate communication.
+{\pkg} is unique in that it offers users asynchronous communication and parallelism that is flexible enough to support any communication pattern, implicitly handles conversion between native data types in the supported languages, and does not require any additional time investment on the part of the user to achieve a speedup.
 	\item[No New Language] Unlike DSLs or language specific frameworks, {\pkg} doesn't require a model to be written in one specific language.
@@
 While {\pkg} is a powerful tool for integrating scientific models and opening up new avenues for collaboration, it is not applicable to every scenario.
+Before diving into model integration, potential users should determine if their use case falls under one of the following scenarios.
 %
@@
 Although the constituent models will still benefit from integration via {\pkg} in regards to reusability, {\pkg} may not be the best choice in such cases if execution time is a concern.
-	\item[Easily Translatable Models] For relatively simple models ($<100$ lines of code), it may be more efficient for a scientist to translate the model into the language of the model(s) it will be integrated with.
-{\pkg} is still useful as a tool more making the model accessible and more likely to be reused, but the performance advantage of running codes in the same language may be of greater importance in certain situations as discussed above.
-	\item[Unsupported Languages] While {\pkg} currently supports models written in Python, Matlab, C, \& C++, models are written in many other languages.
+Future work on automated creation of wrapper files for function-like (\S\ref{SS:control}) and ODE (\S\ref{SS:ODE}) models may allow {\pkg} to handle this automatically, replacing managed communication with in-language variable passing for models that are written in the same language and co-located on the same machine.
+	\item[Fast \& Tightly Coupled Models] For models involving quick calculations or for tightly couple sets of models such as might occur for two models representing coupled ODEs, the computational cost of rapid communication via {\pkg} can quickly surpass the cost of the calculations themselves causing the {\pkg} to be computationally inefficient in comparison with a manual integration.
+For example, if the sleep statements were removed from the example root and shoot models from \S\ref{S:example}, the calculations become trivial and take significantly less time than the communication between the models.
+Without the sleep statements, the root and shoot models run for 100 time steps in 0.007\,s and 0.095\,s respectively while the integration takes 7.437\,s.
+The speedup offered by {\pkg} is then 0.0137, which is actually a \emph{significant} slow down.
+For such models, large parameter searches or integrations with many time steps can be computationally prohibitive with {\pkg}.
+	\item[Unsupported Languages] While {\pkg} currently supports models written in Python, Matlab, C, \& C++, these languages are not an exhaustive list of those languages in which models are written.
 There are plans to expand support for {\pkg} to other languages (see \S\ref{SS:langsupport}), but in the meantime {\pkg} will not be particularly useful for models that are not written in one of the languages already supported.
+However, if a language can be called from one of the supported languages or vice versa \citep[e.g.
+f2py or SWIG,][]{Peterson2009,Beazley2003}, it may be possible to either write a wrapper or make calls to an interface in a supported language.
 \end{description}
@@
 % FUTURE WORK
-\subsection{Current/Future Improvements}\label{SS:future}
-%
-{\pkg} is being actively developed to expand the number of models 
-that can be used in integration networks and the complexity of integration 
-networks that can be executed. Several improvements are already in progress/planned for {\pkg}.
-\todo{expand?}
+\subsection{Ongoing/Future Improvements}\label{SS:future}
+%
+{\pkg} is being actively developed to expand the number of models that can be used in integration networks and the complexity of integration networks that can be executed.
+Several improvements are already in progress/planned for {\pkg}.
 
@@
 %
-The biggest barrier to running new models is language support. Plant models 
-are written in many different languages. While the current language support 
-covers many of the most popular languages among plant biologists, 
-additional languages must be added to unlock the full set of potential integrations. 
-Surveys of the plant modeling community have helped identify the core languages 
-that models have been written in. Based on these results, we plan to expand 
-{\pkg} to support models written in R, Fortran, Java, and the SBML \citep{Hucka2003} DSL.
-
-We also plan to add support for running Matlab models using Octave \citep{Eaton2002}. Matlab 
-models require a Matlab license to run. Given that plant modelers do not all 
-use Matlab, it cannot be assumed that everyone will have access to a Matlab 
-license. Octave is open source and provides much of the same functionality 
-as Matlab and can run many Matlab codes. Support for Octave will improve 
-modelers ability to collaborate without worrying about access to a Matlab 
-license.
+The biggest barrier to running new models is language support as plant models are written in many different languages.
+While the current language support covers many of the most popular languages among plant biologists, additional languages must be added to unlock the full set of potential integrations.
+Informal surveys of the plant modeling community during recent conferences and workshops have helped identify the core languages being used by model developers that would have the larges impact on the number of models accessible via {\pkg}.
+Based on these results, we plan to expand {\pkg} to support models written in R, Fortran, Java, and the SBML \citep{Hucka2003} DSL during the next major development push with support for additional models in subsequent releases as requested.
+
+We also plan to add support for running Matlab models using Octave \citep{Eaton2002}.
+Matlab models require a Matlab license to run.
+Given that plant modelers do not all use Matlab, it cannot be assumed that everyone will have access to a Matlab license.
+Octave is open source and provides much of the same functionality as Matlab and can run many Matlab codes.
+Support for Octave will improve modelers ability to collaborate without worrying about access to a Matlab license.
 
@@
 %
-High performance computing (HPC) and cloud compute resources are powerful 
-tools in the current computing ecosystem that could be used for running 
-complex integration networks. To this end, we plan to expand {\pkg} 
-support for running integration networks on distributed compute resources. 
-{\pkg} already uses communication tools that can be adapted for use 
-in a distributed pattern. We will leverage tools like the libsubmit package from the 
-Parsl project \citep{babuji18} for 
-automating the submission process to HPC and compute resources. In addition, 
-we will add tools for running models as a service and using RMQ to 
-permit access to these models within integration networks.
+High performance computing (HPC) and cloud compute resources are powerful tools in the current computing ecosystem that could be used for running complex integration networks.
+To this end, we plan to expand {\pkg} support for running integration networks on distributed compute resources.
+{\pkg} already uses communication tools that can be adapted for use in a distributed pattern \citep[ZMQ, RMQ][]{ZMQ, RMQ} and we will leverage tools like the libsubmit package from the Parsl project \citep{babuji18} for automating the submission process to HPC and compute resources.
+This work will also enable {\pkg} to adaptively spawning additional model processes on elastic compute resources when a model becomes a bottle neck to the integration process.
+This feature will be particularly useful for integrations simulating large ensembles of biological units (e.g.
+leaves on a plant or plants in a field) and will enable scientists to explore emergent properties of these systems.
+
+In addition, we will also add tools to {\pkg} for running models as a service and using RMQ to permit access to these models within integration networks.
+This feature will be useful for scientists who are unable or unwilling to share the source code for their model.
+Instead, these models can be hosted on dedicated compute resources managed by the model's home institution and collaborators interested in using the model in an integration can connect to it remotely, submitting requests and receiving results as needed.
 
 % CONTROL FLOW
-\subsubsection{Control Flow}\label{SS:control}
-%
-Currently, modelers must explicitly specify how a model should process input 
-within the model code including things like which input variables are static 
-and received once versus which variables change and should be looped over. 
-We plan on adding options to {\pkg} for dynamically generating model code 
-based a user provided function call and list of static/variable input channels. 
-This will allow better reusability of model code such as during parameter studies.
-Control flow will be part of the {\pkg} 1.0 release.
+\subsubsection{Wrapper Automation \& Control Flow}\label{SS:control}
+%
+Currently, modelers must explicitly decide how a model will be used when writing the model wrapper.
+These decisions can include things like which input and output variables are static versus variable, if there are sets of variables that should be updated at the same time, and how often output is generated, resulting in a hard coded use pattern.
+If someone wishing to use the model in another integration requires a different use patter, they will need to write their own wrapper.
+However, for models that have a functional form, the process of writing the model wrapper can be automated and done dynamically for each integration.
+We plan on adding options to {\pkg} for dynamically generating model wrappers based a user provided function call and list of static/variable input channels.
+This feature will improve the reusability of model code such as during parameter studies and decrease the amount of work required for integrating models that are written or can be exposed in a functional form.
+
+Dynamic generation of model wrappers also opens up possibilities for managing the flow of data through an integration network from the YAML.
+As part of the wrapper automation feature, we will add options for users to specify conditional execution of models via the YAML specification file.
+For example, a user may have two models for the same process that are valid under different conditions.
+This feature will allow users to declare this conditional in the YAML such that {\pkg} automatically switches between the two models as appropriate without the need to write an additional wrapper encoding the conditional.
+
+%Control flow and wrapper automation will be part of the {\pkg} 1.0 release.
 
@@
 scipy.integrate.ode][]{Jones2001, Hairer1993}.
-Using these tools, we will add features to {\pkg} that allow ODE models to be constructed directly from symbolic representation provided in YAML specification files.
+Using these tools, we will add features to {\pkg} that allow ODE models to be constructed directly from symbolic representations provided in YAML specification files.
+These features will allow users to quickly prototype new models and integrate old ones without the need to write a wrapper.
+
+Symbolic integration of ODE based models also has advantages for performance.
+Integrating models containing coupled ODEs is currently one of the hardest problems for {\pkg} to tackle due to the performance cost associated with rapid communications between models (see \S\ref{SS:limitations}).
+However, symbolic ODE representations can be used to represent two coupled ODE models as a single model which is much more computationally efficient as it does not require communication via {\pkg}.
+As such, ODE support for {\pkg} will also include automated coupling at runtime such that, while two ODE models can be declared separately and used independently in integrations, they will be combined into a single code when connected.
+This feature will greatly improve the performance of {\pkg} for integration of coupled ODEs.
 
 % DATA AGGREGATIOIN
-\subsubsection{Data Aggregation}\label{SS:dataagg}
-%
-Much of the original version of {\pkg} centered around the use of 
-tabular input data. While tabular data is used heavily by plant models, it 
-presents several barriers for constructing integration networks. Tabular input 
-data only works when one model outputs the same columns that are expected as 
-input by another model. However, this is unlikely to be the case if two models 
-are developed independently. Future improvements to {\pkg} will allow 
-tabular input to models to be composed by aggregating output data from more than 
-one model and/or file. Data aggregation will be part of the {\pkg} 1.0 release.
+\subsubsection{Data Aggregation \& Forking}\label{SS:dataagg}
+%
+Much of the original version of {\pkg} centered around the use of tabular or groups sets of input data where multiple inputs to a model are updated simultaneously such as during a loop.
+While this type of input is used heavily by plant models, it presents several barriers for constructing integration networks.
+Simultaneously updating multiple variables requires that either all of the variables are received from the same source (e.g.
+a row in a table with inputs taken from each column) or that the variables are updated simultaneously from multiple sources.
+The first case, while useful for input from a correctly formatted table, does not work for integration unless it is between two models where one model outputs exactly the same variables that are expected by the other.
+However, this is unlikely to be the case if two models are developed independently.
+The second case is also not ideal as it makes receiving and updating input variables tedious for more than a few variables as each receive call has to be checked independently for errors.
+
+Future improvements to {\pkg} will allow input and output variables to be grouped based on when they will be received or sent within a model while remaining independent connections internally to {\pkg}.
+In this way, grouped inputs/outputs can be composed of variables received/sent from /to different sources, but will be received/sent simultaneous by the model.
+This feature will reduce the number of send/receive calls used within a model wrapper, streamline the error checking process, and greatly simplify the model wrappers that user need to write.
+
+In addition to aggregation, this line of improvements will also add features to {\pkg} to allow connections to receive/send from/to multiple sources.
+Connection forking will enable users to duplicate data among several destinations, including files, allowing intermediate outputs to be recorded along the way.
+It will also allow single models to process input from multiple source models such as in the case of using multiple instances of a model to speed up performance of rate limiting integration step.
+%
+Data aggregation and forking will be part of the {\pkg} 1.0 release.
 
@@
 %
-While {\pkg} supports serialization of several data types/formats, we 
-would like to make serialization as flexible as possible. To this end, 
-{\pkg} will be adapted to read JSON schema \citep{jsonschema} for user defined types that 
-can be used to automatically create the appropriate data structures alongside
-methods for parsing and serializing them. JSON serialization will be part of the {\pkg} 1.0 release.
+{\pkg} currently supports serialization of several data types/formats, but support for new data types must be added manually and can prevent or slow down the integration of new models.
+Therefore, expanding the flexibility of {\pkg} serialization will allow a greater variety of models to be connected and is a priority for future development.
+{\pkg} will be adapted to read JSON schema \citep{jsonschema} for user defined types that can be used for automatically creating the appropriate data structures alongside methods for parsing and serializing them.
+
+In addition to using JSON schema for adding new data types, JSON schema will be developed for validating YAML files and existing data types will be transitioned to a JSON friendly representation.
+As there are many existing tools in different languages for encoding/decoding JSON \citep[e.g.
+FSON, jsonlite,][]{Levin2019,Ooms2014}, this will greatly decrease the amount of new development required to add support to {\pkg} for new languages.
+%
+JSON serialization will be part of the {\pkg} 1.0 release.
 
@@
 %
-In an effort to make {\pkg} as user friendly as possible, we have also begun work on a graphical user interface (GUI) for entering model information, composing integration networks from an existing palette of models, and displaying basic output from an integration run.
+In an effort to make {\pkg} as user friendly as possible, the Crops in Silico group at the University of Illinois at Urbana-Champaign has also begun work on a graphical user interface (GUI) for entering model information, composing integration networks from an existing palette of models, and displaying basic output from an integration run.
 The current prototype of the GUI can only handle model ingestion and network composition; model execution must be done locally.
-However, our ultimate goal with the GUI is to provide a service where users can register their model, compose and run integration networks on dedicated cloud compute resources, and view output in real time all from the web browser.
+However, the ultimate goal is to provide the GUI as a service where users can register their model, compose and run integration networks on dedicated cloud compute resources, and view output in real time all from the web browser.
+
 This GUI for model integration and execution will also greatly increase the accessibility of models as standalone scientific products.
@@
 Model developers could publish a Jupyter notebook exposing their model as a function via {\pkg} that can be used by those unfamiliar with modeling or the model's particular programming language to explore the model's behavior.
+
 This workflow would be particularly well suited to examining models during the peer review process and allowing on-the-fly parameter adjustment/estimation by experimentalists.
@@
 They can then interact with the model using either widgets or the full set of Python libraries.
+%
 Model execution as Python functions will be part of the {\pkg} 1.0 release.
@@
 Full documentation for {\pkg} can be found \href{https://cropsinsilico.github.io/cis_interface/}{here} including step-by-step directions from the tutorial conducted during the \href{https://cropsinsilico.github.io/cis_interface/hackathon2018/index.html}{2018 Crops in Silico Hackathon}.
+
+\todo{comment about 1.0 release}
 

### 2019-01-05 papers/lang2018/referee_comments.tex
Updated times and speedup for root/shoot example and added response to reviewer 3

@@
 \usepackage{epstopdf}
+\usepackage{color}
 \DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
@@
 
+\newcommand{\todo}[1]{{\color{red}{#1}}}
 \newcommand{\pkg}{{\tt cis\_interface}{}}
 \usepackage{enumitem,amssymb}
+\usepackage{xparse}
+\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
+\let\endchangemargin=\endlist
+
 \newlist{todolist}{itemize}{2}
@@
 \usepackage{pifont}
+\usepackage{calc}
 \newcommand{\cmark}{\ding{51}}%
 \newcommand{\xmark}{\ding{55}}%
-\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
+\newcommand{\done}{\makebox[0pt][l]{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
 \hspace{-2.5pt}}
+%\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
+%\hspace{-2.5pt}}
 \newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}
+\newcommand\NoIndent[1]{%
+  \begingroup
+  \par
+  \parshape0
+  #1\par
+  \endgroup
+}
+
+%\settowidth{\donelength}{\done}
+\newlist{comments}{itemize}{2}
+\setlist[comments]{rightmargin=\leftmargin, labelsep=!,%\leftmargin, %
+labelwidth=\widthof{$\square$},labelindent=0em,itemindent=0em,parsep=0.5\topsep,itemsep=\topsep}
+%font=\bfseries,leftmargin=4em,labelwidth=4em,labelsep=0pt,label=$\square$,labelindent=-4em}
+\NewDocumentCommand\comment{O{$\square$}m+o}{%
+    \item[\hspace{-\leftmargin}#1] #2 \IfValueT{#3}{
+   	%\begin{changemargin}{-\leftmargin}{-\leftmargin}
+    	%{#3}
+    	%\end{changemargin}
+                \NoIndent{#3}
+    		%\vspace{0.25em}
+    }
+}
+
+\newcommand{\Comment}[3][$\square$]{%
+%
+\begin{quote}
+	\makebox[0pt][l]{#1}#2
+\end{quote}
+#3
+}
+\newcommand{\Response}[2][$\square$]{%
+	#1 #2
+}
 
@@
 
-\subsection{Comments to the Author}
-
-This is timely and a really nice piece of work.
+We would like to thank the reviewer for their comments as we believe they helped make the manuscript much stronger and a better representation of the work being described.
+We will address the reviewer's comments in the order they were raised:
+%
+\begin{comments}
+%%%%%
+\comment[]{This is timely and a really nice piece of work.
 This Cis-interface has the potential to become a major tool to support the future development of the plant systems model integration.
-I have a few specific comments on this and would like the author to address.
-\begin{todolist}
-\item First, it would be great to have a more clear description of the novelty of the pipeline developed.
+I have a few specific comments on this and would like the author to address.}[]
+%%%%%
+\comment[\done]{First, it would be great to have a more clear description of the novelty of the pipeline developed.
 Is it mainly a compilation of existing tools, in that case, this needs to be clearly indicated.
-If not, then a clear statement in this might be great.
-\item[\done] Secondly, the work will benefit a lot from a more explicit description of a user case. For example, clearly write out the exact commands needed to finish one round of code integration. 
-\item[\done] Thirdly, I would be keen to show some performance data of integration of different modules. No data are provided on this at this stage. 
-\item Fourthly, the main purpose of this paper is to describe a platform which can help plant modelers of different research areas and using programming languages work together more effectively and easily, it would be great if author can explain the advantage of this new software from the perspective of users.
-\item Fifthly,  many of the software engineering methodologies described in the paper are widely used in the field of software engineering.
-It would be better to provide some more concrete description of how this particular tool help users.
-\item Finally, it would be helpful to provide a flow chart and pseudo-code of the core algorithms used to link different modules.
-\end{todolist}
-
-\subsection{Response from the Author}
-
-We would like to thank the reviewer for their comments as we believe they helped make the paper much stronger and to better represent the work being described.
-We will address the reviewer's comments in the order they were raised:
-\begin{enumerate}
-\item ?
-\item We whole heartedly agree with the reviewer that the manuscript greatly benefits from a concrete use case.
+If not, then a clear statement in this might be great.}[%
+%
+While {\pkg} does make use of several existing tools, it uses them in novel ways and implements several new tools of its own.
+We have added text to the methods introduction and methods subsections, clarifying the new features that are provided by {\pkg}.
+In addition, we have added a new ``Advantages \& Limitations" subsection to the Summary that explicitly describes the features provided by {\pkg} that are not available in other existing packages.]
+%%%%%
+\comment[\done]{Secondly, the work will benefit a lot from a more explicit description of a user case.
+For example, clearly write out the exact commands needed to finish one round of code integration.}[%
+%
+We whole heartedly agree with the reviewer that the manuscript greatly benefits from a concrete use case.
 As such, we have added a new section ``Worked Example" after the methods section that walks through, step-by-step, integration of the example models presented in the introduction.
-The code from the example is available on the paper branch of the {\pkg} GitHub repository and will be published as part of the documentation during the 1.0 release.
-\item We believe that the reviewer is referring to performance data on integration of specific biological \emph{models}, but would appreciate some clarification if this is not the case.
-While we would like to show concrete examples of the performance of some integrations done with {\pkg} using actual models, most of the models that {\pkg} has been used with are still in the pre-publication stage.
+The code from the example is available on the paper branch of the {\pkg} GitHub repository and will be published as part of the documentation during the 1.0 release.]
+%%%%%
+\comment[\done]{Thirdly, I would be keen to show some performance data of integration of different modules. No data are provided on this at this stage.}[%
+%
+We believe that the reviewer is referring to performance data on integration of specific biological \emph{models}, but would appreciate some clarification if this is not the case.
+While we would like to show concrete real-world examples of the performance of some integrations done with {\pkg} using actual models, most of the models that {\pkg} has been used with are still in the pre-publication stage.
 As the code for the models is not yet public, these performance tests would not be informative to those without access.
-In addition, the performance increase due to parallelism is highly dependent on how coupled the models are and how many models are part of the integration.
+In addition, the performance increase due to parallelism and asynchronous communication is highly dependent on how coupled the models are and how many models are part of the integration.
 If two models are integrated that are tightly coupled, one model will always be waiting for the other to complete its calculation and there not be a speedup over scripted integration of the two models via an intermediary file.
@@
 The speedup calculation is even more complex when more than two models are involved.
-To reflect this complex relationship, we have added a subsection to the end of the results section entitled ``Speedup from Parallels" that discusses this and present an equation for calculating the speedup for an integration of two models given the appropriate parameters.
-\item ?
-\item ?
-\item ?
-\end{enumerate}
+To reflect this complex relationship, we have added a subsection to the end of the results section entitled ``Speedup from Parallelism" that discusses this and present equations for calculating the speedup for an integration of two example models given the appropriate parameters.]
+%%%%%
+\comment[\done]{Fourthly, the main purpose of this paper is to describe a platform which can help plant modelers of different research areas and using programming languages work together more effectively and easily, it would be great if author can explain the advantage of this new software from the perspective of users.}[%
+%
+We have expanded the ``CurrentStatus" subsection in the Summary section to more explicitly describe the utility of {\pkg} to domain scientists, including its ability to promote model reuse, acessiblity, and collaboration.]
+%%%%%
+\comment[\done]{Fifthly,  many of the software engineering methodologies described in the paper are widely used in the field of software engineering.
+It would be better to provide some more concrete description of how this particular tool help users.}[%
+%
+The new ``Advantages \& Limitations" subsection in the Summary describes the advantages of {\pkg} over existing software packages targeted at similar problems.
+This is in addition to updated methods sections that are now more explicit in their description of features which are unique to {\pkg} and a worked example.]
+%%%%%
+\comment[\done]{Finally, it would be helpful to provide a flow chart and pseudo-code of the core algorithms used to link different modules.}[%
+%
+Although not exhaustive in their representation of the algorithms, we have added pseudo-code for the core algorithms in an Appendix that are intended to provide the reader with a general picture of the processes involved.]
+\end{comments}
 
@@
 
-\subsection{Comments to the Author}
-This is a very well written manuscript that is accessible to people from different backgrounds. Few minor comments:
-\begin{todolist}
-\item[\done] Background: Pg 4, L8-18: The first paragraph contains broad, bold statements with no citations as examples.
+We would like to thank the reviewer for their comments as we the manuscript is stronger as a result. We have responded to the reviewer's comments in the order they were posed.
+%
+\begin{comments}
+%%%%%
+\comment[]{This is a very well written manuscript that is accessible to people from different backgrounds. Few minor comments:}[]
+%%%%%
+\comment[\done]{Background: Pg 4, L8-18: The first paragraph contains broad, bold statements with no citations as examples.
 To move away from a largely anecdotal paragraph, examples from the literature should be given.
-It would also help to cite the Crops in silico perspective article in Frontiers (https://www.frontiersin.org/articles/10.3389/fpls.2017.00786/full).
-\item[\done] P4, L35-38: give examples here too. What solutions?
-\item[\done] P6 line 17: a to "as"
-\item[\done] P7 line 9: remove extra "tools"
-\item[\done] In the conclusions, it would be helpful to have a few lines suggesting uses or applications of the Cis interface.
+It would also help to cite the Crops in silico perspective article in Frontiers (https://www.frontiersin.org/articles/10.3389/fpls.2017.00786/full).}[%
+%
+We agree with the reviewer that the paragraph in question was vague and have added several references to examples from the literature, including the Crops in silico perspective article.]
+%%%%%
+\comment[\done]{P4, L35-38: give examples here too. What solutions?}[%
+%
+The sentence in question "Within both computational biology and the larger scientific computing community, groups have developed around different solutions to the problem of connecting models." was the last sentence in the introductory part of the background section and was meant to refer to the solutions described from the literature in the following subsections.
+We understand the confusion and have clarified this by explicitly listing the solutions that are described in the subsequent subsections.]
+%%%%%
+\comment[\done]{P6 line 17: a to "as"}[%
+%
+This typo has been corrected.
+]
+%%%%%
+\comment[\done]{P7 line 9: remove extra "tools"}[%
+%
+This typo has been corrected.
+]
+%%%%%
+\comment[\done]{In the conclusions, it would be helpful to have a few lines suggesting uses or applications of the Cis interface.
 For example, researchers could use cis{\_}interface as a user-friendly way to run model code either from the published literature or when reviewing manuscripts.
-Likewise, experimentalists can update parameter files based on in-house experimental data and then run the models, even if they have no prior expertise in modeling.
-\end{todolist}
-
-\subsection{Response from the Author}
-
-We would like to thank the reviewer for their comments as we the manuscript is stronger as a result. We have responded to the reviewer's comments in the order they were posed.
-\begin{enumerate}
-\item We agree with the reviewer that the paragraph in question was vague and have added several references to examples from the literature and including the Crops in silico perspective article.
-\item The sentence in question "Within both computational biology and the larger scientific computing community, groups have developed around different solutions to the problem of connecting models." was the last sentence in the introductory part of the background section and was meant to refer to the solutions described form the literature in the following subsections.
-We have clarified this by explicitly listing the solutions that are described in those sections.
-\item We have added several sentences to the conclusions discussing the use cases for {\pkg} including as a tool for integrating models and as a tool for executing models without knowledge of the model's programming language.
-While we agree with the reviewer that {\pkg} could be used for peer review and collaboration by exposing models via a standard interface, we believe this application will be most powerful when used with the model integration GUI (which is separate form the {\pkg} package itself) or by calling models as functions from Jupyter notebooks (which is a feature that was added after submitting the manuscript and will be released in verison 1.0 of {\pkg}).
-As such, we have added a subsection for "Models as Functions" in the "Current/Future Improvements" section and discuss this application further there and in the "Graphical User Interface" subsection.
-\end{enumerate}
+Likewise, experimentalists can update parameter files based on in-house experimental data and then run the models, even if they have no prior expertise in modeling.}[%
+%
+We have added several sentences to the conclusions discussing the use cases for {\pkg} including as a tool for integrating models and as a tool for executing models without knowledge of the model's programming language.
+While we agree with the reviewer that {\pkg} could be used for peer review and collaboration by exposing models via a standard interface, we believe this application of {\pkg} will be most powerful when used with the model integration GUI (which is separate form the {\pkg} package itself) or by calling models as functions from Jupyter notebooks (which is a feature that was added after submitting the manuscript and will be released in verison 1.0 of {\pkg}).
+As such, we have added a subsection for ``Models as Functions" in the ``Ongoing/Future Improvements" section and discuss this application further there and in the ``Graphical User Interface" subsection.
+]
+\end{comments}
 
@@
 
-\subsection{Comments to the Author}
-This manuscript reports on efforts to create a universal interface toolbox for integrating computational models across languages and scales.
+We would like to thank the reviewer for their comments as we believe they helped make the manuscript much stronger and a better representation of the work being described.
+We will address the reviewer's comments in the order they were raised:
+%
+\begin{comments}
+%%%%%
+\comment[]{This manuscript reports on efforts to create a universal interface toolbox for integrating computational models across languages and scales.
 This topic is without doubt one of the key topics in further development and application of existing and to-be-developed plant models.
@@
 effort in developing an open source Python package cis{\_}interface, which can be installed on 3 different operating systems and support integration of models written in 4 different programming languages.
-I think the paper has great potential but I do have some comments:
-\begin{todolist}
-\item The author had a lengthy discussion on comparison of communication time cost for model integration with different communication mechanisms, programming languages, Python versions and operating systems, actually this is also the main result of current manuscript.
+I think the paper has great potential but I do have some comments:}[]
+%%%%%
+\comment[\done]{The author had a lengthy discussion on comparison of communication time cost for model integration with different communication mechanisms, programming languages, Python versions and operating systems, actually this is also the main result of current manuscript.
 However, without comparison with existed tools, I think this part should be greatly condensed.
@@
 Moreover, how to perform integration of input for one model from different models?
-For example, canopy model only gives 3D mesh data with light level on each facet, however, metabolic model needs both light level and protein levels (calculated from another model) for each facet, how to correctly integrate the protein data and the light data?
-\item I expect more detailed biological modeling examples to be provided, from which the readers are able to know how to use the cis{\_}interface and what?s the advantage of it.
-I also expect the author to suggest some general modeling principles from following which the newly developed model can easily be integrated with other models.
-\end{todolist}
+For example, canopy model only gives 3D mesh data with light level on each facet, however, metabolic model needs both light level and protein levels (calculated from another model) for each facet, how to correctly integrate the protein data and the light data?}[%
+%
+We thank the reviewer for this comment as it has pointed out several areas in which the manuscript was previously lacking and major improvements have been made as a result.
+First, we have added a subsection entitled ``Speedup from Parallelism" to the end of the ``Results" section that discusses the speedup achievable using {\pkg} for a two model integration based on model properties and the presented performance results and we have added a subsection entitled ``Advantages \& Limitations" to the ``Summary" which enumerates the advantages of using {\pkg} over other existing tools targeted at similar problems.
 
-\subsubsection{Minor comments}
-\begin{todolist}
-\item[\done] Page 7 line 9: remove redundant ?tools?
-\end{todolist}
+Second, we have added a section ``Worked Example" which walks through each step required for an integration of the two example models presented in the introduction.
+All of the code and YAML files required for this integration are included in the text, are available on the {\tt paper} branch of the {\pkg} GitHub repository, and will be released with version 1.0 of the code.
+While we agree that a more complex worked example would be more useful to users who are already familiar with the code, we believe that such an example is outside the scope of this paper which serves as a introduction to the goals, methods, and performance of {\pkg}.
+There are several more complex examples provided online in the {\pkg} documentation (including an example from the 2018 Crops in Silico hackathon using toy biological models), which we now reference throughout the paper such that readers can refer to them.
 
-\subsection{Response from the Author}
+Third, many of the topics that the reviewer explicitly mentions (time steps, complex data conversion, data aggregation from multiple sources) are problems that are not completely solved in the current version of {\pkg}, but are solved or ameliorated by ongoing work (discussed in the ``Ongoing/Future Improvements" subsection of the ``Summary" section) that will be released in version 1.0.
+To address these concerns we have dedicated the last subsection of the ``Worked Example" section to ``Advanced Integration Topics" which discusses these use cases, references examples in the documentation, and points to the appropriate topics in the ``Ongoing/Future Improvements" subsection.
+In addition, we have expanded the ``Ongoing/Future Improvements" subsection to provide additional details and also identify those improvements which will be released in version 1.0.
+Those features not yet supported by {\pkg} are also discussed in the new ``Advantages \& Limitations" subsection.]
+%%%%%
+\comment[\done]{I expect more detailed biological modeling examples to be provided, from which the readers are able to know how to use the cis{\_}interface and what's the advantage of it.
+I also expect the author to suggest some general modeling principles from following which the newly developed model can easily be integrated with other models.}[%
+%
+As discussed in our response to the reviewers previous comment, we have added a new ``Worked Example" section which provides a detailed walk through of integrating two example models.
+Throughout this section we provide tips on creating/preparing models such that they can be easily integrated using {\pkg} and follow best practices for reproducibility.
+Although we would also like to include real examples from the integration network shown in the ``Summary" section, most of the models in that network are in the pre-publication stage and their source code cannot be released at this time.
+Although not all biological models, we do reference several more complex examples from the {\pkg} documentation throughout the text.
+In addition, we have expanded the summary to include a more in-depth discussion of use cases and the ``Advantages \& Limitations" subsection discussing the advantages of {\pkg} over other solutions.]
+%%%%%
+\comment[\done]{Page 7 line 9: remove redundant `tools'}[
+%
+This typo has been corrected.]
+\end{comments}
 

### 2019-01-05 cis_interface/examples/root_to_shoot/root_to_shoot.sh

@@ -2,16 +2,32 @@
 
 # ----------------Your Commands------------------- #
 case $1 in
+    --root-iso )
+	echo "Running Root Model w/o Interface"
+	gcc -o root_isolated src/root_isolated.c
+	# time for i in {1..100}; 
+	# do
+	time ./root_isolated ./Input/root_growth_rate.txt ./Input/init_root_mass.txt ./Input/timesteps.txt ./Output/root_output.txt
+	# done
+	rm root_isolated
+	;;
+    --shoot-iso )
+	echo "Running Shoot Model w/o Interface"
+	# time for i in {1..100}; 
+	# do
+	time python src/shoot_isolated.py ./Input/shoot_growth_rate.txt ./Input/init_shoot_mass.txt ./Input/timesteps.txt ./Input/root_output.txt ./Output/shoot_output.txt
+	# done
+	;;
     --root )
 	echo "Running Root Model"
-	cisrun root.yml root_files.yml
+	time cisrun root.yml root_files.yml
 	;;
     --shoot )
 	echo "Running Shoot Model"
-	cisrun shoot.yml shoot_files.yml
+	time cisrun shoot.yml shoot_files.yml
 	;;
     * )
 	echo "Running Integration"
-	cisrun root.yml shoot.yml root_to_shoot.yml
+	time cisrun root.yml shoot.yml root_to_shoot.yml
 	;;
 esac

### 2019-01-05 cis_interface/examples/root_to_shoot/src/root.h

@@ -1,3 +1,13 @@
+#include <unistd.h>
+
+/*!
+  @brief Calculation of root mass following time step.
+  @param[in] r_r double Relative root growth rate.
+  @param[in] dt double Duration of time step.
+  @param[in] R_t double Current root mass.
+  @returns double Root mass after time step.
+ */
 double calc_root_mass(double r_r, double dt, double R_t) {
+  usleep(1e5);  // To simulate a longer calculation
   return R_t + (R_t * r_r * dt);
 };

### 2019-01-05 cis_interface/examples/root_to_shoot/src/root_isolated.c

@@ -1,0 +1,127 @@
+#include <stdio.h>
+#include <stdlib.h>
+// Include C header containing model calculation
+#include "root.h"
+#define LINESZ 1024
+
+
+int main(int argc, char *argv[]) {
+  int i, ret;
+  double r_r, dt, R_t, R_tp1;
+  char buff[LINESZ];
+
+  // Create input/output files
+  if (argc!=5) {
+    printf("3 input files and 1 output file must be specified.\n");
+    return -1;
+  }
+  FILE *RootGrowthRate = fopen(argv[1], "r");
+  if (RootGrowthRate == NULL) {
+    printf("root: Failed to open file: '%s'\n", argv[1]);
+    return -1;
+  }
+  FILE *InitRootMass = fopen(argv[2], "r");
+  if (InitRootMass == NULL) {
+    printf("root: Failed to open file: '%s'\n", argv[2]);
+    fclose(RootGrowthRate);
+    return -1;
+  }
+  FILE *TimeStep = fopen(argv[3], "r");
+  if (TimeStep == NULL) {
+    printf("root: Failed to open file: '%s'\n", argv[3]);
+    fclose(RootGrowthRate);
+    fclose(InitRootMass);
+    return -1;
+  }
+  FILE *NextRootMass = fopen(argv[4], "w");
+  if (NextRootMass == NULL) {
+    printf("root: Failed to open file: '%s'\n", argv[4]);
+    fclose(RootGrowthRate);
+    fclose(InitRootMass);
+    fclose(TimeStep);
+    return -1;
+  }
+
+  // Read root growth rate
+  ret = -1;
+  while (fgets (buff, LINESZ, RootGrowthRate)) {
+    if (buff[0] == '#') continue;
+    ret = sscanf(buff, "%lf", &r_r);
+    break;
+  }
+  fclose(RootGrowthRate);
+  if (ret != 1) {
+    printf("root: Error reading root growth rate.\n");
+    fclose(InitRootMass);
+    fclose(TimeStep);
+    fclose(NextRootMass);
+    return -1;
+  }
+  printf("root: Read root growth rate: %lf\n", r_r);
+
+  // Read initial root mass
+  ret = -1;
+  while (fgets (buff, LINESZ, InitRootMass)) {
+    if (buff[0] == '#') continue;
+    ret = sscanf(buff, "%lf", &R_t);
+    break;
+  }
+  fclose(InitRootMass);
+  if (ret != 1) {
+    printf("root: Error reading initial root mass.\n");
[…]

### 2019-01-05 cis_interface/examples/root_to_shoot/src/shoot.py

@@ -1,3 +1,6 @@
+import time
+
+
 def calc_shoot_mass(r_s, dt, S_t, R_t, R_tp1):
     r"""Calculate the shoot mass.
 
@@ -12,4 +15,5 @@ def calc_shoot_mass(r_s, dt, S_t, R_t, R_tp1):
         float: Shoot mass at the next timestep.
 
     """
+    time.sleep(0.1)  # To simulate a longer calculation
     return (S_t * r_s * dt) + S_t - (R_tp1 - R_t)

### 2019-01-05 cis_interface/examples/root_to_shoot/src/shoot_wrapper.py

@@ -7,7 +7,6 @@ from shoot import calc_shoot_mass
 # Create input/output channels
 ShootGrowthRate = CisInput('shoot_growth_rate')
 InitShootMass = CisInput('init_shoot_mass')
-InitRootMass = CisInput('init_root_mass_shoot')
 TimeStep = CisInput('shoot_time_step')
 NextRootMass = CisInput('next_root_mass')
 NextShootMass = CisOutput('next_shoot_mass', '%lf\n')
@@ -27,7 +26,7 @@ S_t = input[0]
 print('shoot: Received initial shoot mass: %f' % S_t)
 
 # Receive inital root mass
-flag, input = InitRootMass.recv()
+flag, input = NextRootMass.recv()
 if not flag:
     raise RuntimeError('shoot: Error receiving initial root mass.')
 R_t = input[0]

### 2019-01-21 papers/lang2018/ms.tex
Matts feedback

@@
 \newcommand{\TB}{\,Tbytes}
-\newcommand{\pkg}{{\tt \todo{cis\_interface}}{}}
+\newcommand{\pkg}{{\tt yggdrasil}{}}
+\newcommand{\cis}{{\tt cis\_interface}{}}
+\newcommand{\pkglink}{\todo{cis\_interface}}
+%\newcommand{\pkg}{{\tt \todo{cis\_interface}}{}}
 \newcommand{\pkgrun}{{\tt \todo{cisrun}}}
 \newcommand{\A}{Appendix{ }}
+\newcommand{\hrefgit}[1]{\href{https://github.com/cropsinsilico/cis_interface}{\todo{#1}}}
+\newcommand{\hrefdoc}[1]{\href{https://cropsinsilico.github.io/cis_interface/}{\todo{#1}}}
+\newcommand{\hrefhack}[1]{\href{https://cropsinsilico.github.io/cis_interface/hackathon2018/index.html}{\todo{#1}}}
 
@@
 \begin{abstract}
-Thousands of computational models have been created within the plant biology community and broader scientific communities in the past two decades that have the potential to be combined into complex integration networks capable of capturing more complex biological processes than possible with isolated models.
+Thousands of computational models have been created within both the plant biology community and broader scientific communities in the past two decades that have the potential to be combined into complex integration networks capable of capturing more complex biological processes than possible with isolated models.
 However, the technological barriers introduced by differences in language and data formats has slowed this progress.
-We present \href{https://github.com/cropsinsilico/cis_interface}{\pkg}, a Python package for running integration networks with connections between models across languages and scales.
+We present \hrefgit{\pkg} (previously {\cis}), a Python package for running integration networks with connections between models across languages and scales.
 {\pkg} coordinates parallel execution of models in Python, C, C++, and Matlab on Linux, Mac OS, and Windows operating systems, and handles communication in a number of data formats common to computational plant modeling.
-{\pkg} is designed to be user-friendly and can be accessed at {\tt https://github.com/cropsinsilico/cis\_interface}.
+{\pkg} is designed to be user-friendly and can be accessed at {\tt https://github.com/cropsinsilico/\pkglink}.
 Although originally developed for plant models, {\pkg} can be used to connect computational models from any domain.
@@
 
-Consider the task of integrating a root growth model with a shoot growth model. The root growth model is written in C (see Listing \ref{lst:src:root}) and can be expressed as
+Consider the task of integrating a simple root growth model with a shoot growth model. Our example root growth model is written in C (see Listing \ref{lst:src:root}) and can be expressed as
 \begin{equation}
@@
 \begin{itemize}
-	\item $R_{t}$: the root mass at time $t$
-	\item $r_{r}$: the relative root growth rate
-	\item $dt$: the time step
+	\item $R_{t}$: the root mass at time $t$ (g)
+	\item $r_{r}$: the relative root growth rate (hr$^{-1}$)
+	\item $dt$: the time step (hr)
 \end{itemize}
@@
 \begin{itemize}
-	\item $S_{t}$: the shoot mass at time $t$
-	\item $r_{s}$: the relative shoot growth rate
-	\item $dt$: the time step
-	\item $R_{t}$: the root mass at time $t$ 
-	\item $R_{t+1}$: the root mass at time $t+1$ 
+	\item $S_{t}$: the shoot mass at time $t$ (kg)
+	\item $r_{s}$: the relative shoot growth rate (d$^{-1}$)
+	\item $dt$: the time step (d)
+	\item $R_{t}$: the root mass at time $t$ (kg)
+	\item $R_{t+1}$: the root mass at time $t+1$ (kg)
 \end{itemize}
@@
 For a tightly coupled set of models such as these where one model depends directly on the result from the other at the current time step, manual integration is very inefficient for more than a handful of time steps even when done using a script.
-Although translating one or both of these simple toy models would be relatively straight forward, actual plant models are much more complex and could contain thousands of lines of code that would be time consuming to translate and result in unnecessary replication of model algorithms.
+Although translating one or both of these simple toy models would be relatively straight forward, actual plant models are much more complex and could contain thousands of lines of code that would be time consuming to translate and result in unnecessary duplication of model algorithms.
 
@@
 {\pkg} was developed as part of the Crops \emph{in Silico} \citep{Marshall-Colon2017} initiative to build a complete crop \emph{in silico} from the level of the genes to the level of the field.
-{\pkg} is available on Github\footnote{{\tt https://github.com/cropsinsilico/cis\_interface}} with full documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/}} and can be installed from PyPI\footnote{{\tt https://pypi.org/project/cis\_interface/}} via {\tt pip install {\pkg}} or conda-forge\footnote{{\tt https://github.com/conda-forge/cis\_interface-feedstock}} via {\tt conda -c conda-forge install {\pkg}}.
+{\pkg} is available on Github\footnote{{\tt https://github.com/cropsinsilico/\pkglink}} with full documentation\footnote{{\tt https://cropsinsilico.github.io/\pkglink/}} and can be installed from PyPI\footnote{{\tt https://pypi.org/project/\pkglink/}} via {\tt pip install {\pkg}} or conda-forge\footnote{{\tt https://github.com/conda-forge/\pkglink-feedstock}} via {\tt conda -c conda-forge install {\pkg}}.
 Although developed for the purpose of integrating plant biology models, {\pkg} can be applied to any situation requiring coordination between models in the supported languages.
@@
 running the root model, converting the output root mass into the units expected by the shoot model, and then running the shoot model).
-However, manual integration is 1) computationally inefficient when many iterations are necessary, such as during parameter searches or steady state convergence tests, 2) unique to every integration, requiring the production of new scripts, and 3) complex when more than three models are being integrated.
+However, manual integration is 1) computationally inefficient when many iterations are necessary, such as during parameter searches or steady state convergence tests, 2) unique to every integration, requiring the production of new scripts, 3) time consuming, 4) prone to error, and 5) complex when more than three models are being integrated.
 Within both computational biology and the larger scientific computing community, groups have developed around different solutions to the problem of connecting models.
@@
 Although written in Python, {\pkg} was written such that users need not have any knowledge of the Python language and can interact with {\pkg} solely through the CLI.
+In addition to the CLI, more advanced capabilities are also exposed via the Python API including access to more detailed information about the running integration.
 %At execution, {\pkg} launches models on new processes based on information contained in a YAML specification file (See \S\ref{SSS:yaml} below) and establishes
@@
 
-Model YAMLs include information about the location of the model source code, the language the model is written in, how the model should be run, and any input or output variables including their data type and units.
+Model YAMLs include information about the location of the model source code, the language the model is written in, how the model should be run, and any input or output variables including their data type (e.g.
+array, scalar, mesh) and physical units.
 %
@@
 %
-{\pkg} launches each model in an integration on its own process, allowing models to complete independent operations in parallel and complete tasks more quickly.
+{\pkg} launches each model in an integration in its own process, allowing models to complete independent operations in parallel and complete tasks more quickly.
 For example, the root and shoot growth models from \S\ref{S:intro} (described further in \S\ref{S:example}) require 10.28\,s and 10.40\,s respectively to run for 100 time steps in isolation with direct input/output from/to files in their native language.
 If these two models were manually integrated in serial via the command line, the integration would require a total of 20.68\,s.
-However, the same integration requires only 16.03\,s using {\pkg}, a speed-up of 1.29.
+However, because {\pkg} offers parallel execution of the models, the same integration requires only 16.03\,s when using {\pkg}, a speed-up of 1.29.
 The exact speedup provided by parallel integration using {\pkg} depends on how independent the models are and how equally distributed the work is between the models.
@@
 
-While every model is executed on a new process, how the model is handled depends on the language it is written in.
+While every model is executed in a new process, how the model is handled depends on the language it is written in.
 {\pkg} has a dedicated driver for each of the supported languages with utilities that allow {\pkg} to launch and monitor executables written in that language from Python.
 Models written in interpreted languages (Python and Matlab) are executed on the command line via the interpreter.
-In the case of Matlab, where a significant amount of time is required to start the Matlab interpreter (See \S\ref{SS:results_language}), Matlab shared engines are used to execute Matlab models.
+In the case of Matlab, where a significant amount of time is required to start the Matlab interpreter (see \S\ref{SS:results_language}), Matlab shared engines are used to execute Matlab models.
 Matlab shared engines are Matlab instances that other processes can submit Matlab code to for execution.
@@
 %
-In order for integrations to be possible, models must be able to send and receive information to and from other models written in different languages via some communication mechanism.
+Integrating models requires that they are able to send and receive information to and from other models written in different languages via some communication mechanism.
 Communication within {\pkg} integration networks was designed to be flexible in terms of the languages, platforms, and data types available.
@@
 The particular communication mechanism used by {\pkg} is determined by the platform, available libraries, and integration strategy.
-However, regardless of the communication mechanisms, the user will always use the same interface in the language of their model (See \S\ref{SSS:interface}).
+However, regardless of the communication mechanisms, the user will always use the same interface in the language of their model, simplifying the number of routines that users need to use for integration (see \S\ref{SSS:interface}).
 
@@
 IPC queues are not natively supported by Windows operating systems and do not allow communication between remote processes.
-In addition, IPC queues also have relatively low default message size limits on Mac OS X systems (2048 bytes).
-Once the queue is full or if the message is larger than the limit, any process attempting to send an additional message will block until a sufficient number of messages has been removed from the queue to accommodate the new message.
-For messages larger than the limit, the sending process will block indefinitely.
-This can be handled by splitting large messages into multiple smaller messages (See \S\ref{SSS:large}), however, the time required to send a message increases with the number of message it must be broken into (See \S\ref{SS:results_commtype}).
+In addition, IPC queues also have relatively low default message size limits on Mac OS X systems (2048 bytes or 256 64 bit numbers).
+Once the queue is full or if the message is larger than the limit, any process attempting to send an additional message will stop until a sufficient number of messages has been removed from the queue to accommodate the new message.
+For messages larger than the limit, the sending process will stop indefinitely.
+This can be handled by splitting large messages into multiple smaller messages (see \S\ref{SSS:large}); however, the time required to send a message increases with the number of message it must be broken into (See \S\ref{SS:results_commtype}).
 These limits make sending large messages relatively inefficient when compared with other communications mechanisms.
-As a result, IPC queues are used by {\pkg} as a fallback on Posix systems if none of the other supported communication libraries have not been installed.
+As a result, IPC queues are used by {\pkg} only as a fallback on Posix systems if none of the other supported communication libraries have not been installed.
 
@@
 Models do not block on sending messages to output channels; the model is free to continue working on its task while an \emph{output driver} waits for the message to be routed and received on a separate master process thread.
-Similarly, an \emph{input driver} continuously pings input channels on another thread, moving received messages into a buffer queue so that they are ready and waiting for the receiving model when it asks for input.
+Similarly, an \emph{input driver} continuously checks input channels on another thread, moving received messages into a intermediate buffer queue so that they are ready and waiting for the receiving model when it asks for input.
 These drivers work together to move message along from one model to the next like a conveyor belt.
@@
 compilation flags or data types), but they are beyond the scope of this limited introduction.
-Additional examples and descriptions of these options can be found in the documentation \footnote{{\tt https://cropsinsilico.github.io/cis\_interface/yaml.html}}.
+Additional examples and descriptions of these options can be found in the documentation \footnote{{\tt https://cropsinsilico.github.io/\pkglink/yaml.html}}.
 
@@
 Both models follow a similar pattern.
-They first receive static variables that will not change over the course of the run and then send the initial mass to output so that the output record is a complete history of the mass.
+They first receive ``static" variables that will not change over the course of the run and then send the initial mass to output so that the output record is a complete history of the mass.
 Then both entire a while loop that is only broken when there is not a new time step available.
@@
 In addition, outputs can also be provided with a format string that tells {\pkg} interface what data type to expect how the output should be formatted if it is sent to a file.
-Additional information about how these strings are processed can be found in the ``C-Style Format Strings" section of the documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/c\_format\_strings.html}}.
+Additional information about how these strings are processed can be found in the ``C-Style Format Strings" section of the documentation\footnote{{\tt https://cropsinsilico.github.io/\pkglink/c\_format\_strings.html}}.
 
@@
 There are many file formats that {\pkg} supports which have additional YAML options.
-Information about these file formats and their YAML options can be found in the ``Connection Options" subsection of the ``YAML Files" section of the documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/yaml.html\#connection-options}}.
+Information about these file formats and their YAML options can be found in the ``Connection Options" subsection of the ``YAML Files" section of the documentation\footnote{{\tt https://cropsinsilico.github.io/\pkglink/yaml.html\#connection-options}}.
 
@@
 In addition, work is underway to add generic support for any data type that can be expressed as a JSON object (see \S\ref{SS:json}).
-Examples using these different data types can be found in the ``Formatted I/O" section of the documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/formatted\_io.html}}.
+Examples using these different data types can be found in the ``Formatted I/O" section of the documentation\footnote{{\tt https://cropsinsilico.github.io/\pkglink/formatted\_io.html}}.
 \item[Model Networks] It is possible to build up complex integration networks using {\pkg} one connection at a time.
@@
 Except for Matlab models, the time required per message is not affected by the language of either model for smaller messages.
-There is a very small increase in the time per message when there is a Python model involved, but this is much lower than the standard deviation resulting from background activity on the test machine and so is in conclusive.
+There is a very small increase in the time per message when there is a Python model involved, but this is much lower than the standard deviation resulting from background activity on the test machine and so is inconclusive.
 Matlab models have a much higher time per message than any other language.
@@
 Jack2,][]{Magoules2018} requires explicit definition of the communication pattern in advance, prohibiting dynamic communication networks as discussed above.
-Work/data flow tools like Parsl \citep{babuji18} allow dynamic execution of processes along a dependency chain in parallel, but usually require addition code, are limited to passing data as files or data object native to the tool's implementation language, and are ill suited to models with persistent internal states or intermediate communication.
+Work/data flow tools like Parsl \citep{babuji18} allow dynamic execution of processes along a dependency chain in parallel, but usually require additional code, are limited to passing data as files or data object native to the tool's implementation language, and are ill-suited to models with persistent internal states or intermediate communication.
 {\pkg} is unique in that it offers users asynchronous communication and parallelism that is flexible enough to support any communication pattern, implicitly handles conversion between native data types in the supported languages, and does not require any additional time investment on the part of the user to achieve a speedup.
@@
 \section*{Code}\label{S:code}
-The {\pkg} package is available publicly on \href{https://github.com/cropsinsilico/cis_interface}{Github} and can be 
+The {\pkg} package is available publicly on \hrefgit{Github} and can be 
 installed using {\tt pip} or {\tt conda}.
 To ensure code health, {\pkg} boasts 100\% code coverage with automated testing on Linux, Mac OS X, and Windows for Python version 2.7, 3.4, 3.5, 3.6, and 3.7 via continuous integration with TravisCI \citep{travisci} and Appveyor \citep{appveyor}.
-Full documentation for {\pkg} can be found \href{https://cropsinsilico.github.io/cis_interface/}{here} including step-by-step directions from the tutorial conducted during the \href{https://cropsinsilico.github.io/cis_interface/hackathon2018/index.html}{2018 Crops in Silico Hackathon}.
-
-\todo{comment about 1.0 release}
+Full documentation for {\pkg} can be found \hrefdoc{here} including step-by-step directions from the tutorial conducted during the \hrefhack{2018 Crops in Silico Hackathon}.
+
+%Subsequent to publication of this manuscript, version 1.0 of {\pkg} will be released.
+%\todo{comment about 1.0 release}
 
@@
 %
-The authors would like to thank Mike Lambert and Craig Willis for their ongoing work on a graphical user interface for visually composing and running integration networks for {\pkg} as well as David Raila for his work on the original integration code that {\pkg} was based on.
+The authors would like to thank Mike Lambert and Craig Willis for their ongoing work on a graphical user interface for visually composing and running integration networks for {\pkg}, David Raila for his work on the original integration code that {\pkg} was based on, Amy Marshall-Colon for her leadership of the CiS project, and the referees for their feedback which have greatly strengthened this manuscript.
 
@@
 % COMMENTS
-\newpage
-\todo{To do:
-\begin{itemize}
-\item Replace links with those referencing new package name
-\item Forking example
-\item Transformation example
-\item Add JSON to documentation
-\item Add forking to documentation
-\item Add transformation to documentation
-\item Add schema to documentation
-\item Regenerate documentation
-\item Fix matlab autodocs for interface
-\item Check that final version of algorithms is split between pages correctly
-\end{itemize}
-}
+%\newpage
+%\todo{To do:
+%\begin{itemize}
+%\item Replace links with those referencing new package name
+%\item Forking example
+%\item Transformation example
+%\item Add JSON to documentation
+%\item Add forking to documentation
+%\item Add transformation to documentation
+%\item Add schema to documentation
+%\item Regenerate documentation
+%\item Fix matlab autodocs for interface
+%\item Check that final version of algorithms is split between pages correctly
+%\end{itemize}
+%}
 

### 2019-01-21 yggdrasil/communication/CommBase.py
Change schema names and fix some small errors created during initial name change

@@ -5,7 +5,7 @@ import atexit
 import threading
 from logging import info
 from yggdrasil import backwards, tools, serialize
-from yggdrasil.tools import get_CIS_MSG_MAX, CIS_MSG_EOF
+from yggdrasil.tools import get_YGG_MSG_MAX, YGG_MSG_EOF
 from yggdrasil.communication import (
     new_comm, get_comm, get_comm_class, determine_suffix)
 from yggdrasil.schema import register_component

### 2019-01-21 yggdrasil/metaschema/datatypes/MetaschemaType.py

@@ -6,7 +6,7 @@ import jsonschema
 from yggdrasil import backwards, tools
 from yggdrasil.metaschema import get_metaschema, get_validator
 from yggdrasil.metaschema.datatypes import (
-    MetaschemaTypeError, compare_schema, CIS_MSG_HEAD, get_type_class,
+    MetaschemaTypeError, compare_schema, YGG_MSG_HEAD, get_type_class,
     conversions)
 from yggdrasil.metaschema.properties import get_metaschema_property
 

### 2019-01-21 yggdrasil/communication/tests/test_CommBase.py
Change schema names and fix some small errors created during initial name change

@@ -1,6 +1,6 @@
 import os
 import uuid
-from yggdrasil.tests import CisTestClassInfo, assert_equal
+from yggdrasil.tests import YggTestClassInfo, assert_equal
 from yggdrasil.communication import new_comm, get_comm, CommBase
 
 
@@ -21,7 +21,7 @@ def test_registry():
     assert(not CommBase.unregister_comm(comm_class, key))
 
 
-class TestCommBase(CisTestClassInfo):
+class TestCommBase(YggTestClassInfo):
     r"""Tests for CommBase communication class.
 
     Attributes:

### 2019-01-21 yggdrasil/metaschema/datatypes/tests/__init__.py

@@ -40,7 +40,7 @@ def test_get_type_from_def():
 def test_guess_type_from_msg():
     r"""Test guess_type_from_msg."""
     assert_raises(ValueError, datatypes.guess_type_from_msg,
-                     b'fake message')
+                  b'fake message')
 
 
 def test_guess_type_from_obj():

### 2019-01-21 yggdrasil/metaschema/datatypes/tests/test_MetaschemaType.py

@@ -126,7 +126,7 @@ class TestMetaschemaType(YggTestClassInfo):
         r"""Test update_typedef raises error on non-matching typename."""
         self.instance.update_typedef(**self.typedef)
         assert_raises(MetaschemaTypeError, self.instance.update_typedef,
-                         type='invalid')
+                      type='invalid')
         if self._explicit:
             typedef_base = self.import_cls.typedef_fixed2base(self.typedef)
             self.instance.update_typedef(**typedef_base)

### 2019-01-21 papers/lang2018/ms.tex
Update text with new commands, minor bug fix in server test, todo with links in red

@@
 %\newcommand{\pkg}{{\tt \todo{cis\_interface}}{}}
-\newcommand{\pkgrun}{{\tt \todo{cisrun}}}
+\newcommand{\pkgrun}{{\tt yggrun}}
 \newcommand{\A}{Appendix{ }}
@@
 % SOURCE CODE
-\newcommand{\rootsrc}{\lstinputlisting[language=C,label={lst:src:root},caption={root.h: Source code for root model in C.}]{../../cis_interface/examples/root_to_shoot/src/root.h}}
-\newcommand{\shootsrc}{\lstinputlisting[language=Python,label={lst:src:shoot},caption={shoot.py: Source code for shoot model in Python.}]{../../cis_interface/examples/root_to_shoot/src/shoot.py}}
+\newcommand{\rootsrc}{\lstinputlisting[language=C,label={lst:src:root},caption={root.h: Source code for root model in C.}]{../../yggdrasil/examples/root_to_shoot/src/root.h}}
+\newcommand{\shootsrc}{\lstinputlisting[language=Python,label={lst:src:shoot},caption={shoot.py: Source code for shoot model in Python.}]{../../yggdrasil/examples/root_to_shoot/src/shoot.py}}
 
@@
 % WRAPPER CODE
-\newcommand{\rootwrap}{\lstinputlisting[language=C,label={lst:wrap:root},caption={root\_wrapper.c: Wrapper source code for root model in C.}]{../../cis_interface/examples/root_to_shoot/src/root_wrapper.c}}
-\newcommand{\shootwrap}{\lstinputlisting[language=Python,label={lst:wrap:shoot},caption={shoot\_wrapper.py: Wrapper source code for shoot model in Python}]{../../cis_interface/examples/root_to_shoot/src/shoot_wrapper.py}}
+\newcommand{\rootwrap}{\lstinputlisting[language=C,label={lst:wrap:root},caption={root\_wrapper.c: Wrapper source code for root model in C.}]{../../yggdrasil/examples/root_to_shoot/src/root_wrapper.c}}
+\newcommand{\shootwrap}{\lstinputlisting[language=Python,label={lst:wrap:shoot},caption={shoot\_wrapper.py: Wrapper source code for shoot model in Python}]{../../yggdrasil/examples/root_to_shoot/src/shoot_wrapper.py}}
 
@@
 % YAML FILES
-\newcommand{\rootyml}{\lstinputlisting[label={lst:yml:root},caption={root.yml: Model YAML specification file for root model.}]{../../cis_interface/examples/root_to_shoot/root.yml}}
-\newcommand{\shootyml}{\lstinputlisting[label={lst:yml:shoot},caption={shoot.yml: Model YAML specification file for shoot model.}]{../../cis_interface/examples/root_to_shoot/shoot.yml}}
-\newcommand{\rootymlfiles}{\lstinputlisting[label={lst:yml:root_files},caption={root\_files.yml: Connection YAML specification file for running the root model in isolation.}]{../../cis_interface/examples/root_to_shoot/root_files.yml}}
-\newcommand{\shootymlfiles}{\lstinputlisting[label={lst:yml:shoot_files},caption={shoot\_files.yml: Model YAML specification file for running the shoot model in isolation.}]{../../cis_interface/examples/root_to_shoot/shoot_files.yml}}
-\newcommand{\intyml}{\lstinputlisting[label={lst:yml:int},caption={root\_to\_shoot.yml: Connection YAML specification file for the root/shoot integration.}]{../../cis_interface/examples/root_to_shoot/root_to_shoot.yml}}
+\newcommand{\rootyml}{\lstinputlisting[label={lst:yml:root},caption={root.yml: Model YAML specification file for root model.}]{../../yggdrasil/examples/root_to_shoot/root.yml}}
+\newcommand{\shootyml}{\lstinputlisting[label={lst:yml:shoot},caption={shoot.yml: Model YAML specification file for shoot model.}]{../../yggdrasil/examples/root_to_shoot/shoot.yml}}
+\newcommand{\rootymlfiles}{\lstinputlisting[label={lst:yml:root_files},caption={root\_files.yml: Connection YAML specification file for running the root model in isolation.}]{../../yggdrasil/examples/root_to_shoot/root_files.yml}}
+\newcommand{\shootymlfiles}{\lstinputlisting[label={lst:yml:shoot_files},caption={shoot\_files.yml: Model YAML specification file for running the shoot model in isolation.}]{../../yggdrasil/examples/root_to_shoot/shoot_files.yml}}
+\newcommand{\intyml}{\lstinputlisting[label={lst:yml:int},caption={root\_to\_shoot.yml: Connection YAML specification file for the root/shoot integration.}]{../../yggdrasil/examples/root_to_shoot/root_to_shoot.yml}}
 
@@
 {\pkg} was developed as part of the Crops \emph{in Silico} \citep{Marshall-Colon2017} initiative to build a complete crop \emph{in silico} from the level of the genes to the level of the field.
-{\pkg} is available on Github\footnote{{\tt https://github.com/cropsinsilico/\pkglink}} with full documentation\footnote{{\tt https://cropsinsilico.github.io/\pkglink/}} and can be installed from PyPI\footnote{{\tt https://pypi.org/project/\pkglink/}} via {\tt pip install {\pkg}} or conda-forge\footnote{{\tt https://github.com/conda-forge/\pkglink-feedstock}} via {\tt conda -c conda-forge install {\pkg}}.
+{\pkg} is available on Github\footnote{{\tt https://github.com/cropsinsilico/\pkglink}} with full documentation\footnote{{\tt https://cropsinsilico.github.io/\pkglink/}} and can be installed from PyPI\footnote{{\tt https://pypi.org/project/\pkglink/}} via \todo{\tt pip install {\pkg}} or conda-forge\footnote{{\tt https://github.com/conda-forge/\pkglink-feedstock}} via \todo{\tt conda -c conda-forge install {\pkg}}.
 Although developed for the purpose of integrating plant biology models, {\pkg} can be applied to any situation requiring coordination between models in the supported languages.

### 2019-01-21 yggdrasil/drivers/tests/test_ServerDriver.py
Update text with new commands, minor bug fix in server test, todo with links in red

@@ -1,5 +1,5 @@
 import uuid
-from yggdrasill.tests import assert_raises, assert_equal
+from yggdrasil.tests import assert_raises, assert_equal
 import yggdrasil.drivers.tests.test_ConnectionDriver as parent
 from yggdrasil import runner, tools
 

### 2019-02-01 papers/lang2018/ms.tex
Change names but don't remove todo flags

@@
 \newcommand{\TB}{\,Tbytes}
+\newcommand{\pippkg}{\todo{\tt yggdrasil-framework}}
+\newcommand{\condapkg}{\todo{\tt yggdrasil}}
 \newcommand{\pkg}{{\tt yggdrasil}{}}
@@
 \newcommand{\A}{Appendix{ }}
-\newcommand{\hrefgit}[1]{\href{https://github.com/cropsinsilico/cis_interface}{\todo{#1}}}
-\newcommand{\hrefdoc}[1]{\href{https://cropsinsilico.github.io/cis_interface/}{\todo{#1}}}
-\newcommand{\hrefhack}[1]{\href{https://cropsinsilico.github.io/cis_interface/hackathon2018/index.html}{\todo{#1}}}
+\newcommand{\hrefgit}[1]{\href{https://github.com/cropsinsilico/yggdrasil}{\todo{#1}}}
+\newcommand{\hrefdoc}[1]{\href{https://cropsinsilico.github.io/yggdrasil/}{\todo{#1}}}
+\newcommand{\hrefhack}[1]{\href{https://cropsinsilico.github.io/yggdrasil/hackathon2018/index.html}{\todo{#1}}}
 
@@
 {\pkg} was developed as part of the Crops \emph{in Silico} \citep{Marshall-Colon2017} initiative to build a complete crop \emph{in silico} from the level of the genes to the level of the field.
-{\pkg} is available on Github\footnote{{\tt https://github.com/cropsinsilico/\pkglink}} with full documentation\footnote{{\tt https://cropsinsilico.github.io/\pkglink/}} and can be installed from PyPI\footnote{{\tt https://pypi.org/project/\pkglink/}} via \todo{\tt pip install {\pkg}} or conda-forge\footnote{{\tt https://github.com/conda-forge/\pkglink-feedstock}} via \todo{\tt conda -c conda-forge install {\pkg}}.
+{\pkg} is available on Github\footnote{{\tt https://github.com/cropsinsilico/\pkglink}} with full documentation\footnote{{\tt https://cropsinsilico.github.io/\pkglink/}} and can be installed from PyPI\footnote{{\tt https://pypi.org/project/\pkglink/}} via {\tt pip install {\pippkg}} or conda-forge\footnote{{\tt https://github.com/conda-forge/\pkglink-feedstock}} via {\tt conda -c conda-forge install {\condapkg}}.
 Although developed for the purpose of integrating plant biology models, {\pkg} can be applied to any situation requiring coordination between models in the supported languages.

### 2019-02-01 cis_interface/communication/tests/test_CommBase.py
Merge rmq test into base test for new merged case

@@ -203,6 +203,9 @@ class TestCommBase(CisTestClassInfo):
         close returns false."""
         self.send_instance.open()
         self.recv_instance.open()
+        if self.comm in ['RMQComm', 'RMQAsyncComm']:
+            self.send_instance.bind()
+            self.recv_instance.bind()
         self.send_instance.close()
         self.recv_instance.close()
         assert(self.send_instance.is_closed)

### 2019-02-01 cis_interface/communication/tests/test_RMQComm.py

@@ -14,12 +14,6 @@ class TestRMQComm(test_AsyncComm.TestAsyncComm):
     timeout = 10.0
     attr_list = (copy.deepcopy(test_AsyncComm.TestAsyncComm.attr_list)
                  + ['connection', 'channel'])
-    
-    def test_double_open(self):
-        r"""test that opening/binding twice dosn't cause errors."""
-        super(TestRMQComm, self).test_double_open()
-        self.send_instance.bind()
-        self.recv_instance.bind()
 
 
 @unittest.skipIf(_rmq_server_running, "RMQ Server running")

### 2019-02-01 cis_interface/serialize/DefaultSerialize.py
Add to list of metadata keywords that should be ignored.

@@ -403,9 +403,10 @@ class DefaultSerialize(tools.CisClass):
         # Remove metadata keywords unrelated to serialization
         # TODO: Find a better way of tracking these
         _remove_kws = ['body', 'address', 'size', 'id', 'incomplete', 'raw',
-                       'commtype', 'filetype',
+                       'commtype', 'filetype', 'response_address', 'request_id',
                        'append', 'in_temp', 'is_series', 'working_dir', 'fmts',
-                       'model_driver', 'env', 'send_converter', 'recv_converter']
+                       'model_driver', 'env', 'send_converter', 'recv_converter',
+                       'typedef_base']
         kws = list(kwargs.keys())
         for k in kws:
             if (k in _remove_kws) or k.startswith('zmq'):
@@ -422,6 +423,7 @@ class DefaultSerialize(tools.CisClass):
         # Update extra keywords
         if (len(kwargs) > 0):
             self.extra_kwargs.update(kwargs)
+            self.info("Extra kwargs: %s" % str(self.extra_kwargs))
         # Update type
         if not skip_type:
             # Update typedef from oldstyle keywords in extra_kwargs
@@ -559,7 +561,7 @@ class DefaultSerialize(tools.CisClass):
         self.initialize_from_message(args, **header_kwargs)
         metadata = {'no_metadata': no_metadata}
         if add_serializer_info:
-            self.info("serializer_info = %s", str(self.serializer_info))
+            self.debug("serializer_info = %s", str(self.serializer_info))
             metadata.update(self.serializer_info)
             metadata['typedef_base'] = self.typedef
         if header_kwargs is not None:

### 2019-02-02 papers/lang2018/ms.tex
Updated links

@@
 \newcommand{\TB}{\,Tbytes}
-\newcommand{\pippkg}{\todo{\tt yggdrasil-framework}}
-\newcommand{\condapkg}{\todo{\tt yggdrasil}}
+\newcommand{\pippkg}{{\tt yggdrasil-framework}}
+\newcommand{\condapkg}{{\tt yggdrasil}}
 \newcommand{\pkg}{{\tt yggdrasil}{}}
 \newcommand{\cis}{{\tt cis\_interface}{}}
-\newcommand{\pkglink}{\todo{cis\_interface}}
+\newcommand{\pkglink}{yggdrasil}
 %\newcommand{\pkg}{{\tt \todo{cis\_interface}}{}}
@@
 \newcommand{\A}{Appendix{ }}
-\newcommand{\hrefgit}[1]{\href{https://github.com/cropsinsilico/yggdrasil}{\todo{#1}}}
-\newcommand{\hrefdoc}[1]{\href{https://cropsinsilico.github.io/yggdrasil/}{\todo{#1}}}
-\newcommand{\hrefhack}[1]{\href{https://cropsinsilico.github.io/yggdrasil/hackathon2018/index.html}{\todo{#1}}}
+\newcommand{\hrefgit}[1]{\href{https://github.com/cropsinsilico/yggdrasil}{{#1}}}
+\newcommand{\hrefdoc}[1]{\href{https://cropsinsilico.github.io/yggdrasil/}{{#1}}}
+\newcommand{\hrefhack}[1]{\href{https://cropsinsilico.github.io/yggdrasil/hackathon2018/index.html}{{#1}}}
 
@@
 If an output channel is closed, a send call will return immediately with a flag indicating failure.
-Additional information about the interface calls for sending and receiving can be found in the ``Model Interface" section of the documentation\footnote{{\tt \todo{link}}}.
+Additional information about the interface calls for sending and receiving can be found in the ``Model Interface" section of the documentation\footnote{{\tt https://cropsinsilico.github.io/\pkglink/model\_interface.html}}.
 

### 2019-04-22 papers/lang2018/proof_corrections/listing2.py
Update to paper during proofs

@@ -1,0 +1,19 @@
+import time
+
+
+def calc_shoot_mass(r_s, dt, S_t, R_t, R_tp1):
+    r"""Calculate the shoot mass.
+
+    Args:
+        r_s (float): Relative shoot growth rate.
+        dt (float): The time step.
+        S_t (float): Previous shoot mass.
+        R_t (float): Previous root mass.
+        R_tp1 (float): Root mass at the next timestep.
+
+    Returns:
+        float: Shoot mass at the next timestep.
+
+    """
+    time.sleep(0.1)  # To simulate a longer calculation
+    return (S_t * r_s * dt) + S_t - (R_tp1 - R_t)

### 2019-04-22 papers/lang2018/proof_corrections/listing5.c

@@ -1,0 +1,73 @@
+#include <stdio.h>
+#include <stdlib.h>
+// Include C interface
+#include "YggInterface.h"
+// Include C header containing model calculation
+#include "root.h"
+
+
+int main(int argc, char *argv[]) {
+  int i, ret;
+  double r_r, dt, R_t, R_tp1;
+
+  // Create input/output channels
+  yggInput_t RootGrowthRate = yggInput("root_growth_rate");
+  yggInput_t InitRootMass = yggInput("init_root_mass");
+  yggInput_t TimeStep = yggInput("root_time_step");
+  yggOutput_t NextRootMass = yggOutputFmt("next_root_mass", "%lf\n");
+
+  // Receive root growth rate
+  ret = yggRecv(RootGrowthRate, &r_r);
+  if (ret < 0) {
+    printf("root: Error receiving root growth rate.\n");
+    return -1;
+  }
+  printf("root: Received root growth rate: %lf\n", r_r);
+
+  // Receive initial root mass
+  ret = yggRecv(InitRootMass, &R_t);
+  if (ret < 0) {
+    printf("root: Error receiving initial root mass.\n");
+    return -1;
+  }
+  printf("root: Received initial root mass: %lf\n", R_t);
+
+  // Send initial root mass
+  ret = yggSend(NextRootMass, R_t);
+  if (ret < 0) {
+    printf("root: Error sending initial root mass.\n");
+    return -1;
+  }
+
+  // Keep advancing until there arn't any new input times
+  i = 0;
+  while (1) {
+
+    // Receive the time step
+    ret = yggRecv(TimeStep, &dt);
+    if (ret < 0) {
+      printf("root: No more time steps.\n");
+      break;
+    }
+    printf("root: Received next time step: %lf\n", dt);
+
+    // Calculate root mass
+    R_tp1 = calc_root_mass(r_r, dt, R_t);
+    printf("root: Calculated next root mass: %lf\n", R_tp1);
+
+    // Output root mass
+    ret = yggSend(NextRootMass, R_tp1);
+    if (ret < 0) {
+      printf("root: Error sending root mass for timestep %d.\n", i+1);
+      return -1;
+    }
+
+    // Advance root mass to next timestep
+    R_t = R_tp1;
+    i++;
+
+  }
+
+  return 0;
+
+}

### 2019-04-22 papers/lang2018/proof_corrections/listing6.py

@@ -1,0 +1,67 @@
+# Import Python interface
+from yggdrasil.interface import YggInput, YggOutput
+# Import Python module containing model calculation
+from shoot import calc_shoot_mass
+
+
+# Create input/output channels
+ShootGrowthRate = YggInput('shoot_growth_rate')
+InitShootMass = YggInput('init_shoot_mass')
+TimeStep = YggInput('shoot_time_step')
+NextRootMass = YggInput('next_root_mass')
+NextShootMass = YggOutput('next_shoot_mass', '%lf\n')
+
+# Receive shoot growth rate
+flag, r_s = ShootGrowthRate.recv()
+if not flag:
+    raise RuntimeError('shoot: Error receiving shoot growth rate.')
+print('shoot: Received shoot growth rate: %s' % str(r_s))
+
+# Receive initial shoot mass
+flag, S_t = InitShootMass.recv()
+if not flag:
+    raise RuntimeError('shoot: Error receiving initial shootmass.')
+print('shoot: Received initial shoot mass: %s' % str(S_t))
+
+# Receive inital root mass
+flag, R_t = NextRootMass.recv()
+if not flag:
+    raise RuntimeError('shoot: Error receiving initial root mass.')
+print('shoot: Received initial root mass: %s' % str(R_t))
+
+# Send initial shoot mass
+flag = NextShootMass.send(S_t)
+if not flag:
+    raise RuntimeError('shoot: Error sending initial shoot mass.')
+
+# Keep advancing until there arn't any new input times
+i = 0
+while True:
+    
+    # Receive the time step
+    flag, dt = TimeStep.recv()
+    if not flag:
+        print('shoot: No more time steps.')
+        break
+    print('shoot: Received next time step: %s' % str(dt))
+
+    # Receive the next root mass
+    flag, R_tp1 = NextRootMass.recv()
+    if not flag:
+        # This raises an error because there must be a root mass for each time step
+        raise RuntimeError('shoot: Error receiving root mass for timestep %d.' % (i + 1))
+    print('shoot: Received next root mass: %s' % str(R_tp1))
+
+    # Calculate shoot mass
+    S_tp1 = calc_shoot_mass(r_s, dt, S_t, R_t, R_tp1)
+    print('shoot: Calculated next shoot mass: %s' % str(S_tp1))
+
+    # Output shoot mass
+    flag = NextShootMass.send(S_tp1)
+    if not flag:
[…]

### 2019-04-22 yggdrasil/drivers/CModelDriver.py
Initial work on automated file wrapping

@@ -172,8 +172,25 @@ class CModelDriver(CompiledModelDriver):
                       'external_dependencies': ['rapidjson'],
                       'include_dirs': []}}
     function_param = {
+        'import': '#include \"{filename}\"',
+        'interface': '#include \"{interface_library}\"',
+        'input': 'yggInput_t {channel} = yggInput(\"{channel_name}\");',
+        'output': 'yggOutput_t {channel} = yggOutput(\"{channel_name}\");',
+        'recv': '{flag_var} = yggRecv({channel}, {recv_var});',
+        'send': '{flag_var} = yggSend({channel}, {send_var});',
+        'flag_cond': '{flag_var} >= 0',
+        # Model functions should return non-zero integer codes to indicate errors
+        'function_call': '{flag_var} = {function_name}({input_var}, {output_var});',
+        'declare': '{type_name} {variable};',
+        'define': '{variable} = {value};',
         'comment': '//',
         'indent': 2 * ' ',
+        'quote': '\"',
+        'true': '1',
+        'break': 'break;',
+        'error': 'printf(\"{message}\"); return -1;',
+        'print': 'printf(\"{message}\");',
+        'fprintf': 'printf(\"{message}\", {variables});',
         'block_end': '}',
         'if_begin': 'if ({cond}) {',
         'for_begin': ('for ({iter_var} = {iter_begin}; {iter_var} < {iter_end}; '
@@ -244,3 +261,19 @@ class CModelDriver(CompiledModelDriver):
                 base = os.path.splitext(x)[0]
                 self.products += [base + ext for ext in ['.ilk', '.pdb', '.obj']]
         return out
+
+    @classmethod
+    def prepare_output_variables(cls, vars_list):
+        r"""Concatenate a set of output variables such that it can be passed as
+        a single string to the function_call parameter.
+
+        Args:
+            vars_list (list): List of variable names to concatenate as output
+                from a function call.
+
+        Returns:
+            str: Concatentated variables list.
+
+        """
+        return super(CModelDriver, cls).prepare_output_variables(
+            ['&' + x for x in vars_list])

### 2019-04-22 yggdrasil/drivers/CPPModelDriver.py

@@ -63,6 +63,11 @@ class CPPModelDriver(CModelDriver):
     default_compiler = None
     default_linker = None
     function_param = dict(CModelDriver.function_param,
+                          input='YggInput {channel}(\"{channel_name}\");',
+                          output='YggOutput {channel}(\"{channel_name}\");',
+                          recv='{flag_var} = {channel}.recv({recv_num}, {recv_var});',
+                          send='{flag_var} = {channel}.send({send_num}, {send_var});',
+                          error='error(\"{message}\");',
                           try_begin='try {',
                           try_error_type='const std::exception&',
                           try_except='} catch ({try_error} {error_var}) {')

### 2019-04-22 yggdrasil/drivers/MatlabModelDriver.py

@@ -429,8 +429,20 @@ class MatlabModelDriver(InterpretedModelDriver):  # pragma: matlab
                                  '-nojvm', '-batch']
     version_flags = ["fprintf('R%s', version('-release')); exit();"]
     function_param = {
+        'input': '{channel} = YggInterface(\'YggInput\', \'{channel_name}\');',
+        'output': '{channel} = YggInterface(\'YggOutput\', \'{channel_name}\');',
+        'recv': '[{flag_var}, {recv_var}] = {channel}.recv();',
+        'send': '{flag_var} = {channel}.send({send_var});',
+        'function_call': '{output_var} = {function_name}({input_var});',
+        'define': '{variable} = {value};',
         'comment': '%',
         'indent': 2 * ' ',
+        'quote': '\'',
+        'true': 'true',
+        'break': 'break;',
+        'error': 'error(\'{message}\');',
+        'print': 'disp(\'{message}\');',
+        'fprintf': 'fprintf(\'{message}\', {variables});',
         'block_end': 'end;',
         'if_begin': 'if ({cond})',
         'for_begin': 'for {iter_var} = {iter_begin}:{iter_end}',
@@ -825,3 +837,19 @@ class MatlabModelDriver(InterpretedModelDriver):  # pragma: matlab
                 + "'exit' call which will exit the MATLAB engine "
                 + "such that it cannot be reused. Please replace 'exit' "
                 + "with a return or error.")
+
+    @classmethod
+    def prepare_output_variables(cls, vars_list):
+        r"""Concatenate a set of output variables such that it can be passed as
+        a single string to the function_call parameter.
+
+        Args:
+            vars_list (list): List of variable names to concatenate as output
+                from a function call.
+
+        Returns:
+            str: Concatentated variables list.
+
+        """
+        return '[%s]' % super(MatlabModelDriver, cls).prepare_output_variables(
+            vars_list)

### 2019-04-22 yggdrasil/drivers/ModelDriver.py

@@ -831,6 +831,198 @@ class ModelDriver(Driver):
     #     self.debug('')
     #     self.kill_process()
     #     super(ModelDriver, self).do_terminate()
+
+    @classmethod
+    def get_mapped_type(cls, type_dict):
+        r"""Determine the native type that represents the type described by
+        the provided JSON schema.
+
+        Args:
+            type_dict (dict): JSON schema describing a type.
+
+        Returns:
+            str: The native type in the language of this class.
+
+        """
+        raise NotImplementedError("get_mapped_type undefined for language '%s'"
+                                  % cls.language)
+
+    @classmethod
+    def prepare_input_variables(cls, vars_list):
+        r"""Concatenate a set of input variables such that it can be passed as a
+        single string to the function_call parameter.
+
+        Args:
+            vars_list (list): List of variable names to concatenate as input to
+                a function call.
+
+        Returns:
+            str: Concatentated variables list.
+
+        """
+        return ', '.join(vars_list)
+
+    @classmethod
+    def prepare_output_variables(cls, vars_list):
+        r"""Concatenate a set of output variables such that it can be passed as
+        a single string to the function_call parameter.
+
+        Args:
+            vars_list (list): List of variable names to concatenate as output
+                from a function call.
+
+        Returns:
+            str: Concatentated variables list.
+
+        """
+        return ', '.join(vars_list)
+
+    @classmethod
+    def write_model_interface(cls, model_file, model_function,
+                              inputs=[], outputs=[],
+                              static_inputs=[], static_outputs=[]):
+        r"""Return the lines required to run a model looping over specified
+        variables.
+
+        Args:
+            model_file (str): Full path to the file containing the model
+                function.
+            model_function (str): Name of the model function.
+            inputs (list, optional): List of JSON documents defining inputs to
+                the model function. Defaults to [] and the model is assumed to
+                not take any input.
+            outputs (list, optional): List of JSON documents defining outputs
+                from the model function. Defaults to [] and the model is
+                assumed to not return any input.
+            static_inputs (list, optional): List of the names of input channels
+                that should only be recieved from once. Defaults to [] and is
+                ignored.
+            static_outputs (list, optional): List of the names of output channels
+                that should only be output to once. Defaults to [] and is
+                ignored.
+
+        Returns:
+            list: Lines of code wrapping a model function such that it is called
[…]

### 2019-04-22 yggdrasil/drivers/PythonModelDriver.py

@@ -25,8 +25,22 @@ class PythonModelDriver(InterpretedModelDriver):
         'zmq': {'libraries': ['zmq']},
         'rmq': {'libraries': ['pika']}}
     function_param = {
+        'import': 'import {function}',
+        'interface': 'import {interface_library} as ygg',
+        'input': '{channel} = ygg.YggInput(\'{channel_name}\')',
+        'output': '{channel} = ygg.YggOutput(\'{channel_name}\')',
+        'recv': '{flag_var}, {recv_var} = {channel}.recv()',
+        'send': '{flag_var} = {channel}.send({send_var})',
+        'function_call': '{output_var} = {function_name}({input_var})',
+        'define': '{variable} = {value}',
         'comment': '#',
         'indent': 4 * ' ',
+        'quote': '\"',
+        'true': 'True',
+        'break': 'break',
+        'error': 'raise Exception(\"{message}\")',
+        'print': 'print(\"{message}\")',
+        'fprintf': 'print(\"{message}\" % ({variables}))',
         'block_end': '',
         'if_begin': 'if ({cond}):',
         'for_begin': 'for {iter_var} in range({iter_begin}, {iter_end}):',
